### 2018-01-16 Paper/jaccardbaetica.tex
new paper

\documentclass[review]{elsarticle}
\usepackage{hyperref,lineno}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\modulolinenumbers[5]

\newcommand{\memo}[2]{\textcolor{#1}{#2}}
\newcommand{\maria}[1]{\memo{red}{#1\\}}
\newcommand{\revise}[1]{\memo{blue}{#1\\}}

\journal{Journal of Archaeological Science}

\bibliographystyle{model2-names.bst}\biboptions{authoryear}

\begin{document}

\section{Material and Methods}

\subsection{Baetica province}

hola

\end{document}

### 2018-01-16 Paper/jaccardbaetica.tex
add paper jaccard

@@
 
-\subsection{Baetica province}
+We analysed a dataset of 131 stamps from 81 workshops.
+All the dataset was collected from CEIPAC database (citar).
+The workshops were located in Baetica area along the river Guadalquivir and its tributary Genil.
 
-hola
+\subsection{Jaccard distance}
+%%usar Jaccard disimilarity and similarity
+
+The dataset was analysed using a statistic method as Jaccard index. Jaccard distance was used to calculate the distance among the stamps.
 

### 2018-01-17 Paper/jaccardbaetica.tex
update introduction

@@
 
+\section{Introduction}
+
+All the amphora stamps belong to Dressel 20 types.
+Dressel 20 was used to transport olive oil along the province during the Roman Empire.
+There is evidence that stamps were used from three centuries.
+(economía oleica betica).
+Frequently, stamps were marked mostly in handles but rarely in rims and body.
+The meaning of stamp is still under debate. Some authors suggest that they were identified as the owner of the olive land. Other authors propose that stamp could be the owner of the amphorae workshop.
+
+The information of the stamps is known as tria nomina (HABLAR SOBRE EL DEBATE DE LOS INDIVIDUOS DE ACEITE O YO QUE SE)and contend
+
 \section{Material and Methods}
@@
 All the dataset was collected from CEIPAC database (citar).
-The workshops were located in Baetica area along the river Guadalquivir and its tributary Genil.
+The workshops were located in different locations in Baetica province, along the river Guadalquivir and its tributary Genil.
+
+%los sellos fueron marcados a partir del siglo II d.C (la economía oleícola bética Remesal)
 
@@
 
-The dataset was analysed using a statistic method as Jaccard index. Jaccard distance was used to calculate the distance among the stamps.
+The dataset was analysed using a statistic method as Jaccard index. Jaccard distance was used to calculate the distance among different stamps.
 

### 2018-01-18 Paper/jaccardbaetica.tex
update method

@@
 Frequently, stamps were marked mostly in handles but rarely in rims and body.
-The meaning of stamp is still under debate. Some authors suggest that they were identified as the owner of the olive land. Other authors propose that stamp could be the owner of the amphorae workshop.
+The meaning of stamp is still under debate.
+Some authors suggest that they were identified as the owner of the olive land \citep{rodriguez_economioleicola_1977}.
+Other authors propose that stamp could be the owner of the amphorae workshop.
+In any case, the use of these stamps defined the system of working in the workshops.
 
 The information of the stamps is known as tria nomina (HABLAR SOBRE EL DEBATE DE LOS INDIVIDUOS DE ACEITE O YO QUE SE)and contend
+
+%habría que citar a Dressel
 
@@
 All the dataset was collected from CEIPAC database (citar).
-The workshops were located in different locations in Baetica province, along the river Guadalquivir and its tributary Genil.
+The workshops were located in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance.
+The chronology detected in the workshop is widely diverse from I-III B.C.
+However, all the workshops presented a long activity of production being difficult to specify a accurate chronology.
+
+%hablar de la cronologia
 
@@
 
-The dataset was analysed using a statistic method as Jaccard index. Jaccard distance was used to calculate the distance among different stamps.
+The dataset was analysed using a statistic method as Jaccard distance.
+This method allows to measure the similarity between sets.
+In our case, Jaccard distance was used to calculate the mutual presence of traits in the stamps.
+
+%should i include morisita o horn
+
+\subsection{Dissimilarity}
+
+This method is used to measure the overlap between different samples of sets. 
+%hablar sobre Horn (Morisita)
+
+\bibliography{bibliotex}
 

### 2018-01-19 Paper/jaccardbaetica.tex
material and method update

@@
 
-\journal{Journal of Archaeological Science}
+%\journal{Journal of Archaeological Science}
 
-\bibliographystyle{model2-names.bst}\biboptions{authoryear}
+%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
 
@@
 Frequently, stamps were marked mostly in handles but rarely in rims and body.
-The meaning of stamp is still under debate.
+The information of the stamps is known as \textit{tria nomina} (HABLAR SOBRE EL DEBATE DE LOS INDIVIDUOS DE ACEITE O YO QUE SE). \textit{Tria Nomina} was mostly formed for a code of three letters.
+
+The meaning of the amphora stamps is still under debate.
 Some authors suggest that they were identified as the owner of the olive land \citep{rodriguez_economioleicola_1977}.
@@
 In any case, the use of these stamps defined the system of working in the workshops.
-
-The information of the stamps is known as tria nomina (HABLAR SOBRE EL DEBATE DE LOS INDIVIDUOS DE ACEITE O YO QUE SE)and contend
-
 %habría que citar a Dressel
@@
 The dataset was analysed using a statistic method as Jaccard distance.
-This method allows to measure the similarity between sets.
+This method allows to measure the dissimilarity between sets.
 In our case, Jaccard distance was used to calculate the mutual presence of traits in the stamps.
+A comparison was done with the distance of the workshops to know whether there was an association between stamps and spatial distance.
 
-%should i include morisita o horn
+\subsection{Dissimilarity correlation}
 
-\subsection{Dissimilarity}
-
-This method is used to measure the overlap between different samples of sets. 
+This method was performed to measure the overlap between different samples of sets \citep{horn_measurement_1966}.
+Unlike Jaccard distance, Morisita-Horn variant can be used to handle large samples with different sizes and diversity.
+The index can be described  where 0 is no presence of overlap and 1 totally similarity.
 %hablar sobre Horn (Morisita)

### 2018-01-22 Paper/jaccardbaetica.tex
material and method 1st part

@@
 
-We analysed a dataset of 131 stamps from 81 workshops.
-All the dataset was collected from CEIPAC database (citar).
+We analysed a dataset of 990 stamps from 81 workshops.
+From the dataset, a sample of 131 different stamps were identified and collected from from CEIPAC database (citar).
 The workshops were located in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance.
-The chronology detected in the workshop is widely diverse from I-III B.C.
-However, all the workshops presented a long activity of production being difficult to specify a accurate chronology.
+The chronology in the workshops is widely diverse from I-III B.C.
+However, some stamps show a more specific chronology while the majority of them display a large activity of production being difficult to specify an accurate chronology.
+A reason can be that most of them were partially excavated focused on archaeological surveys to collect the maximum stamps as possible (CITAR)
 
-%hablar de la cronologia
-
-%los sellos fueron marcados a partir del siglo II d.C (la economía oleícola bética Remesal)
+%los sellos fueron marcados a partir del siglo II d.C (la economía oleícola bética Remesal) hablar de stams más concreto
 
@@
 Unlike Jaccard distance, Morisita-Horn variant can be used to handle large samples with different sizes and diversity.
+In our case, our dataset shows a sample not uniform so this method is useful when sets show a diversity amongst samples.
+
+%escribir un poco más 
 The index can be described  where 0 is no presence of overlap and 1 totally similarity. 
-%hablar sobre Horn (Morisita)
+%hablar sobre Horn (Morisita)hablar sobre que el sample no estaba uniforme por eso usamos el morisita horn
+
+\section{Results}
+
+\section{Discussion and Conclusion}
 

### 2018-01-25 Paper/jaccardbaetica.tex
add text dissimilarity

@@
 From the dataset, a sample of 131 different stamps were identified and collected from from CEIPAC database (citar).
-The workshops were located in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance.
+The workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance.
 The chronology in the workshops is widely diverse from I-III B.C.
@@
 The dataset was analysed using a statistic method as Jaccard distance.
-This method allows to measure the dissimilarity between sets.
-In our case, Jaccard distance was used to calculate the mutual presence of traits in the stamps.
-A comparison was done with the distance of the workshops to know whether there was an association between stamps and spatial distance.
+This method allows to measure the dissimilarity by calculating the presence of sets (CITAR).
+In our case, Jaccard distance was used to compute the mutual presence of traits in the amphora stamps but it does not consider the number of absences.
+A comparison was done with the distance of the workshops to identify whether there was an association between stamps and spatial distance amongst workshops.
+
+%indicar si se usa en archaeology el jaccard distance
 
@@
 
-This method was performed to measure the overlap between different samples of sets \citep{horn_measurement_1966}.
-Unlike Jaccard distance, Morisita-Horn variant can be used to handle large samples with different sizes and diversity.
-In our case, our dataset shows a sample not uniform so this method is useful when sets show a diversity amongst samples.
+As a complement of Jaccard distance, we use the statistical technique Morisita-Horn index.
+Unlike Jaccard distance, this method was performed to measure the overlap between different samples of sets \citep{horn_measurement_1966}.
+In ecology, it describes the dissimilarity between the system of two communities.
 
-%escribir un poco más 
-The index can be described  where 0 is no presence of overlap and 1 totally similarity. 
+Considering our dataset as non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity.
+Index can be describe considering 0 as no presence of similarity of stamps and 1 a totally similarity between stamps.
+
 %hablar sobre Horn (Morisita)hablar sobre que el sample no estaba uniforme por eso usamos el morisita horn
@@
 
+No similarity between stamps and amphora workshops were founded.
+
+\section{Acknowledgements}
+
+The research was funded by European Research Council Advanced Grant
+EPNet (340828).
+
 \bibliography{bibliotex}

### 2018-02-02 Paper/jaccardbaetica.tex
introduction

@@
 
-%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
+\bibliographystyle{model2-names.bst}\biboptions{authoryear}
 
@@
 \section{Introduction}
+
+Material culture allows us to understand a part of mechanism of production... (aquí creo que me repito como siempre)
+
+The economy of Roman Empire have been an object of study in the last centuries. (cambiar)
+
+This paper aims to highlight the production dynamics in relation to a specific area within Roman Empire. We explore here the economic relation between stamps and workshops.
+
+%this paper aims to highlight the production dynamic in connection with the proximity between amphora workshops and stamps.
 
@@
 Frequently, stamps were marked mostly in handles but rarely in rims and body.
-The information of the stamps is known as \textit{tria nomina} (HABLAR SOBRE EL DEBATE DE LOS INDIVIDUOS DE ACEITE O YO QUE SE). \textit{Tria Nomina} was mostly formed for a code of three letters.
+The information of the stamps is known as \textit{tria nomina} (HABLAR SOBRE EL DEBATE DE LOS INDIVIDUOS DE ACEITE O YO QUE SE).
+\textit{Tria Nomina} was mostly formed for a code three of letters.
+These letters can appear in a abbreviated form or complete \citep{berni_millet_amphora_1996}
 
@@
 Other authors propose that stamp could be the owner of the amphorae workshop.
-In any case, the use of these stamps defined the system of working in the workshops.
+In any case, the use of these stamps defined somehow the system of working in the workshops.
 %habría que citar a Dressel
@@
 
+\section{References}
+
 \bibliography{bibliotex}

### 2018-02-06 Paper/jaccardbaetica.tex
add morisita

@@
 
-This paper aims to highlight the production dynamics in relation to a specific area within Roman Empire. We explore here the economic relation between stamps and workshops.
+This paper aims to highlight the production dynamics in relation to a specific area within Roman Empire. We explore here the economic relationship between stamps and production centres.
 
@@
 Unlike Jaccard distance, this method was performed to measure the overlap between different samples of sets \citep{horn_measurement_1966}.
-In ecology, it describes the dissimilarity between the system of two communities.
+In Ecology, it describes the dissimilarity between the system of two communities.
 
 Considering our dataset as non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity.
-Index can be describe considering 0 as no presence of similarity of stamps and 1 a totally similarity between stamps.
+Morisita-Horn index can be expressed considering 0 as total presence of similarity of stamps and 1 a totally dissimilarity between stamps.
 
@@
 \section{Results}
+
+\subsection{Jaccard distance}
+
+Result of Jaccard distance can be The coefficients range from x to x.
+
+The dendrogram shown in Fig. Bla bla was obtained with the Jaccard distance measure. The dendrogram suggests
+
+\subsection{Dissimilarity correlation}
 

### 2018-02-08 Paper/jaccardbaetica.tex
update paper

@@
 
+\subsection{Case study}
+
+Our case study ....
 We analysed a dataset of 990 stamps from 81 workshops.
-From the dataset, a sample of 131 different stamps were identified and collected from from CEIPAC database (citar).
+A sample of 131 different stamps were identified and collected from CEIPAC database (citar).
 The workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance.
@@
 
-Considering our dataset as non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity.
+Considering our dataset as non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity \citep{wolda_similarity_1981}.
 Morisita-Horn index can be expressed considering 0 as total presence of similarity of stamps and 1 a totally dissimilarity between stamps.
@@
 
+There is not connection between stamps and amphora workshops
+
+There is not similar stamps in closest workshop
+
+Similar stamps were not found in closest amphorae workshops so it could be interpreted as no direct connection between stamps and the location of amphora workshops were found.
+
 \section{Acknowledgements}
@@
 The research was funded by European Research Council Advanced Grant
-EPNet (340828).
+EPNet (340828). We are grateful to 
+All data has been analysed and conducted in R program version 3.2.4, using vegan package (citar)
 

### 2018-02-16 Paper/jaccardbaetica.tex
update case study
last update

@@
 
-Our case study ....
+Our case study examines the relation between the distribution of amphoric stamps and the workshops.
 We analysed a dataset of 990 stamps from 81 workshops.
@@
 
-\subsection{Jaccard distance}
+%%\subsection{Jaccard distance}
 %%usar Jaccard disimilarity and similarity
-
-The dataset was analysed using a statistic method as Jaccard distance.
+%%
+%The dataset was analysed using a statistic method as Jaccard distance.
 This method allows to measure the dissimilarity by calculating the presence of sets (CITAR).

### 2018-02-22 Paper/jaccardbaetica.tex
change horn-morisita

@@
 
+\begin{frontmatter}
+
+\begin{abstract}
+
+The aim of this study is to detect the patterns of olive oil production that link amphora workshops and amphoric stamps.
+Roman provinces such as Baetica became important production and distribution centers during the Roman Empire.
+However, it remains under debate how this province was organized and whether it is possible to identify patterns in the olive oil market.
+
+Our case of study has been focused on the production processes located in Baetica province (currently Andalusia) from 1st to 3rd AD.
+In particular, we want to explore economic dynamics that include the production and distribution of olive oil trade.
+Amphoric stamps are used to identify the presence of different producer groups that might share similar stamps.
+To achieve this goal, we analyse a set of stamps from different workshops in Baetica province in order to detect a relation between the distribution of amphoric stamps and the economic structure in this province.
+Here we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance.
+
+The analysis explores how quantitative approach provides a useful tool for the interpretation of the economic processes.
+Finally, results pretend to highlight the organization of olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.
+
+\end{abstract}
+
+\end{frontmatter}
+
 \section{Introduction}
@@
 
-This paper aims to highlight the production dynamics in relation to a specific area within Roman Empire. We explore here the economic relationship between stamps and production centres.
+This paper aims to highlight the production dynamics in relation to a specific area within Roman Empire.
+We want to detect the pattern of olive oil production that link amphora workshops and amphoric stamps.
+We focus here on exploring the economic relation between stamps and amphora production centres.
+
+Roman provinces such as Baetica became important production and distribution centers during the Roman Empire.
+However, it remains under debate how this province was organized and whether it is possible to identify patterns in the olive oil market.
 
@@
 %%\subsection{Jaccard distance}
-%%usar Jaccard disimilarity and similarity
-%%
 %The dataset was analysed using a statistic method as Jaccard distance.
@@
 
-%indicar si se usa en archaeology el jaccard distance
-
 \subsection{Dissimilarity correlation}
 
-As a complement of Jaccard distance, we use the statistical technique Morisita-Horn index.
-Unlike Jaccard distance, this method was performed to measure the overlap between different samples of sets \citep{horn_measurement_1966}.
+We use the statistical technique Morisita-Horn index.
+This method was performed to measure the overlap between different samples of sets \citep{horn_measurement_1966}.
 In Ecology, it describes the dissimilarity between the system of two communities.
@@
 Morisita-Horn index can be expressed considering 0 as total presence of similarity of stamps and 1 a totally dissimilarity between stamps.
+In our case, it will be calculated the number of times that one stamp appear in a amphora workshop.
+This method allows to bear in mind the similar number of times for each repeated stamp per workshop.
 
@@
 
-\subsection{Jaccard distance}
+%%\subsection{Jaccard distance}
 
-Result of Jaccard distance can be The coefficients range from x to x.
-
-The dendrogram shown in Fig. Bla bla was obtained with the Jaccard distance measure. The dendrogram suggests
+%Result of Jaccard distance can be The coefficients range from x to x.The dendrogram shown in Fig. Bla bla was obtained with the Jaccard distance measure. The dendrogram suggests
 

### 2018-02-23 Paper/jaccardbaetica.tex
update material

@@
 The meaning of the amphora stamps is still under debate.
-Some authors suggest that they were identified as the owner of the olive land \citep{rodriguez_economioleicola_1977}.
-Other authors propose that stamp could be the owner of the amphorae workshop.
+Some authors suggest that they were identified as the land-owners of the olive groves \citep{rodriguez_economioleicola_1977}.
+Other authors propose that stamp could be the owners of the making-amphorae workshop.
 In any case, the use of these stamps defined somehow the system of working in the workshops.
 %habría que citar a Dressel
+%olive producing
 
@@
 The workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance.
+
 The chronology in the workshops is widely diverse from I-III B.C.
 However, some stamps show a more specific chronology while the majority of them display a large activity of production being difficult to specify an accurate chronology.
-A reason can be that most of them were partially excavated focused on archaeological surveys to collect the maximum stamps as possible (CITAR)
+A reason can be that most of them were partially excavated and only focused on archaeological surveys to collect the maximum stamps as possible (CITAR)
 
@@
 
+A measure of dissimilarity has been chosen to analyse the dataset.
 We use the statistical technique Morisita-Horn index.
@@
 %Result of Jaccard distance can be The coefficients range from x to x.The dendrogram shown in Fig. Bla bla was obtained with the Jaccard distance measure. The dendrogram suggests
-
-\subsection{Dissimilarity correlation}
 

### 2018-02-27 Paper/jaccardbaetica.tex
update material and method

@@
 Dressel 20 was used to transport olive oil along the province during the Roman Empire.
+Most Dressel 20 were marked in stamps and tituli picti and inked in graffities with different information but there is not a general consensus about the meaning of them \citep{rodriguez_baetican_1998}.
+Stamps are the most studied in this type of amphorae.
 There is evidence that stamps were used from three centuries.
 (economía oleica betica).
-Frequently, stamps were marked mostly in handles but rarely in rims and body.
+Frequently, stamps were marked mainly in handles but rarely in rims and body.
+
 The information of the stamps is known as \textit{tria nomina} (HABLAR SOBRE EL DEBATE DE LOS INDIVIDUOS DE ACEITE O YO QUE SE).
@@
 
+%quizás habría que hablar también del filtrado de códigos que se hizo en python porque antes había 3783 stamps pero se filtró por el número de letras...
+
+The approach proposed here is based on the idea of measuring the similarity between amphora workshops by quantifying similar stamps.
 A measure of dissimilarity has been chosen to analyse the dataset.
 We use the statistical technique Morisita-Horn index.
-This method was performed to measure the overlap between different samples of sets \citep{horn_measurement_1966}.
+This method was performed to measure the overlap between different samples of sets \citep{morisita_measuring_1959, horn_measurement_1966}.
 In Ecology, it describes the dissimilarity between the system of two communities.
@@
 This method allows to bear in mind the similar number of times for each repeated stamp per workshop.
+%quizas ampliar con el método de xavielin
 

### 2018-02-28 Paper/jaccardbaetica.tex
update conclusion

@@
 Our case study examines the relation between the distribution of amphoric stamps and the workshops.
-We analysed a dataset of 990 stamps from 81 workshops.
-A sample of 131 different stamps were identified and collected from CEIPAC database (citar).
+We studied a dataset of 3787 stamps collected from CEIPAC database (CITAR).
+However, the 70 \% of stamps cannot be tested due to fragmentation or imcomplete information.
+Consequently, we discard integrate the fragmented stamps in our dataset.
+
+We finally filter a total sample of 990 stamps composed by 131 different stamps from 81 workshops.
 The workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance.
@@
 
-\subsection{Dissimilarity correlation}
+\subsection{Quantifying the diversity Dissimilarity correlation}
 
@@
 
+The analysis shows that amphoric stamps are not correlated with spatial distance. A matrix distance of similarity between workshops can be seen in Fig.1. The matrix displays
+
+Morisita-Horn dissimilarities
+
+A dendrogram was generated
+
 %%\subsection{Jaccard distance}
@@
 
-No similarity between stamps and amphora workshops were founded.
+No strong similarity between stamps and amphora workshops were found.
 
@@
 
+It could be due to several reason.
+On the one hand, the use of this amphoric stamps were exclusively running by the owner or family to distinguish the amphora workshop (CITAR).
+On the other, it could be somehow a batch sistematic organization to prepare and distribute the commodity, considering that Dressel 20 was not marked in several cases.
+
+just 30-40 \% of amphora Dressel 20 was marked 
+An any case, 
+Considering,
+
 \section{Acknowledgements}
 
-The research was funded by European Research Council Advanced Grant
-EPNet (340828). We are grateful to 
+The research was funded by European Research Council Advanced Grant EPNet (340828). We are grateful to 
 All data has been analysed and conducted in R program version 3.2.4, using vegan package (citar)

### 2018-03-01 Paper/jaccardbaetica.tex
add some bib

@@
 
-All the amphora stamps belong to Dressel 20 types.
-Dressel 20 was used to transport olive oil along the province during the Roman Empire.
+All the amphora stamps belong to Dressel 20 types \citep{martin-kilcher_romischen_1994}.
+Dressel 20 was used to transport olive oil along the province during the Roman Empire \citep{berni_millet_epigrafianforica_2008}.
 Most Dressel 20 were marked in stamps and tituli picti and inked in graffities with different information but there is not a general consensus about the meaning of them \citep{rodriguez_baetican_1998}.

### 2018-03-01 Jaccard distance/codeCoto.R
add some lines

@@ -19,6 +19,7 @@ plot(hclust(baeDistance))
 ##plot matrix
 
 library(reshape)
+library(ggplot2)
 roo <- melt(as.matrix(baeDistance))
 ggplot(roo, aes(x=X1, y=X2, fill=value, label=round(value, 2))) + geom_raster() + geom_text() #always use X1, X2
 
@@ -26,3 +27,15 @@ ggplot(roo, aes(x=X1, y=X2, fill=value, label=round(value, 2))) + geom_raster() 
 pdf('resultadosnew.pdf', width=15, height=8)    
 plot(hclust(baeDistance))
 dev.off()
+
+#print the matrix
+pdf('resultadosmatrix.pdf', width=35, height=10) 
+ggplot(roo, aes(x=X1, y=X2, fill=value, label=round(value, 2))) + geom_raster() + geom_text()
+dev.off()
+
+#para contar de que datos se disponen
+
+count(mydata$site) 
+names(mydata) #para saber el nombre de los valores (columnas)
+
+

### 2018-03-02 Paper/jaccardbaetica.tex
add new changes
modify introduction

@@
 All the amphora stamps belong to Dressel 20 types \citep{martin-kilcher_romischen_1994}.
-Dressel 20 was used to transport olive oil along the province during the Roman Empire \citep{berni_millet_epigrafianforica_2008}.
-Most Dressel 20 were marked in stamps and tituli picti and inked in graffities with different information but there is not a general consensus about the meaning of them \citep{rodriguez_baetican_1998}.
+Dressel 20 is commonly associated with transportation of Baetican olive oil through the provinces during the Roman Empire \citep{berni_millet_epigrafianforica_2008}.
+Most Dressel 20 were marked in stamps and \textit{tituli picti} and inked in graffities with different information but there is not a general consensus about the meaning of them \citep{rodriguez_baetican_1998}.
 Stamps are the most studied in this type of amphorae.
-There is evidence that stamps were used from three centuries.
+There is evidence that stamps were used for almost three centuries.
 (economía oleica betica).
@@
 Some authors suggest that they were identified as the land-owners of the olive groves \citep{rodriguez_economioleicola_1977}.
-Other authors propose that stamp could be the owners of the making-amphorae workshop.
+Other authors propose that stamp could be the owners of the making-amphorae workshop (CITAR) or even a production counting system \citep{berni_millet_epigrafianforica_2008}.
 In any case, the use of these stamps defined somehow the system of working in the workshops.
@@
 
-The analysis shows that amphoric stamps are not correlated with spatial distance. A matrix distance of similarity between workshops can be seen in Fig.1. The matrix displays
+The analysis shows that amphoric stamps are not correlated with spatial distance.
+
+A matrix distance of similarity between workshops can be seen in Fig.1. The matrix displays
 

### 2018-03-05 Paper/jaccardbaetica.tex
modify material and method

@@
 
-All the amphora stamps belong to Dressel 20 types \citep{martin-kilcher_romischen_1994}.
+All the amphora stamps belong to Dressel 20 types \citep{dressel,martin-kilcher_romischen_1994}.
 Dressel 20 is commonly associated with transportation of Baetican olive oil through the provinces during the Roman Empire \citep{berni_millet_epigrafianforica_2008}.
@@
 
-The information of the stamps is known as \textit{tria nomina} (HABLAR SOBRE EL DEBATE DE LOS INDIVIDUOS DE ACEITE O YO QUE SE).
-\textit{Tria Nomina} was mostly formed for a code three of letters.
-These letters can appear in a abbreviated form or complete \citep{berni_millet_amphora_1996}
+The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criteria (CITAR).
+Stamps was mostly formed for a code of three letters.
+There letters can appear in a abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
 
@@
 Our case study examines the relation between the distribution of amphoric stamps and the workshops.
-We studied a dataset of 3787 stamps collected from CEIPAC database (CITAR).
-However, the 70 \% of stamps cannot be tested due to fragmentation or imcomplete information.
-Consequently, we discard integrate the fragmented stamps in our dataset.
-
-We finally filter a total sample of 990 stamps composed by 131 different stamps from 81 workshops.
+We studied a dataset of 3787 stamps collected from different amphora workshops in Baetica provice (see CEIPAC database).
 The workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance.
 
-The chronology in the workshops is widely diverse from I-III B.C.
+However, the 70 \% of stamps cannot be tested due to fragmentation or incomplete information.
+Consequently, we discard integrate the fragmented stamps in our dataset.
+We finally filter a total sample of 990 stamps composed by 131 different stamps from 81 workshops.
+
+The chronology in the workshops is widely diverse from the first to the third centuries AD .
 However, some stamps show a more specific chronology while the majority of them display a large activity of production being difficult to specify an accurate chronology.
-A reason can be that most of them were partially excavated and only focused on archaeological surveys to collect the maximum stamps as possible (CITAR)
+A reason may be that most of them were partially excavated and only focused on archaeological surveys to collect the maximum stamps as possible (CITAR)
 
@@
 On the one hand, the use of this amphoric stamps were exclusively running by the owner or family to distinguish the amphora workshop (CITAR).
-On the other, it could be somehow a batch sistematic organization to prepare and distribute the commodity, considering that Dressel 20 was not marked in several cases.
+On the other, it could be somehow a batch systematic organization to prepare and distribute the commodity, considering that Dressel 20 was not marked in several cases.
 

### 2018-03-06 Paper/jaccardbaetica.tex
change methods and results

@@
 The analysis explores how quantitative approach provides a useful tool for the interpretation of the economic processes.
-Finally, results pretend to highlight the organization of olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.
+Finally, results pretend to highlight the organization of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.
 
@@
 
+Here, therefore, this study aim to explore the effect of Baetican olive oil production in the Roman Empire
+of production organization over Baetican olive oil production.
+
 \section{Material and Methods}
@@
 Consequently, we discard integrate the fragmented stamps in our dataset.
-We finally filter a total sample of 990 stamps composed by 131 different stamps from 81 workshops.
+We finally filter a total sample of 987 stamps composed by 131 different stamps from 81 workshops.
 
@@
 We use the statistical technique Morisita-Horn index.
+The formula can be described as follows
+
+(FORMULA Y DESCRIPCIÓN)
+
 This method was performed to measure the overlap between different samples of sets \citep{morisita_measuring_1959, horn_measurement_1966}.
@@
 A dendrogram was generated
+
+Some closest workshops show a similarity on the stamps such as BLA BLA BLA . This could be due to several reasons.
+
+doesn't support
+
+As summary, these results
 

### 2018-03-06 Jaccard distance/codeCoto.R
deleted ILEGIBLE stamps

@@ -14,6 +14,7 @@ baeDistance = vegdist(baeMatrix, method="horn")
 #baeDistance = vegdist(baeMatrix, method= "horn", na.rm= TRUE) na.rm when data is missing. Read ?vedist
 baeDistance
 hclust(baeDistance)
+#use summary to see the results
 plot(hclust(baeDistance))
 
 ##plot matrix

### 2018-03-07 Paper/jaccardbaetica.tex
update results' git status

@@
 
-The analysis shows that amphoric stamps are not correlated with spatial distance.
+The analysis shows that amphoric stamps could not be correlated with spatial distance.
 
-A matrix distance of similarity between workshops can be seen in Fig.1. The matrix displays
-
-Morisita-Horn dissimilarities
-
-A dendrogram was generated
-
-Some closest workshops show a similarity on the stamps such as BLA BLA BLA . This could be due to several reasons.
-
-doesn't support
-
-As summary, these results
+The correlation coefficients range from a minimum to a maximum.
+The dendrogram shown in Fig.
+2 was obtained with Morisita-Horn index.
+This dendrogram suggests that amphora workshops used different stamps for their prodution system.
+Accordingly only few closer workshops show a similarity on the stamps while most of them seem to display different stamps roles.
 
@@
 
-No strong similarity between stamps and amphora workshops were found.
+No strong relation between stamps and spatial distance have been detected in amphora workshops.  
+The analysis suggests that there is not connection between stamps and the same amphora workshops so similar stamps between closer amphora workshops were not found, excluding certain exception.
+Consequently the majority of stamps are located in different amphora workshops.
 
-There is not connection between stamps and amphora workshops
+The hypothesis about groups of amphora workshops sharing the same stamps seems not match with the results of analysis.
+Rather it could be interpreted due to several reason.
+On the one hand, the use of this amphoric stamps were exclusively running by the owner or family to distinguish the amphora workshop (CITAR).
+(EXPLICAR)
 
-There is not similar stamps in closest workshop
+On the other, it could be somehow a batch systematic organization to prepare and distribute the commodity, considering that Dressel 20 was not marked in several cases (CITAR).
+(EXPLICAR)
 
-Similar stamps were not found in closest amphorae workshops so it could be interpreted as no direct connection between stamps and the location of amphora workshops were found.
+%Neither it seems a group of amphora workshops sharing the same stamps.
 
-It could be due to several reason.
-On the one hand, the use of this amphoric stamps were exclusively running by the owner or family to distinguish the amphora workshop (CITAR).
-On the other, it could be somehow a batch systematic organization to prepare and distribute the commodity, considering that Dressel 20 was not marked in several cases.
+An any case,
 
-just 30-40 \% of amphora Dressel 20 was marked 
-An any case, 
-Considering,
+As summary, this method presented here highlights
+
+%Dressel 20 marked in the 30-40\%
 

### 2018-03-12 Paper/jaccardbaetica.tex
update results and dis
change in conclusion
add new lines conclusion

@@
 However, it remains under debate how this province was organized and whether it is possible to identify patterns in the olive oil market.
+
+CAMBIAR
 
@@
 The dendrogram shown in Fig.
-2 was obtained with Morisita-Horn index.
+2 \ref{dendro} was obtained with Morisita-Horn index.
 This dendrogram suggests that amphora workshops used different stamps for their prodution system.
 Accordingly only few closer workshops show a similarity on the stamps while most of them seem to display different stamps roles.
+
+\begin{figure}[htp]
+	\centering
+\includegraphics[width=\linewidth]{figs/dendro}
+\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in Baetica area}
+\label{dendro}
+\end{figure}
 
@@
 No strong relation between stamps and spatial distance have been detected in amphora workshops.  
-The analysis suggests that there is not connection between stamps and the same amphora workshops so similar stamps between closer amphora workshops were not found, excluding certain exception.
-Consequently the majority of stamps are located in different amphora workshops.
+Indeed, the analysis suggests that there is not connection between stamps and the same amphora workshops, excluding certain exceptions.
+Consequently the majority of stamps are located in different amphora workshops and only few similar stamps between closer amphora workshops were found.
 
-The hypothesis about groups of amphora workshops sharing the same stamps seems not match with the results of analysis.
-Rather it could be interpreted due to several reason.
-On the one hand, the use of this amphoric stamps were exclusively running by the owner or family to distinguish the amphora workshop (CITAR).
-(EXPLICAR)
+The hypothesis about groups of amphora workshops sharing the same stamps seems not match with the results of analysis even though there are shareable stamps in closer workshops.
+Probably same stamps detected in closer amphora workshop could have being wrongly catalogued twice in different way in spite of belonging to the same place or being closer between each other.
 
-On the other, it could be somehow a batch systematic organization to prepare and distribute the commodity, considering that Dressel 20 was not marked in several cases (CITAR).
-(EXPLICAR)
+%los que aparecen cerca son porque están cerca o puede haber confusión entre talleres
 
-%Neither it seems a group of amphora workshops sharing the same stamps.
+By contrast, we have located different stamps in a same amphora workshop.
+Neither it seems that stamps were used to specify the quality or value of the product \citep{callender}.
+Rather our results could be interpreted due to several reason according to previous hypotheses.
+On the one hand, the use of these amphoric stamps could be exclusively running by the owner or family to distinguish the amphora workshop (CITAR).
+
+(aumentar)
+Considering that Dressel 20 type was not marked in several cases (CITAR) it can be interpreted somehow a batch systematic organization to prepare and distribute the commodity (CITAR)
+
+it could be somehow a batch systematic organization to prepare and distribute the commodity, considering that Dressel 20 was not marked in several cases. This method was used of counting BLA BAL
 

### 2018-03-13 Paper/jaccardbaetica.tex
updating

@@
 
+%hablar un poco del comercio del aceite de oliva quizás citando el paper de Williams, Peacok y mattlingly el colega
+
 %this paper aims to highlight the production dynamic in connection with the proximity between amphora workshops and stamps.
@@
 (economía oleica betica).
-Frequently, stamps were marked mainly in handles but rarely in rims and body.
+Frequently, stamps were marked mainly in handles but rarely in rims and body \citep{millet_anforas_1998}
 
@@
 
-However, the 70 \% of stamps cannot be tested due to fragmentation or incomplete information.
+However, approximately the 70 \% of stamps cannot be tested due to fragmentation or incomplete information.
 Consequently, we discard integrate the fragmented stamps in our dataset.
@@
 
-The chronology in the workshops is widely diverse from the first to the third centuries AD .
+The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{garcia_vargas_enrique_formal_2010}.
 However, some stamps show a more specific chronology while the majority of them display a large activity of production being difficult to specify an accurate chronology.
@@
 The hypothesis about groups of amphora workshops sharing the same stamps seems not match with the results of analysis even though there are shareable stamps in closer workshops.
-Probably same stamps detected in closer amphora workshop could have being wrongly catalogued twice in different way in spite of belonging to the same place or being closer between each other.
+Probably same stamps detected in closer amphora workshop could have been wrongly catalogued twice in different way in spite of belonging to the same place or being closer between each other.
 

### 2018-03-15 Paper/jaccardbaetica.tex
add new changes bib
add new changes conclusion

@@
 %hablar un poco del comercio del aceite de oliva quizás citando el paper de Williams, Peacok y mattlingly el colega
+%hablar de las tres áreas de producción: hispalis, corduba y astigi
 
@@
 %olive producing
+
+Aquí poner algo sobre esto \citep{remesal_anforas_2004}
 
@@
 Consequently the majority of stamps are located in different amphora workshops and only few similar stamps between closer amphora workshops were found.
+In any case, our results show that most similar amphoric stamps were detected in closer workshops and not in other areas.
+These stamps tend to share the same area of production but there is not relation of similarity in general.
 
-The hypothesis about groups of amphora workshops sharing the same stamps seems not match with the results of analysis even though there are shareable stamps in closer workshops.
-Probably same stamps detected in closer amphora workshop could have been wrongly catalogued twice in different way in spite of belonging to the same place or being closer between each other.
-
-%los que aparecen cerca son porque están cerca o puede haber confusión entre talleres
+The hypothesis about groups of amphora workshops sharing the same stamps seems not match with the results of analysis even though there are similar stamps in closer workshops.
+Rather the stamps detected in closer amphorae workshops seem not move from other farthest workshops.
+Additionally they could have been "wrongly" catalogued twice in different mode in spite of belonging to the same place or being closer between each other.
+%los que aparecen cerca son porque están cerca o puede haber confusion entre talleres
 
@@
 Neither it seems that stamps were used to specify the quality or value of the product \citep{callender}.
-Rather our results could be interpreted due to several reason according to previous hypotheses.
-On the one hand, the use of these amphoric stamps could be exclusively running by the owner or family to distinguish the amphora workshop (CITAR).
+Probably our results could be interpreted due to several reason according to previous hypotheses.
+The use of these amphoric stamps could be exclusively running by the owner or family to distinguish the amphora workshop (CITAR).
 
 (aumentar)
-Considering that Dressel 20 type was not marked in several cases (CITAR) it can be interpreted somehow a batch systematic organization to prepare and distribute the commodity (CITAR)
+On the other hand, it can be interpreted somehow a batch systematic organization to prepare and distribute the commodity considering that Dressel 20 type was not marked in several cases (CITAR)
 
@@
 
-As summary, this method presented here highlights
+As summary, this method presented here provides a potential tool to understand mechanisms of production based on the similarity of artefacts.  This method have identified
+
+highlight
 
@@
 
-The research was funded by European Research Council Advanced Grant EPNet (340828). We are grateful to 
+The research was funded by European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon. 
 All data has been analysed and conducted in R program version 3.2.4, using vegan package (citar)

### 2018-03-16 Paper/jaccardbaetica.tex
add new lines in conclusion

@@
 Indeed, the analysis suggests that there is not connection between stamps and the same amphora workshops, excluding certain exceptions.
-Consequently the majority of stamps are located in different amphora workshops and only few similar stamps between closer amphora workshops were found.
+Consequently the majority of stamps are located in different amphora workshops and only similar stamps between closer amphora workshops were found.
 In any case, our results show that most similar amphoric stamps were detected in closer workshops and not in other areas.
-These stamps tend to share the same area of production but there is not relation of similarity in general.
+These stamps tend to share the same area of production but there is not a general relation between groups of amphora workshops and area.
 
@@
 Rather the stamps detected in closer amphorae workshops seem not move from other farthest workshops.
-Additionally they could have been "wrongly" catalogued twice in different mode in spite of belonging to the same place or being closer between each other.
+Additionally they could have been catalogued with different names by authors (METER PONSICH Y DRESSEL) in spite of belonging to the same place or being closer between each other.
+We also identify different stamps in other closer workshops.
 %los que aparecen cerca son porque están cerca o puede haber confusion entre talleres
@@
 Probably our results could be interpreted due to several reason according to previous hypotheses.
-The use of these amphoric stamps could be exclusively running by the owner or family to distinguish the amphora workshop (CITAR).
+The use of these amphoric stamps could be exclusively running by the owner or families to distinguish the amphora workshop (CITAR).
 
@@
 
-As summary, this method presented here provides a potential tool to understand mechanisms of production based on the similarity of artefacts.  This method have identified
-
-highlight
+As summary, this method presented here provides a potential tool to understand mechanisms of production based on the similarity of artefacts.
+This method have identified differences in the case of the amphoric production within Roman Empire.
+Accordingly the results have highlighted to interpreter the complex economical processes based on  the archaeological evidence.
 

### 2018-03-19 Paper/jaccardbaetica.tex
update introduction

@@
 
-Material culture allows us to understand a part of mechanism of production... (aquí creo que me repito como siempre)
+%primer párrafo más general---- especifico (¿dónde se ha aplicado?)
+Material culture is one of the most frequent indicator of trade in the archaeological record.
+They allows us to highlight a part of mechanism of production and distribution of goods between different communities.
 
-The economy of Roman Empire have been an object of study in the last centuries. (cambiar)
+%Roman trade has been an object of study in the last centuries. 
+%hablar de que diferentes modelos se han utilizado como el de brumel, el ikasen de las redes, xavi pero qeu lo vamos a probar a microescala para ver si rula o no.
 
+%This paper explore--- resumir en dos líneas el objeto de estudio y definir que ha sido usado para ello
 This paper aims to highlight the production dynamics in relation to a specific area within Roman Empire.
-We want to detect the pattern of olive oil production that link amphora workshops and amphoric stamps.
+We want to detect the pattern of olive oil production that link amphora workshops and amphoric stamps used to mark them.
 We focus here on exploring the economic relation between stamps and amphora production centres.
+An ecology approach has been used to analyse the dispersion of stamps between amphora workshops.
 
-Roman provinces such as Baetica became important production and distribution centers during the Roman Empire.
-However, it remains under debate how this province was organized and whether it is possible to identify patterns in the olive oil market.
+% hipotesis
+
+% Debate
+
+%más específico--- olive oil
+
+Roman province such as Baetica (currently Andalusia, south Spain) became important production centers of olive oil in the Roman Age.
+Olive oil was an important product during the Roman Empire used in almost all the aspect of daily life as cooking, hygiene or lightning \citep{mattingly_d.j._oil_1988}.
+
+The good condition of the Baetican lands stimulated by the high demand of Roman provinces allowed to develop a massive infrastructure of olive oil production.
+This product was distributed in large amount of amphorae along the province, mostly to supply the Roman Army (citar Carreras o Blazquez illo).
+
+The production and distribution of olive oil in this ancient province were growing exponentially during almost three centuries (cronologiaaa CITAR).
+As a results, hundred of amphora workshops were created to support this high demand.
+These workshops were located along the river Guadalquivir and its tributary Genil supplying from the riverine connectivity to the Mediterranean and Atlantic (CITAR GARCIA VARGAS)
+
+The majority of olive oil amphorae produced in this province and shipped thought the provinces belong to Dressel 20 type \citep{dressel,martin-kilcher_romischen_1994}.
+This amphora is commonly associated with transportation of Baetican olive oil during the Roman Empire \citep{berni_millet_epigrafianforica_2008}.
+Most Dressel 20 were marked in stamps and \textit{tituli picti} and inked in graffities with different information but there is not a general consensus about the meaning of them \citep{rodriguez_baetican_1998}.
+Stamps are the most studied in this type of amphorae.
+There is evidence that stamps were used for almost three centuries.
+(economía oleica betica).
+Frequently, stamps were marked mainly in handles but rarely in rims and body \citep{millet_anforas_1998}.
+The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criteria (CITAR).
+Stamps was mostly formed for a code of three letters.
+There letters can appear in a abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
+The meaning of the amphora stamps is still under debate.
+Some authors suggest that they were identified as the land-owners of the olive groves \citep{rodriguez_economioleicola_1977}.
+Other authors propose that stamp could be the owners of the making-amphorae workshop (CITAR) or even a production counting system \citep{berni_millet_epigrafianforica_2008}.
+In any case, the use of these stamps defined somehow the system of working in the workshops.
+
+%Hyphotesis
+However, it remains under debate how this province was organized and whether it is possible to identify patterns in the olive oil trade.
 
@@
 %this paper aims to highlight the production dynamic in connection with the proximity between amphora workshops and stamps.
-
-All the amphora stamps belong to Dressel 20 types \citep{dressel,martin-kilcher_romischen_1994}.
-Dressel 20 is commonly associated with transportation of Baetican olive oil through the provinces during the Roman Empire \citep{berni_millet_epigrafianforica_2008}.
-Most Dressel 20 were marked in stamps and \textit{tituli picti} and inked in graffities with different information but there is not a general consensus about the meaning of them \citep{rodriguez_baetican_1998}.
-Stamps are the most studied in this type of amphorae.
-There is evidence that stamps were used for almost three centuries.
-(economía oleica betica).
-Frequently, stamps were marked mainly in handles but rarely in rims and body \citep{millet_anforas_1998}
-
-The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criteria (CITAR).
-Stamps was mostly formed for a code of three letters.
-There letters can appear in a abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
-
-The meaning of the amphora stamps is still under debate.
-Some authors suggest that they were identified as the land-owners of the olive groves \citep{rodriguez_economioleicola_1977}.
-Other authors propose that stamp could be the owners of the making-amphorae workshop (CITAR) or even a production counting system \citep{berni_millet_epigrafianforica_2008}.
-In any case, the use of these stamps defined somehow the system of working in the workshops.
-%habría que citar a Dressel
-%olive producing
 

### 2018-03-20 Paper/jaccardbaetica.tex
updated introduction

@@
 % Debate
+In archaeology, this debate has been widely studied in the last centuries (CITAR). Specifically, our work pretend to identify production processes focus on the olive oil trade.
 
@@
 
-The good condition of the Baetican lands stimulated by the high demand of Roman provinces allowed to develop a massive infrastructure of olive oil production.
-This product was distributed in large amount of amphorae along the province, mostly to supply the Roman Army (citar Carreras o Blazquez illo).
+The high demand of Roman provinces stimulated by the good condition of the Baetica lands allowed to develop a massive infrastructure of olive oil production.
+This product was distributed in large amount of amphorae along the province, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.
 
@@
 As a results, hundred of amphora workshops were created to support this high demand.
-These workshops were located along the river Guadalquivir and its tributary Genil supplying from the riverine connectivity to the Mediterranean and Atlantic (CITAR GARCIA VARGAS)
+These workshops were located along the river Guadalquivir and its tributary Genil supplying from the riverine connectivity to the Mediterranean and Atlantic \citep{garcia_vargas_enrique_formal_2010}.
 
@@
 %Hyphotesis
-However, it remains under debate how this province was organized and whether it is possible to identify patterns in the olive oil trade.
+However, some challenges remain under discussion such as how this province was organized and whether it is possible to identify production patterns in the olive oil trade (AMPLIAR Y METER SELLOS).
+Neither written evidences have been found that it could explain the economic role of Baetica province in the Roman organization.
+On the other hand, archaeological evidence show a highly specialized production with a long activity in a this specific area with apparently few changes \citep{remesal_anforas_2004}.
 
-%hablar un poco del comercio del aceite de oliva quizás citando el paper de Williams, Peacok y mattlingly el colega
+%hipotesis: si esto es así entonces es asá
+
+Here, therefore, this study propose an approach to explore the effect of Baetican olive oil production by estimating of similarity index (COPIADO CAMBIAR).
+The assumption of this analysis is that closer workshops concentrate similar amphoric stamps in a specific area than the farthest workshops.
+By contrast, if we found similar amphoric stamps in different areas between workshops then the correlation between spatial distance would not be valued.
+
+%resumen del paper
+
 %hablar de las tres áreas de producción: hispalis, corduba y astigi
@@
 
-Aquí poner algo sobre esto \citep{remesal_anforas_2004}
-
-Here, therefore, this study aim to explore the effect of Baetican olive oil production in the Roman Empire
-of production organization over Baetican olive oil production.
+Aquí poner algo sobre esto
 
@@
 Our case study examines the relation between the distribution of amphoric stamps and the workshops.
-We studied a dataset of 3787 stamps collected from different amphora workshops in Baetica provice (see CEIPAC database).
+We studied a dataset of 3787 stamps collected from different amphora workshops in Baetica provice by CEIPAC database \citep{remesal_centro_2015} (see CEIPAC database).
 The workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance.
@@
 
-The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{garcia_vargas_enrique_formal_2010}.
+The chronology in the workshops is widely diverse from the first to the third centuries AD (CITAR).
 However, some stamps show a more specific chronology while the majority of them display a large activity of production being difficult to specify an accurate chronology.

### 2018-03-21 Paper/jaccardbaetica.tex
update intro

@@
 %Roman trade has been an object of study in the last centuries.
-%hablar de que diferentes modelos se han utilizado como el de brumel, el ikasen de las redes, xavi pero qeu lo vamos a probar a microescala para ver si rula o no.
 
@@
 % Debate
-In archaeology, this debate has been widely studied in the last centuries (CITAR). Specifically, our work pretend to identify production processes focus on the olive oil trade.
+The discussion about the production processes in Baetican area due to the lack of written sources
+In archaeology, this debate has been widely studied in the last centuries (CITAR).
+Despite this knowledge the application of new methodology to understand the Roman trade has allowed us to deal with the complexity of Roman Market combined with archaeological evidence \citep{orengo_seeds_2016,brughmans_roman_2016,coto-sarmiento_maria_bayesian_????}.
+Specifically, our work pretend to detect economic processes in a microescala focused in a commercial product from a specific province \citep{isaksen_network_2006}.
 
@@
 %Hyphotesis
-However, some challenges remain under discussion such as how this province was organized and whether it is possible to identify production patterns in the olive oil trade (AMPLIAR Y METER SELLOS).
+However, some challenges remain under discussion such as how this province was organized and whether it is possible to distinguish production patterns in the olive oil trade.
+Our questions will be focused on the distribution of amphoric stamps.
+Did they follow some distribution pattern?
+Did stamps share the same workshop?
 Neither written evidences have been found that it could explain the economic role of Baetica province in the Roman organization.
@@
 %this paper aims to highlight the production dynamic in connection with the proximity between amphora workshops and stamps.
-
-Aquí poner algo sobre esto
 

### 2018-03-22 Paper/jaccardbaetica.tex
update introduction

@@
 Material culture is one of the most frequent indicator of trade in the archaeological record.
-They allows us to highlight a part of mechanism of production and distribution of goods between different communities.
+They allows us to highlight a part of mechanism of production and distribution of goods between different communities \citep{bevan_mediterranean_2014}.
 
@@
 
-The production and distribution of olive oil in this ancient province were growing exponentially during almost three centuries (cronologiaaa CITAR).
-As a results, hundred of amphora workshops were created to support this high demand.
+The production and distribution of olive oil in this ancient province were growing exponentially during almost three centuries \citep{remesal_concierto}.
+As a result, hundred of amphora workshops were created to support this high demand.
 These workshops were located along the river Guadalquivir and its tributary Genil supplying from the riverine connectivity to the Mediterranean and Atlantic \citep{garcia_vargas_enrique_formal_2010}.
@@
 Frequently, stamps were marked mainly in handles but rarely in rims and body \citep{millet_anforas_1998}.
-The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criteria (CITAR).
+The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criteria (CITAR                                           ).
 Stamps was mostly formed for a code of three letters.

### 2018-03-23 Paper/jaccardbaetica.tex
add new changes introduction

@@
 Material culture is one of the most frequent indicator of trade in the archaeological record.
-They allows us to highlight a part of mechanism of production and distribution of goods between different communities \citep{bevan_mediterranean_2014}.
-
-%Roman trade has been an object of study in the last centuries.
+They allow us to highlight a part of mechanism of production and distribution of goods between different communities \citep{bevan_mediterranean_2014}.
 
 %This paper explore--- resumir en dos líneas el objeto de estudio y definir que ha sido usado para ello
-This paper aims to highlight the production dynamics in relation to a specific area within Roman Empire.
+This paper aims to understand the production dynamics in relation to a specific area within Roman Empire.
 We want to detect the pattern of olive oil production that link amphora workshops and amphoric stamps used to mark them.
@@
 In archaeology, this debate has been widely studied in the last centuries (CITAR).
-Despite this knowledge the application of new methodology to understand the Roman trade has allowed us to deal with the complexity of Roman Market combined with archaeological evidence \citep{orengo_seeds_2016,brughmans_roman_2016,coto-sarmiento_maria_bayesian_????}.
+Despite this knowledge the application of new methodology to understand the Roman trade has allowed us to deal with the complexity of Roman Market combined with the archaeological evidence \citep{orengo_seeds_2016,brughmans_roman_2016,coto-sarmiento_maria_bayesian_????}.
 Specifically, our work pretend to detect economic processes in a microescala focused in a commercial product from a specific province \citep{isaksen_network_2006}.
@@
 The production and distribution of olive oil in this ancient province were growing exponentially during almost three centuries \citep{remesal_concierto}.
-As a result, hundred of amphora workshops were created to support this high demand.
+As a result, hundred of amphora workshops were created to support the high demand of Roman Empire.
 These workshops were located along the river Guadalquivir and its tributary Genil supplying from the riverine connectivity to the Mediterranean and Atlantic \citep{garcia_vargas_enrique_formal_2010}.

### 2018-03-23 Jaccard distance/codeCoto.R
add new lines code split

@@ -39,4 +39,5 @@ dev.off()
 count(mydata$site) 
 names(mydata) #para saber el nombre de los valores (columnas)
 
-
+#check by different region: Hispalis, Corduba and Astigi. Download the baezone.csv
+myData= subset(foo, region %in% c("Hispalis")) 

### 2018-03-27 Paper/jaccardbaetica.tex
add new changes introduction

@@
 % Debate
-The discussion about the production processes in Baetican area due to the lack of written sources
-In archaeology, this debate has been widely studied in the last centuries (CITAR).
-Despite this knowledge the application of new methodology to understand the Roman trade has allowed us to deal with the complexity of Roman Market combined with the archaeological evidence \citep{orengo_seeds_2016,brughmans_roman_2016,coto-sarmiento_maria_bayesian_????}.
-Specifically, our work pretend to detect economic processes in a microescala focused in a commercial product from a specific province \citep{isaksen_network_2006}.
+Despite Roman studies have widely discussed in the last centuries, the knowledge about the production processes is still under debate mostly due to the lack of written sources (CAMBIAR PORQUE LO REPITO OCHENTA VECES).
+The application of different approaches combined with the archaeological evidence has allowed us partly to deal with the complexity of understanding the Roman trade \citep{orengo_seeds_2016,brughmans_roman_2016,coto-sarmiento_maria_bayesian_????}.
+Specifically, our work pretend to detect economic processes in a microescala focuses on a commercial product from a specific province \citep{isaksen_network_2006}.
 

### 2018-04-18 Jaccard distance/codeCoto.R
modify code

@@ -41,3 +41,54 @@ names(mydata) #para saber el nombre de los valores (columnas)
 
 #check by different region: Hispalis, Corduba and Astigi. Download the baezone.csv
 myData= subset(foo, region %in% c("Hispalis")) 
+
+
+###plot de puntitos para hacer frequencia
+
+
+library(ggplot2)
+
+ggplot(stamp, aes(y = region, x = code)) + geom_point()
+
+#library plyr (no confuse with dplyr)
+count(stamp$code)
+
+sasa = count(stamp$code)
+idiot = sasa$freq
+
+matrix 
+
+#list of all the sites to make a matrix
+sites=sort(unique(foo$site))
+#to add the regions
+sites = unique(foo[,c("site","region")])
+
+#to order the lsi tof sites with the regions
+sites = sites[order(sites$site),]
+
+
+foo=as.matrix(table(unique(stamp[,c("site","code")])))
+write.csv(foo,"matrixfoo.csv")
+
+how to make a matrix
+
+sites$foo = rowSums(matrix)
+
+ggplot(sites, aes(y=reorder(region, foo, FUN=sum), x=foo, fill=region)) + geom_jitter(col="grey50", alpha=0.5, shape=21, height=0.2, width=0.3, size=3) + scale_colour_manual(values=myPalette) + theme_bw() + theme(legend.position="None") + xlab("number of code stamps") + ylab("region")
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+

### 2018-04-18 Jaccard distance/codesimonsito.R

@@ -1,0 +1,49 @@
+
+library(stats)
+library(dplyr, warn.conflicts = FALSE)
+library(reshape2)
+library(vegan)
+library(reshape)
+library(ggplot2)
+library(dendextend)
+
+foo = read.csv("baezone.csv")
+#matrix
+baeMatrix = acast(foo, site ~ code)
+library(vegan)
+baeDistance = vegdist(baeMatrix, method="horn")
+hclust(baeDistance)
+plot(hclust(baeDistance))
+
+
+model = hclust(baeDistance)
+pepedend = as.dendrogram(model)
+
+regioncolors = c("red","blue","yellow")
+
+ all_workshop_label=labels(pepedend)
+ #all_workshop_label=labels(pepedend)
+
+all_regions = c()
+color_regions=c("olivedrab","steelblue4","tomato")
+for (leaf in 1:length(all_workshop_label)){
+print(leaf)
+reg=unique(foo$region[ foo$site == all_workshop_label[leaf]])
+all_regions=c(all_regions,reg)
+print(reg)
+
+}
+
+labels_colors(pepedend,cex=1.5) = color_regions[all_regions]
+
+par(mar=c(15,2,1,1))
+plot(pepedend)
+
+
+#to keep it 
+
+pdf('dendro.pdf', width= 15, height=5)
+#define the margen
+par(mar=c(12,2,1,1))
+plot(pepedend)
+dev.off()

### 2018-04-18 Paper/jaccardbaetica.tex
add new changes

@@
 Frequently, stamps were marked mainly in handles but rarely in rims and body \citep{millet_anforas_1998}.
-The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criteria (CITAR                                           ).
+The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criteria (CITAR).
 Stamps was mostly formed for a code of three letters.
@@
 Our case study examines the relation between the distribution of amphoric stamps and the workshops.
-We studied a dataset of 3787 stamps collected from different amphora workshops in Baetica provice by CEIPAC database \citep{remesal_centro_2015} (see CEIPAC database).
 The workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance.
 
+The chronology in the workshops is widely diverse from the first to the third centuries AD (CITAR).
+Some stamps show a more specific chronology while the majority of them display a large activity of production being difficult to specify an accurate chronology.
+A reason may be that most of them were partially excavated and only focused on archaeological surveys to collect the maximum stamps as possible (CITAR)
+
+We studied a dataset of 3787 stamps collected from different Dressel 20 amphora workshops in Baetica province.
+The stamp database was compiled by CEIPAC database \citep{remesal_centro_2015} (see CEIPAC database).
 However, approximately the 70 \% of stamps cannot be tested due to fragmentation or incomplete information.
@@
 
-The chronology in the workshops is widely diverse from the first to the third centuries AD (CITAR).
-However, some stamps show a more specific chronology while the majority of them display a large activity of production being difficult to specify an accurate chronology.
-A reason may be that most of them were partially excavated and only focused on archaeological surveys to collect the maximum stamps as possible (CITAR)
+From the database, we collected the following information: a) the site where it was found and b) the stamp code.
+We also created a new column with the area where stamps were found.
+Those areas, known as \textit{conventus}, were administrative centers for territorial organization in the Roman Empire (CITAR).
+Dressel 20 stamps were found in three different \textit{conventus}: Hispalis (currently Seville), Corduba (currently C\'ordoba) and Astigi (currently Écija, Sevilla) \citep{rodriguez_economioleicola_1977,chic_datos_2001,berni_millet_epigrafianforica_2008} .
+AQUÍ QUIZÁS HABLAR DE SELLOS NO?
 
@@
 \includegraphics[width=\linewidth]{figs/dendro}
-\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in Baetica area}
+\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in Baetica area. Colors are represented by areas}
 \label{dendro}

### 2018-04-19 Paper/jaccardbaetica.tex
add new changes material

@@
 Some stamps show a more specific chronology while the majority of them display a large activity of production being difficult to specify an accurate chronology.
-A reason may be that most of them were partially excavated and only focused on archaeological surveys to collect the maximum stamps as possible (CITAR)
+This could be due to two reasons.
+First, most of workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible.
+They presented a long chronology.
+Second, this amphora type was produced during almost three centuries with apparently few changes (PODRÍA CITARME AQUÍ MUAHA)
 
@@
 From the database, we collected the following information: a) the site where it was found and b) the stamp code.
-We also created a new column with the area where stamps were found.
+We also created a new column with the area where stamps were found (decir berni).
 Those areas, known as \textit{conventus}, were administrative centers for territorial organization in the Roman Empire (CITAR).
 Dressel 20 stamps were found in three different \textit{conventus}: Hispalis (currently Seville), Corduba (currently C\'ordoba) and Astigi (currently Écija, Sevilla) \citep{rodriguez_economioleicola_1977,chic_datos_2001,berni_millet_epigrafianforica_2008} .
+
+The distribution of amphora stamps in different conventus can be seen in Fig.
+2.
+The majority of stamps found are concentrated in Hispalis with ...
+stamps while Astigi and Corduba concentrated  respectivamently.
+Mostly workshops show a homogeneity of distribution except La Catria or bla bla that show a big amount of stamps.
+%explicar por qué: mayor intensifación en las excavaciones
+
+explicar las razones del conventus no?
 AQUÍ QUIZÁS HABLAR DE SELLOS NO?
@@
 \includegraphics[width=\linewidth]{figs/dendro}
-\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in Baetica area. Colors are represented by areas}
+\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in Baetica area. Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
 \label{dendro}

### 2018-04-20 Jaccard distance/codeCoto.R
some lines of code

@@ -50,33 +50,32 @@ library(ggplot2)
 
 ggplot(stamp, aes(y = region, x = code)) + geom_point()
 
-#library plyr (no confuse with dplyr)
+#library plyr (no confuse with dplyr) to calculate the frequency of stamps
 count(stamp$code)
 
-sasa = count(stamp$code)
-idiot = sasa$freq
+##doing a matrix
 
-matrix 
+stamp = read.csv("baezone.csv")
+
 
 #list of all the sites to make a matrix
-sites=sort(unique(foo$site))
+sites=sort(unique(stamp$site))
 #to add the regions
-sites = unique(foo[,c("site","region")])
+sites = unique(stamp[,c("site","region")])
 
 #to order the lsi tof sites with the regions
 sites = sites[order(sites$site),]
-
 
 foo=as.matrix(table(unique(stamp[,c("site","code")])))
 write.csv(foo,"matrixfoo.csv")
 
 how to make a matrix
 
-sites$foo = rowSums(matrix)
+sites$foo = rowSums(foo)
 
+pdf('frequency.pdf', width= 15, height=5)
 ggplot(sites, aes(y=reorder(region, foo, FUN=sum), x=foo, fill=region)) + geom_jitter(col="grey50", alpha=0.5, shape=21, height=0.2, width=0.3, size=3) + scale_colour_manual(values=myPalette) + theme_bw() + theme(legend.position="None") + xlab("number of code stamps") + ylab("region")
-
-
+dev.off()
 
 
 

### 2018-04-20 Paper/jaccardbaetica.tex
add new pic distribution

@@
 The distribution of amphora stamps in different conventus can be seen in Fig.
-2.
-The majority of stamps found are concentrated in Hispalis with ...
-stamps while Astigi and Corduba concentrated  respectivamently.
-Mostly workshops show a homogeneity of distribution except La Catria or bla bla that show a big amount of stamps.
-%explicar por qué: mayor intensifación en las excavaciones
+\ref{frequency}.
+The majority of stamps found are concentrated in Hispalis with 574 stamps while Corduba and Astigi with 267 and 146 stamps respectively.
+Mostly workshops show a homogeneity on frequency of stamps except La Catria and Arva (both show 29 different stamps) that show a big amount of different stamps.
+QUIZAS HABLAR DE LA IMPORTANCIA DE LA CATRIA Y ARVA
+
+\begin{figure}[htp]
+	\centering
+\includegraphics[width=\linewidth]{figs/frequency}
+\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in Baetica area. Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
+\label{frequency}
+\end{figure} 
+%explicar por qué: mayor intensifación en las excavaciones quizás en sevilla
 

### 2018-04-23 Paper/jaccardbaetica.tex
add new changes bib

@@
 
-Here, therefore, this study propose an approach to explore the effect of Baetican olive oil production by estimating of similarity index (COPIADO CAMBIAR).
+Here, therefore, this study proposes an approach to explore the effect of Baetican olive oil production by estimating of similarity index (COPIADO CAMBIAR).
 The assumption of this analysis is that closer workshops concentrate similar amphoric stamps in a specific area than the farthest workshops.
@@
 First, most of workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible.
-They presented a long chronology.
-Second, this amphora type was produced during almost three centuries with apparently few changes (PODRÍA CITARME AQUÍ MUAHA)
+Second, this amphora type was produced during almost three centuries with apparently few changes so they presented a long chronology.
+(PODRÍA CITARME AQUÍ MUAHA)
 
@@
 
-From the database, we collected the following information: a) the site where it was found and b) the stamp code.
+From the database, we collected the site where it was found and the stamp code.
 We also created a new column with the area where stamps were found (decir berni).
 Those areas, known as \textit{conventus}, were administrative centers for territorial organization in the Roman Empire (CITAR).
-Dressel 20 stamps were found in three different \textit{conventus}: Hispalis (currently Seville), Corduba (currently C\'ordoba) and Astigi (currently Écija, Sevilla) \citep{rodriguez_economioleicola_1977,chic_datos_2001,berni_millet_epigrafianforica_2008} .
+Dressel 20 stamps were found in three different \textit{conventus}: \textit{Hispalis} (currently Seville), \textit{Corduba} (currently C\'ordoba) and \textit{Astigi} (currently Écija, Sevilla) \citep{rodriguez_economioleicola_1977,chic_datos_2001,berni_millet_epigrafianforica_2008} .
 
@@
 \ref{frequency}.
-The majority of stamps found are concentrated in Hispalis with 574 stamps while Corduba and Astigi with 267 and 146 stamps respectively.
+The majority of stamps found are concentrated in \textit{Hispalis} with 574 stamps while \textit{Corduba} and \textit{Astigi} with 267 and 146 stamps, respectively.
 Mostly workshops show a homogeneity on frequency of stamps except La Catria and Arva (both show 29 different stamps) that show a big amount of different stamps.
-QUIZAS HABLAR DE LA IMPORTANCIA DE LA CATRIA Y ARVA
+According to some researchers, those workshops became in one of the most important centres of amphora production (CITAR) although it could have been intensely prospected.
+\citep{arva_1997}
 
@@
 The dendrogram shown in Fig.
-2 \ref{dendro} was obtained with Morisita-Horn index.
+\ref{dendro} was obtained with Morisita-Horn index.
 This dendrogram suggests that amphora workshops used different stamps for their prodution system.

### 2018-04-25 Paper/jaccardbaetica.tex
add form

@@
 A measure of dissimilarity has been chosen to analyse the dataset.
-We use the statistical technique Morisita-Horn index.
-The formula can be described as follows
+We use the statistical technique Morisita-Horn index \citep{morisita_measuring_1959, horn_measurement_1966}.
+This method was performed to measure the overlap between different samples of sets.
+Generally, it describes the dissimilarity between the system of two communities based on the idea of inverse correlation between diversity and species (CITAR AQUI MANGURRIAN)
 
-(FORMULA Y DESCRIPCIÓN)
+The formula can be described as follows \citep{magurran_measuring_2013}:
 
-This method was performed to measure the overlap between different samples of sets \citep{morisita_measuring_1959, horn_measurement_1966}.
-In Ecology, it describes the dissimilarity between the system of two communities.
+\begin{equation}
+D(MH) = 1- \frac{2 \sum(a_{i} \cdot b_{i})}{(d_{a} + d_{b}) \cdot (N_{a} \cdot N_{b})}
+\end{equation} \\
+
+$d_{a}$ and $d_{b}$ are given by the following equation:
+
+\begin{equation}
+d_{a} = \frac{\sum a_{i}^{2}}{N_{a}^{2}} 
+\end{equation} \\
+
+where $N_{a}$ is the total number of stamps in workshop A; $N_{b}$ is the total number of stamps in workshop B; $a_{i}$ is the number of different stamps for workshop A and $b_{i}$ is the number of different stamps for workhop B.
 

### 2018-04-26 Paper/jaccardbaetica.tex
add introduction

@@
 Material culture is one of the most frequent indicator of trade in the archaeological record.
-They allow us to highlight a part of mechanism of production and distribution of goods between different communities \citep{bevan_mediterranean_2014}.
+In archaeology, they allow us to highlight a part of mechanism of production and distribution of goods along the Mediterranean \citep{bevan_mediterranean_2014}.
+Particularly, the spread of those factors have an important impact during the Roman Age, when the progressive exploitation of communication networks allowed the a major interaction between communities \citep{orengo_seeds_2016}.
+As a result, an important mechanism of production control under Roman government was created spreading by different areas attending to the richness of different places.
 
-%This paper explore--- resumir en dos líneas el objeto de estudio y definir que ha sido usado para ello
+Despite this issue has been widely discussed over many decades, the understanding about the production processes is still under debate, mostly due to the lack of written sources.
+The application of different approaches combined with the archaeological evidence has allowed us partly to deal with the complexity of understanding the Roman production \citep{orengo_seeds_2016,brughmans_roman_2016,coto-sarmiento_maria_bayesian_????}.
+
+%This paper explore--- resumir en dos líneas el objeto de estudio y definir que ha sido usado para ello dispersion of culture
 This paper aims to understand the production dynamics in relation to a specific area within Roman Empire.
+Specifically, our work pretend to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}.
 We want to detect the pattern of olive oil production that link amphora workshops and amphoric stamps used to mark them.
 We focus here on exploring the economic relation between stamps and amphora production centres.
-An ecology approach has been used to analyse the dispersion of stamps between amphora workshops.
+To do this, an Ecology approach has been used to analyse the dispersion of stamps between amphora workshops.
 
@@
 % Debate
-Despite Roman studies have widely discussed in the last centuries, the knowledge about the production processes is still under debate mostly due to the lack of written sources (CAMBIAR PORQUE LO REPITO OCHENTA VECES).
-The application of different approaches combined with the archaeological evidence has allowed us partly to deal with the complexity of understanding the Roman trade \citep{orengo_seeds_2016,brughmans_roman_2016,coto-sarmiento_maria_bayesian_????}.
-Specifically, our work pretend to detect economic processes in a microescala focuses on a commercial product from a specific province \citep{isaksen_network_2006}.
 
 %más específico--- olive oil
+%incluir algo que enlace con roman province
+
+\section{The amphora production in the Roman Empire}
 
@@
 As a result, hundred of amphora workshops were created to support the high demand of Roman Empire.
-These workshops were located along the river Guadalquivir and its tributary Genil supplying from the riverine connectivity to the Mediterranean and Atlantic \citep{garcia_vargas_enrique_formal_2010}.
+These workshops were located along the rivers supplying from the riverine connectivity to the Mediterranean and Atlantic \citep{garcia_vargas_enrique_formal_2010}.
 
@@
 (economía oleica betica).
-Frequently, stamps were marked mainly in handles but rarely in rims and body \citep{millet_anforas_1998}.
+Frequently, they were marked mainly in handles but rarely in rims and body \citep{millet_anforas_1998}.
 The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criteria (CITAR).
@@
 There letters can appear in a abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
-The meaning of the amphora stamps is still under debate.
-Some authors suggest that they were identified as the land-owners of the olive groves \citep{rodriguez_economioleicola_1977}.
-Other authors propose that stamp could be the owners of the making-amphorae workshop (CITAR) or even a production counting system \citep{berni_millet_epigrafianforica_2008}.
-In any case, the use of these stamps defined somehow the system of working in the workshops.
+
+%Many hypothesis involved the meaning of stamps. 
+The meaning of the amphora stamps is not clear.
+On the one hand, stamps are identified as the land owner of the olive groves \citep{rodriguez_economioleicola_1977}.
+On the other hand, they could be the owner of the making amphorae workshops or even a production counting system  \citep{berni_millet_epigrafianforica_2008}.
+In any case, the use of these stamps became in a good proxy to define somehow the system of working in the workshops.
 
@@
 Did stamps share the same workshop?
-Neither written evidences have been found that it could explain the economic role of Baetica province in the Roman organization.
+Neither written sources have been found that it could explain the economic role of Baetica province in the Roman organization.
 On the other hand, archaeological evidence show a highly specialized production with a long activity in a this specific area with apparently few changes \citep{remesal_anforas_2004}.

### 2018-04-30 Paper/jaccardbaetica.tex
add change discussion

@@
 We focus here on exploring the economic relation between stamps and amphora production centres.
-To do this, an Ecology approach has been used to analyse the dispersion of stamps between amphora workshops.
+To do this, an Ecology approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}.
 
@@
 
-No strong relation between stamps and spatial distance have been detected in amphora workshops.
+%crear como una introducción
+
+In this study, we aimed to analyse whether amphoric stamps could play an important role in the organization of the workshops along rivers. For that, dissimilarity index was used to detect differences.
+
+In our analysis, no strong relation between stamps and geographic distance were detected in amphora workshops.
 Indeed, the analysis suggests that there is not connection between stamps and the same amphora workshops, excluding certain exceptions.
 Consequently the majority of stamps are located in different amphora workshops and only similar stamps between closer amphora workshops were found.
-In any case, our results show that most similar amphoric stamps were detected in closer workshops and not in other areas.
+In any case, our results show that most similar stamps were detected in closer workshops and not in other areas.
 These stamps tend to share the same area of production but there is not a general relation between groups of amphora workshops and area.
 
-The hypothesis about groups of amphora workshops sharing the same stamps seems not match with the results of analysis even though there are similar stamps in closer workshops.
-Rather the stamps detected in closer amphorae workshops seem not move from other farthest workshops.
-Additionally they could have been catalogued with different names by authors (METER PONSICH Y DRESSEL) in spite of belonging to the same place or being closer between each other.
-We also identify different stamps in other closer workshops.
+The hypothesis about groups of amphora workshops sharing the same stamps seems do not match with the results of the analysis even though there are similar stamps in closer workshops.
+Rather, it seems that each workshop were organized independently with different stamps.
+Those stamps detected in closer workshops do not move from other farthest workshops.
+In other words, the stamps tend to keep in the same area and different stamps were located in a same amphora workshops.
+This could be defined by several factors.
+Firstly, each workshop had a different organization involved to the use of stamps and they were not used in other workshops.
+Secondly, stamps similarity in closer workshops could be linked to a spatial pattern.
+It is more probably than closer workshops tend to share more traits than farthest.
+While the role of river was significant for the distribution of amphorae in consumption places, river connection amongst workshops was not relevant for the distribution of stamps.
+Finally, the distribution of stamps could have showed some research bias.
+In some cases, workshops have been catalogued with different names despite belonging to the same workshops or being closer between each other.
+Additionally, most workshops were not widely excavated.
+
 %los que aparecen cerca son porque están cerca o puede haber confusion entre talleres
+%By contrast, we have located different stamps in a same amphora workshop.
 
-By contrast, we have located different stamps in a same amphora workshop.
+%soltar aquí las hipótesis
+
 Neither it seems that stamps were used to specify the quality or value of the product \citep{callender}.
@@
 On the other hand, it can be interpreted somehow a batch systematic organization to prepare and distribute the commodity considering that Dressel 20 type was not marked in several cases (CITAR)
-
 it could be somehow a batch systematic organization to prepare and distribute the commodity, considering that Dressel 20 was not marked in several cases. This method was used of counting BLA BAL
 
-An any case,
+An any case, no podemos precisar a qué se tratase
 

### 2018-05-01 Paper/jaccardbaetica.tex
add new pic

@@
 Our case study examines the relation between the distribution of amphoric stamps and the workshops.
-The workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance.
+The workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance (see Fig.\ref{workshop})
+
+\begin{figure}[htp]
+	\centering
+\includegraphics[width=\linewidth]{figs/workshop}
+\caption{Distribution of the workshops}
+\label{workshop}
+\end{figure}
 
@@
 
-Neither it seems that stamps were used to specify the quality or value of the product \citep{callender}.
  Probably our results could be interpreted due to several reason according to previous hypotheses.
-The use of these amphoric stamps could be exclusively running by the owner or families to distinguish the amphora workshop (CITAR).
 
-(aumentar)
-On the other hand, it can be interpreted somehow a batch systematic organization to prepare and distribute the commodity considering that Dressel 20 type was not marked in several cases (CITAR)
-it could be somehow a batch systematic organization to prepare and distribute the commodity, considering that Dressel 20 was not marked in several cases. This method was used of counting BLA BAL
+On the one hand, the use of these amphoric stamps could have been exclusively running by the owner of families to distinguisth the amphora workshop (CITAR).
+The kind of distintion does not seem to be used to specify the quality or value of the product \citep{callender}.
+On the other hand, it can be interpreted somehow a batch systematic organization (CITAR).
+Considering that Dressel 20 was not marked in several cases (CITAR), some authors point out that potters marked amphorae to prepare and distribute the commodity to be shipped.
+This method would be used as an identifier to count the number of amphorae of a banch (CITAR).
+This is would be explain why we found different stamps in a same workshops.
 

### 2018-05-10 Paper/jaccardbaetica.tex
add new lines intro

@@
 In archaeology, they allow us to highlight a part of mechanism of production and distribution of goods along the Mediterranean \citep{bevan_mediterranean_2014}.
-Particularly, the spread of those factors have an important impact during the Roman Age, when the progressive exploitation of communication networks allowed the a major interaction between communities \citep{orengo_seeds_2016}.
+Particularly, the spread of those factors have an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}.
 As a result, an important mechanism of production control under Roman government was created spreading by different areas attending to the richness of different places.
@@
 
-Here, therefore, this study proposes an approach to explore the effect of Baetican olive oil production by estimating of similarity index (COPIADO CAMBIAR).
+Here, therefore, this study proposes an approach to explore the effect of Baetican olive oil production by computing of dissimilarity index.
 The assumption of this analysis is that closer workshops concentrate similar amphoric stamps in a specific area than the farthest workshops.
@@
 
-The chronology in the workshops is widely diverse from the first to the third centuries AD (CITAR).
+The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
 Some stamps show a more specific chronology while the majority of them display a large activity of production being difficult to specify an accurate chronology.
@@
 The research was funded by European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon. 
-All data has been analysed and conducted in R program version 3.2.4, using vegan package (citar)
+All data has been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}.
 

### 2018-05-28 Paper/jaccardbaetica.tex
new changes intro

@@
 
-The analysis shows that amphoric stamps could not be correlated with spatial distance.
-
+The analysis shows that amphoric stamps could be correlated with the spatial distance.
 The correlation coefficients range from a minimum to a maximum.
@@
 \ref{dendro} was obtained with Morisita-Horn index.
-This dendrogram suggests that amphora workshops used different stamps for their prodution system.
-Accordingly only few closer workshops show a similarity on the stamps while most of them seem to display different stamps roles.
+This dendrogram suggests that amphora workshops used different stamps for their production system.
+No groups of similar stamps were found in the cluster.
+On the other hand, closer workshops show a similarity on the stamps while most of them seem to display different stamps roles.
+Indeed workshops that shared more similar amphoric stamps belonged to the same \textit{conventus} area, as for instance Picachos, Cerro de los Pesebres and El Castillejo.
 
@@
 
-On the one hand, the use of these amphoric stamps could have been exclusively running by the owner of families to distinguisth the amphora workshop (CITAR).
-The kind of distintion does not seem to be used to specify the quality or value of the product \citep{callender}.
+On the one hand, the use of these amphoric stamps could have been exclusively running by the owner of the workshop to distinguish the amphora workshop (CITAR).
+The kind of distinction does not seem to be used to specify the quality or value of the product \citep{callender}.
 On the other hand, it can be interpreted somehow a batch systematic organization (CITAR).
 Considering that Dressel 20 was not marked in several cases (CITAR), some authors point out that potters marked amphorae to prepare and distribute the commodity to be shipped.
-This method would be used as an identifier to count the number of amphorae of a banch (CITAR).
-This is would be explain why we found different stamps in a same workshops.
+This method would be used as an identifier to count the number of amphorae of a branch (CITAR).
+This it could be explain wherefore we found different stamps in a same workshop.
 
-An any case, no podemos precisar a qué se tratase
+In any case, no podemos precisar a qué se tratase
 

### 2018-05-29 Paper/jaccardbaetica.tex
add retoques

@@
 In archaeology, they allow us to highlight a part of mechanism of production and distribution of goods along the Mediterranean \citep{bevan_mediterranean_2014}.
-Particularly, the spread of those factors have an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}.
+Particularly, the spread of these factors have an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}.
 As a result, an important mechanism of production control under Roman government was created spreading by different areas attending to the richness of different places.
@@
 
-Roman province such as Baetica (currently Andalusia, south Spain) became important production centers of olive oil in the Roman Age.
-Olive oil was an important product during the Roman Empire used in almost all the aspect of daily life as cooking, hygiene or lightning \citep{mattingly_d.j._oil_1988}.
-
-The high demand of Roman provinces stimulated by the good condition of the Baetica lands allowed to develop a massive infrastructure of olive oil production.
+Roman province such as Baetica (currently Andalusia, south Spain) became important production centers of olive oil during the Roman Empire.
+Olive oil was considered as the liquid gold since it was used in different aspect of the daily life as cooking, hygiene or lighting  \citep{mattingly_d.j._oil_1988}.
+As a consequence, the high demand of Roman provinces stimulated by the good condition of the Baetica lands allowed to develop a massive infrastructure of olive oil production.
 This product was distributed in large amount of amphorae along the province, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.
@@
 As a result, hundred of amphora workshops were created to support the high demand of Roman Empire.
-These workshops were located along the rivers supplying from the riverine connectivity to the Mediterranean and Atlantic \citep{garcia_vargas_enrique_formal_2010}.
+These workshops were located along the rivers Guadalquivir and Genil, supplying from the riverine connectivity to the Mediterranean and Atlantic \citep{garcia_vargas_enrique_formal_2010}.
 
 The majority of olive oil amphorae produced in this province and shipped thought the provinces belong to Dressel 20 type \citep{dressel,martin-kilcher_romischen_1994}.
-This amphora is commonly associated with transportation of Baetican olive oil during the Roman Empire \citep{berni_millet_epigrafianforica_2008}.
-Most Dressel 20 were marked in stamps and \textit{tituli picti} and inked in graffities with different information but there is not a general consensus about the meaning of them \citep{rodriguez_baetican_1998}.
-Stamps are the most studied in this type of amphorae.
-There is evidence that stamps were used for almost three centuries.
-(economía oleica betica).
+This amphora is commonly associated with the transportation of Baetican olive oil during the Roman Empire \citep{berni_millet_epigrafianforica_2008}.
+Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in graffities with different information.
+However, graffities and \textit{tituli picti} have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
+By contrast, stamps are the most studied in this type of amphorae.
 Frequently, they were marked mainly in handles but rarely in rims and body \citep{millet_anforas_1998}.
-The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criteria (CITAR).
+The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criteria.
 Stamps was mostly formed for a code of three letters.
-There letters can appear in a abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
+These letters can appear in a abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
+
+INCLUIR AQUÍ QUE LAS STAMPS VAN CAMBIANDO DE FECHA Y SIGLOS
 
 %Many hypothesis involved the meaning of stamps. 
-The meaning of the amphora stamps is not clear.
+ There is not a general consensus about the meaning of them \citep{rodriguez_baetican_1998}.
 On the one hand, stamps are identified as the land owner of the olive groves \citep{rodriguez_economioleicola_1977}.
-On the other hand, they could be the owner of the making amphorae workshops or even a production counting system  \citep{berni_millet_epigrafianforica_2008}.
+On the other, they could be the owner of the making amphorae workshops or even a production counting system  \citep{berni_millet_epigrafianforica_2008}.
 In any case, the use of these stamps became in a good proxy to define somehow the system of working in the workshops.
@@
 %Hyphotesis
-However, some challenges remain under discussion such as how this province was organized and whether it is possible to distinguish production patterns in the olive oil trade.
+Nevertheless, some challenges remain under discussion such as how this province was organized and whether it is possible to distinguish production patterns in the olive oil trade.
 Our questions will be focused on the distribution of amphoric stamps.
@@
 Neither written sources have been found that it could explain the economic role of Baetica province in the Roman organization.
-On the other hand, archaeological evidence show a highly specialized production with a long activity in a this specific area with apparently few changes \citep{remesal_anforas_2004}.
+On the other hand, archaeological evidence shows a highly specialized production with a long activity in a this specific area with apparently few changes \citep{remesal_anforas_2004}.
 
@@
 We also created a new column with the area where stamps were found (decir berni).
-Those areas, known as \textit{conventus}, were administrative centers for territorial organization in the Roman Empire (CITAR).
+Those areas, known as \textit{conventus}, were administrative centers for territorial organization in the Roman Empire.
 Dressel 20 stamps were found in three different \textit{conventus}: \textit{Hispalis} (currently Seville), \textit{Corduba} (currently C\'ordoba) and \textit{Astigi} (currently Écija, Sevilla) \citep{rodriguez_economioleicola_1977,chic_datos_2001,berni_millet_epigrafianforica_2008} .
@@
 Mostly workshops show a homogeneity on frequency of stamps except La Catria and Arva (both show 29 different stamps) that show a big amount of different stamps.
-According to some researchers, those workshops became in one of the most important centres of amphora production (CITAR) although it could have been intensely prospected.
+According to some researchers, those workshops became in one of the most important centres of amphora production although it could have been more intensely prospected then others.
 \citep{arva_1997}
@@
 This method allows to bear in mind the similar number of times for each repeated stamp per workshop.
-%quizas ampliar con el método de xavielin
+If two workshops have similar stamp codes then the probability would be 0 while the results would be 1 when stamp codes are totally different.
 
@@
 This dendrogram suggests that amphora workshops used different stamps for their production system.
-No groups of similar stamps were found in the cluster.
-On the other hand, closer workshops show a similarity on the stamps while most of them seem to display different stamps roles.
+Nearby workshops show a similarity on the stamps while most of them seem to display different stamps roles.
+Additionally, no groups of similar stamps were found in the cluster.
+The majority of stamp grouping was composed by no more than three workshops.
 Indeed workshops that shared more similar amphoric stamps belonged to the same \textit{conventus} area, as for instance Picachos, Cerro de los Pesebres and El Castillejo.
@@
 
-In this study, we aimed to analyse whether amphoric stamps could play an important role in the organization of the workshops along rivers. For that, dissimilarity index was used to detect differences.
+In this study, we aimed to analyse whether amphoric stamps could play an important role in the organization of the workshops along rivers.
+For that, dissimilarity index was used to detect differences among workshops and stamps.
 
 In our analysis, no strong relation between stamps and geographic distance were detected in amphora workshops.
-Indeed, the analysis suggests that there is not connection between stamps and the same amphora workshops, excluding certain exceptions.
+Indeed, the analysis suggests that there is not connection between stamps and the same amphora workshops, excluding certain exceptions when nearby workshops share the same amphoric stamps.
 Consequently the majority of stamps are located in different amphora workshops and only similar stamps between closer amphora workshops were found.
-In any case, our results show that most similar stamps were detected in closer workshops and not in other areas.
+In any case, our results show that most similar stamps were detected in the same conventus area.
 These stamps tend to share the same area of production but there is not a general relation between groups of amphora workshops and area.
@@
 Additionally, most workshops were not widely excavated.
-
-%los que aparecen cerca son porque están cerca o puede haber confusion entre talleres
-%By contrast, we have located different stamps in a same amphora workshop.
 
@@
 
-The research was funded by European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon. 
+The research was funded by European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon and Nachete. 
 All data has been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}.

### 2018-05-30 Paper/jaccardbaetica.tex
lines of conclusion

@@
 \begin{document}
+
+\title{The markings of the trade: exploring the patterns of olive oil production in Roman Baetica}
 
@@
 
-Probably our results could be interpreted due to several reason according to previous hypotheses.
-
-On the one hand, the use of these amphoric stamps could have been exclusively running by the owner of the workshop to distinguish the amphora workshop (CITAR).
-The kind of distinction does not seem to be used to specify the quality or value of the product \citep{callender}.
+The results of the case study could be interpreted due to several reason according to previous hypotheses.
+On the one hand, the use of these amphoric stamps could have been exclusively running by the owner or owners of the workshop to distinguish the amphora workshop (CITAR).
+This hypotheses would explain the fact that we do not find similar stamps in different workshops, however, we found different amphora stamps in a same workshops that they would be barely difficult to assign different owners.
+Neither a kind of quality distinction does seem to be used to specify the quality or value of the product \citep{callender}.
 On the other hand, it can be interpreted somehow a batch systematic organization (CITAR).
-Considering that Dressel 20 was not marked in several cases (CITAR), some authors point out that potters marked amphorae to prepare and distribute the commodity to be shipped.
+Considering that Dressel 20 was not marked in several cases, some authors point out that potters marked amphorae to prepare and distribute the commodity to be shipped (CITAR JUAN MOROS).
 This method would be used as an identifier to count the number of amphorae of a branch (CITAR).
 This it could be explain wherefore we found different stamps in a same workshop.
-
-In any case, no podemos precisar a qué se tratase
+In any case, we do not have enough archaeological evidence that can validated the interpretations presented here and our results are certainly valid only with the context of our case study.
 

### 2018-05-31 Paper/jaccardbaetica.tex
add magu

@@
 This method was performed to measure the overlap between different samples of sets.
-Generally, it describes the dissimilarity between the system of two communities based on the idea of inverse correlation between diversity and species (CITAR AQUI MANGURRIAN)
+Generally, it describes the dissimilarity between the system of two communities based on the idea of inverse correlation between diversity and species \citep{magurran_why_1988}.
 

### 2018-06-04 Paper/jaccardbaetica.tex
review

@@
 By contrast, stamps are the most studied in this type of amphorae.
+During almost three centuries, stamps were used to mark amphorae with a different chronological frequency \citep{remesal_sellar_2016}.
 Frequently, they were marked mainly in handles but rarely in rims and body \citep{millet_anforas_1998}.
@@
 These letters can appear in a abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
-
-INCLUIR AQUÍ QUE LAS STAMPS VAN CAMBIANDO DE FECHA Y SIGLOS
 
@@
 From the database, we collected the site where it was found and the stamp code.
-We also created a new column with the area where stamps were found (decir berni).
-Those areas, known as \textit{conventus}, were administrative centers for territorial organization in the Roman Empire.
+We also created a new column with the area where stamps were found.
+These areas, known as \textit{conventus}, were administrative centers for territorial organization in the Roman Empire.
 Dressel 20 stamps were found in three different \textit{conventus}: \textit{Hispalis} (currently Seville), \textit{Corduba} (currently C\'ordoba) and \textit{Astigi} (currently Écija, Sevilla) \citep{rodriguez_economioleicola_1977,chic_datos_2001,berni_millet_epigrafianforica_2008} .
@@
 Neither a kind of quality distinction does seem to be used to specify the quality or value of the product \citep{callender}.
-On the other hand, it can be interpreted somehow a batch systematic organization (CITAR).
-Considering that Dressel 20 was not marked in several cases, some authors point out that potters marked amphorae to prepare and distribute the commodity to be shipped (CITAR JUAN MOROS).
+On the other hand, it can be interpreted somehow a batch systematic organization.
+Considering that Dressel 20 was not marked in several cases, some authors point out that potters marked amphorae to prepare and distribute the commodity to be shipped \citep{berni_millet_epigrafianforica_2008}.
 This method would be used as an identifier to count the number of amphorae of a branch (CITAR).
 This it could be explain wherefore we found different stamps in a same workshop.
-In any case, we do not have enough archaeological evidence that can validated the interpretations presented here and our results are certainly valid only with the context of our case study.
+In any case, we do not have enough archaeological evidence that can validate the interpretations presented here and our results are certainly valid only with the context of our case study.
 

### 2018-06-04 Paper/jaccardbaetica.tex
review

@@
 Mostly workshops show a homogeneity on frequency of stamps except La Catria and Arva (both show 29 different stamps) that show a big amount of different stamps.
-According to some researchers, those workshops became in one of the most important centres of amphora production although it could have been more intensely prospected then others.
+According to some researchers, those workshops became in most important centres of amphora production although it could have been more intensely prospected then others.
 \citep{arva_1997}

### 2018-06-05 Paper/jaccardbaetica.tex
fix bug

@@
 \includegraphics[width=\linewidth]{figs/frequency}
-\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in Baetica area. Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
+\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in Baetica area. Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue).}
 \label{frequency}
@@
 
-explicar las razones del conventus no?
-AQUÍ QUIZÁS HABLAR DE SELLOS NO?
+%explicar las razones del conventus no?
+%aquí hablar de sellos?
 
@@
 
-\subsection{Quantifying the diversity Dissimilarity correlation}
+\subsection{Quantifying the diversity}
 
@@
 This method would be used as an identifier to count the number of amphorae of a branch (CITAR).
+This method could also have served to identify different groups of potter workers working in the same amphora workshop.
+Potters could have marked the amphorae to distinguish  different groups working in parallel \citep{li_crossbows_2014}.
 This it could be explain wherefore we found different stamps in a same workshop.
@@
 
-The research was funded by European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon and Nachete. 
-All data has been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}.
+The research was funded by European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros and Ignacio Morer for their useful suggestions.  
+All data has been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}. Source and code are available
 

### 2018-06-06 Paper/jaccardbaetica.tex
modify some error

@@
 These areas, known as \textit{conventus}, were administrative centers for territorial organization in the Roman Empire.
-Dressel 20 stamps were found in three different \textit{conventus}: \textit{Hispalis} (currently Seville), \textit{Corduba} (currently C\'ordoba) and \textit{Astigi} (currently Écija, Sevilla) \citep{rodriguez_economioleicola_1977,chic_datos_2001,berni_millet_epigrafianforica_2008} .
+Dressel 20 stamps were found in three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter Hispalis), \textit{Cordubensis} (currently C\'ordoba, hereafter Corduba) and \textit{Astigi} (currently Écija, Sevilla, hereafter Astigi) \citep{rodriguez_economioleicola_1977,chic_datos_2001,berni_millet_epigrafianforica_2008} .
 
@@
 \includegraphics[width=\linewidth]{figs/frequency}
-\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in Baetica area. Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue).}
+\caption{Distribution of the number of stamps for each area. Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue).}
 \label{frequency}
@@
 Consequently the majority of stamps are located in different amphora workshops and only similar stamps between closer amphora workshops were found.
-In any case, our results show that most similar stamps were detected in the same conventus area.
+In any case, our results show that most similar stamps were detected in the same \textit{conventus} area.
 These stamps tend to share the same area of production but there is not a general relation between groups of amphora workshops and area.
@@
 In some cases, workshops have been catalogued with different names despite belonging to the same workshops or being closer between each other.
-Additionally, most workshops were not widely excavated.
+Additionally, most of the workshops were not widely excavated.
 

### 2018-06-07 Paper/jaccardbaetica.tex
modify some error
review english

@@
 
+%cambiar quizás
 \begin{abstract}
@@
 However, it remains under debate how this province was organized and whether it is possible to identify patterns in the olive oil market.
-
-CAMBIAR
 
@@
 In archaeology, they allow us to highlight a part of mechanism of production and distribution of goods along the Mediterranean \citep{bevan_mediterranean_2014}.
-Particularly, the spread of these factors have an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}.
+Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}.
 As a result, an important mechanism of production control under Roman government was created spreading by different areas attending to the richness of different places.
@@
 This paper aims to understand the production dynamics in relation to a specific area within Roman Empire.
-Specifically, our work pretend to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}.
+Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}.
 We want to detect the pattern of olive oil production that link amphora workshops and amphoric stamps used to mark them.
@@
 As a result, hundred of amphora workshops were created to support the high demand of Roman Empire.
-These workshops were located along the rivers Guadalquivir and Genil, supplying from the riverine connectivity to the Mediterranean and Atlantic \citep{garcia_vargas_enrique_formal_2010}.
+These workshops were located along the rivers Guadalquivir and Genil, supplying from the riverine connectivity to the Mediterranean and Atlantic routes \citep{garcia_vargas_enrique_formal_2010}.
 
-The majority of olive oil amphorae produced in this province and shipped thought the provinces belong to Dressel 20 type \citep{dressel,martin-kilcher_romischen_1994}.
+The majority of olive oil amphorae produced in this province and shipped through the provinces belong to Dressel 20 type \citep{dressel,martin-kilcher_romischen_1994}.
 This amphora is commonly associated with the transportation of Baetican olive oil during the Roman Empire \citep{berni_millet_epigrafianforica_2008}.
-Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in graffities with different information.
-However, graffities and \textit{tituli picti} have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
+Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti} with different information.
+However, inscriptions and \textit{tituli picti} have not enough analysed by researchers due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
 By contrast, stamps are the most studied in this type of amphorae.
@@
 Frequently, they were marked mainly in handles but rarely in rims and body \citep{millet_anforas_1998}.
-The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criteria.
+The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion.
 Stamps was mostly formed for a code of three letters.
@@
 
-Here, therefore, this study proposes an approach to explore the effect of Baetican olive oil production by computing of dissimilarity index.
+Here, therefore, this study proposes an evolutionary approach to explore the effect of Baetican olive oil production by computing of dissimilarity index.
 The assumption of this analysis is that closer workshops concentrate similar amphoric stamps in a specific area than the farthest workshops.
@@
 Our case study examines the relation between the distribution of amphoric stamps and the workshops.
-The workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance (see Fig.\ref{workshop})
+The amphora workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance (see Fig.\ref{workshop})
 
@@
 This could be due to two reasons.
-First, most of workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible.
-Second, this amphora type was produced during almost three centuries with apparently few changes so they presented a long chronology.
-(PODRÍA CITARME AQUÍ MUAHA)
+Firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible.
+Secondly, this amphora type was produced during almost three centuries with apparently few changes so they presented a long chronology.
+(podría citar nuestro paper por toda la face).
 
 We studied a dataset of 3787 stamps collected from different Dressel 20 amphora workshops in Baetica province.
-The stamp database was compiled by CEIPAC database \citep{remesal_centro_2015} (see CEIPAC database).
+The stamp database was compiled by CEIPAC database \citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}).
 However, approximately the 70 \% of stamps cannot be tested due to fragmentation or incomplete information.
@@
 
-The distribution of amphora stamps in different conventus can be seen in Fig.
+The distribution of amphora stamps in different \textit{conventus} can be seen in Fig.
 \ref{frequency}.
@@
 Mostly workshops show a homogeneity on frequency of stamps except La Catria and Arva (both show 29 different stamps) that show a big amount of different stamps.
-According to some researchers, those workshops became in most important centres of amphora production although it could have been more intensely prospected then others.
+According to some researchers, those workshops became in most important centres of amphora production although they could have been more intensely prospected then others.
 \citep{arva_1997}
@@
 We use the statistical technique Morisita-Horn index \citep{morisita_measuring_1959, horn_measurement_1966}.
-This method was performed to measure the overlap between different samples of sets.
+This method was performed to measure the dissimilarity between different samples of sets.
 Generally, it describes the dissimilarity between the system of two communities based on the idea of inverse correlation between diversity and species \citep{magurran_why_1988}.
@@
 
-where $N_{a}$ is the total number of stamps in workshop A; $N_{b}$ is the total number of stamps in workshop B; $a_{i}$ is the number of different stamps for workshop A and $b_{i}$ is the number of different stamps for workhop B.
+where $N_{a}$ is the total number of stamps in workshop A; $N_{b}$ is the total number of stamps in workshop B; $a_{i}$ is the number of different stamps for workshop A and $b_{i}$ is the number of different stamps for workshop B.
 
@@
 \ref{dendro} was obtained with Morisita-Horn index.
-This dendrogram suggests that amphora workshops used different stamps for their production system.
+The dendrogram suggests that amphora workshops used different stamps for their production system.
 Nearby workshops show a similarity on the stamps while most of them seem to display different stamps roles.
@@
 The majority of stamp grouping was composed by no more than three workshops.
-Indeed workshops that shared more similar amphoric stamps belonged to the same \textit{conventus} area, as for instance Picachos, Cerro de los Pesebres and El Castillejo.
+Indeed workshops that shared more similar amphoric stamps belonged to the same \textit{conventus} area, such as Picachos, Cerro de los Pesebres and El Castillejo.
 
@@
 \includegraphics[width=\linewidth]{figs/dendro}
-\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in Baetica area. Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
+\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in Baetica area. Colours are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
 \label{dendro}
@@
 
-In our analysis, no strong relation between stamps and geographic distance were detected in amphora workshops.
+In our analysis, no strong relation between stamps and spatial distance were detected in amphora workshops.
 Indeed, the analysis suggests that there is not connection between stamps and the same amphora workshops, excluding certain exceptions when nearby workshops share the same amphoric stamps.
-Consequently the majority of stamps are located in different amphora workshops and only similar stamps between closer amphora workshops were found.
+Consequently, the majority of stamps are located in different amphora workshops and only similar stamps between closer amphora workshops were found.
 In any case, our results show that most similar stamps were detected in the same \textit{conventus} area.
@@
 In other words, the stamps tend to keep in the same area and different stamps were located in a same amphora workshops.
+
 This could be defined by several factors.
-Firstly, each workshop had a different organization involved to the use of stamps and they were not used in other workshops.
-Secondly, stamps similarity in closer workshops could be linked to a spatial pattern.
+First, each workshop had a different organization involved to the use of stamps and they were not used in other workshops.
+Second, stamp similarity in closer workshops could be linked to a spatial pattern.
 It is more probably than closer workshops tend to share more traits than farthest.
-While the role of river was significant for the distribution of amphorae in consumption places, river connection amongst workshops was not relevant for the distribution of stamps.
+While the role of river was significant for the distribution of amphorae in consumption places, river connection amongst workshops do not seem show a relevancy for the distribution of stamps.
 Finally, the distribution of stamps could have showed some research bias.
@@
 
-%soltar aquí las hipótesis
+%soltar aquí las hipótesis repito aquí several reasons no repetir
 
@@
 On the one hand, the use of these amphoric stamps could have been exclusively running by the owner or owners of the workshop to distinguish the amphora workshop (CITAR).
-This hypotheses would explain the fact that we do not find similar stamps in different workshops, however, we found different amphora stamps in a same workshops that they would be barely difficult to assign different owners.
-Neither a kind of quality distinction does seem to be used to specify the quality or value of the product \citep{callender}.
+This hypothesis would explain the fact that we do not find similar stamps in different workshops, however, we found different amphora stamps in the same workshops that they would be barely difficult to assign different owners.
+Neither a  quality distinction does seem to be used to specify the value of the product \citep{callender}.
 On the other hand, it can be interpreted somehow a batch systematic organization.
@@
 This method have identified differences in the case of the amphoric production within Roman Empire.
-Accordingly the results have highlighted to interpreter the complex economical processes based on the archaeological evidence.
+Accordingly the results have highlighted the interpretation of the complex economic processes based on the archaeological evidence.
 
@@
 \section{Acknowledgements}
-
+%completar 
 The research was funded by European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros and Ignacio Morer for their useful suggestions.  

### 2018-06-20 Paper/jaccardbaetica.tex
changes

@@
 The aim of this study is to detect the patterns of olive oil production that link amphora workshops and amphoric stamps.
-Roman provinces such as Baetica became important production and distribution centers during the Roman Empire.
+Roman provinces such as Baetica became important production and distribution centres during the Roman Empire.
 However, it remains under debate how this province was organized and whether it is possible to identify patterns in the olive oil market.
@@
 
-Despite this issue has been widely discussed over many decades, the understanding about the production processes is still under debate, mostly due to the lack of written sources.
+Despite this issue has been widely discussed over many decades, the understanding about the production processes is still under debate, mostly due to the lack of written records.
 The application of different approaches combined with the archaeological evidence has allowed us partly to deal with the complexity of understanding the Roman production \citep{orengo_seeds_2016,brughmans_roman_2016,coto-sarmiento_maria_bayesian_????}.
@@
 %This paper explore--- resumir en dos líneas el objeto de estudio y definir que ha sido usado para ello dispersion of culture
-This paper aims to understand the production dynamics in relation to a specific area within Roman Empire.
+This paper aims to understand the production dynamics in connection with a specific area within Roman Empire.
 Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}.
@@
 The production and distribution of olive oil in this ancient province were growing exponentially during almost three centuries \citep{remesal_concierto}.
-As a result, hundred of amphora workshops were created to support the high demand of Roman Empire.
+As a result, hundred of amphora workshops were made to support this high demand of Roman Empire.
 These workshops were located along the rivers Guadalquivir and Genil, supplying from the riverine connectivity to the Mediterranean and Atlantic routes \citep{garcia_vargas_enrique_formal_2010}.
@@
 The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion.
-Stamps was mostly formed for a code of three letters.
-These letters can appear in a abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
+Stamps was mostly formed by a code of three letters and they can appear in a abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
 
 %Many hypothesis involved the meaning of stamps. 
- There is not a general consensus about the meaning of them \citep{rodriguez_baetican_1998}.
+ There is not a general consensus about the meaning of the stamps \citep{rodriguez_baetican_1998}.
 On the one hand, stamps are identified as the land owner of the olive groves \citep{rodriguez_economioleicola_1977}.
@@
 We also created a new column with the area where stamps were found.
-These areas, known as \textit{conventus}, were administrative centers for territorial organization in the Roman Empire.
+These areas, known as \textit{conventus}, were administrative centres for territorial organization in the Roman Empire.
 Dressel 20 stamps were found in three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter Hispalis), \textit{Cordubensis} (currently C\'ordoba, hereafter Corduba) and \textit{Astigi} (currently Écija, Sevilla, hereafter Astigi) \citep{rodriguez_economioleicola_1977,chic_datos_2001,berni_millet_epigrafianforica_2008} .
@@
 According to some researchers, those workshops became in most important centres of amphora production although they could have been more intensely prospected then others.
-\citep{arva_1997}
+\citep{arva_1997}.
 
@@
 This method allows to bear in mind the similar number of times for each repeated stamp per workshop.
-If two workshops have similar stamp codes then the probability would be 0 while the results would be 1 when stamp codes are totally different.
+If two workshops have similar stamp codes then the probability would be 0 whereas stamps codes are totally different when the results would be 1.
 
@@
 
-In this study, we aimed to analyse whether amphoric stamps could play an important role in the organization of the workshops along rivers.
-For that, dissimilarity index was used to detect differences among workshops and stamps.
+In this work, we aimed to analyse whether amphoric stamps could play an important role in the organization of the workshops along rivers.
+For this reason, dissimilarity index was used to detect differences among workshops and stamps.
 
 In our analysis, no strong relation between stamps and spatial distance were detected in amphora workshops.
-Indeed, the analysis suggests that there is not connection between stamps and the same amphora workshops, excluding certain exceptions when nearby workshops share the same amphoric stamps.
+The analysis suggests that there is not connection between stamps and the same amphora workshops, excluding certain exceptions when nearby workshops share the same amphoric stamp.
 Consequently, the majority of stamps are located in different amphora workshops and only similar stamps between closer amphora workshops were found.
@@
 Rather, it seems that each workshop were organized independently with different stamps.
-Those stamps detected in closer workshops do not move from other farthest workshops.
+Those stamps detected in closer workshops do not move from other distant workshops.
 In other words, the stamps tend to keep in the same area and different stamps were located in a same amphora workshops.
@@
 Second, stamp similarity in closer workshops could be linked to a spatial pattern.
-It is more probably than closer workshops tend to share more traits than farthest.
-While the role of river was significant for the distribution of amphorae in consumption places, river connection amongst workshops do not seem show a relevancy for the distribution of stamps.
-Finally, the distribution of stamps could have showed some research bias.
-In some cases, workshops have been catalogued with different names despite belonging to the same workshops or being closer between each other.
+It is more probably than closer workshops tend to share more traits than distant workshops.
+While the role of river was significant for the distribution of amphorae in consumption places, river connection amongst workshops does not seem show a relevancy for the distribution of stamps.
+Finally, the distribution of stamps could have shown some research bias.
+In some cases, workshops have been catalogued with different names in spite of belonging to the same workshops or being closer between each other.
 Additionally, most of the workshops were not widely excavated.
@@
 Neither a quality distinction does seem to be used to specify the value of the product \citep{callender}.
-On the other hand, it can be interpreted somehow a batch systematic organization.
-Considering that Dressel 20 was not marked in several cases, some authors point out that potters marked amphorae to prepare and distribute the commodity to be shipped \citep{berni_millet_epigrafianforica_2008}.
+On the other hand, it could be interpreted somehow a batch systematic organization.
+Considering that Dressel 20 was not marked in most cases, some authors point out that potters marked amphorae to prepare and distribute the commodity to be shipped \citep{berni_millet_epigrafianforica_2008}.
 This method would be used as an identifier to count the number of amphorae of a branch (CITAR).
@@
 This method have identified differences in the case of the amphoric production within Roman Empire.
-Accordingly the results have highlighted the interpretation of the complex economic processes based on the archaeological evidence.
+Accordingly the results have highlighted for the interpretation of the complex economic processes connected with the archaeological evidence.
 

### 2018-06-22 Paper/jaccardbaetica.tex
add new changes

@@
 
-Roman province such as Baetica (currently Andalusia, south Spain) became important production centers of olive oil during the Roman Empire.
+Roman provinces such as Baetica (currently Andalusia, south Spain) became important production centres of olive oil during the Roman Empire.
 Olive oil was considered as the liquid gold since it was used in different aspect of the daily life as cooking, hygiene or lighting  \citep{mattingly_d.j._oil_1988}.
@@
 The production and distribution of olive oil in this ancient province were growing exponentially during almost three centuries \citep{remesal_concierto}.
-As a result, hundred of amphora workshops were made to support this high demand of Roman Empire.
+As a result, thousands of amphora workshops were made to support this high demand of Roman Empire.
 These workshops were located along the rivers Guadalquivir and Genil, supplying from the riverine connectivity to the Mediterranean and Atlantic routes \citep{garcia_vargas_enrique_formal_2010}.
@@
 Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti} with different information.
-However, inscriptions and \textit{tituli picti} have not enough analysed by researchers due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
+However, inscriptions and \textit{tituli picti} have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
 By contrast, stamps are the most studied in this type of amphorae.
 During almost three centuries, stamps were used to mark amphorae with a different chronological frequency \citep{remesal_sellar_2016}.
-Frequently, they were marked mainly in handles but rarely in rims and body \citep{millet_anforas_1998}.
+Frequently, they were marked mainly in handles but seldom in rims and body \citep{millet_anforas_1998}.
 The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion.
-Stamps was mostly formed by a code of three letters and they can appear in a abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
+Stamps are mostly formed by a code of three letters and they can appear in a abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
 
@@
  There is not a general consensus about the meaning of the stamps \citep{rodriguez_baetican_1998}.
-On the one hand, stamps are identified as the land owner of the olive groves \citep{rodriguez_economioleicola_1977}.
-On the other, they could be the owner of the making amphorae workshops or even a production counting system  \citep{berni_millet_epigrafianforica_2008}.
+It seems that stamps could have been identified as the land owner of the olive groves \citep{rodriguez_economioleicola_1977}.
+On the other hand, they could belong to the owner of the making amphorae workshops or even a production counting system  \citep{berni_millet_epigrafianforica_2008}.
 In any case, the use of these stamps became in a good proxy to define somehow the system of working in the workshops.
@@
 Our questions will be focused on the distribution of amphoric stamps.
-Did they follow some distribution pattern?
+Did they follow a distribution pattern?
 Did stamps share the same workshop?
-Neither written sources have been found that it could explain the economic role of Baetica province in the Roman organization.
-On the other hand, archaeological evidence shows a highly specialized production with a long activity in a this specific area with apparently few changes \citep{remesal_anforas_2004}.
+Neither written records have been found that it could explain the economic role of Baetica province in the Roman organization.
+Indeed, archaeological evidence shows a highly specialized production with a long activity in a this specific area with apparently few changes \citep{remesal_anforas_2004}.
 
 %hipotesis: si esto es así entonces es asá
-
 Here, therefore, this study proposes an evolutionary approach to explore the effect of Baetican olive oil production by computing of dissimilarity index.
-The assumption of this analysis is that closer workshops concentrate similar amphoric stamps in a specific area than the farthest workshops.
+The assumption of this analysis is that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops.
 By contrast, if we found similar amphoric stamps in different areas between workshops then the correlation between spatial distance would not be valued.
@@
 The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
-Some stamps show a more specific chronology while the majority of them display a large activity of production being difficult to specify an accurate chronology.
-This could be due to two reasons.
-Firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible.
-Secondly, this amphora type was produced during almost three centuries with apparently few changes so they presented a long chronology.
+Some stamps show a more specific chronology while the majority of them display a large activity of production that it can be difficult for specifying an accurate chronology.
+This could be due to two reasons; firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible; secondly, Dressel 20 was produced during almost three centuries with apparently few changes.
 (podría citar nuestro paper por toda la face).
@@
 Consequently, we discard integrate the fragmented stamps in our dataset.
-We finally filter a total sample of 987 stamps composed by 131 different stamps from 81 workshops.
+We finally filter a total sample of 987 stamps comprised of 131 different stamps from 81 workshops.
 
-From the database, we collected the site where it was found and the stamp code.
-We also created a new column with the area where stamps were found.
+From the database, we collected the site where stamps were found and the stamp code.
+We also created a new column with the area of the stamps.
 These areas, known as \textit{conventus}, were administrative centres for territorial organization in the Roman Empire.
@@
 The majority of stamps found are concentrated in \textit{Hispalis} with 574 stamps while \textit{Corduba} and \textit{Astigi} with 267 and 146 stamps, respectively.
-Mostly workshops show a homogeneity on frequency of stamps except La Catria and Arva (both show 29 different stamps) that show a big amount of different stamps.
-According to some researchers, those workshops became in most important centres of amphora production although they could have been more intensely prospected then others.
+Mostly, workshops show a homogeneity on the frequency of stamps except both La Catria and Arva that show a big amount of stamps with 29 different stamps.
+According to previous studies, those workshops became in most important centres of amphora production although they could have been more intensely prospected then others.
 \citep{arva_1997}.
@@
 
-Considering our dataset as non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity \citep{wolda_similarity_1981}.
+Considering our dataset as a non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity \citep{wolda_similarity_1981}.
 Morisita-Horn index can be expressed considering 0 as total presence of similarity of stamps and 1 a totally dissimilarity between stamps.
-In our case, it will be calculated the number of times that one stamp appear in a amphora workshop.
+In our case, it will be calculated the number of times that one stamp appear in an amphora workshop.
 This method allows to bear in mind the similar number of times for each repeated stamp per workshop.
@@
 Nearby workshops show a similarity on the stamps while most of them seem to display different stamps roles.
-Additionally, no groups of similar stamps were found in the cluster.
+Additionally, groupings of similar stamps were not found in the cluster.
 The majority of stamp grouping was composed by no more than three workshops.
@@
 
-In our analysis, no strong relation between stamps and spatial distance were detected in amphora workshops.
+In our analysis, no strong relation between stamps and spatial distance have been detected in amphora workshops.
 The analysis suggests that there is not connection between stamps and the same amphora workshops, excluding certain exceptions when nearby workshops share the same amphoric stamp.
@@
 
-The results of the case study could be interpreted due to several reason according to previous hypotheses.
+The results of the case study could be interpreted by several reasons.
 On the one hand, the use of these amphoric stamps could have been exclusively running by the owner or owners of the workshop to distinguish the amphora workshop (CITAR).
@@
 Potters could have marked the amphorae to distinguish different groups working in parallel \citep{li_crossbows_2014}.
-This it could be explain wherefore we found different stamps in a same workshop.
+This it would explain wherefore why we detect different stamps in a same workshop.
 In any case, we do not have enough archaeological evidence that can validate the interpretations presented here and our results are certainly valid only with the context of our case study.
@@
 This method have identified differences in the case of the amphoric production within Roman Empire.
-Accordingly the results have highlighted for the interpretation of the complex economic processes connected with the archaeological evidence.
+Accordingly, the results have highlighted for the interpretation of the complex economic processes connected with the archaeological evidence.
 

### 2018-07-06 Paper/jaccardbaetica.tex
xavicomments

@@
 \newcommand{\revise}[1]{\memo{blue}{#1\\}}
+
+\newcommand{\xavi}[1]{\memo{magenta}{XRC: #1\\}}
 
@@
 Finally, results pretend to highlight the organization of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.
+
+\xavi{el abstract es demasiado espécifico al inicio y demasiado genérico después en relación a lo que se ha hecho}
 
@@
 
+\xavi{no digo nada porque entiendo que es una primera versión de intro no?}
+
 \section{The amphora production in the Roman Empire}
+
+\xavi{y si esta sección detalla directamente la producción amfórica de Dressel 20 en el Guadalquivir?}
 
@@
 
+\xavi{veo cierta confusión de secciones...si todo esto hasta ahora es intro, entonces deberían ser subsecciones?}
+
+\xavi{en todo caso debe quedar mucho más claro el objetivo del paper}
+
 \section{Material and Methods}
@@
 \end{figure}
+
+\xavi{guai el mapa pero el mapa pequeño debería ser como mínimo a nivel de Península Ibérica y, a ser posible, el imperio con provincias?}
 
@@
 \end{figure}
+
+\xavi{gráfico molón, y útil; haría los puntos y letras de los ejes más gordos solo}
+
+\xavi{entiendo que són códigos, no sellos? Estaría bien también mostrar cuántos sellos se encuentran por yacimiento.}
 %explicar por qué: mayor intensifación en las excavaciones quizás en sevilla
@@
 
+\xavi{este titulín es mejorable :-D}
+
 %quizás habría que hablar también del filtrado de códigos que se hizo en python porque antes había 3783 stamps pero se filtró por el número de letras...
+\xavi{lo del filtrado lo puedes poner de supplementary information con referencia al antiquity}
 
@@
 \end{figure}
+
+\xavi{no pillo demasiado lo que comentas de "groupings of similar stamps were not found in the cluster".}
+\xavi{Si usas esos colores aquí y en la figura anterior entonces deberías usarlos también en el mapa.}
 

### 2018-07-12 Paper/jaccardbaetica.tex
add new pic

@@
 The aim of this study is to detect the patterns of olive oil production that link amphora workshops and amphoric stamps.
-Roman provinces such as Baetica became important production and distribution centres during the Roman Empire.
-However, it remains under debate how this province was organized and whether it is possible to identify patterns in the olive oil market.
-
 Our case of study has been focused on the production processes located in Baetica province (currently Andalusia) from 1st to 3rd AD.
 In particular, we want to explore economic dynamics that include the production and distribution of olive oil trade.
+Baetica became important production and distribution centres during the Roman Empire.
+However, it remains under debate how this province was organized and whether it is possible to identify patterns in the olive oil market.
 Amphoric stamps are used to identify the presence of different producer groups that might share similar stamps.
@@
 Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}.
-As a result, an important mechanism of production control under Roman government was created spreading by different areas attending to the richness of different places.
+Given these factors, an important mechanism of production control under Roman government was created spreading by different areas attending to the richness of different places.
+Roman provinces such as Baetica (currently Andalusia, south Spain) became important production centres of olive oil during the Roman Empire.
+Olive oil was considered as the liquid gold since it was used in different aspect of the daily life as cooking, hygiene or lighting  \citep{mattingly_d.j._oil_1988}.
+As a consequence, this product was distributed in large amount of amphorae along the province, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.
 
@@
 
-\section{The amphora production in the Roman Empire}
+\section{The Amphoric production in Baetica province}
 
@@
 
-Roman provinces such as Baetica (currently Andalusia, south Spain) became important production centres of olive oil during the Roman Empire.
-Olive oil was considered as the liquid gold since it was used in different aspect of the daily life as cooking, hygiene or lighting  \citep{mattingly_d.j._oil_1988}.
-As a consequence, the high demand of Roman provinces stimulated by the good condition of the Baetica lands allowed to develop a massive infrastructure of olive oil production.
-This product was distributed in large amount of amphorae along the province, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.
-
+The high demand of Roman provinces stimulated by the good condition of the Baetica lands allowed to develop a massive infrastructure of olive oil production.
 The production and distribution of olive oil in this ancient province were growing exponentially during almost three centuries \citep{remesal_concierto}.
@@
 
-\subsection{Quantifying the diversity}
+\subsection{Quantifying the cultural diversity: Morisita-Horn index}
 
@@
 %quizás habría que hablar también del filtrado de códigos que se hizo en python porque antes había 3783 stamps pero se filtró por el número de letras...
-\xavi{lo del filtrado lo puedes poner de supplementary information con referencia al antiquity}
+%\xavi{lo del filtrado lo puedes poner de supplementary information con referencia al antiquity}
 
@@
 Nearby workshops show a similarity on the stamps while most of them seem to display different stamps roles.
-Additionally, groupings of similar stamps were not found in the cluster.
-The majority of stamp grouping was composed by no more than three workshops.
+Additionally, groups of workshops sharing similar stamps were not found in the cluster: the majority of stamp grouping was composed by no more than three workshops.
 Indeed workshops that shared more similar amphoric stamps belonged to the same \textit{conventus} area, such as Picachos, Cerro de los Pesebres and El Castillejo.

### 2018-07-12 Jaccard distance/codeCoto.R
add new line plot

@@ -79,7 +79,8 @@ pdf('frequency.pdf', width= 15, height=5)
 ggplot(sites, aes(y=reorder(region, foo, FUN=sum), x=foo, fill=region)) + geom_jitter(col="grey50", alpha=0.5, shape=21, height=0.2, width=0.3, size=3) + scale_colour_manual(values=myPalette) + theme_bw() + theme(legend.position="None") + xlab("number of code stamps") + ylab("region")
 dev.off()
 
-
+#para adaptarlo al gráfico dendrograma con los mismos colores. No se puede usar si tenemos la paleta scale_colour_manual
+ggplot(sites, aes(y=reorder(region, foo, FUN=sum), x=foo, fill=region)) + geom_jitter(col="grey50", alpha=0.5, shape=21, height=0.2, width=0.3, size=5) + scale_fill_manual(values=c("olivedrab","steelblue4","tomato")) + theme_bw() + theme(axis.text=element_text(size=13),legend.position="None") + xlab("Number of code stamps") + ylab("")
 
 
 

### 2018-07-13 Jaccard distance/codeCoto.R
add some lines

@@ -45,9 +45,9 @@ myData= subset(foo, region %in% c("Hispalis"))
 
 ###plot de puntitos para hacer frequencia
 
-
 library(ggplot2)
 
+stamp = read.csv("baezone.csv")
 ggplot(stamp, aes(y = region, x = code)) + geom_point()
 
 #library plyr (no confuse with dplyr) to calculate the frequency of stamps
@@ -55,7 +55,9 @@ count(stamp$code)
 
 #function table to see how many items have a variable table(stamp$code) for example
 
-##doing a matrix
+##doing a matrix 
+
+library(ggplot2)
 
 stamp = read.csv("baezone.csv")
 

### 2018-07-16 Paper/jaccardbaetica.tex
correct errors
modify number stamps

@@
 
-We studied a dataset of 3787 stamps collected from different Dressel 20 amphora workshops in Baetica province.
+We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in Baetica province.
 The stamp database was compiled by CEIPAC database \citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}).
@@
 Consequently, we discard integrate the fragmented stamps in our dataset.
-We finally filter a total sample of 987 stamps comprised of 131 different stamps from 81 workshops.
+We finally filter a total sample of 987 stamps comprised of 130 different stamps from 81 workshops.
+
+%3798 = base de datos sin limpiar
+%3791 = base de datos limpiada con cleanstamp.py
+%3787 = no sé a qué corresponde pero es archivo baetica.csv
 
@@
 
-%%\subsection{Jaccard distance}
-%The dataset was analysed using a statistic method as Jaccard distance.
+\subsection{Quantifying the cultural diversity}
+
+\subsubsection{Jaccard distance}
+
+The dataset was analysed using a statistic method as Jaccard distance.
 This method allows to measure the dissimilarity by calculating the presence of sets (CITAR).
@@
 
-\subsection{Quantifying the cultural diversity: Morisita-Horn index}
+(EXTENDER)
+
+\subsubsection{Morisita-Horn index}
 
@@
 
-\xavi{no pillo demasiado lo que comentas de "groupings of similar stamps were not found in the cluster".}
-\xavi{Si usas esos colores aquí y en la figura anterior entonces deberías usarlos también en el mapa.}
+%\xavi{no pillo demasiado lo que comentas de "groupings of similar stamps were not found in the cluster".}
+%\xavi{Si usas esos colores aquí y en la figura anterior entonces deberías usarlos también en el mapa.}
 

### 2018-07-25 Jaccard distance/codeCoto.R
add new lines to code

@@ -85,7 +85,25 @@ dev.off()
 ggplot(sites, aes(y=reorder(region, foo, FUN=sum), x=foo, fill=region)) + geom_jitter(col="grey50", alpha=0.5, shape=21, height=0.2, width=0.3, size=5) + scale_fill_manual(values=c("olivedrab","steelblue4","tomato")) + theme_bw() + theme(axis.text=element_text(size=13),legend.position="None") + xlab("Number of code stamps") + ylab("")
 
 
+##how to make a histogram with frequency of stamps 
 
+
+
+library(ggplot2)
+library(dplyr)
+
+foo = read.csv("baezone.csv")
+data = table(foo$site)
+#doing without ggplot
+hist(data,breaks=50)
+df = as.data.frame(data)
+ggplot(df, aes(x=Freq)) + geom_histogram()
+#to scale (poner más junto básicamente)
+ggplot(df, aes(x=Freq)) + geom_histogram()+scale_x_log10()
+
+#add color, alpha is the transparency
+
+ggplot(df, aes(x=Freq)) + geom_histogram(fill="blue", col="grey", alpha=.4) + scale_x_log10()
 
 
 

### 2018-07-26 Jaccard distance/codeCoto.R
add new lines to graphic

@@ -103,7 +103,10 @@ ggplot(df, aes(x=Freq)) + geom_histogram()+scale_x_log10()
 
 #add color, alpha is the transparency
 
-ggplot(df, aes(x=Freq)) + geom_histogram(fill="blue", col="grey", alpha=.4) + scale_x_log10()
+ggplot(df, aes(x=Freq)) + geom_histogram(bins=30, fill="blue", col="grey", alpha=.4) + scale_x_log10()
+ggplot(df, aes(x=Freq)) + geom_histogram(bins=30, fill="blue", col="grey", alpha=.3) + scale_x_log10() + labs(x="Number of Stamps", y="Frequency") + theme(panel.background = element_rect(fill = "white", colour = "grey50"))
+#other example is use + theme_classic()
+ggplot(df, aes(x=Freq)) + geom_histogram(bins=30, fill="lightcoral", col="gray25") + scale_x_log10() + labs(x="Number of Stamps", y="Frequency") + theme_classic()
 
 
 

### 2018-07-26 Paper/jaccardbaetica.tex
correct fig lines
add new graphic

@@
 We finally filter a total sample of 987 stamps comprised of 130 different stamps from 81 workshops.
+The frequency of stamps per workshop can be seen in Fig.
+\ref{stamps}.
+Most stamps are distributed in only one workshop whereas few workshops concentrate a high frequency of different amphoric stamps.
+
+\begin{figure}[htp]
+	\centering
+\includegraphics[width=\linewidth]{figs/frequencystamp.pdf}
+\caption{Distribution of the number of stamps}
+\label{stamps}
+\end{figure}
 
@@
 \includegraphics[width=\linewidth]{figs/frequency}
-\caption{Distribution of the number of stamps for each area. Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue).}
+\caption{Distribution of the number of different code stamps for each area. Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue).}
 \label{frequency}
@@
 
-\xavi{gráfico molón, y útil; haría los puntos y letras de los ejes más gordos solo}
+%\xavi{gráfico molón, y útil; haría los puntos y letras de los ejes más gordos solo}
 
-\xavi{entiendo que són códigos, no sellos? Estaría bien también mostrar cuántos sellos se encuentran por yacimiento.}
+%\xavi{entiendo que són códigos, no sellos? Estaría bien también mostrar cuántos sellos se encuentran por yacimiento.}
 %explicar por qué: mayor intensifación en las excavaciones quizás en sevilla

### 2018-10-09 Paper/jaccardbaetica.tex
add corrections home

@@
 The production and distribution of olive oil in this ancient province were growing exponentially during almost three centuries \citep{remesal_concierto}.
-As a result, thousands of amphora workshops were made to support this high demand of Roman Empire.
+As a result, hundreds of amphora workshops were made to support this high demand of Roman Empire.
 These workshops were located along the rivers Guadalquivir and Genil, supplying from the riverine connectivity to the Mediterranean and Atlantic routes \citep{garcia_vargas_enrique_formal_2010}.
@@
 The assumption of this analysis is that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops.
-By contrast, if we found similar amphoric stamps in different areas between workshops then the correlation between spatial distance would not be valued.
+By contrast, if we found similar amphoric stamps in different areas between workshops then the correlation between spatial distance and workshops would not be valued.
 
@@
 Morisita-Horn index can be expressed considering 0 as total presence of similarity of stamps and 1 a totally dissimilarity between stamps.
-In our case, it will be calculated the number of times that one stamp appear in an amphora workshop.
+In our case, it will be calculated the number of times that one stamp appears in an amphora workshop.
 This method allows to bear in mind the similar number of times for each repeated stamp per workshop.

### 2018-10-23 Province_data/military.py
add new folder new data work

@@ -1,0 +1,67 @@
+#!/usr/bin/python3
+
+import re
+
+# this script removes everything except for lat, long and code. The code only has alphabetic characters and all of them are finally transformed to uppercase.
+
+sep = ';'
+stamps = open("cleangerman.csv", "r")
+output = open("cleanedStamps1.csv", "w")
+output.write("id"+sep+"lat"+sep+"long"+sep+"type"+sep+"site"+sep+"code\n")
+
+listIds = list()
+listStamps = dict()    
+listTypes = list()
+listSites = list()
+
+
+
+for stamp in stamps:
+    stampParsed = stamp.split(',')
+    idStamp = stampParsed[1]
+    latStamp = stampParsed[5]
+    longStamp = stampParsed[6]
+    typeStamp = stampParsed[2]
+    siteStamp = stampParsed[3]
+
+    code = None
+    # if family
+    if len(stampParsed)==9:
+        code = stampParsed[8]
+    # else pick the base code        
+    else:
+        code = stampParsed[7]
+    # transform to upper
+    code = code.upper()
+    # remove everything except for A-Z
+    codeCleaned = re.sub(r'[^A-Z]+', '', code)
+    # remove empty codes
+    print("id",idStamp,"code cleaned ",codeCleaned)
+    if not codeCleaned:
+        continue
+
+    if idStamp in listIds:
+        previousCodes = listStamps[idStamp]
+        if codeCleaned in previousCodes:
+            continue
+    
+    # Attention! Only Dressel 20
+#    if typeStamp != "Dressel 20":
+#        continue
+
+
+    listIds.append(idStamp)
+    
+    if not idStamp in listStamps:
+        listStamps[idStamp] = list()
+
+    listStamps[idStamp].append(codeCleaned)
+
+    listTypes.append(typeStamp)
+    listSites.append(siteStamp)
+    output.write(idStamp+sep+latStamp+sep+longStamp+sep+typeStamp+sep+siteStamp+sep+codeCleaned+'\n')
+
+stamps.close()
+output.close()
+
+

### 2018-10-26 Paper/jaccardbaetica.tex
add some lines consumption cent

@@
 By contrast, stamps are the most studied in this type of amphorae.
-During almost three centuries, stamps were used to mark amphorae with a different chronological frequency \citep{remesal_sellar_2016}.
+The fascination aroused by stamps has caused it to be one of the inscriptions with the most publications about its origin and meaning.
+Its endure in time become it an essential element that were used to mark amphorae during almost three centuries \citep{remesal_sellar_2016}.
 Frequently, they were marked mainly in handles but seldom in rims and body \citep{millet_anforas_1998}.
@@
 There is not a general consensus about the meaning of the stamps \citep{rodriguez_baetican_1998}.
-It seems that stamps could have been identified as the land owner of the olive groves \citep{rodriguez_economioleicola_1977}.
-On the other hand, they could belong to the owner of the making amphorae workshops or even a production counting system  \citep{berni_millet_epigrafianforica_2008}.
+The meaning of these stamps could be identified based on two premises: content (olive oil) and context (amphora workshop).
+On the one hand, it seems that stamps could have been identified as the land owner of the olive groves \citep{rodriguez_economioleicola_1977}.
+On the other, they could belong to the owner of the making amphorae workshops or even a group of amphora workers (CITAR).
+Additionally, they have been related to a quality system \citep{callender} or a productin counting system \citep{berni_millet_epigrafianforica_2008}.
 In any case, the use of these stamps became in a good proxy to define somehow the system of working in the workshops.
@@
 %Hyphotesis
-Nevertheless, some challenges remain under discussion such as how this province was organized and whether it is possible to distinguish production patterns in the olive oil trade.
+Nevertheless, some challenges remain under discussion such as how this production was organized and whether it is possible to distinguish production patterns in the olive oil trade.
 Our questions will be focused on the distribution of amphoric stamps.
@@
 %hipotesis: si esto es así entonces es asá
-Here, therefore, this study proposes an evolutionary approach to explore the effect of Baetican olive oil production by computing of dissimilarity index.
-The assumption of this analysis is that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops.
-By contrast, if we found similar amphoric stamps in different areas between workshops then the correlation between spatial distance and workshops would not be valued.
+Here, therefore, this study proposes an evolutionary approach to explore the effect of Baetican olive oil production by computing the distribution of stamps.
+To do that, three case study have been studied to analyse the relation between production centres (Baetica) and consumption centres (Britannia and Germania).
+
+The assumption of this study is analysing the effect of the production patterns between different centres.
+This effect could affect due to different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamp in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province.
+In the case of Baetica province, we want to identify the role of the stamps in the organization of the workshop.
+In Roman provinces, our aim is detecting groups of stamps concentrated in a area or if some groups have an important role for the exportation of olive oil in those provinces.
 
@@
 
-\subsection{Case study}
+\subsection{Production centres: Baetica province}
 
 Our case study examines the relation between the distribution of amphoric stamps and the workshops.
-The amphora workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance (see Fig.\ref{workshop})
+In particular, we study the distribution of amphoric stamp with the aim of identifying a correlation between geographical distance and amphoric stamps.
+Based on this assumption, we proposed three hypotheses: a) it exists a correlation between spatial distance and the distribution of stamps, b) stamps located in close workshops share similar traits and c) Low mobility of amphoric stamps to other region: stamps always stay in the same region.
+
+The amphora workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance (see Fig.\ref{workshop}).
 
@@
 
-\subsection{Quantifying the cultural diversity}
+\subsection{Consumption centres}
+
+The conquest of new provinces allowed to the Roman Empire the arrival of new resources.
+This lead to a progressive change in the economic and social structure.
+As a result, Augusto's administration created the figure of the \textit{praefectura annonae} for the supply of wheat.
+The role of the praefectura annonae was mainly to provide though \textit{frumentationes} a quantity per month of wheat to Roman citizens CITAR (Remesal, 1986; 2011).
+
+\subsection{Britannia}
 

### 2018-10-26 Province_data/military.py
no change code just modify csv

@@ -5,8 +5,8 @@ import re
 # this script removes everything except for lat, long and code. The code only has alphabetic characters and all of them are finally transformed to uppercase.
 
 sep = ';'
-stamps = open("cleangerman.csv", "r")
-output = open("cleanedStamps1.csv", "w")
+stamps = open("allUK.csv", "r")
+output = open("cleanedStamps2.csv", "w")
 output.write("id"+sep+"lat"+sep+"long"+sep+"type"+sep+"site"+sep+"code\n")
 
 listIds = list()

### 2020-01-31 Paper/jaccardbaetica.tex
add new changes

@@
 
-\section{Material and Methods}
+\section{Material and methods}
 
@@
 
+\subsubsection{Material}
+
 We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in Baetica province.
@@
 We finally filter a total sample of 987 stamps comprised of 130 different stamps from 81 workshops.
-The frequency of stamps per workshop can be seen in Fig.
-\ref{stamps}.
-Most stamps are distributed in only one workshop whereas few workshops concentrate a high frequency of different amphoric stamps.
-
-\begin{figure}[htp]
-	\centering
-\includegraphics[width=\linewidth]{figs/frequencystamp.pdf}
-\caption{Distribution of the number of stamps}
-\label{stamps}
-\end{figure}
 
@@
 Dressel 20 stamps were found in three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter Hispalis), \textit{Cordubensis} (currently C\'ordoba, hereafter Corduba) and \textit{Astigi} (currently Écija, Sevilla, hereafter Astigi) \citep{rodriguez_economioleicola_1977,chic_datos_2001,berni_millet_epigrafianforica_2008} .
+
+\subsubsection{Exploratory Analysis}
+
+We calculated the frequency of distribution in amphora stamps with the aims to carry out a exploratory analysis (EDA) to know the distribution and the number of stamps per each centre.
+
+The frequency of stamps per workshop can be seen in Fig.
+\ref{stamps}.
+Most stamps are distributed in only one workshop whereas few workshops concentrate a high frequency of different amphoric stamps.
+The type of distribution is also frequent in Roman economy where we observed a self-organized complex system patterns typical of a free market \citep{coto-sarmiento_maria_bayesian_????}.
+
+\begin{figure}[htp]
+	\centering
+\includegraphics[width=\linewidth]{figs/frequencystamp.pdf}
+\caption{Distribution of the number of stamps}
+\label{stamps}
+\end{figure}
 
@@
 
-%\xavi{gráfico molón, y útil; haría los puntos y letras de los ejes más gordos solo}
-
-%\xavi{entiendo que són códigos, no sellos? Estaría bien también mostrar cuántos sellos se encuentran por yacimiento.}
-%explicar por qué: mayor intensifación en las excavaciones quizás en sevilla
-
-%explicar las razones del conventus no?
-%aquí hablar de sellos?
-
-%los sellos fueron marcados a partir del siglo II d.C (la economía oleícola bética Remesal) hablar de stams más concreto
-
-\subsection{Consumption centres}
+\subsection{Consumption centres: Britannia and Germania}
 
@@
 
-\subsection{Britannia}
+\subsubsection{Britannia}
 
-\subsubsection{Jaccard distance}
+\subsubsection{Material}
 
-The dataset was analysed using a statistic method as Jaccard distance.
+We analysed a dataset of 2219 stamps from different centres in Britannia. 
+%CITAR (Callender, 1965; Carreras Monfort y Funari,1998; Ayllón-Martı́ et al., 2018). 
+The stamp database was also compiled by CEIPAC database.
+The database contained anomalies that were eliminated using the same criteria cited above.
+We finally selected the centres with more or equal than 5 stamps.
+As a result, we studied a total of 1765 stamps comprised of 968 different stamps from 46 centres.
+The centres can be seen in Fig.\ref{britannia}.
+
+\begin{figure}[htp]
+	\centering
+\includegraphics[width=\linewidth]{figs/britmap.pdf}
+\caption{Distribution of centres in Britannia}
+\label{britannia}
+\end{figure}
+
+%\subsubsection{Jaccard distance}
+%The dataset was analysed using a statistic method as Jaccard distance.
 This method allows to measure the dissimilarity by calculating the presence of sets (CITAR).
@@
 
-(EXTENDER)
+\subsection{Measuring the dissimilarity in stamps}
 
-\subsubsection{Morisita-Horn index}
-
-\xavi{este titulín es mejorable :-D}
+%\xavi{este titulín es mejorable :-D}
 

### 2020-02-04 Paper/jaccardbaetica.tex
add changes germania brit

@@
 
+The consumption of olive oil in Britannia was residual until the start of the Roman conquest \citep{funari_corpus_1996,
+carreras_abastecimiento_2003}.
+
+It is well known that this product was not frequently consumed by autonomous population.
+The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of high percentage of Dressel 20 amphorae to Britannia \citep[ 1]{carreras_britannia_1998}.
+Additionally, the land of Britannia showed inadequate for the olive oil production due to low environmental conditions.
+This issue was necessary for creating an important export apparatus from Baetica to supply Roman legions.
+In this moment, we detect an increase of the olive oil exportation concurring with the displacement of Roman legions during the military campaigns \citep[ 161]{monfort_britanniaen_1998}.
+This fact will be spatial incidence in sites close at Hadrian Wall.
+Olive oil production in Baetica would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic points \citep{carreras_atlantic_2012}.
+The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges.
+Thus, the network was mainly focused on the support of soldiers during military campaigns.
+%debate intercambios puntuales o redes comerciales de abastecimiento
+
+The presence of Dressel 20 stamps in military camps in Britannia have been widely studied in Roman archaeology \citep{carreras_britannia_1998}.
+This fact also indicates a possible governmental organisation allocate to reforce and supply of olive oil in military camps, such as Germania (Remesal Rodrıguez, 1986).
+However, it is unknown the type of economical system when managing or control the redistribution of olive oil in different sites in Britannia but it seems that it could have been in the main cities \citep[45]{funari_economic_2005}.
+
+Finally, The increase in olive oil exports would experience a progressive slowdown
+from the third century A.D., coinciding with the change in market strategy in the
+Empire. In that moment, a progressive dicrease of Dressel 20 is documented being gradually replace for Dressel 23 amphorae.
+
 \subsubsection{Material}
@@
 As a result, we studied a total of 1765 stamps comprised of 968 different stamps from 46 centres.
-The centres can be seen in Fig.\ref{britannia}.
+The centres located in Britannia can be seen in Fig.\ref{britannia}.
 
@@
 \label{britannia}
+\end{figure}
+
+\subsubsection{Germania}
+
+The Roman conquest in Germania dates from the end of the 1st century BC, during Caesar campaign and subsequently continued by Augusto.
+The advance of the Roman legions toward Gallia opened a way though the Atlantic Ocean to Germania \citep{remesal_annona_1986,
+remesal_baetica_2002}.
+
+It is thought that this route would have been initiated previously through 
+terrestrial transport, although recent studies question this position giving an important role to the Atlantic route for both the conquests of Germania and those of Britannia \citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
+
+As the case of Britain, the consumption of olive oil in Germania was not detected before the conquest.
+
+The need to supply roman legions created a logistic network to supply to the army.
+In this moment, we detect the arrival of olive oil amphorae to Germania evidenced by the presence of Dressel 20 amphora in different sites close to limes.
+
+The roman studies about the presence of Dressel 20 amphorae has not been the same repercusion than the rest of the province. This can be explained due to the lack of data \citep{horacio2010llegada}.
+
+As a consequence, we do not the role that commerce agents in the participation of the distribution and exportation of olive oil in Germania or, on the contrary, the Roman army was the responsible to supply the product in the own province \citep[156]{remesal_germn_2010}.
+
+The presence of Roman army encouraged the exchange in the province showed by the arrival of this product both civil settlement and military sites with a mayor concentration at german limes.
+It seems that some baetican centres would be assigned to the support of olive oil.
+However, this hypothesis can not be proven by the lack of archaeological sources \citep[125]{remesal_concierto}.
+
+\subsubsection{Material}
+
+We analysed a dataset of 2052 stamps placed in Germania.
+All the data was also compiled by CEIPAC database.
+We finally selected a total of 1621 stamps consist of 850 different stamps from 46 sites (see Fig.
+\ref{germania}).
+As previously before, the centres were selected with more or equal than 5 stamps.
+
+\begin{figure}[htp]
+	\centering
+\includegraphics[width=\linewidth]{figs/germania5.pdf}
+\caption{Distribution of the sites in Germania. Most sites were located in limes during the Roman Empire}
+\label{germania}
 \end{figure}
@@
 \subsection{Measuring the dissimilarity in stamps}
-
-%\xavi{este titulín es mejorable :-D}
 
@@
 
+\subsection{Production centres: Baetica Province}
+
 The analysis shows that amphoric stamps could be correlated with the spatial distance.
@@
 
-%\xavi{no pillo demasiado lo que comentas de "groupings of similar stamps were not found in the cluster".}
-%\xavi{Si usas esos colores aquí y en la figura anterior entonces deberías usarlos también en el mapa.}
+\subsection{Reception centres: Britannia and Germania}
 
-%%\subsection{Jaccard distance}
+\subsubsection{Britannia}
 
-%Result of Jaccard distance can be The coefficients range from x to x.The dendrogram shown in Fig. Bla bla was obtained with the Jaccard distance measure. The dendrogram suggests
+The results show a correlation similar to the obtained in Baetica province. However, the similarity is less pronounced as we observed in the dendrogram in Fig. \ref{britmap}
+
+\begin{figure}[htp]
+	\centering
+\includegraphics[width=\linewidth]{figs/dendrobrit5.pdf}
+\caption{Dendrogram obtained by Morisita-Horn algorthm of different sites in Britannia.
+Colours are represented by type of sites divided into military sites (red), civil sites (green) and no specific (grey)}
+\label{britmap}
+\end{figure}
+
+A minor correlation can be explained by the fact that Britannia spatial area is much wide than \textit{Baetica} where the centres were more concentrated.
+
+Most military sites that showed greater similarity in the results were geographically close.
+We identified two factors in the results: 1) nearby sites tend to share similar stamps and 2) most sites have different stamps.
+By contrast, most sites did not show a strong similarity in stamp correlation .
+Neither we found a grouping of similar stamps in a specific place.
+In general, we did not observe any production pattern that indicates a clear organisation between production centres and reception centres.
+In other words, our results did not indicate the presence of a specific production center from Baetica in Britannia province.
+Rather BLA BLA
+
+\subsection{Germania}
+
+The similarity of stamps show a results less significative than Britannia province as it can be seen in the dendrogram Fig. \ref{germap}.
+
+\begin{figure}[htp]
+	\centering
+\includegraphics[width=\linewidth]{figs/dendroger5.pdf}
+\caption{Dendrogram obtained by Morisita-Horn algorthm of different sites in Germania.
+Colours are represented by type of sites divided into military sites (red), civil sites (green) and no specific (grey)}
+\label{germap}
+\end{figure}
+
+The sites with more similarity in stamps were closely located in limes.
+In this case, most sites have different stamps  which means that we did not detect any pattern.
+On the other hand, there seems to be a greater concentration of the seals in areas eminently militarized and close to the german limes.
+However, It is still no possible to determine with the data that the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.
 
@@
 
-\bibliography{bibliotex}
+%\bibliography{bibliotex}
+\bibliography{bibtesis}
 

### 2020-02-07 Paper/jaccardbaetica.tex
new changes intro final

@@
 
-% hipotesis
+\subsection{The Amphoric production in Baetica province}
 
-% Debate
-
-%más específico--- olive oil
-%incluir algo que enlace con roman province
-
-\xavi{no digo nada porque entiendo que es una primera versión de intro no?}
-
-\section{The Amphoric production in Baetica province}
-
-\xavi{y si esta sección detalla directamente la producción amfórica de Dressel 20 en el Guadalquivir?}
+%\xavi{y si esta sección detalla directamente la producción amfórica de Dressel 20 en el Guadalquivir?}
 
@@
 
-The majority of olive oil amphorae produced in this province and shipped through the provinces belong to Dressel 20 type \citep{dressel,martin-kilcher_romischen_1994}.
+The majority of olive oil amphorae produced in this province and shipped through the provinces belong to Dressel 20 type \citep{dressel_ricerche_1878,
+millet_anforas_1998}.
 This amphora is commonly associated with the transportation of Baetican olive oil during the Roman Empire \citep{berni_millet_epigrafianforica_2008}.
@@
 However, inscriptions and \textit{tituli picti} have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
-By contrast, stamps are the most studied in this type of amphorae.
+
+\subsection{A potential indicator of Roman economy}
+
+Stamps are the most studied in this type of amphorae.
 The fascination aroused by stamps has caused it to be one of the inscriptions with the most publications about its origin and meaning.
@@
 %hipotesis: si esto es así entonces es asá
-Here, therefore, this study proposes an evolutionary approach to explore the effect of Baetican olive oil production by computing the distribution of stamps.
-To do that, three case study have been studied to analyse the relation between production centres (Baetica) and consumption centres (Britannia and Germania).
 
 The assumption of this study is analysing the effect of the production patterns between different centres.
-This effect could affect due to different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamp in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province.
-In the case of Baetica province, we want to identify the role of the stamps in the organization of the workshop.
+We are specially interested in identifying links between producers and consumption centres using amphoric stamps.
+To do that, we use the CEIPAC database to collect the stamps from different places.
+The CEIPAC dataset contains over 50.000 of epigraphy records found in amphorae, mostly from \textit{Monte Testaccio}.
+
+Three case study have been studied in order to analyse the relation between production centres (Baetica) and consumption centres (Britannia and Germania).
+In the case of Baetica province, we want to identify the role of the stamps in the organisation of the workshop.
 In Roman provinces, our aim is detecting groups of stamps concentrated in a area or if some groups have an important role for the exportation of olive oil in those provinces.
+This economic connection could affect due to different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamp in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province.
 
-%resumen del paper
-
-%hablar de las tres áreas de producción: hispalis, corduba y astigi
+Therefore, this study proposes an evolutionary baseline to explore the distribution of Baetican olive oil production by computing spatial correlation between stamps.
+A way to analyse is by using a quantitative framework to measure the similarity.
+Here we use a ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potential spatial correlation and c)to establish a correlation between similarity of stamp and spatial distance.
 

### 2020-02-12 Paper/jaccardbaetica.tex
new changes bib

@@
 
+\begin{frontmatter}
+
 \title{The markings of the trade: exploring the patterns of olive oil production in Roman Baetica}
 
-\begin{frontmatter}
+\author[ceipacadress]{Maria Coto-Sarmiento\corref{mycorrespondingauthor}}
+\cortext[mycorrespondingauthor]{Corresponding author}
+\ead{mcotsar@gmail.com}
+
+\author[edadress]{Xavier Rubio-Campillo}
+
+\address[edadress]{School of History, Classic \& Archaeology, Room OOM.33, William Robertson Wing, Old Medical School, Teviot Place, University of Edinburgh, UK}
+\address[ceipacadress]{CEIPAC, Department of Prehistory and Archaeology, Montalegre, 6-8, 08001, University of Barcelona, Barcelona, Spain}
+
+\begin{keyword}
+Roman Empire; amphora workshops; Dressel 20; social learning; cultural evolution
+\end{keyword}
+
+\end{frontmatter}
 
@@
 
-The aim of this study is to detect the patterns of olive oil production that link amphora workshops and amphoric stamps.
+The aim of this study is to explore economics dynamics in the production and distribution of olive oil trade. 
 Our case of study has been focused on the production processes located in Baetica province (currently Andalusia) from 1st to 3rd AD.
-In particular, we want to explore economic dynamics that include the production and distribution of olive oil trade.
+In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps.
 Baetica became important production and distribution centres during the Roman Empire.
-However, it remains under debate how this province was organized and whether it is possible to identify patterns in the olive oil market.
+However, it remains under debate how this province was organised and whether it is possible to identify patterns in the olive oil market.
 Amphoric stamps are used to identify the presence of different producer groups that might share similar stamps.
@@
 The analysis explores how quantitative approach provides a useful tool for the interpretation of the economic processes.
-Finally, results pretend to highlight the organization of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.
-
-\xavi{el abstract es demasiado espécifico al inicio y demasiado genérico después en relación a lo que se ha hecho}
+Finally, results pretend to highlight the organisation of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.
 
 \end{abstract}
-
-\end{frontmatter}
 
@@
 Despite this issue has been widely discussed over many decades, the understanding about the production processes is still under debate, mostly due to the lack of written records.
-The application of different approaches combined with the archaeological evidence has allowed us partly to deal with the complexity of understanding the Roman production \citep{orengo_seeds_2016,brughmans_roman_2016,coto-sarmiento_maria_bayesian_????}.
+The application of different quantitative approaches combined with the archaeological evidence has allowed us partly to deal with the complexity of understanding the Roman production \citep{orengo_seeds_2016,
+brughmans_roman_2016,coto-sarmiento_identifying_2018,bayesian_2018}.
 
@@
 
-\xavi{guai el mapa pero el mapa pequeño debería ser como mínimo a nivel de Península Ibérica y, a ser posible, el imperio con provincias?}
-
 The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
 Some stamps show a more specific chronology while the majority of them display a large activity of production that it can be difficult for specifying an accurate chronology.
-This could be due to two reasons; firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible; secondly, Dressel 20 was produced during almost three centuries with apparently few changes.
-(podría citar nuestro paper por toda la face).
+This could be due to two reasons; firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible; secondly, Dressel 20 was produced during almost three centuries with apparently few changes \citep{berni_piero_chapter_2017}.
 
@@
 These areas, known as \textit{conventus}, were administrative centres for territorial organization in the Roman Empire.
-Dressel 20 stamps were found in three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter Hispalis), \textit{Cordubensis} (currently C\'ordoba, hereafter Corduba) and \textit{Astigi} (currently Écija, Sevilla, hereafter Astigi) \citep{rodriguez_economioleicola_1977,chic_datos_2001,berni_millet_epigrafianforica_2008} .
+Dressel 20 stamps were found in three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter Hispalis), \textit{Cordubensis} (currently C\'ordoba, hereafter Corduba) and \textit{Astigi} (currently Écija, Sevilla, hereafter Astigi) \citep{rodriguez_economioleicola_1977,
+chicdatos2001,berni_millet_epigrafianforica_2008} .
 
@@
 Most stamps are distributed in only one workshop whereas few workshops concentrate a high frequency of different amphoric stamps.
-The type of distribution is also frequent in Roman economy where we observed a self-organized complex system patterns typical of a free market \citep{coto-sarmiento_maria_bayesian_????}.
+The type of distribution is also frequent in Roman economy where we observed a self-organized complex system patterns typical of a free market \citep{bayesian_2018}.
 
@@
 This issue was necessary for creating an important export apparatus from Baetica to supply Roman legions.
-In this moment, we detect an increase of the olive oil exportation concurring with the displacement of Roman legions during the military campaigns \citep[ 161]{monfort_britanniaen_1998}.
+In this moment, we detect an increase of the olive oil exportation concurring with the displacement of Roman legions during the military campaigns \citep[161]{monfort_britanniaen_1998}.
 This fact will be spatial incidence in sites close at Hadrian Wall.
@@
 
-The similarity of stamps show a results less significative than Britannia province as it can be seen in the dendrogram Fig. \ref{germap}.
+The similarity of stamps show a results less significant than Britannia province as it can be seen in the dendrogram (Fig. \ref{germap}).
 
@@
 \caption{Dendrogram obtained by Morisita-Horn algorthm of different sites in Germania.
-Colours are represented by type of sites divided into military sites (red), civil sites (green) and no specific (grey)}
+Colours are represented by types of sites divided into military sites (red), civil sites (green) and no specific (grey)}
 \label{germap}
@@
 On the one hand, the use of these amphoric stamps could have been exclusively running by the owner or owners of the workshop to distinguish the amphora workshop (CITAR).
-This hypothesis would explain the fact that we do not find similar stamps in different workshops, however, we found different amphora stamps in the same workshops that they would be barely difficult to assign different owners.
-Neither a quality distinction does seem to be used to specify the value of the product \citep{callender}.
-On the other hand, it could be interpreted somehow a batch systematic organization.
+This hypothesis would explain the fact that we do not find similar stamps in different workshops; however, we found different amphora stamps in the same workshops that they would be barely difficult to assign different owners.
+Neither a quality distinction does seem to be used to specify the value of the product \citep{callenderindex}.
+On the other hand, it could be interpreted somehow a batch systematic organisation.
 Considering that Dressel 20 was not marked in most cases, some authors point out that potters marked amphorae to prepare and distribute the commodity to be shipped \citep{berni_millet_epigrafianforica_2008}.
-This method would be used as an identifier to count the number of amphorae of a branch (CITAR).
+This method would be used as an identifier to count the number of amphorae of a branch \citep{juanmorostesis}.
 This method could also have served to identify different groups of potter workers working in the same amphora workshop.

### 2020-02-13 Paper/jaccardbaetica.tex
add new lines

@@
 
-\subsection{A potential indicator of Roman economy}
+\subsection{A potential indicator of Roman economy: Dressel 20 stamps}
 
@@
 Its endure in time become it an essential element that were used to mark amphorae during almost three centuries \citep{remesal_sellar_2016}.
+(Fig.\ref{amphora}).
+
+\begin{figure}[htp]
+	\centering
+\includegraphics[scale=0.5]{figs/dressel20}
+\caption{Dressel 20 were mostly marked with stamps of three letters called \textit{tria nomina}}
+\label{amphora}
+\end{figure}
+
 Frequently, they were marked mainly in handles but seldom in rims and body \citep{millet_anforas_1998}. 
@@
 Stamps are mostly formed by a code of three letters and they can appear in a abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
+Despite researchers identify stamps as a identity marks, there is not a general consensus about the meaning of the stamps \citep{rodriguez_baetican_1998}.
+Alike most Dressel 20 containers were not stamped being difficult to determinate.
 
-%Many hypothesis involved the meaning of stamps. 
-There is not a general consensus about the meaning of the stamps \citep{rodriguez_baetican_1998}.
-The meaning of these stamps could be identified based on two premises: content (olive oil) and context (amphora workshop).
+They could be identified based on three premises: content (olive oil), context (amphora workshop) and subject (individuals involved).
 On the one hand, it seems that stamps could have been identified as the land owner of the olive groves \citep{rodriguez_economioleicola_1977}.
 On the other, they could belong to the owner of the making amphorae workshops or even a group of amphora workers (CITAR).
-Additionally, they have been related to a quality system \citep{callender} or a productin counting system \citep{berni_millet_epigrafianforica_2008}.
+Additionally, they have been related to a quality system \citep{callender} or a production counting system \citep{berni_millet_epigrafianforica_2008}.
 In any case, the use of these stamps became in a good proxy to define somehow the system of working in the workshops.

### 2020-02-26 Paper/jaccardbaetica.tex
update wrong map

@@
 The fascination aroused by stamps has caused it to be one of the inscriptions with the most publications about its origin and meaning.
-Its endure in time become it an essential element that were used to mark amphorae during almost three centuries \citep{remesal_sellar_2016}.
-(Fig.\ref{amphora}).
+Its endure in time become it an essential element that were used to mark amphorae during almost three centuries \citep{remesal_sellar_2016} (Fig.\ref{amphora}).
 
@@
 %this paper aims to highlight the production dynamic in connection with the proximity between amphora workshops and stamps.
-
-\xavi{veo cierta confusión de secciones...si todo esto hasta ahora es intro, entonces deberían ser subsecciones?}
-
-\xavi{en todo caso debe quedar mucho más claro el objetivo del paper}
 
@@
 \includegraphics[width=\linewidth]{figs/dendroger5.pdf}
-\caption{Dendrogram obtained by Morisita-Horn algorthm of different sites in Germania.
+\caption{Dendrogram obtained by Morisita-Horn algorithm of different sites in Germania.
 Colours are represented by types of sites divided into military sites (red), civil sites (green) and no specific (grey)}
@@
 In this case, most sites have different stamps  which means that we did not detect any pattern.
-On the other hand, there seems to be a greater concentration of the seals in areas eminently militarized and close to the german limes.
+On the other hand, there seems to be a greater concentration of the seals in areas eminently militarized and close to the German limes.
 However, It is still no possible to determine with the data that the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.
@@
 
-%soltar aquí las hipótesis repito aquí several reasons no repetir
-
 The results of the case study could be interpreted by several reasons.
@@
 The research was funded by European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros and Ignacio Morer for their useful suggestions.  
-All data has been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}. Source and code are available
+All data has been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}.
+Source and code are available at \maria{incluir github}.
 

### 2020-03-05 Paper/jaccardbaetica.tex
add new changes

@@
 \newcommand{\maria}[1]{\memo{red}{#1\\}}
-\newcommand{\revise}[1]{\memo{blue}{#1\\}}
-
 \newcommand{\xavi}[1]{\memo{magenta}{XRC: #1\\}}
@@
 On the one hand, it seems that stamps could have been identified as the land owner of the olive groves \citep{rodriguez_economioleicola_1977}.
-On the other, they could belong to the owner of the making amphorae workshops or even a group of amphora workers (CITAR).
-Additionally, they have been related to a quality system \citep{callender} or a production counting system \citep{berni_millet_epigrafianforica_2008}.
+On the other, they could belong to the owner of the making amphorae workshops or even a group of amphora workers \citep{berni_millet_epigrafianforica_2008}.
 In any case, the use of these stamps became in a good proxy to define somehow the system of working in the workshops.
@@
 As a result, Augusto's administration created the figure of the \textit{praefectura annonae} for the supply of wheat.
-The role of the praefectura annonae was mainly to provide though \textit{frumentationes} a quantity per month of wheat to Roman citizens CITAR (Remesal, 1986; 2011).
+The role of the \textit{praefectura annonae} was mainly to provide though \textit{frumentationes} a quantity per month of wheat to Roman citizens \citep{remesal_annona_1986,remesal_concierto}.
 
@@
 
-The consumption of olive oil in Britannia was residual until the start of the Roman conquest \citep{funari_corpus_1996,
+The consumption of olive oil in Britannia was residual until starting the Roman conquest \citep{funari_corpus_1996,
 carreras_abastecimiento_2003}.
@@
 We identified two factors in the results: 1) nearby sites tend to share similar stamps and 2) most sites have different stamps.
-By contrast, most sites did not show a strong similarity in stamp correlation .
+By contrast, most sites did not show a strong similarity in stamp correlation.
 Neither we found a grouping of similar stamps in a specific place.
@@
 In other words, our results did not indicate the presence of a specific production center from Baetica in Britannia province.
-Rather BLA BLA
+Rather it seems that olive oil production was distributed by non specific production centres.
 
@@
 
-The similarity of stamps show a results less significant than Britannia province as it can be seen in the dendrogram (Fig. \ref{germap}).
+The similarity of stamps shows a results less significant than Britannia province as it can be seen in the dendrogram (Fig. \ref{germap}).
 
@@
 In this case, most sites have different stamps  which means that we did not detect any pattern.
-On the other hand, there seems to be a greater concentration of the seals in areas eminently militarized and close to the German limes.
+On the other hand, there seems to be a greater concentration of the seals in areas eminently militarised and close to the German limes.
 However, It is still no possible to determine with the data that the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.
@@
 The results of the case study could be interpreted by several reasons.
-On the one hand, the use of these amphoric stamps could have been exclusively running by the owner or owners of the workshop to distinguish the amphora workshop (CITAR).
+On the one hand, the use of these amphoric stamps could have been exclusively running by the owner or owners of the workshop to distinguish the amphora workshop.
 This hypothesis would explain the fact that we do not find similar stamps in different workshops; however, we found different amphora stamps in the same workshops that they would be barely difficult to assign different owners.
-Neither a quality distinction does seem to be used to specify the value of the product \citep{callenderindex}.
 On the other hand, it could be interpreted somehow a batch systematic organisation.

### 2020-03-10 Paper/jaccardbaetica.tex
review abstract and intro

@@
 
-\end{frontmatter}
-
 %cambiar quizás
@@
 In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps.
-Baetica became important production and distribution centres during the Roman Empire.
-However, it remains under debate how this province was organised and whether it is possible to identify patterns in the olive oil market.
-Amphoric stamps are used to identify the presence of different producer groups that might share similar stamps.
-To achieve this goal, we analyse a set of stamps from different workshops in Baetica province in order to detect a relation between the distribution of amphoric stamps and the economic structure in this province.
+Baetica became important production and distribution centre during the Roman Empire.
+However, it remains under debate how this province was organised and whether it could be possible to identify patterns in the olive oil market.
+Amphoric stamps are used to identify the presence of different groups that might share similar stamps.
+To achieve this goal, we analyse a set of stamps from two centres: 1) producer centres by analysing different workshops in Baetica province and 2) two Roman provinces such as Germania and Britannia as a receptor centres.
+They will be used to detect a relation between the distribution of amphoric stamps and the economic structure in both centres.
 Here we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance.
@@
 
+\end{frontmatter}
+
 \section{Introduction}
@@
 Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}.
-Given these factors, an important mechanism of production control under Roman government was created spreading by different areas attending to the richness of different places.
+Given this fact, an important mechanism of production control under Roman government was created spreading by different areas attending to the richness of different places.
 Roman provinces such as Baetica (currently Andalusia, south Spain) became important production centres of olive oil during the Roman Empire.
 Olive oil was considered as the liquid gold since it was used in different aspect of the daily life as cooking, hygiene or lighting  \citep{mattingly_d.j._oil_1988}.
-As a consequence, this product was distributed in large amount of amphorae along the province, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.
+As a consequence, this product was distributed in a large number of amphorae along the province, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.
 
@@
 %This paper explore--- resumir en dos líneas el objeto de estudio y definir que ha sido usado para ello dispersion of culture
-This paper aims to understand the production dynamics in connection with a specific area within Roman Empire.
+This paper aims to understand the production dynamics in connection with a different areas within Roman Empire.
 Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}.
 We want to detect the pattern of olive oil production that link amphora workshops and amphoric stamps used to mark them.
-We focus here on exploring the economic relation between stamps and amphora production centres.
+We focus here on exploring the economic relation between stamps and amphora production and distribution centres.
 To do this, an Ecology approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}.
+
+\maria{deberia cerrar esto}
 
@@
 
-%\xavi{y si esta sección detalla directamente la producción amfórica de Dressel 20 en el Guadalquivir?}
-
 The high demand of Roman provinces stimulated by the good condition of the Baetica lands allowed to develop a massive infrastructure of olive oil production.
-The production and distribution of olive oil in this ancient province were growing exponentially during almost three centuries \citep{remesal_concierto}.
+The production and distribution of olive oil were growing exponentially during almost three centuries \citep{remesal_concierto}.
 As a result, hundreds of amphora workshops were made to support this high demand of Roman Empire.
@@
 This amphora is commonly associated with the transportation of Baetican olive oil during the Roman Empire \citep{berni_millet_epigrafianforica_2008}.
-Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti} with different information.
-However, inscriptions and \textit{tituli picti} have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
+Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti} with different information still unknown.
+Indeed most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
 

### 2020-03-19 Paper/jaccardbaetica.tex
add new changes

@@
 The aim of this study is to explore economics dynamics in the production and distribution of olive oil trade. 
-Our case of study has been focused on the production processes located in Baetica province (currently Andalusia) from 1st to 3rd AD.
+Our case of study has been focused on the production processes located in \textit{Baetica} province (currently Andalusia) from 1st to 3rd AD.
 In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps.
-Baetica became important production and distribution centre during the Roman Empire.
+\textit{Baetica} became important production and distribution centre during the Roman Empire.
 However, it remains under debate how this province was organised and whether it could be possible to identify patterns in the olive oil market.
 Amphoric stamps are used to identify the presence of different groups that might share similar stamps.
-To achieve this goal, we analyse a set of stamps from two centres: 1) producer centres by analysing different workshops in Baetica province and 2) two Roman provinces such as Germania and Britannia as a receptor centres.
+To achieve this goal, we analyse a set of stamps from two centres: 1) producer centres by analysing different workshops in \textit{Baetica} province and 2) two Roman provinces such as \textit{Germania} and \textit{Britannia} as a receptor centres.
 They will be used to detect a relation between the distribution of amphoric stamps and the economic structure in both centres.
@@
 Given this fact, an important mechanism of production control under Roman government was created spreading by different areas attending to the richness of different places.
-Roman provinces such as Baetica (currently Andalusia, south Spain) became important production centres of olive oil during the Roman Empire.
+Roman provinces such as \textit{Baetica} (currently Andalusia, south Spain) became important production centres of olive oil during the Roman Empire.
 Olive oil was considered as the liquid gold since it was used in different aspect of the daily life as cooking, hygiene or lighting  \citep{mattingly_d.j._oil_1988}.
@@
 
-The high demand of Roman provinces stimulated by the good condition of the Baetica lands allowed to develop a massive infrastructure of olive oil production.
+The high demand of Roman provinces stimulated by the good condition of the \textit{Baetica} lands allowed to develop a massive infrastructure of olive oil production.
 The production and distribution of olive oil were growing exponentially during almost three centuries \citep{remesal_concierto}.
@@
 %Hyphotesis
-Nevertheless, some challenges remain under discussion such as how this production was organized and whether it is possible to distinguish production patterns in the olive oil trade.
+Nevertheless, some challenges remain under discussion such as how this production was organised and whether it is possible to distinguish production patterns in the olive oil trade.
 Our questions will be focused on the distribution of amphoric stamps.
@@
 Did stamps share the same workshop?
-Neither written records have been found that it could explain the economic role of Baetica province in the Roman organization.
-Indeed, archaeological evidence shows a highly specialized production with a long activity in a this specific area with apparently few changes \citep{remesal_anforas_2004}.
+Neither written records have been found that it could explain the economic role of \textit{Baetica} province in the Roman organisation.
+Indeed, archaeological evidence shows a highly specialised production with a long activity in a this specific area with apparently few changes \citep{remesal_anforas_2004}.
 
@@
 
-Three case study have been studied in order to analyse the relation between production centres (Baetica) and consumption centres (Britannia and Germania).
-In the case of Baetica province, we want to identify the role of the stamps in the organisation of the workshop.
+Three case study have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}).
+In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop.
 In Roman provinces, our aim is detecting groups of stamps concentrated in a area or if some groups have an important role for the exportation of olive oil in those provinces.
@@
 
-The amphora workshops were situated in different locations in Baetica province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance (see Fig.\ref{workshop}).
+The amphora workshops were situated in different locations in \textit{Baetica} province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance (see Fig.\ref{workshop}).
 
@@
 	\centering
-\includegraphics[width=\linewidth]{figs/workshop}
+\includegraphics[width=\linewidth]{figs/workshop1}
 \caption{Distribution of the workshops}
@@
 
-We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in Baetica province.
+We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in \textit{Baetica} province.
 The stamp database was compiled by CEIPAC database \citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}).
@@
 
-The consumption of olive oil in Britannia was residual until starting the Roman conquest \citep{funari_corpus_1996,
+The consumption of olive oil in \textit{Britannia} was residual until starting the Roman conquest \citep{funari_corpus_1996,
 carreras_abastecimiento_2003}.
@@
 It is well known that this product was not frequently consumed by autonomous population.
-The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of high percentage of Dressel 20 amphorae to Britannia \citep[ 1]{carreras_britannia_1998}.
-Additionally, the land of Britannia showed inadequate for the olive oil production due to low environmental conditions.
-This issue was necessary for creating an important export apparatus from Baetica to supply Roman legions.
+The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}.
+Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions.
+This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman legions.
 In this moment, we detect an increase of the olive oil exportation concurring with the displacement of Roman legions during the military campaigns \citep[161]{monfort_britanniaen_1998}.
 This fact will be spatial incidence in sites close at Hadrian Wall.
-Olive oil production in Baetica would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic points \citep{carreras_atlantic_2012}.
+Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic points \citep{carreras_atlantic_2012}.
 The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges.
@@
 
-The presence of Dressel 20 stamps in military camps in Britannia have been widely studied in Roman archaeology \citep{carreras_britannia_1998}.
-This fact also indicates a possible governmental organisation allocate to reforce and supply of olive oil in military camps, such as Germania (Remesal Rodrıguez, 1986).
-However, it is unknown the type of economical system when managing or control the redistribution of olive oil in different sites in Britannia but it seems that it could have been in the main cities \citep[45]{funari_economic_2005}.
+The presence of Dressel 20 stamps in military camps in \textit{Britannia} have been widely studied in Roman archaeology \citep{carreras_britannia_1998}.
+This fact also indicates a possible governmental organisation allocate to reforce and supply of olive oil in military camps, such as \textit{Germania} (Remesal Rodrıguez, 1986).
+However, it is unknown the type of economical system when managing or control the redistribution of olive oil in different sites in \textit{Britannia} but it seems that it could have been in the main cities \citep[45]{funari_economic_2005}.
 
@@
 from the third century A.D., coinciding with the change in market strategy in the
-Empire. In that moment, a progressive dicrease of Dressel 20 is documented being gradually replace for Dressel 23 amphorae.
+Empire. In that moment, a progressive decrease of Dressel 20 is documented being gradually replace for Dressel 23 amphorae.
 
@@
 
-We analysed a dataset of 2219 stamps from different centres in Britannia. 
+We analysed a dataset of 2219 stamps from different centres in \textit{Britannia}. 
 %CITAR (Callender, 1965; Carreras Monfort y Funari,1998; Ayllón-Martı́ et al., 2018). 
@@
 As a result, we studied a total of 1765 stamps comprised of 968 different stamps from 46 centres.
-The centres located in Britannia can be seen in Fig.\ref{britannia}.
+The centres located in \textit{Britannia} can be seen in Fig.\ref{britannia}.
 
@@
 	\centering
-\includegraphics[width=\linewidth]{figs/britmap.pdf}
-\caption{Distribution of centres in Britannia}
+\includegraphics[width=\linewidth]{figs/britmap2.pdf}
+\caption{Distribution of centres in \textit{Britannia}}
 \label{britannia}
@@
 
-The Roman conquest in Germania dates from the end of the 1st century BC, during Caesar campaign and subsequently continued by Augusto.
-The advance of the Roman legions toward Gallia opened a way though the Atlantic Ocean to Germania \citep{remesal_annona_1986,
+The Roman conquest in \textit{Germania} dates from the end of the 1st century BC, during Caesar campaign and subsequently continued by Augusto.
+The advance of the Roman legions toward \textit{Gallia} opened a way though the Atlantic Ocean to \textit{Germania} \citep{remesal_annona_1986,
 remesal_baetica_2002}.
 
-It is thought that this route would have been initiated previously through 
-terrestrial transport, although recent studies question this position giving an important role to the Atlantic route for both the conquests of Germania and those of Britannia \citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
+It is thought that this route would have been initiated previously through terrestrial transport, although recent studies question this position giving an important role to the Atlantic route for both the conquests of \textit{Germania} and those of \textit{Britannia} \citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
 
-As the case of Britain, the consumption of olive oil in Germania was not detected before the conquest.
+As the case of Britain, the consumption of olive oil in \textit{Germania} was not detected before the conquest.
 
 The need to supply roman legions created a logistic network to supply to the army.
-In this moment, we detect the arrival of olive oil amphorae to Germania evidenced by the presence of Dressel 20 amphora in different sites close to limes.
+In this moment, we detect the arrival of olive oil amphorae to \textit{Germania} evidenced by the presence of Dressel 20 amphora in different sites close to limes.
 
-The roman studies about the presence of Dressel 20 amphorae has not been the same repercusion than the rest of the province. This can be explained due to the lack of data \citep{horacio2010llegada}.
+The roman studies about the presence of Dressel 20 amphorae has not been the same repercussion than the rest of the province. This can be explained due to the lack of data \citep{horacio2010llegada}.
 
-As a consequence, we do not the role that commerce agents in the participation of the distribution and exportation of olive oil in Germania or, on the contrary, the Roman army was the responsible to supply the product in the own province \citep[156]{remesal_germn_2010}.
+As a consequence, we do not the role that commerce agents in the participation of the distribution and exportation of olive oil in \textit{Germania} or, on the contrary, the Roman army was the responsible to supply the product in the own province \citep[156]{remesal_germn_2010}.
 
@@
 
-We analysed a dataset of 2052 stamps placed in Germania.
+We analysed a dataset of 2052 stamps placed in \textit{Germania}.
 All the data was also compiled by CEIPAC database.
@@
 	\centering
-\includegraphics[width=\linewidth]{figs/germania5.pdf}
-\caption{Distribution of the sites in Germania. Most sites were located in limes during the Roman Empire}
+\includegraphics[width=\linewidth]{figs/germania5_2.pdf}
+\caption{Distribution of the sites in \textit{Germania}. Most sites were located in limes during the Roman Empire}
 \label{germania}
@@
 \includegraphics[width=\linewidth]{figs/dendro}
-\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in Baetica area. Colours are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
+\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in \textit{Baetica} area.
+Colours are represented by areas divided into \textit{Hispalis} (red), \textit{Astigi} (green) and \textit{Corduba} (blue)}
 \label{dendro}
@@
 
-The results show a correlation similar to the obtained in Baetica province. However, the similarity is less pronounced as we observed in the dendrogram in Fig. \ref{britmap}
+The results show a correlation similar to the obtained in \textit{Baetica} province. However, the similarity is less pronounced as we observed in the dendrogram in Fig. \ref{britmap}
 
@@
 \includegraphics[width=\linewidth]{figs/dendrobrit5.pdf}
-\caption{Dendrogram obtained by Morisita-Horn algorthm of different sites in Britannia.
+\caption{Dendrogram obtained by Morisita-Horn algorithm of different sites in \textit{Britannia}.
 Colours are represented by type of sites divided into military sites (red), civil sites (green) and no specific (grey)}
@@
 
-A minor correlation can be explained by the fact that Britannia spatial area is much wide than \textit{Baetica} where the centres were more concentrated.
+A minor correlation can be explained by the fact that \textit{Britannia} spatial area is much wide than \textit{Baetica} where the centres were more concentrated.
 
@@
 In general, we did not observe any production pattern that indicates a clear organisation between production centres and reception centres.
-In other words, our results did not indicate the presence of a specific production center from Baetica in Britannia province.
+In other words, our results did not indicate the presence of a specific production center from \textit{Baetica} in \textit{Britannia} province.
 Rather it seems that olive oil production was distributed by non specific production centres.
@@
 
-The similarity of stamps shows a results less significant than Britannia province as it can be seen in the dendrogram (Fig. \ref{germap}).
+The similarity of stamps shows a results less significant than \textit{Britannia} province as it can be seen in the dendrogram (Fig. \ref{germap}).
 
@@
 \includegraphics[width=\linewidth]{figs/dendroger5.pdf}
-\caption{Dendrogram obtained by Morisita-Horn algorithm of different sites in Germania.
+\caption{Dendrogram obtained by Morisita-Horn algorithm of different sites in \textit{Germania}.
 Colours are represented by types of sites divided into military sites (red), civil sites (green) and no specific (grey)}
@@
 
-In this work, we aimed to analyse whether amphoric stamps could play an important role in the organization of the workshops along rivers.
+In this work, we aimed to analyse whether amphoric stamps could play an important role in the organisation of the workshops along rivers.
 For this reason, dissimilarity index was used to detect differences among workshops and stamps.
@@
 The hypothesis about groups of amphora workshops sharing the same stamps seems do not match with the results of the analysis even though there are similar stamps in closer workshops.
-Rather, it seems that each workshop were organized independently with different stamps.
+Rather, it seems that each workshop were organised independently with different stamps.
 Those stamps detected in closer workshops do not move from other distant workshops.
@@
 This could be defined by several factors.
-First, each workshop had a different organization involved to the use of stamps and they were not used in other workshops.
+First, each workshop had a different organisation involved to the use of stamps and they were not used in other workshops.
 Second, stamp similarity in closer workshops could be linked to a spatial pattern.

### 2020-03-20 Paper/jaccardbaetica.tex
close introduction

@@
 Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}.
-We want to detect the pattern of olive oil production that link amphora workshops and amphoric stamps used to mark them.
+We want to understand the pattern of olive oil production that link amphora workshops and amphoric stamps used to mark them.
 We focus here on exploring the economic relation between stamps and amphora production and distribution centres.
 To do this, an Ecology approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}.
+Stamps will be used to identify economic patterns by analysing their similarity.
+If workshops and provinces share stamps with similar traits then we can identify connections.
+By contrast, if we do not detect similar stamps then workshops worked independently.
 
 \maria{deberia cerrar esto}
+
+The paper is structured as follows: the next section introduces the historical context.
+Section two displays the dataset and the methods used for the analysis.
+Section three presents the results and the last section shows the discussion and the main conclusions of this work.
 
@@
 A way to analyse is by using a quantitative framework to measure the similarity.
-Here we use a ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potential spatial correlation and c)to establish a correlation between similarity of stamp and spatial distance.
+Here we use a ecological approach based on three steps: a) to detect similarities between stamp codes, b)     to explore a potential spatial correlation and c) to establish a correlation between similarity of stamp and spatial distance.
 

### 2020-03-24 Paper/jaccardbaetica.tex
discussion britannia

@@
 
-The paper is structured as follows: the next section introduces the historical context.
+The paper addresses these questions as follows: the next section introduces the historical context.
 Section two displays the dataset and the methods used for the analysis.
@@
 
-\section{Discussion and Conclusion}
+\section{Discussion}
 
@@
 
-In this work, we aimed to analyse whether amphoric stamps could play an important role in the organisation of the workshops along rivers.
-For this reason, dissimilarity index was used to detect differences among workshops and stamps.
+In this work, we aimed to analyse the distribution of the amphoric stamps in the organisation of the olive oil market both in the production and consumption areas.
+For this reason, a index of dissimilarity was used to detect differences between the distribution of the amphoric stamps and the spatial distance of producers and consumption centres.
+The general purpose was to explore if such differences found in the stamp codes could play an important role in the Roman market.
 
-In our analysis, no strong relation between stamps and spatial distance have been detected in amphora workshops.
-The analysis suggests that there is not connection between stamps and the same amphora workshops, excluding certain exceptions when nearby workshops share the same amphoric stamp.
+\subsection{Production centres}
+
+The analysis of the amphora workshops in Baetica province resulted a significance correlation between spatial distance and the similarity of the amphoric stamps.
+
+The analysis suggests that there is  a connection between the same stamp code and the nearby amphora workshops, excluding certain exceptions.
 Consequently, the majority of stamps are located in different amphora workshops and only similar stamps between closer amphora workshops were found.
-In any case, our results show that most similar stamps were detected in the same \textit{conventus} area.
-These stamps tend to share the same area of production but there is not a general relation between groups of amphora workshops and area.
+Indeed, our results show that most similar stamps were detected in the same \textit{conventus} area.
+This results can indicated a grouping of stamps with similar codes in the same \textit{conventus}.
+However,these stamps tend to share the same area of production but there is not a general relation between groups of amphora workshops and connection areas.
+In other words, we do not identify groups of more than three workshops or less sharing the same amphoric stamps as the dendrogram showed.
+In general, the majority of stamps were located in different amphora workshops.
 
-The hypothesis about groups of amphora workshops sharing the same stamps seems do not match with the results of the analysis even though there are similar stamps in closer workshops.
+Our results indicate that the hypothesis about groups of amphora workshops sharing the same stamps seems do not match with the results of the analysis even though there are similar stamps in closer workshops.
 Rather, it seems that each workshop were organised independently with different stamps.
 Those stamps detected in closer workshops do not move from other distant workshops.
-In other words, the stamps tend to keep in the same area and different stamps were located in a same amphora workshops.
+In other words, the stamps tend to remain in the same area and different stamps were located in a same amphora workshops.
 
@@
 Additionally, most of the workshops were not widely excavated.
+
+\subsection{Consumption centres}
+
+Both consumption centres showed a correlation between spatial distance and similarity in the amphoric stamps. In the case of\textit{Britannia}, the correlation was higher than \textit{Germania}.
+
+In \textit{Britannia}, the majority of similarity stamps were mainly found in military centres.
+This could be also interpreted by a intense bias where military centres have been mostly excavated than civil centres.
+
+It is worth to mention that centres with a similarity correlation between  spatial distance and stamps were located close to the Atlantic coast around the North Sea and the Celtic Sea.
+This could indicate that the Atlantic route could have played an essential role in transporting olive oil to the area of \textit{Britannia}, since in the places where there is greater similarity they are found in different strategic points near the sea.
+
+This fact seems to concur with the latest research that reinforce the idea of a more important Atlantic route that would connect the provinces militarized.
+
+(Remesal Rodrı́guez, 1986; 2008; Carreras Monfort y Morais, 2012; Morillo
+et al., 2016; Rubio-Campillo et al., 2018b).
+
+\maria{falta germania}
+
+\section{Conclusion remarks}
 

### 2020-03-25 Paper/jaccardbaetica.tex
add germania part

@@
 
-This fact seems to concur with the latest research that reinforce the idea of a more important Atlantic route that would connect the provinces militarized.
+This fact seems to concur with the latest research that suggest the idea of a more important Atlantic route that would connect the provinces militarised\citep{remesal_annona_1986,
+remesal_provincial_2008,
+carreras_atlantic_2012,
+morillo_hispania_2016,rubio-campillo_provincias_2018}.
 
-(Remesal Rodrı́guez, 1986; 2008; Carreras Monfort y Morais, 2012; Morillo
-et al., 2016; Rubio-Campillo et al., 2018b).
+Therefore, most part of the centres where similar stamps are shared can correspond to eminently military areas, which may mean that the transport could reach the military areas from the beginning and then be distributed by land routes along other civil areas \citep{carreras_britannia_1998,
+ayllon_olive_2018}.
+
+In \textit{Germania}, the results followed the same pattern than Britannia but with a minor correlation as we can observed in the dendrogram.
+The areas mostly militarised share similar stamps than civil areas.
+However, we do not detect a concrete pattern regarding the distribution of the stamps in the german limes.
+\maria{citar remesal xanten}.
+
+It is worth to know that it was not detected a robust model of organisation in both cases with regard to the distribution of stamps in consumption centres. This means
+that it is unknown whether certain production centres went to one province or another or, at the
+less, that it can be clearly reflected in the data with a greater similarity in the amphoric stamps.
+Thus, production centres could have distributed randomly olive oil both \textit{Britannia} and \textit{Germania}.
+We do not detect production centres dedicated to the distribution of olive oil in a certain province.
+
+In general, we observed a distribution of the stamps to military stamps.
+By contrast, judging by the results obtained, there does not seem to be a specific pattern in terms of geographical distribution.
+Nor it was detected consumption areas where stamps are specified from a amphora workshop.
 

### 2020-03-30 Paper/jaccardbaetica.tex
add correction english

@@
 \textit{Baetica} became important production and distribution centre during the Roman Empire.
-However, it remains under debate how this province was organised and whether it could be possible to identify patterns in the olive oil market.
+However, it remains under debate about how this province was organised and whether it could be possible to identify patterns in the olive oil market.
 Amphoric stamps are used to identify the presence of different groups that might share similar stamps.
 To achieve this goal, we analyse a set of stamps from two centres: 1) producer centres by analysing different workshops in \textit{Baetica} province and 2) two Roman provinces such as \textit{Germania} and \textit{Britannia} as a receptor centres.
-They will be used to detect a relation between the distribution of amphoric stamps and the economic structure in both centres.
+They will be used to detect a connection between the distribution of amphoric stamps and the economic structure in both centres.
 Here we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance.
 
-The analysis explores how quantitative approach provides a useful tool for the interpretation of the economic processes.
+The analysis explores how the quantitative approach provides a useful tool for the interpretation of the economic processes.
 Finally, results pretend to highlight the organisation of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.
@@
 %primer párrafo más general---- especifico (¿dónde se ha aplicado?)
-Material culture is one of the most frequent indicator of trade in the archaeological record.
-In archaeology, they allow us to highlight a part of mechanism of production and distribution of goods along the Mediterranean \citep{bevan_mediterranean_2014}.
+Material culture is one of the most frequent indicators of trade in the archaeological record.
+In archaeology, they allow us to highlight a part of the mechanism of production and distribution of goods along the Mediterranean \citep{bevan_mediterranean_2014}.
 Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}.
-Given this fact, an important mechanism of production control under Roman government was created spreading by different areas attending to the richness of different places.
+Given this fact, an important mechanism of production control under the Roman government was created spreading by different areas attending to the richness of different places.
 Roman provinces such as \textit{Baetica} (currently Andalusia, south Spain) became important production centres of olive oil during the Roman Empire.
-Olive oil was considered as the liquid gold since it was used in different aspect of the daily life as cooking, hygiene or lighting  \citep{mattingly_d.j._oil_1988}.
+Olive oil was considered the liquid gold since it was used in different aspect of daily life for cooking, hygiene or lighting  \citep{mattingly_d.j._oil_1988}.
 As a consequence, this product was distributed in a large number of amphorae along the province, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.
 
-Despite this issue has been widely discussed over many decades, the understanding about the production processes is still under debate, mostly due to the lack of written records.
-The application of different quantitative approaches combined with the archaeological evidence has allowed us partly to deal with the complexity of understanding the Roman production \citep{orengo_seeds_2016,
+Despite this issue has been widely discussed over many decades, the understanding of the olive oil production processes is still under debate, mostly due to the lack of written records.
+The application of different quantitative approaches combined with the archaeological evidence has allowed us partly to deal with the complexity of the Roman production \citep{orengo_seeds_2016,
 brughmans_roman_2016,coto-sarmiento_identifying_2018,bayesian_2018}.
@@
 %This paper explore--- resumir en dos líneas el objeto de estudio y definir que ha sido usado para ello dispersion of culture
-This paper aims to understand the production dynamics in connection with a different areas within Roman Empire.
+This paper aims to understand the production dynamics in connection with different areas within the Roman Empire.
 Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}.
-We want to understand the pattern of olive oil production that link amphora workshops and amphoric stamps used to mark them.
+We want to understand the pattern of olive oil production linked to amphora workshops and amphoric stamps used to mark them.
 We focus here on exploring the economic relation between stamps and amphora production and distribution centres.
@@
 Stamps will be used to identify economic patterns by analysing their similarity.
-If workshops and provinces share stamps with similar traits then we can identify connections.
-By contrast, if we do not detect similar stamps then workshops worked independently.
+If workshops and provinces share stamps with similar traits, then we can identify connections.
+By contrast, if we do not detect similar stamps between workshops, then workshops worked independently.
 
@@
 The high demand of Roman provinces stimulated by the good condition of the \textit{Baetica} lands allowed to develop a massive infrastructure of olive oil production.
-The production and distribution of olive oil were growing exponentially during almost three centuries \citep{remesal_concierto}.
-As a result, hundreds of amphora workshops were made to support this high demand of Roman Empire.
+The production and distribution of olive oil were growing exponentially for almost three centuries \citep{remesal_concierto}.
+As a result, hundreds of amphora workshops were made to support this high demand of the Roman Empire.
 These workshops were located along the rivers Guadalquivir and Genil, supplying from the riverine connectivity to the Mediterranean and Atlantic routes \citep{garcia_vargas_enrique_formal_2010}.
@@
 The fascination aroused by stamps has caused it to be one of the inscriptions with the most publications about its origin and meaning.
-Its endure in time become it an essential element that were used to mark amphorae during almost three centuries \citep{remesal_sellar_2016} (Fig.\ref{amphora}).
+Its endure in time become an essential element that were used to mark amphorae during almost three centuries \citep{remesal_sellar_2016} (Fig.\ref{amphora}).
 
@@
 
-Frequently, they were marked mainly in handles but seldom in rims and body \citep{millet_anforas_1998}. 
+Frequently, they were marked mainly in handles, but seldom in rims and body \citep{millet_anforas_1998}. 
 The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion.
-Stamps are mostly formed by a code of three letters and they can appear in a abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
-Despite researchers identify stamps as a identity marks, there is not a general consensus about the meaning of the stamps \citep{rodriguez_baetican_1998}.
+Stamps are mostly formed by a code of three letters and they can appear in abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
+Despite researchers identify stamps as identity marks, there is not a general consensus about the meaning of the stamps \citep{rodriguez_baetican_1998}.
 Alike most Dressel 20 containers were not stamped being difficult to determinate.
@@
 Neither written records have been found that it could explain the economic role of \textit{Baetica} province in the Roman organisation.
-Indeed, archaeological evidence shows a highly specialised production with a long activity in a this specific area with apparently few changes \citep{remesal_anforas_2004}.
+Indeed, archaeological evidence shows a highly specialised production with a long activity in this specific area with apparently few changes \citep{remesal_anforas_2004}.
 
@@
 
-Three case study have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}).
-In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop.
-In Roman provinces, our aim is detecting groups of stamps concentrated in a area or if some groups have an important role for the exportation of olive oil in those provinces.
-This economic connection could affect due to different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamp in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province.
+Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}).
+In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces.
+This economic connection could be affected by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamp in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province.
 
-Therefore, this study proposes an evolutionary baseline to explore the distribution of Baetican olive oil production by computing spatial correlation between stamps.
-A way to analyse is by using a quantitative framework to measure the similarity.
-Here we use a ecological approach based on three steps: a) to detect similarities between stamp codes, b)     to explore a potential spatial correlation and c) to establish a correlation between similarity of stamp and spatial distance.
+Therefore, this study proposes an evolutionary baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps.
+A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps.
+Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
 
@@
 
-Our case study examines the relation between the distribution of amphoric stamps and the workshops.
-In particular, we study the distribution of amphoric stamp with the aim of identifying a correlation between geographical distance and amphoric stamps.
-Based on this assumption, we proposed three hypotheses: a) it exists a correlation between spatial distance and the distribution of stamps, b) stamps located in close workshops share similar traits and c) Low mobility of amphoric stamps to other region: stamps always stay in the same region.
+Our case study examines the relationship between the distribution of amphoric stamps and the workshops.
+In particular, we study the distribution of amphoric stamp in order to identify a correlation between geographical distance and amphoric stamps.
+Based on this assumption, we proposed three hypotheses: a) it exists a correlation between spatial distance and the distribution of stamps, b) stamps located in close workshops share similar traits and c) Low mobility of amphoric stamps to other regions: stamps always stay in the same region.
 
@@
 \includegraphics[width=\linewidth]{figs/workshop1}
-\caption{Distribution of the workshops}
+\caption{Distribution of the workshops in Baetica province}
 \label{workshop}
@@
 Consequently, we discard integrate the fragmented stamps in our dataset.
-We finally filter a total sample of 987 stamps comprised of 130 different stamps from 81 workshops.
+We finally filter a total sample of 987 stamps composed of 130 different stamps from 81 workshops.
 
@@
 
-We calculated the frequency of distribution in amphora stamps with the aims to carry out a exploratory analysis (EDA) to know the distribution and the number of stamps per each centre.
+We calculated the frequency of distribution in amphora stamps with the aim to carry out an exploratory analysis (EDA) to know the distribution and the number of stamps per each centre.
 
@@
 Most stamps are distributed in only one workshop whereas few workshops concentrate a high frequency of different amphoric stamps.
-The type of distribution is also frequent in Roman economy where we observed a self-organized complex system patterns typical of a free market \citep{bayesian_2018}.
+The type of distribution is also frequent in the Roman economy where we observed a self-organised complex system pattern typical of a free market \citep{bayesian_2018}.
 
@@
 The majority of stamps found are concentrated in \textit{Hispalis} with 574 stamps while \textit{Corduba} and \textit{Astigi} with 267 and 146 stamps, respectively.
-Mostly, workshops show a homogeneity on the frequency of stamps except both La Catria and Arva that show a big amount of stamps with 29 different stamps.
-According to previous studies, those workshops became in most important centres of amphora production although they could have been more intensely prospected then others.
+Mostly, workshops show a homogeneity on the frequency of stamps except for both La Catria and Arva that exhibit a big amount of stamps with 29 different stamps.
+According to previous studies, those workshops became in most important centres of amphora production although they could have more intensely prospected than others.
 \citep{arva_1997}.
@@
 
-It is well known that this product was not frequently consumed by autonomous population.
-The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}.
+It is well known that this product was not frequently consumed by indigenous population.
+The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}.
 Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions.
 This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman legions.
-In this moment, we detect an increase of the olive oil exportation concurring with the displacement of Roman legions during the military campaigns \citep[161]{monfort_britanniaen_1998}.
+
+At this moment, we detect an increase of the olive oil exportation concurring with the displacement of Roman legions during the military campaigns \citep[161]{monfort_britanniaen_1998}.
 This fact will be spatial incidence in sites close at Hadrian Wall.
@@
 The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges.
-Thus, the network was mainly focused on the support of soldiers during military campaigns.
+Thus, the network was mainly focused on the support of soldiers during the military campaigns.
 %debate intercambios puntuales o redes comerciales de abastecimiento
@@
 The presence of Dressel 20 stamps in military camps in \textit{Britannia} have been widely studied in Roman archaeology \citep{carreras_britannia_1998}.
-This fact also indicates a possible governmental organisation allocate to reforce and supply of olive oil in military camps, such as \textit{Germania} (Remesal Rodrıguez, 1986).
+This fact also indicates a possible governmental organisation allocate to reinforce and supply of olive oil in military camps, such as \textit{Germania} (Remesal Rodrıguez, 1986).
 However, it is unknown the type of economical system when managing or control the redistribution of olive oil in different sites in \textit{Britannia} but it seems that it could have been in the main cities \citep[45]{funari_economic_2005}.
@@
 from the third century A.D., coinciding with the change in market strategy in the
-Empire. In that moment, a progressive decrease of Dressel 20 is documented being gradually replace for Dressel 23 amphorae.
+Empire. As a result, a progressive decrease of Dressel 20 is documented being gradually replaced for Dressel 23 amphorae. \maria {citation needed}
 
@@
 The database contained anomalies that were eliminated using the same criteria cited above.
-We finally selected the centres with more or equal than 5 stamps.
-As a result, we studied a total of 1765 stamps comprised of 968 different stamps from 46 centres.
+We finally selected the centres with more or equal than five stamps.
+As a result, we studied a total of 1765 stamps composed of 968 different stamps from 46 centres.
 The centres located in \textit{Britannia} can be seen in Fig.\ref{britannia}.
@@
 The Roman conquest in \textit{Germania} dates from the end of the 1st century BC, during Caesar campaign and subsequently continued by Augusto.
-The advance of the Roman legions toward \textit{Gallia} opened a way though the Atlantic Ocean to \textit{Germania} \citep{remesal_annona_1986,
+The advance of the Roman legions toward \textit{Gallia} opened a way through the Atlantic Ocean to \textit{Germania} \citep{remesal_annona_1986,
 remesal_baetica_2002}.
@@
 
-As the case of Britain, the consumption of olive oil in \textit{Germania} was not detected before the conquest.
+As the case of \textit{Britannia}, the consumption of olive oil in \textit{Germania} was not detected before the conquest.
+This is evidenced by the presence of Dressel 20 amphora in different sites close to limes.
 
-The need to supply roman legions created a logistic network to supply to the army.
-In this moment, we detect the arrival of olive oil amphorae to \textit{Germania} evidenced by the presence of Dressel 20 amphora in different sites close to limes.
-
-The roman studies about the presence of Dressel 20 amphorae has not been the same repercussion than the rest of the province. This can be explained due to the lack of data \citep{horacio2010llegada}.
+However, the Roman studies about the presence of Dressel 20 amphorae in textit{Germania} has not had the same repercussion as the rest of the province.
+This could be explained due to the lack of archaeological sources citep{horacio2010llegada}.
 
@@
 
-The presence of Roman army encouraged the exchange in the province showed by the arrival of this product both civil settlement and military sites with a mayor concentration at german limes.
-It seems that some baetican centres would be assigned to the support of olive oil.
+The presence of the Roman army encouraged the exchange in the province showed by the arrival of this product both civil settlement and military sites with a mayor concentration at German limes.
+It seems that some Baetican centres would be assigned to the support of olive oil.
 However, this hypothesis can not be proven by the lack of archaeological sources \citep[125]{remesal_concierto}.
@@
 All the data was also compiled by CEIPAC database.
-We finally selected a total of 1621 stamps consist of 850 different stamps from 46 sites (see Fig.
+We finally selected a total of 1621 stamps consisted of 850 different stamps from 46 sites (see Fig.
 \ref{germania}).
-As previously before, the centres were selected with more or equal than 5 stamps.
+As previously before, the centres were selected with more or equal than five stamps.
 
@@
 Considering our dataset as a non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity \citep{wolda_similarity_1981}.
-Morisita-Horn index can be expressed considering 0 as total presence of similarity of stamps and 1 a totally dissimilarity between stamps.
+Morisita-Horn index can be expressed considering 0 as the total presence of similarity of stamps and 1 a total dissimilarity between stamps.
 In our case, it will be calculated the number of times that one stamp appears in an amphora workshop.
-This method allows to bear in mind the similar number of times for each repeated stamp per workshop.
-If two workshops have similar stamp codes then the probability would be 0 whereas stamps codes are totally different when the results would be 1.
+This method allows to bear in mind a similar number of times for each repeated stamp per workshop.
+If two workshops have similar stamp codes, then the probability would be 0 whereas stamps codes are totally different when the results would be 1.
 
@@
 
-The analysis shows that amphoric stamps could be correlated with the spatial distance.
+The analysis shows that amphoric stamps could be correlated with spatial distance.
 The correlation coefficients range from a minimum to a maximum.
-The dendrogram shown in Fig.
+The dendrogram in Fig.
 \ref{dendro} was obtained with Morisita-Horn index.
 The dendrogram suggests that amphora workshops used different stamps for their production system.
-Nearby workshops show a similarity on the stamps while most of them seem to display different stamps roles.
-Additionally, groups of workshops sharing similar stamps were not found in the cluster: the majority of stamp grouping was composed by no more than three workshops.
-Indeed workshops that shared more similar amphoric stamps belonged to the same \textit{conventus} area, such as Picachos, Cerro de los Pesebres and El Castillejo.
+Nearby workshops show similarity on the stamps while most of them seem to display different stamps roles.
+Additionally, groups of workshops sharing similar stamps were not found in the cluster: the majority of stamp grouping was composed of no more than three workshops.
+Indeed, workshops that shared more similar amphoric stamps belonged to the same \textit{conventus} area, such as Picachos, Cerro de los Pesebres and El Castillejo.
 
@@
 In other words, our results did not indicate the presence of a specific production center from \textit{Baetica} in \textit{Britannia} province.
-Rather it seems that olive oil production was distributed by non specific production centres.
+Rather, it seems that olive oil production was distributed by non-specific production centres.
 
-\subsection{Germania}
+\subsubsection{Germania}
 
-The similarity of stamps shows a results less significant than \textit{Britannia} province as it can be seen in the dendrogram (Fig. \ref{germap}).
+The similarity of stamps shows results less significant than \textit{Britannia} province as it can be seen in the dendrogram (Fig. \ref{germap}).
 
@@
 On the other hand, there seems to be a greater concentration of the seals in areas eminently militarised and close to the German limes.
-However, It is still no possible to determine with the data that the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.
+However, It is still no possibility to determine the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.
 
@@
 In this work, we aimed to analyse the distribution of the amphoric stamps in the organisation of the olive oil market both in the production and consumption areas.
-For this reason, a index of dissimilarity was used to detect differences between the distribution of the amphoric stamps and the spatial distance of producers and consumption centres.
+For this reason, an index of dissimilarity was used to detect differences between the distribution of the amphoric stamps and the spatial distance of producers and consumption centres.
 The general purpose was to explore if such differences found in the stamp codes could play an important role in the Roman market.
@@
 
-The analysis of the amphora workshops in Baetica province resulted a significance correlation between spatial distance and the similarity of the amphoric stamps.
+The analysis of the amphora workshops in Baetica province resulted in a significant correlation between spatial distance and the similarity of the amphoric stamps.
 
-The analysis suggests that there is  a connection between the same stamp code and the nearby amphora workshops, excluding certain exceptions.
+The analysis suggests that there is  a connection between the same stamp code and nearby amphora workshops, excluding certain exceptions.
 Consequently, the majority of stamps are located in different amphora workshops and only similar stamps between closer amphora workshops were found.
 Indeed, our results show that most similar stamps were detected in the same \textit{conventus} area.
-This results can indicated a grouping of stamps with similar codes in the same \textit{conventus}.
-However,these stamps tend to share the same area of production but there is not a general relation between groups of amphora workshops and connection areas.
+This results can indicate a grouping of stamps with similar codes in the same \textit{conventus}.
+However, these stamps tend to share the same area of production but there is not a general relation between groups of amphora workshops and connection areas.
 In other words, we do not identify groups of more than three workshops or less sharing the same amphoric stamps as the dendrogram showed.
@@
 
-Our results indicate that the hypothesis about groups of amphora workshops sharing the same stamps seems do not match with the results of the analysis even though there are similar stamps in closer workshops.
-Rather, it seems that each workshop were organised independently with different stamps.
+Our results indicate that the hypothesis about groups of amphora workshops sharing the same stamps seems do not match with the results of the analysis, even though there are similar stamps in closer workshops.
+Rather, it seems that each workshop was organised independently with different stamps.
 Those stamps detected in closer workshops do not move from other distant workshops.
-In other words, the stamps tend to remain in the same area and different stamps were located in a same amphora workshops.
+In other words, the stamps tend to remain in the same area and different stamps were located in a same amphora workshop.
 
 This could be defined by several factors.
-First, each workshop had a different organisation involved to the use of stamps and they were not used in other workshops.
+First, each workshop had a different organisation involved in the use of stamps and they were not used in other workshops.
 Second, stamp similarity in closer workshops could be linked to a spatial pattern.
 It is more probably than closer workshops tend to share more traits than distant workshops.
-While the role of river was significant for the distribution of amphorae in consumption places, river connection amongst workshops does not seem show a relevancy for the distribution of stamps.
+While the role of the river was significant for the distribution of amphorae in consumption places, river connection amongst workshops does not seem to show relevancy for the distribution of stamps.
 Finally, the distribution of stamps could have shown some research bias.
@@
 
-It is worth to mention that centres with a similarity correlation between  spatial distance and stamps were located close to the Atlantic coast around the North Sea and the Celtic Sea.
+It is worth to mention that centres with a similarity correlation were located close to the Atlantic coast around the North Sea and the Celtic Sea.
 This could indicate that the Atlantic route could have played an essential role in transporting olive oil to the area of \textit{Britannia}, since in the places where there is greater similarity they are found in different strategic points near the sea.
@@
 
-Therefore, most part of the centres where similar stamps are shared can correspond to eminently military areas, which may mean that the transport could reach the military areas from the beginning and then be distributed by land routes along other civil areas \citep{carreras_britannia_1998,
+Therefore, most part of the centres where similar stamps are shared can correspond to eminently military areas, which may mean that the transport could reach the military areas from the beginning and then be distributed by land routes along with other civil areas \citep{carreras_britannia_1998,
 ayllon_olive_2018}.
 
-In \textit{Germania}, the results followed the same pattern than Britannia but with a minor correlation as we can observed in the dendrogram.
+In \textit{Germania}, the results followed the same pattern as \textit{Britannia} but with a minor correlation as we can observe in the dendrogram.
 The areas mostly militarised share similar stamps than civil areas.
-However, we do not detect a concrete pattern regarding the distribution of the stamps in the german limes.
+However, we do not detect a concrete pattern regarding the distribution of the stamps in the German limes.
 \maria{citar remesal xanten}.
 
-It is worth to know that it was not detected a robust model of organisation in both cases with regard to the distribution of stamps in consumption centres. This means
+It is worth to mention that it was not detected a robust model of organisation in both cases with regard to the distribution of stamps in consumption centres. This means
 that it is unknown whether certain production centres went to one province or another or, at the
@@
 By contrast, judging by the results obtained, there does not seem to be a specific pattern in terms of geographical distribution.
-Nor it was detected consumption areas where stamps are specified from a amphora workshop.
-
-\maria{falta germania}
+Nor it was detected consumption areas where stamps are specified from an amphora workshop.
 
@@
 This hypothesis would explain the fact that we do not find similar stamps in different workshops; however, we found different amphora stamps in the same workshops that they would be barely difficult to assign different owners.
-On the other hand, it could be interpreted somehow a batch systematic organisation.
-Considering that Dressel 20 was not marked in most cases, some authors point out that potters marked amphorae to prepare and distribute the commodity to be shipped \citep{berni_millet_epigrafianforica_2008}.
+
+On the other hand,the existence of different stamps in the same workshop but which are only repeated in very close places it would imply some kind of organisation from within workshop affecting the closest centres.
+This could be interpreted such as batch systematic organisation between potters \maria{citar tesis de moros}.
+This method allowed to potters organise the production with the batch stamping for the posterior commercialisation.
+
+Considering that Dressel 20 was not marked in most cases, some researchers point out that potters marked amphorae to prepare and distribute the commodity to be shipped \citep{berni_millet_epigrafianforica_2008}.
 This method would be used as an identifier to count the number of amphorae of a branch \citep{juanmorostesis}.
-This method could also have served to identify different groups of potter workers working in the same amphora workshop.
+This organisation could also have served to identify different groups of potters working in the same amphora workshop.
 Potters could have marked the amphorae to distinguish different groups working in parallel \citep{li_crossbows_2014}.
-This it would explain wherefore why we detect different stamps in a same workshop.
+This fact would explain wherefore why we detect different stamps in the same workshop.
 In any case, we do not have enough archaeological evidence that can validate the interpretations presented here and our results are certainly valid only with the context of our case study.
 
-As summary, this method presented here provides a potential tool to understand mechanisms of production based on the similarity of artefacts.
-This method have identified differences in the case of the amphoric production within Roman Empire.
+As a summary, this method presented here provides a potential tool to understand mechanisms of production based on the similarity of artefacts.
+This work has identified differences in the case of the amphoric production within the Roman Empire.
 Accordingly, the results have highlighted for the interpretation of the complex economic processes connected with the archaeological evidence.
 
-%Dressel 20 marked in the 30-40\%
+\section{Acknowledgements}
 
-\section{Acknowledgements}
-%completar 
 The research was funded by European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros and Ignacio Morer for their useful suggestions.  

### 2020-03-31 Paper/jaccardbaetica.tex
add bib

@@
 from the third century A.D., coinciding with the change in market strategy in the
-Empire. As a result, a progressive decrease of Dressel 20 is documented being gradually replaced for Dressel 23 amphorae. \maria {citation needed}
+Empire. As a result, a progressive decrease of Dressel 20 is documented being gradually replaced for Dressel 23 amphorae \citep{rodriguez1991aceite,
+millet_anforas_1998}.
 
@@
 However, the Roman studies about the presence of Dressel 20 amphorae in textit{Germania} has not had the same repercussion as the rest of the province.
-This could be explained due to the lack of archaeological sources citep{horacio2010llegada}.
+This could be explained due to the lack of archaeological sources \citep{horacio2010llegada}.
 
@@
 The areas mostly militarised share similar stamps than civil areas.
-However, we do not detect a concrete pattern regarding the distribution of the stamps in the German limes.
-\maria{citar remesal xanten}.
+However, we do not detect a concrete pattern regarding the distribution of the stamps in the German limes \citep{xanten2018}.
 
@@
 On the other hand,the existence of different stamps in the same workshop but which are only repeated in very close places it would imply some kind of organisation from within workshop affecting the closest centres.
-This could be interpreted such as batch systematic organisation between potters \maria{citar tesis de moros}.
+This could be interpreted such as batch systematic organisation between potters \citep{juanmorostesis}.
 This method allowed to potters organise the production with the batch stamping for the posterior commercialisation.

### 2020-04-01 Paper/jaccardbaetica.tex
1st draft

@@
 
-\maria{deberia cerrar esto}
-
 The paper addresses these questions as follows: the next section introduces the historical context.
@@
 Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti} with different information still unknown.
-Indeed most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
+Indeed, most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
 
@@
 
-Frequently, they were marked mainly in handles, but seldom in rims and body \citep{millet_anforas_1998}. 
+Frequently, they were mainly marked  in handles, but seldom in rims and body \citep{millet_anforas_1998}. 
 The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion.
@@
 In particular, we study the distribution of amphoric stamp in order to identify a correlation between geographical distance and amphoric stamps.
-Based on this assumption, we proposed three hypotheses: a) it exists a correlation between spatial distance and the distribution of stamps, b) stamps located in close workshops share similar traits and c) Low mobility of amphoric stamps to other regions: stamps always stay in the same region.
+Based on this assumption, we proposed three hypotheses: a) we can identify a correlation between spatial distance and the distribution of stamps, b) stamps located in close workshops share similar traits and c) Low mobility of amphoric stamps to other regions: stamps always stay in the same region.
 
@@
 Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions.
-This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman legions.
+This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman army.
 
-At this moment, we detect an increase of the olive oil exportation concurring with the displacement of Roman legions during the military campaigns \citep[161]{monfort_britanniaen_1998}.
+At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}.
 This fact will be spatial incidence in sites close at Hadrian Wall.
@@
 The Roman conquest in \textit{Germania} dates from the end of the 1st century BC, during Caesar campaign and subsequently continued by Augusto.
-The advance of the Roman legions toward \textit{Gallia} opened a way through the Atlantic Ocean to \textit{Germania} \citep{remesal_annona_1986,
+The advance of the Roman legions toward \textit{Gallia} opened a path through the Atlantic Ocean to \textit{Germania} \citep{remesal_annona_1986,
 remesal_baetica_2002}.
@@
 Nearby workshops show similarity on the stamps while most of them seem to display different stamps roles.
-Additionally, groups of workshops sharing similar stamps were not found in the cluster: the majority of stamp grouping was composed of no more than three workshops.
-Indeed, workshops that shared more similar amphoric stamps belonged to the same \textit{conventus} area, such as Picachos, Cerro de los Pesebres and El Castillejo.
+Groups of workshops sharing similar stamps were not found in the cluster: the majority of stamp grouping was composed of no more than three workshops.
+Additionally, workshops that shared more similar amphoric stamps belonged to the same \textit{conventus} area, such as Picachos, Cerro de los Pesebres and El Castillejo.
 
@@
 
-\section{Conclusion remarks}
+\section{Conclusion}
 
@@
 
-On the other hand,the existence of different stamps in the same workshop but which are only repeated in very close places it would imply some kind of organisation from within workshop affecting the closest centres.
-This could be interpreted such as batch systematic organisation between potters \citep{juanmorostesis}.
+On the other hand, the existence of different stamps in the same workshop but which are only repeated in very close places it would imply some kind of organisation from within workshop affecting the closest centres.
+This could be explained such as batch systematic organisation between potters \citep{juanmorostesis}.
 This method allowed to potters organise the production with the batch stamping for the posterior commercialisation.

### 2020-04-21 Paper/jaccardbaetica.tex
comentarios

@@
 %primer párrafo más general---- especifico (¿dónde se ha aplicado?)
+\xavi{queda raro esto de "en arqueología la cultura material nos permite ver la distribución en el Mediterráneo". Yo iría directa a hablar del Mediterráneo sin la primera frase.}
+
 Material culture is one of the most frequent indicators of trade in the archaeological record.
 In archaeology, they allow us to highlight a part of the mechanism of production and distribution of goods along the Mediterranean \citep{bevan_mediterranean_2014}.
-Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}.
+Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}\xavi{aquí faltan citas de obras más generales; esto es un ejemplo pero muy específico noÇ?}.
 Given this fact, an important mechanism of production control under the Roman government was created spreading by different areas attending to the richness of different places.
@@
 %This paper explore--- resumir en dos líneas el objeto de estudio y definir que ha sido usado para ello dispersion of culture
+\xavi{este aim es demasiado general}
 This paper aims to understand the production dynamics in connection with different areas within the Roman Empire.
@@
 We focus here on exploring the economic relation between stamps and amphora production and distribution centres.
-To do this, an Ecology approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}.
+To do this, an Ecology approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}\xavi{lo de Ecology approach es raro no?
+Yo lo usé pero en el título...quizás population approach?}.
 Stamps will be used to identify economic patterns by analysing their similarity.
@@
 
+\xavi{esta frase tendría que ir al final de la Intro; si pones subsecciones entonces queda muy raro creo}
+
 The paper addresses these questions as follows: the next section introduces the historical context.
@@
 Section three presents the results and the last section shows the discussion and the main conclusions of this work.
+
+\xavi{creo que le falta chicha aquí, sobre qué se ha hecho hasta ahora, qué quieres saber específicamente, el caso de estudio...}
+
+\xavi{seguro que tiene subsecciones este journal?}
 
@@
 This economic connection could be affected by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamp in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province.
+\xavi{affected?
+No sería "identified"?
+Es decir, la correlación es la evidencia, pero la causa sería la interacción económica...no?}
 
@@
 Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
+\xavi{no veo la parte evolutiva en el paper; por otra parte, si es relevante no la lanzaría aquí al vuelo; hace falta explicar por qué es necesaria, qué implica, etc.}
 
@@
 
+\xavi{Yo aplicaría aquí el feedback de los mapas para lo del libro...no creo que colorearlo basándose en el DEM aporte demasiado no?}
+
 The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
@@
 We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in \textit{Baetica} province.
-The stamp database was compiled by CEIPAC database \citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}).
+The stamp database was compiled by CEIPAC database \citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}) \xavi{no fue la database la que compiló los sellos, no?
+:-D}}.
 However, approximately the 70 \% of stamps cannot be tested due to fragmentation or incomplete information.
@@
 
+\xavi{a ver, esto queda raro...si esto es una exploración de la base de datos entonces tiene que ir dentro de "Materials", así mientras vas comentando los sesgos y cosas generales de los datos (como soporte, pero no hace falta subsubsección "EDA")}
+
 We calculated the frequency of distribution in amphora stamps with the aim to carry out an exploratory analysis (EDA) to know the distribution and the number of stamps per each centre.
@@
 The type of distribution is also frequent in the Roman economy where we observed a self-organised complex system pattern typical of a free market \citep{bayesian_2018}.
+
+\xavi{quieres decir que esto es una power base? yo creo que no, no es tan bestia no? Yo lo matizaría}
 
@@
 \subsection{Consumption centres: Britannia and Germania}
+
+\xavi{Esta parte más histórica está chachi pero creo que debería estar en la Intro.
+Ahí podrías explicar la organización, y a partir de estas fuentes escritas plantear preguntas e hipótesis que justifiquen el paper y tu análisis usando arqueología.}
 
@@
 
+\xavi{pos eso, los mapas...te parece si los hago yo...? andaaa vaaaa déjame hacer algo}
+
 \subsubsection{Germania}
@@
 However, this hypothesis can not be proven by the lack of archaeological sources \citep[125]{remesal_concierto}.
+
+\xavi{yo creo que solo hay 1 nivel en el JAMT, ni sub ni subsub}
 
@@
 
+\xavi{pero este es el único método que usas?}
+
 %hablar sobre Horn (Morisita)hablar sobre que el sample no estaba uniforme por eso usamos el morisita horn
@@
 
+\xavi{no puede ser vertical la figura? se leería mejor}
+
 \subsection{Reception centres: Britannia and Germania}
@@
 The results show a correlation similar to the obtained in \textit{Baetica} province. However, the similarity is less pronounced as we observed in the dendrogram in Fig. \ref{britmap}
+
+\xavi{no tengo claro que le puedas decir correlación, porque eso es un valor específico entre 2 distribuciones numéricas. Te refieres a similaridad?}
 

### 2020-04-23 Paper/jaccardbaetica.tex
first correction xavi

--- 2020-04-01 Paper/jaccardbaetica.tex
@@
 \usepackage{hyperref,lineno}
+\usepackage{wrapfig}
+\usepackage{lscape}
+\usepackage{rotating}
 \usepackage{xcolor}
@@
 We focus here on exploring the economic relation between stamps and amphora production and distribution centres.
-To do this, an Ecology approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}.
+To do this, an population approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}.
 Stamps will be used to identify economic patterns by analysing their similarity.
@@
 By contrast, if we do not detect similar stamps between workshops, then workshops worked independently.
+
+%\xavi{esta frase tendría que ir al final de la Intro; si pones subsecciones entonces queda muy raro creo}
 
@@
 
-\subsection{The Amphoric production in Baetica province}
+\xavi{creo que le falta chicha aquí, sobre qué se ha hecho hasta ahora, qué quieres saber específicamente, el caso de estudio...}
+
+\section{The Amphoric production in Baetica province}
 
@@
 
-\subsection{A potential indicator of Roman economy: Dressel 20 stamps}
+\section{Britannia and Germania}
+
+%\xavi{Esta parte más histórica está chachi pero creo que debería estar en la Intro.
+Ahí podrías explicar la organización, y a partir de estas fuentes escritas plantear preguntas e hipótesis que justifiquen el paper y tu análisis usando arqueología.}
+
+The conquest of new provinces allowed to the Roman Empire the arrival of new resources.
+This lead to a progressive change in the economic and social structure.
+As a result, Augusto's administration created the figure of the \textit{praefectura annonae} for the supply of wheat.
+The role of the \textit{praefectura annonae} was mainly to provide though \textit{frumentationes} a quantity per month of wheat to Roman citizens \citep{remesal_annona_1986,remesal_concierto}.
+
+The consumption of olive oil in \textit{Britannia} was residual until starting the Roman conquest \citep{funari_corpus_1996,
+carreras_abastecimiento_2003}.
+
+It is well known that this product was not frequently consumed by indigenous population.
+The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}.
+Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions.
+This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman army.
+
+At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}.
+This fact will be spatial incidence in sites close at Hadrian Wall.
+Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic points \citep{carreras_atlantic_2012}.
+The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges.
+Thus, the network was mainly focused on the support of soldiers during the military campaigns.
+%debate intercambios puntuales o redes comerciales de abastecimiento
+
+The presence of Dressel 20 stamps in military camps in \textit{Britannia} have been widely studied in Roman archaeology \citep{carreras_britannia_1998}.
+This fact also indicates a possible governmental organisation allocate to reinforce and supply of olive oil in military camps, such as \textit{Germania} (Remesal Rodrıguez, 1986).
+However, it is unknown the type of economical system when managing or control the redistribution of olive oil in different sites in \textit{Britannia} but it seems that it could have been in the main cities \citep[45]{funari_economic_2005}.
+
+Finally, The increase in olive oil exports would experience a progressive slowdown
+from the third century A.D., coinciding with the change in market strategy in the
+Empire. As a result, a progressive decrease of Dressel 20 is documented being gradually replaced for Dressel 23 amphorae \citep{rodriguez1991aceite,
+millet_anforas_1998}.
+
+The Roman conquest in \textit{Germania} dates from the end of the 1st century BC, during Caesar campaign and subsequently continued by Augusto.
+The advance of the Roman legions toward \textit{Gallia} opened a path through the Atlantic Ocean to \textit{Germania} \citep{remesal_annona_1986,
+remesal_baetica_2002}.
+It is thought that this route would have been initiated previously through terrestrial transport, although recent studies question this position giving an important role to the Atlantic route for both the conquests of \textit{Germania} and those of \textit{Britannia} \citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
+
+As the case of \textit{Britannia}, the consumption of olive oil in \textit{Germania} was not detected before the conquest.
+This is evidenced by the presence of Dressel 20 amphora in different sites close to limes.
+
+However, the Roman studies about the presence of Dressel 20 amphorae in \textit{Germania} has not had the same repercussion as the rest of the province.
+This could be explained due to the lack of archaeological sources \citep{horacio2010llegada}.
+
+As a consequence, we do not the role that commerce agents in the participation of the distribution and exportation of olive oil in \textit{Germania} or, on the contrary, the Roman army was the responsible to supply the product in the own province \citep[156]{remesal_germn_2010}.
+
+The presence of the Roman army encouraged the exchange in the province showed by the arrival of this product both civil settlement and military sites with a mayor concentration at German limes.
+It seems that some Baetican centres would be assigned to the support of olive oil.
+However, this hypothesis can not be proven by the lack of archaeological sources \citep[125]{remesal_concierto}.
+
+\section{A potential indicator of Roman economy: Dressel 20 stamps}
 
@@
 In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces.
-This economic connection could be affected by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamp in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province.
+This economic connection could be identified by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamp in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province.
+%\xavi{affected? No sería "identified"? Es decir, la correlación es la evidencia, pero la causa sería la interacción económica...no?}
 
-Therefore, this study proposes an evolutionary baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps.
+Therefore, this study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps.
 A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps.
 Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
+%\xavi{no veo la parte evolutiva en el paper; por otra parte, si es relevante no la lanzaría aquí al vuelo; hace falta explicar por qué es necesaria, qué implica, etc.}
 
@@
 
+%\xavi{Yo aplicaría aquí el feedback de los mapas para lo del libro...no creo que colorearlo basándose en el DEM aporte demasiado no?}
+
 The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
@@
 
-\subsubsection{Material}
+%\subsubsection{Jaccard distance}
+%The dataset was analysed using a statistic method as Jaccard distance.
+This method allows to measure the dissimilarity by calculating the presence of sets (CITAR).
+In our case, Jaccard distance was used to compute the mutual presence of traits in the amphora stamps but it does not consider the number of absences.
+A comparison was done with the distance of the workshops to identify whether there was an association between stamps and spatial distance amongst workshops.
+
+\section{Method and Analysis}
+
+\subsection{Production centres: Baetica province}
 
 We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in \textit{Baetica} province.
-The stamp database was compiled by CEIPAC database \citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}).
+The stamp database was compiled by CEIPAC group \citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}).
 However, approximately the 70 \% of stamps cannot be tested due to fragmentation or incomplete information.
@@
 
-\subsubsection{Exploratory Analysis}
+%\xavi{a ver, esto queda raro...si esto es una exploración de la base de datos entonces tiene que ir dentro de "Materials", así mientras vas comentando los sesgos y cosas generales de los datos (como soporte, pero no hace falta subsubsección "EDA")}
 
@@
 
+\xavi{quieres decir que esto es una power base? yo creo que no, no es tan bestia no? Yo lo matizaría}
+
 \begin{figure}[htp]
@@
 \includegraphics[width=\linewidth]{figs/frequencystamp.pdf}
-\caption{Distribution of the number of stamps}
+\caption{Histogram with the frequency of the number of stamps per workshop. The distribution is widely diverse with most stamps concentrated in one workshop}
 \label{stamps}
@@
 \includegraphics[width=\linewidth]{figs/frequency}
-\caption{Distribution of the number of different code stamps for each area. Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue).}
+\caption{Distribution of the number of different code stamps for each area. Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
 \label{frequency}
@@
 \subsection{Consumption centres: Britannia and Germania}
-
-The conquest of new provinces allowed to the Roman Empire the arrival of new resources.
-This lead to a progressive change in the economic and social structure.
-As a result, Augusto's administration created the figure of the \textit{praefectura annonae} for the supply of wheat.
-The role of the \textit{praefectura annonae} was mainly to provide though \textit{frumentationes} a quantity per month of wheat to Roman citizens \citep{remesal_annona_1986,remesal_concierto}.
-
-\subsubsection{Britannia}
-
-The consumption of olive oil in \textit{Britannia} was residual until starting the Roman conquest \citep{funari_corpus_1996,
-carreras_abastecimiento_2003}.
-
-It is well known that this product was not frequently consumed by indigenous population.
-The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}.
-Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions.
-This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman army.
-
-At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}.
-This fact will be spatial incidence in sites close at Hadrian Wall.
-Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic points \citep{carreras_atlantic_2012}.
-The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges.
-Thus, the network was mainly focused on the support of soldiers during the military campaigns.
-%debate intercambios puntuales o redes comerciales de abastecimiento
-
-The presence of Dressel 20 stamps in military camps in \textit{Britannia} have been widely studied in Roman archaeology \citep{carreras_britannia_1998}.
-This fact also indicates a possible governmental organisation allocate to reinforce and supply of olive oil in military camps, such as \textit{Germania} (Remesal Rodrıguez, 1986).
-However, it is unknown the type of economical system when managing or control the redistribution of olive oil in different sites in \textit{Britannia} but it seems that it could have been in the main cities \citep[45]{funari_economic_2005}.
-
-Finally, The increase in olive oil exports would experience a progressive slowdown
-from the third century A.D., coinciding with the change in market strategy in the
-Empire. As a result, a progressive decrease of Dressel 20 is documented being gradually replaced for Dressel 23 amphorae \citep{rodriguez1991aceite,
-millet_anforas_1998}.
-
-\subsubsection{Material}
 
@@
 
-\subsubsection{Germania}
-
-The Roman conquest in \textit{Germania} dates from the end of the 1st century BC, during Caesar campaign and subsequently continued by Augusto.
-The advance of the Roman legions toward \textit{Gallia} opened a path through the Atlantic Ocean to \textit{Germania} \citep{remesal_annona_1986,
-remesal_baetica_2002}.
-
-It is thought that this route would have been initiated previously through terrestrial transport, although recent studies question this position giving an important role to the Atlantic route for both the conquests of \textit{Germania} and those of \textit{Britannia} \citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
-
-As the case of \textit{Britannia}, the consumption of olive oil in \textit{Germania} was not detected before the conquest.
-This is evidenced by the presence of Dressel 20 amphora in different sites close to limes.
-
-However, the Roman studies about the presence of Dressel 20 amphorae in textit{Germania} has not had the same repercussion as the rest of the province.
-This could be explained due to the lack of archaeological sources \citep{horacio2010llegada}.
-
-As a consequence, we do not the role that commerce agents in the participation of the distribution and exportation of olive oil in \textit{Germania} or, on the contrary, the Roman army was the responsible to supply the product in the own province \citep[156]{remesal_germn_2010}.
-
-The presence of the Roman army encouraged the exchange in the province showed by the arrival of this product both civil settlement and military sites with a mayor concentration at German limes.
-It seems that some Baetican centres would be assigned to the support of olive oil.
-However, this hypothesis can not be proven by the lack of archaeological sources \citep[125]{remesal_concierto}.
-
-\subsubsection{Material}
-
 We analysed a dataset of 2052 stamps placed in \textit{Germania}.
@@
 \end{figure}
-
-%\subsubsection{Jaccard distance}
-%The dataset was analysed using a statistic method as Jaccard distance.
-This method allows to measure the dissimilarity by calculating the presence of sets (CITAR).
-In our case, Jaccard distance was used to compute the mutual presence of traits in the amphora stamps but it does not consider the number of absences.
-A comparison was done with the distance of the workshops to identify whether there was an association between stamps and spatial distance amongst workshops.
 
@@
 
-%hablar sobre Horn (Morisita)hablar sobre que el sample no estaba uniforme por eso usamos el morisita horn
+\subsection{Hieralchical clustering}
+
+\maria{que pongo aquí}
+
+%\xavi{pero este es el único método que usas?}
 
@@
 
-\begin{figure}[htp]
+\begin{sidewaysfigure}[htp]
 	\centering
@@
 \label{dendro}
-\end{figure}
+\end{sidewaysfigure}
 
@@
 
-\subsubsection{Britannia}
+The results show a similarity with the results obtained in \textit{Baetica} province. However, the similarity is less pronounced as we observed in the dendrogram in Fig. \ref{britmap}
 
-The results show a correlation similar to the obtained in \textit{Baetica} province. However, the similarity is less pronounced as we observed in the dendrogram in Fig. \ref{britmap}
+%\xavi{no tengo claro que le puedas decir correlación, porque eso es un valor específico entre 2 distribuciones numéricas. Te refieres a similaridad?}
 
-\begin{figure}[htp]
+\begin{sidewaysfigure}[htp]
 	\centering
@@
 \label{britmap}
-\end{figure}
+\end{sidewaysfigure}
 
@@
 
-\subsubsection{Germania}
-
 The similarity of stamps shows results less significant than \textit{Britannia} province as it can be seen in the dendrogram (Fig. \ref{germap}).
 
-\begin{figure}[htp]
+\begin{sidewaysfigure}[htp]
 	\centering
@@
 \label{germap}
-\end{figure}
+\end{sidewaysfigure}
 

### 2020-04-24 Paper/jaccardbaetica.tex
new changes xavi

@@
 
-%This paper explore--- resumir en dos líneas el objeto de estudio y definir que ha sido usado para ello dispersion of culture
-\xavi{este aim es demasiado general}
-This paper aims to understand the production dynamics in connection with different areas within the Roman Empire.
+%\xavi{este aim es demasiado general}
+%This paper aims to understand the production dynamics in connection with different areas within the Roman Empire. 
+This paper aims to study the olive oil market connection between province by calculating the similarity index.
 Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}.
@@
 We focus here on exploring the economic relation between stamps and amphora production and distribution centres.
-To do this, an population approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}.
+
+Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}).
+In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces.
+This economic connection could be identified by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province.
+%\xavi{affected? No sería "identified"? Es decir, la correlación es la evidencia, pero la causa sería la interacción económica...no?}
+
+%Our case study examines the relationship between the distribution of amphoric stamps and the workshops.
+In particular, we study the distribution of amphoric stamp in order to identify a correlation between geographical distance and amphoric stamps.
+Based on this assumption, we proposed three hypotheses: a) we can identify a correlation between spatial distance and the distribution of stamps, b) stamps located in close workshops share similar traits and c) Low mobility of amphoric stamps to other regions: stamps always stay in the same region.
+
+To do this, a population approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}.
 Stamps will be used to identify economic patterns by analysing their similarity.
@@
 As a result, hundreds of amphora workshops were made to support this high demand of the Roman Empire.
-These workshops were located along the rivers Guadalquivir and Genil, supplying from the riverine connectivity to the Mediterranean and Atlantic routes \citep{garcia_vargas_enrique_formal_2010}.
+The amphora workshops were situated in different locations in \textit{Baetica} province, along the rivers Guadalquivir and Genil (see Fig.\ref{workshop}).
+
+The location of the workshops with a riverine connectivity allowed to set up a connection with the Mediterranean and Atlantic routes \citep{garcia_vargas_enrique_formal_2010}.
+
+\begin{figure}[htp]
+	\centering
+\includegraphics[width=\linewidth]{figs/workshop1}
+\caption{Distribution of the workshops in Baetica province}
+\label{workshop}
+\end{figure}
+
+%\xavi{Yo aplicaría aquí el feedback de los mapas para lo del libro...no creo que colorearlo basándose en el DEM aporte demasiado no?}
+
+The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
 
@@
 Indeed, most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
+
+Some stamps show a more specific chronology while the majority of them display a large activity of production that it can be difficult for specifying an accurate chronology.
+This could be due to two reasons; firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible; secondly, Dressel 20 was produced during almost three centuries with apparently few changes \citep{berni_piero_chapter_2017}.
 
@@
 
-%Hyphotesis
 Nevertheless, some challenges remain under discussion such as how this production was organised and whether it is possible to distinguish production patterns in the olive oil trade.
@@
 Indeed, archaeological evidence shows a highly specialised production with a long activity in this specific area with apparently few changes \citep{remesal_anforas_2004}.
-
-%hipotesis: si esto es así entonces es asá
-
-The assumption of this study is analysing the effect of the production patterns between different centres.
-We are specially interested in identifying links between producers and consumption centres using amphoric stamps.
-To do that, we use the CEIPAC database to collect the stamps from different places.
-The CEIPAC dataset contains over 50.000 of epigraphy records found in amphorae, mostly from \textit{Monte Testaccio}.
-
-Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}).
-In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces.
-This economic connection could be identified by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamp in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province.
-%\xavi{affected? No sería "identified"? Es decir, la correlación es la evidencia, pero la causa sería la interacción económica...no?}
-
-Therefore, this study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps.
-A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps.
-Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
-%\xavi{no veo la parte evolutiva en el paper; por otra parte, si es relevante no la lanzaría aquí al vuelo; hace falta explicar por qué es necesaria, qué implica, etc.}
-
-%this paper aims to highlight the production dynamic in connection with the proximity between amphora workshops and stamps.
-
-\section{Material and methods}
-
-\subsection{Production centres: Baetica province}
-
-Our case study examines the relationship between the distribution of amphoric stamps and the workshops.
-In particular, we study the distribution of amphoric stamp in order to identify a correlation between geographical distance and amphoric stamps.
-Based on this assumption, we proposed three hypotheses: a) we can identify a correlation between spatial distance and the distribution of stamps, b) stamps located in close workshops share similar traits and c) Low mobility of amphoric stamps to other regions: stamps always stay in the same region.
-
-The amphora workshops were situated in different locations in \textit{Baetica} province, along the river Guadalquivir and its tributary Genil in order to detect similarities between stamps from workshops and spatial distance (see Fig.\ref{workshop}).
-
-\begin{figure}[htp]
-	\centering
-\includegraphics[width=\linewidth]{figs/workshop1}
-\caption{Distribution of the workshops in Baetica province}
-\label{workshop}
-\end{figure}
-
-%\xavi{Yo aplicaría aquí el feedback de los mapas para lo del libro...no creo que colorearlo basándose en el DEM aporte demasiado no?}
-
-The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
-Some stamps show a more specific chronology while the majority of them display a large activity of production that it can be difficult for specifying an accurate chronology.
-This could be due to two reasons; firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible; secondly, Dressel 20 was produced during almost three centuries with apparently few changes \citep{berni_piero_chapter_2017}.
 
@@
 \section{Method and Analysis}
+
+The assumption of this study is analysing the effect of the production patterns between different centres.
+We are specially interested in identifying links between producers and consumption centres using amphoric stamps.
+To do that, we use the CEIPAC database to collect the stamps from different places.
+The CEIPAC dataset contains over 50.000 of epigraphy records found in amphorae, mostly from \textit{Monte Testaccio}.
+Therefore, this study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps.
+A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps.
+Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
+%\xavi{no veo la parte evolutiva en el paper; por otra parte, si es relevante no la lanzaría aquí al vuelo; hace falta explicar por qué es necesaria, qué implica, etc.}
 
@@
 Most stamps are distributed in only one workshop whereas few workshops concentrate a high frequency of different amphoric stamps.
-The type of distribution is also frequent in the Roman economy where we observed a self-organised complex system pattern typical of a free market \citep{bayesian_2018}.
+The type of distribution is also frequent in amphora production where we observed a self-organized complex system pattern with a major concentration of the number of stamps in a few workshops \citep{bayesian_2018,coto-sarmiento_identifying_2018}.
 
-\xavi{quieres decir que esto es una power base? yo creo que no, no es tan bestia no? Yo lo matizaría}
+%The type of distribution is also frequent in the Roman economy where we observed a self-organised complex system pattern typical of a free market 
+%\xavi{quieres decir que esto es una power base? yo creo que no, no es tan bestia no? Yo lo matizaría}
 
@@
 \includegraphics[width=\linewidth]{figs/frequencystamp.pdf}
-\caption{Histogram with the frequency of the number of stamps per workshop. The distribution is widely diverse with most stamps concentrated in one workshop}
+\caption{Histogram with the number of stamps (X axis) and the frequency of distribution per workshop (Y axis). The distribution is widely diverse with most workshops concentrated one stamp}
 \label{stamps}
@@
 \includegraphics[width=\linewidth]{figs/frequency}
-\caption{Distribution of the number of different code stamps for each area. Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
+\caption{Distribution of the number of different code stamps (X axis) for each \textit{conventus} (Y axis).
+Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
 \label{frequency}
@@
 
-\subsection{Hieralchical clustering}
+\subsection{Hierarchical clustering}
 
-\maria{que pongo aquí}
+Morisita-Horn index was used to compute a matrix to group sites with the stamps codes that share similarities.
+The matrix was then visualised using hierarchical clustering.
+The algorithm was selected to cluster similar groups in order to analyse the relationship between groups of sites and the distribution of similar stamp codes.
+
+\maria{mejorar}
 

### 2020-04-28 Paper/jaccardbaetica.tex
new changes intro xavi

@@
 
-Material culture is one of the most frequent indicators of trade in the archaeological record.
+%spread of goods and ideas had a enormous impact during the Roman Empire. The progressive exploitation of communication networks allowed a major interaction between communities
+%This can be seen by the fact that material culture is found in different regions along the Mediterranean as a frequent indicator of trade exchange in the archaeological record.
+
+The role of the Mediterranean sea for the trade had an enormous impact during the Roman Empire.
+The spread of thousands of different goods along maritime shipping allowed a progressive exploitation of communication networks and a major interaction between communities \citep{rodriguez_baetican_1998,
+temin_market_2001,
+bevan_mediterranean_2014}. The exchange of different products between regions can be seen in both archaeological and written records as a frequent indicator of Roman trade.
+
+%Material culture is one of the most frequent indicators of trade in the archaeological record.
 In archaeology, they allow us to highlight a part of the mechanism of production and distribution of goods along the Mediterranean \citep{bevan_mediterranean_2014}.
-Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}\xavi{aquí faltan citas de obras más generales; esto es un ejemplo pero muy específico noÇ?}.
+Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}\xavi{aquí faltan citas de obras más generales; esto es un ejemplo pero muy específico no?}.
+
 Given this fact, an important mechanism of production control under the Roman government was created spreading by different areas attending to the richness of different places.
@@
 Olive oil was considered the liquid gold since it was used in different aspect of daily life for cooking, hygiene or lighting  \citep{mattingly_d.j._oil_1988}.
-As a consequence, this product was distributed in a large number of amphorae along the province, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.
+As a consequence, this product was distributed in a large number of amphorae along the provinces, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.
 
-Despite this issue has been widely discussed over many decades, the understanding of the olive oil production processes is still under debate, mostly due to the lack of written records.
-The application of different quantitative approaches combined with the archaeological evidence has allowed us partly to deal with the complexity of the Roman production \citep{orengo_seeds_2016,
-brughmans_roman_2016,coto-sarmiento_identifying_2018,bayesian_2018}.
+The understanding of the olive oil production processes has been widely discussed over many decades\citep{rodriguez_economioleicola_1977, Chic_hispania_1997,millet_anforas_1998}.
+Despite this topic remains under debate: what patterns were followed for the distribution of olive oil to the different provinces?
+Did each province follow a different pattern for its distribution? Nor the lack of written records has made possible to detect any indication of patterns in the olive oil market.
+
+Advances in the research of the Roman world have currently led an environment with more diverse commercial dynamics \citep{duncan1982economy,
+temin_economy_2006,
+quantifyingwilson2009}.
+The application of different quantitative approaches combined with the archaeological evidence has allowed us partly to deal with the complexity of the Roman production \citep{brughmans_roman_2016,
+orengo_seeds_2016,bayesian_2018,
+coto-sarmiento_identifying_2018,
+rubio-campillo_ecology_2018}.
 

### 2020-04-30 Paper/jaccardbaetica.tex
first draft

@@
 
-%cambiar quizás
 \begin{abstract}
@@
 
-%primer párrafo más general---- especifico (¿dónde se ha aplicado?)
-\xavi{queda raro esto de "en arqueología la cultura material nos permite ver la distribución en el Mediterráneo". Yo iría directa a hablar del Mediterráneo sin la primera frase.}
-
-%spread of goods and ideas had a enormous impact during the Roman Empire. The progressive exploitation of communication networks allowed a major interaction between communities
-%This can be seen by the fact that material culture is found in different regions along the Mediterranean as a frequent indicator of trade exchange in the archaeological record.
+%\xavi{queda raro esto de "en arqueología la cultura material nos permite ver la distribución en el Mediterráneo". Yo iría directa a hablar del Mediterráneo sin la primera frase.}
 
@@
 Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}\xavi{aquí faltan citas de obras más generales; esto es un ejemplo pero muy específico no?}.
+%spread of goods and ideas had a enormous impact during the Roman Empire. The progressive exploitation of communication networks allowed a major interaction between communities
+%This can be seen by the fact that material culture is found in different regions along the Mediterranean as a frequent indicator of trade exchange in the archaeological record.
 
@@
 
-\xavi{creo que le falta chicha aquí, sobre qué se ha hecho hasta ahora, qué quieres saber específicamente, el caso de estudio...}
+%\xavi{creo que le falta chicha aquí, sobre qué se ha hecho hasta ahora, qué quieres saber específicamente, el caso de estudio...}
 
@@
 Consequently, we discard integrate the fragmented stamps in our dataset.
-We finally filter a total sample of 987 stamps composed of 130 different stamps from 81 workshops.
+We finally filtered a total sample of 987 stamps composed of 130 different stamps from 81 workshops.
 
@@
 
-Morisita-Horn index was used to compute a matrix to group sites with the stamps codes that share similarities.
-The matrix was then visualised using hierarchical clustering.
+Morisita-Horn index resulted in a numeric matrix of dissimilarities containing the distance value between the sites and the code stamps.
+The method was used to group sites with similar stamps codes.
+The matrix was then computed using hierarchical clustering.
 The algorithm was selected to cluster similar groups in order to analyse the relationship between groups of sites and the distribution of similar stamp codes.
-
-\maria{mejorar}
+The results were visualised using a dendrogram to detect groups of sites sharing similar stamp codes.
 
@@
 \section{Discussion}
-
-%crear como una introducción
 

### 2020-05-06 Paper/jaccardbaetica.tex
más comentarios

@@
 %This paper aims to understand the production dynamics in connection with different areas within the Roman Empire.
+
+\xavi{le das demasiada importancia al similarity index. Es una simple herramienta de tu trabajo, y no un objetivo}.
 This paper aims to study the olive oil market connection between province by calculating the similarity index.
@@
 
+\xavi{Estás segura de esto?
+Porque aquí hay 2 cosas: 1.
+datar bien y 2 .
+período de uso.
+Yo no tengo claro que la mayoría de sellos se usen durante mucho tiempo.
+En qué está basada la idea?
+Deberías detallar un poco más este temita cronológico y explicar como sabemos el período de uso.}
 Some stamps show a more specific chronology while the majority of them display a large activity of production that it can be difficult for specifying an accurate chronology.
@@
 \section{Britannia and Germania}
+\xavi{este título de sección debería cambiarse por algo más descriptivo. Case studies o algo asínnn? Por otra parte, igual podrías dividirlo en 2 subsections}
+
+\xavi{en general, echo a faltar un poco más de detalle en esta discusión y en los datos que tienes; en qué se parecen y diferencian las 2 provincias?
+Si hablas del uso por las campañas o las guarniciones, donde están?
+Qué strategic points de redistribución hay?
+Tampoco a saco, pero entender mejor cómo llega el aceite, cómo se redistribuye y quién y dónde se consume}
 
@@
 
-It is well known that this product was not frequently consumed by indigenous population.
+It is well known that this product was not frequently consumed by indigenous population \xavi{CITATION NEEDED}.
 The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}.
@@
 
-Finally, The increase in olive oil exports would experience a progressive slowdown
-from the third century A.D., coinciding with the change in market strategy in the
-Empire. As a result, a progressive decrease of Dressel 20 is documented being gradually replaced for Dressel 23 amphorae \citep{rodriguez1991aceite,
-millet_anforas_1998}.
+Finally, The increase in olive oil exports would experience a progressive slowdown from the third century A.D., coinciding with the change in market strategy in the Empire.
+As a result, a progressive decrease of Dressel 20 is documented being gradually replaced for Dressel 23 amphorae \citep{rodriguez1991aceite,millet_anforas_1998}.
 
@@
 \section{A potential indicator of Roman economy: Dressel 20 stamps}
+\xavi{no sería mejor que esta sección esté antes de hablar de los casos de estudios? Así, podrías ya hablar de sellos encontrados en germania y britannia}
 
@@
 \section{Method and Analysis}
+\xavi{estas 2 cosas nunca van juntas. En Method explicas las técnicas, y Analysis en general no se usa porque es el todo del trabajo}
 
-The assumption of this study is analysing the effect of the production patterns between different centres.
+The assumption of this study is analysing the effect of the production patterns between different centres \xavi{esto no es ninguna asunción no??}.
 We are specially interested in identifying links between producers and consumption centres using amphoric stamps.
@@
 
+\xavi{que me quede claro...la mayoría de workshops solo tienen 1 sello??? Y hay un workshop con como 700 sellos distintos??? }
+
 \begin{figure}[htp]
@@
 \end{figure}
+
+\xavi{si el eje es logarítmico deberías mencionarlo}
 
@@
 
+\xavi{vale pero qué es cada punto? un workshop?}
+
 \subsection{Consumption centres: Britannia and Germania}
+
+\xavi{Esta estructura no me convence...tenemos que pensar algo mejor, porque ahora es: 1.
+hablo de britania/germania, 2.
+hablo de sellos en general, 3.
+hablo de sellos en britania/germania.
+Quizás juntar esto con lo de los casos de estudio, o quizás hablar de los 2 casos de estudio en la intro, y dividir la sección anterior en 2 (lo más general a la intro, lo más concreto aquí)}
 
@@
 \subsection{Measuring the dissimilarity in stamps}
+
+\xavi{en stamps o en yacimientos? o en provincias?}
 
@@
 \subsection{Reception centres: Britannia and Germania}
+\xavi{mejor consumption?}
 

### 2020-05-07 Paper/jaccardbaetica.tex
review xavi 2draft

@@
 
-%\xavi{queda raro esto de "en arqueología la cultura material nos permite ver la distribución en el Mediterráneo". Yo iría directa a hablar del Mediterráneo sin la primera frase.}
-
 The role of the Mediterranean sea for the trade had an enormous impact during the Roman Empire.
@@
 
-%\xavi{este aim es demasiado general}
-%This paper aims to understand the production dynamics in connection with different areas within the Roman Empire.
-
-\xavi{le das demasiada importancia al similarity index. Es una simple herramienta de tu trabajo, y no un objetivo}.
-This paper aims to study the olive oil market connection between province by calculating the similarity index.
+%\xavi{le das demasiada importancia al similarity index. Es una simple herramienta de tu trabajo, y no un objetivo}.
+This paper aims to study the olive oil market connection between provinces by calculating the similarity of stamps.
 Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}.
@@
 This economic connection could be identified by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province.
-%\xavi{affected? No sería "identified"? Es decir, la correlación es la evidencia, pero la causa sería la interacción económica...no?}
 
-%Our case study examines the relationship between the distribution of amphoric stamps and the workshops.
 In particular, we study the distribution of amphoric stamp in order to identify a correlation between geographical distance and amphoric stamps.
@@
 
-%\xavi{esta frase tendría que ir al final de la Intro; si pones subsecciones entonces queda muy raro creo}
-
 The paper addresses these questions as follows: the next section introduces the historical context.
@@
 
-%\xavi{creo que le falta chicha aquí, sobre qué se ha hecho hasta ahora, qué quieres saber específicamente, el caso de estudio...}
-
-\section{The Amphoric production in Baetica province}
-
-The high demand of Roman provinces stimulated by the good condition of the \textit{Baetica} lands allowed to develop a massive infrastructure of olive oil production.
-The production and distribution of olive oil were growing exponentially for almost three centuries \citep{remesal_concierto}.
-As a result, hundreds of amphora workshops were made to support this high demand of the Roman Empire.
-The amphora workshops were situated in different locations in \textit{Baetica} province, along the rivers Guadalquivir and Genil (see Fig.\ref{workshop}).
-
-The location of the workshops with a riverine connectivity allowed to set up a connection with the Mediterranean and Atlantic routes \citep{garcia_vargas_enrique_formal_2010}.
-
-\begin{figure}[htp]
-	\centering
-\includegraphics[width=\linewidth]{figs/workshop1}
-\caption{Distribution of the workshops in Baetica province}
-\label{workshop}
-\end{figure}
-
-%\xavi{Yo aplicaría aquí el feedback de los mapas para lo del libro...no creo que colorearlo basándose en el DEM aporte demasiado no?}
-
-The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
-
-The majority of olive oil amphorae produced in this province and shipped through the provinces belong to Dressel 20 type \citep{dressel_ricerche_1878,
-millet_anforas_1998}.
-This amphora is commonly associated with the transportation of Baetican olive oil during the Roman Empire \citep{berni_millet_epigrafianforica_2008}.
-Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti} with different information still unknown.
-Indeed, most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
-
-\xavi{Estás segura de esto?
-Porque aquí hay 2 cosas: 1.
-datar bien y 2 .
-período de uso.
-Yo no tengo claro que la mayoría de sellos se usen durante mucho tiempo.
-En qué está basada la idea?
-Deberías detallar un poco más este temita cronológico y explicar como sabemos el período de uso.}
-Some stamps show a more specific chronology while the majority of them display a large activity of production that it can be difficult for specifying an accurate chronology.
-This could be due to two reasons; firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible; secondly, Dressel 20 was produced during almost three centuries with apparently few changes \citep{berni_piero_chapter_2017}.
-
-\section{Britannia and Germania}
-\xavi{este título de sección debería cambiarse por algo más descriptivo. Case studies o algo asínnn? Por otra parte, igual podrías dividirlo en 2 subsections}
-
-\xavi{en general, echo a faltar un poco más de detalle en esta discusión y en los datos que tienes; en qué se parecen y diferencian las 2 provincias?
-Si hablas del uso por las campañas o las guarniciones, donde están?
-Qué strategic points de redistribución hay?
-Tampoco a saco, pero entender mejor cómo llega el aceite, cómo se redistribuye y quién y dónde se consume}
-
-%\xavi{Esta parte más histórica está chachi pero creo que debería estar en la Intro.
-Ahí podrías explicar la organización, y a partir de estas fuentes escritas plantear preguntas e hipótesis que justifiquen el paper y tu análisis usando arqueología.}
-
-The conquest of new provinces allowed to the Roman Empire the arrival of new resources.
-This lead to a progressive change in the economic and social structure.
-As a result, Augusto's administration created the figure of the \textit{praefectura annonae} for the supply of wheat.
-The role of the \textit{praefectura annonae} was mainly to provide though \textit{frumentationes} a quantity per month of wheat to Roman citizens \citep{remesal_annona_1986,remesal_concierto}.
-
-The consumption of olive oil in \textit{Britannia} was residual until starting the Roman conquest \citep{funari_corpus_1996,
-carreras_abastecimiento_2003}.
-
-It is well known that this product was not frequently consumed by indigenous population \xavi{CITATION NEEDED}.
-The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}.
-Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions.
-This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman army.
-
-At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}.
-This fact will be spatial incidence in sites close at Hadrian Wall.
-Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic points \citep{carreras_atlantic_2012}.
-The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges.
-Thus, the network was mainly focused on the support of soldiers during the military campaigns.
-%debate intercambios puntuales o redes comerciales de abastecimiento
-
-The presence of Dressel 20 stamps in military camps in \textit{Britannia} have been widely studied in Roman archaeology \citep{carreras_britannia_1998}.
-This fact also indicates a possible governmental organisation allocate to reinforce and supply of olive oil in military camps, such as \textit{Germania} (Remesal Rodrıguez, 1986).
-However, it is unknown the type of economical system when managing or control the redistribution of olive oil in different sites in \textit{Britannia} but it seems that it could have been in the main cities \citep[45]{funari_economic_2005}.
-
-Finally, The increase in olive oil exports would experience a progressive slowdown from the third century A.D., coinciding with the change in market strategy in the Empire.
-As a result, a progressive decrease of Dressel 20 is documented being gradually replaced for Dressel 23 amphorae \citep{rodriguez1991aceite,millet_anforas_1998}.
-
-The Roman conquest in \textit{Germania} dates from the end of the 1st century BC, during Caesar campaign and subsequently continued by Augusto.
-The advance of the Roman legions toward \textit{Gallia} opened a path through the Atlantic Ocean to \textit{Germania} \citep{remesal_annona_1986,
-remesal_baetica_2002}.
-It is thought that this route would have been initiated previously through terrestrial transport, although recent studies question this position giving an important role to the Atlantic route for both the conquests of \textit{Germania} and those of \textit{Britannia} \citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
-
-As the case of \textit{Britannia}, the consumption of olive oil in \textit{Germania} was not detected before the conquest.
-This is evidenced by the presence of Dressel 20 amphora in different sites close to limes.
-
-However, the Roman studies about the presence of Dressel 20 amphorae in \textit{Germania} has not had the same repercussion as the rest of the province.
-This could be explained due to the lack of archaeological sources \citep{horacio2010llegada}.
-
-As a consequence, we do not the role that commerce agents in the participation of the distribution and exportation of olive oil in \textit{Germania} or, on the contrary, the Roman army was the responsible to supply the product in the own province \citep[156]{remesal_germn_2010}.
-
-The presence of the Roman army encouraged the exchange in the province showed by the arrival of this product both civil settlement and military sites with a mayor concentration at German limes.
-It seems that some Baetican centres would be assigned to the support of olive oil.
-However, this hypothesis can not be proven by the lack of archaeological sources \citep[125]{remesal_concierto}.
-
 \section{A potential indicator of Roman economy: Dressel 20 stamps}
-\xavi{no sería mejor que esta sección esté antes de hablar de los casos de estudios? Así, podrías ya hablar de sellos encontrados en germania y britannia}
+%\xavi{no sería mejor que esta sección esté antes de hablar de los casos de estudios? Así, podrías ya hablar de sellos encontrados en germania y britannia}
 
@@
 
-\section{Method and Analysis}
-\xavi{estas 2 cosas nunca van juntas. En Method explicas las técnicas, y Analysis en general no se usa porque es el todo del trabajo}
+\section{The Amphoric production in Baetica province}
 
-The assumption of this study is analysing the effect of the production patterns between different centres \xavi{esto no es ninguna asunción no??}.
+The high demand of Roman provinces stimulated by the good condition of the \textit{Baetica} lands allowed to develop a massive infrastructure of olive oil production.
+The production and distribution of olive oil were growing exponentially for almost three centuries \citep{remesal_concierto}.
+As a result, hundreds of amphora workshops were made to support this high demand of the Roman Empire.
+The amphora workshops were situated in different locations in \textit{Baetica} province, along the rivers Guadalquivir and Genil (see Fig.\ref{workshop}).
+
+The location of the workshops with a riverine connectivity allowed to set up a connection with the Mediterranean and Atlantic routes \citep{garcia_vargas_enrique_formal_2010}.
+
+\begin{figure}[htp]
+	\centering
+\includegraphics[width=\linewidth]{figs/workshop1}
+\caption{Distribution of the workshops in Baetica province}
+\label{workshop}
+\end{figure}
+
+The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
+
+The majority of olive oil amphorae produced in this province and shipped through the provinces belong to Dressel 20 type \citep{dressel_ricerche_1878,
+millet_anforas_1998}.
+This amphora is commonly associated with the transportation of Baetican olive oil during the Roman Empire \citep{berni_millet_epigrafianforica_2008}.
+Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti} with different information still unknown.
+Indeed, most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
+
+\xavi{Estás segura de esto?
+Porque aquí hay 2 cosas: 1.
+datar bien y 2 .
+período de uso.
+Yo no tengo claro que la mayoría de sellos se usen durante mucho tiempo.
+En qué está basada la idea?
+Deberías detallar un poco más este temita cronológico y explicar como sabemos el período de uso.}
+Some stamps show a more specific chronology while the majority of them display a large activity of production that it can be difficult for specifying an accurate chronology.
+This could be due to two reasons; firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible; secondly, Dressel 20 was produced during almost three centuries with apparently few changes \citep{berni_piero_chapter_2017}.
+
+\section{Case study: consumption centers}
+%\xavi{este título de sección debería cambiarse por algo más descriptivo. Case studies o algo asínnn? Por otra parte, igual podrías dividirlo en 2 subsections}
+
+\xavi{en general, echo a faltar un poco más de detalle en esta discusión y en los datos que tienes; en qué se parecen y diferencian las 2 provincias?
+Si hablas del uso por las campañas o las guarniciones, donde están?
+Qué strategic points de redistribución hay?
+Tampoco a saco, pero entender mejor cómo llega el aceite, cómo se redistribuye y quién y dónde se consume}
+
+\subsection{Britannia}
+
+The conquest of new provinces allowed to the Roman Empire the arrival of new resources.
+This lead to a progressive change in the economic and social structure.
+As a result, Augusto's administration created the figure of the \textit{praefectura annonae} for the supply of wheat.
+The role of the \textit{praefectura annonae} was mainly to provide though \textit{frumentationes} a quantity per month of wheat to Roman citizens \citep{remesal_annona_1986,remesal_concierto}.
+
+The consumption of olive oil in \textit{Britannia} was residual until starting the Roman conquest \citep{funari_corpus_1996,
+carreras_abastecimiento_2003}.
+
+It is well known that this product was not frequently consumed by indigenous population \citep[161]{monfort_britanniaen_1998}.
+The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}.
+Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions.
+This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman army.
+
+At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}.
+This fact will be spatial incidence in sites close at Hadrian Wall.
+Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic points \citep{carreras_atlantic_2012}.
+The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges.
+Thus, the network was mainly focused on the support of soldiers during the military campaigns.
+
+The presence of Dressel 20 stamps in military camps in \textit{Britannia} have been widely studied in Roman archaeology \citep{carreras_britannia_1998}.
+This fact also indicates a possible governmental organisation allocate to reinforce and supply of olive oil in military camps, such as \textit{Germania} (Remesal Rodrıguez, 1986).
+However, it is unknown the type of economical system when managing or control the redistribution of olive oil in different sites in \textit{Britannia} but it seems that it could have been in the main cities \citep[45]{funari_economic_2005}.
+
+Finally, The increase in olive oil exports would experience a progressive slowdown from the third century A.D., coinciding with the change in market strategy in the Empire.
+As a result, a progressive decrease of Dressel 20 is documented being gradually replaced for Dressel 23 amphorae \citep{rodriguez1991aceite,millet_anforas_1998}.
+
+\subsection{Germania}
+
+The Roman conquest in \textit{Germania} dates from the end of the 1st century BC, during Caesar campaign and subsequently continued by Augusto.
+The advance of the Roman legions toward \textit{Gallia} opened a path through the Atlantic Ocean to \textit{Germania} \citep{remesal_annona_1986,
+remesal_baetica_2002}.
+It is thought that this route would have been initiated previously through terrestrial transport, although recent studies question this position giving an important role to the Atlantic route for both the conquests of \textit{Germania} and those of \textit{Britannia} \citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
+
+As the case of \textit{Britannia}, the consumption of olive oil in \textit{Germania} was not detected before the conquest.
+This is evidenced by the presence of Dressel 20 amphora in different sites close to limes.
+
+However, the Roman studies about the presence of Dressel 20 amphorae in \textit{Germania} has not had the same repercussion as the rest of the province.
+This could be explained due to the lack of archaeological sources \citep{horacio2010llegada}.
+
+As a consequence, we do not the role that commerce agents in the participation of the distribution and exportation of olive oil in \textit{Germania} or, on the contrary, the Roman army was the responsible to supply the product in the own province \citep[156]{remesal_germn_2010}.
+
+The presence of the Roman army encouraged the exchange in the province showed by the arrival of this product both civil settlement and military sites with a mayor concentration at German limes.
+It seems that some Baetican centres would be assigned to the support of olive oil.
+However, this hypothesis can not be proven by the lack of archaeological sources \citep[125]{remesal_concierto}.
+
+\section{Material}
+
+%\xavi{estas 2 cosas nunca van juntas. En Method explicas las técnicas, y Analysis en general no se usa porque es el todo del trabajo}
+
+The goal of this study is analysing the effect of the production patterns between different centres.
 We are specially interested in identifying links between producers and consumption centres using amphoric stamps.
@@
 Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
-%\xavi{no veo la parte evolutiva en el paper; por otra parte, si es relevante no la lanzaría aquí al vuelo; hace falta explicar por qué es necesaria, qué implica, etc.}
 
@@
 
-%\xavi{a ver, esto queda raro...si esto es una exploración de la base de datos entonces tiene que ir dentro de "Materials", así mientras vas comentando los sesgos y cosas generales de los datos (como soporte, pero no hace falta subsubsección "EDA")}
-
-We calculated the frequency of distribution in amphora stamps with the aim to carry out an exploratory analysis (EDA) to know the distribution and the number of stamps per each centre.
+We calculated the frequency of distribution in amphora stamps with the purpose to carry out an exploratory analysis (EDA) to know the distribution and the number of stamps per each centre.
 
@@
 %The type of distribution is also frequent in the Roman economy where we observed a self-organised complex system pattern typical of a free market
-%\xavi{quieres decir que esto es una power base? yo creo que no, no es tan bestia no? Yo lo matizaría}
 
@@
 \includegraphics[width=\linewidth]{figs/frequencystamp.pdf}
-\caption{Histogram with the number of stamps (X axis) and the frequency of distribution per workshop (Y axis). The distribution is widely diverse with most workshops concentrated one stamp}
+\caption{Histogram with the number of stamps (X axis) and the frequency of workshop (Y axis). The distribution is widely diverse with most workshops have only one stamp}
 \label{stamps}
@@
 \caption{Distribution of the number of different code stamps (X axis) for each \textit{conventus} (Y axis).
+Each dot corresponds to a workshop sorted by different areas.
 Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
@@
 
-\xavi{vale pero qué es cada punto? un workshop?}
+%\xavi{vale pero qué es cada punto? un workshop?}
 
@@
 
-\subsection{Measuring the dissimilarity in stamps}
+\section{Method: Measuring the dissimilarity}
 
-\xavi{en stamps o en yacimientos? o en provincias?}
+%\xavi{en stamps o en yacimientos? o en provincias?}
 
@@
 
-\subsection{Reception centres: Britannia and Germania}
-\xavi{mejor consumption?}
+\subsection{Consumption centres: Britannia and Germania}
+%\xavi{mejor consumption?}
 
@@
 Neither we found a grouping of similar stamps in a specific place.
-In general, we did not observe any production pattern that indicates a clear organisation between production centres and reception centres.
+In general, we did not observe any production pattern that indicates a clear organisation between production centres and consumption centres.
 In other words, our results did not indicate the presence of a specific production center from \textit{Baetica} in \textit{Britannia} province.

### 2020-05-08 Paper/jaccardbaetica.tex
add changes and add image draft2

@@
 
+\address[ceipacadress]{CEIPAC, Department of Prehistory and Archaeology, Montalegre, 6-8, 08001, University of Barcelona, Barcelona, Spain}
 \address[edadress]{School of History, Classic \& Archaeology, Room OOM.33, William Robertson Wing, Old Medical School, Teviot Place, University of Edinburgh, UK}
-\address[ceipacadress]{CEIPAC, Department of Prehistory and Archaeology, Montalegre, 6-8, 08001, University of Barcelona, Barcelona, Spain}
 
@@
 
+\section{The Amphoric production in Baetica province}
+
+The high demand of Roman provinces stimulated by the good condition of the \textit{Baetica} lands allowed to develop a massive infrastructure of olive oil production.
+The production and distribution of olive oil were growing exponentially for almost three centuries \citep{remesal_concierto}.
+As a result, hundreds of amphora workshops were made to support this high demand of the Roman Empire.
+The amphora workshops were situated in different locations in \textit{Baetica} province, along the rivers Guadalquivir and Genil (see Fig.\ref{workshop}).
+
+The location of the workshops with a riverine connectivity allowed to set up a connection with the Mediterranean and Atlantic routes \citep{garcia_vargas_enrique_formal_2010}.
+
+\begin{figure}[htp]
+	\centering
+\includegraphics[width=\linewidth]{figs/workshop1}
+\caption{Distribution of the workshops (dots) in Baetica province}
+\label{workshop}
+\end{figure}
+
+The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
+
+The majority of olive oil amphorae produced in this province and shipped through the provinces belong to Dressel 20 type \citep{dressel_ricerche_1878,
+millet_anforas_1998}.
+In particular, the importance of olive oil production could be explained by the fact that we detected amphorae distributed in different Roman provinces such as \textit{Germania}, \textit{Britannia} and \textit{Gallia}.
+This amphora was commonly associated with the  transportation of Baetican olive oil for supplying military camps and civil settlements \citep{berni_millet_epigrafianforica_2008}.
+
+Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti} with different information still unknown.
+Indeed, most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
+
+%\xavi{Estás segura de esto?
+Porque aquí hay 2 cosas: 1.
+datar bien y 2 .
+período de uso.
+Yo no tengo claro que la mayoría de sellos se usen durante mucho tiempo.
+En qué está basada la idea?
+Deberías detallar un poco más este temita cronológico y explicar como sabemos el período de uso.}
+
+%Some stamps show a more specific chronology while the majority of them display a large activity of production that it can be difficult for specifying an accurate chronology in some cases.
+This could be due to two reasons; firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible; secondly, Dressel 20 was produced during almost three centuries with apparently few changes \citep{berni_piero_chapter_2017}.
+
 \section{A potential indicator of Roman economy: Dressel 20 stamps}
@@
 
-Stamps are the most studied in this type of amphorae.
+Dressel 20 was the amphora most stamped during the Roman Empire \citep[18]{millet_anforas_1998}. The  period of use is associated with a large amphora production for almost three centuries.
+
+Most stamps display a large activity of production that it can be difficult for specifying an accurate chronology in some cases.
+They have been frequently dated with the consultar year by studying \textit{tituli picti} found in \textit{Monte Testaccio} \citep{Testaccio1, berni_millet_epigrafianforica_2008}.
+However, the chronology of stamps by using \textit{tituli picti} could be also biased in other sites taking into account that amphorae were deposited after being produced and consumed.  
+\maria{no sé cómo explicarlo mejor}
+
 The fascination aroused by stamps has caused it to be one of the inscriptions with the most publications about its origin and meaning.
@@
 A comparison was done with the distance of the workshops to identify whether there was an association between stamps and spatial distance amongst workshops.
-
-\section{The Amphoric production in Baetica province}
-
-The high demand of Roman provinces stimulated by the good condition of the \textit{Baetica} lands allowed to develop a massive infrastructure of olive oil production.
-The production and distribution of olive oil were growing exponentially for almost three centuries \citep{remesal_concierto}.
-As a result, hundreds of amphora workshops were made to support this high demand of the Roman Empire.
-The amphora workshops were situated in different locations in \textit{Baetica} province, along the rivers Guadalquivir and Genil (see Fig.\ref{workshop}).
-
-The location of the workshops with a riverine connectivity allowed to set up a connection with the Mediterranean and Atlantic routes \citep{garcia_vargas_enrique_formal_2010}.
-
-\begin{figure}[htp]
-	\centering
-\includegraphics[width=\linewidth]{figs/workshop1}
-\caption{Distribution of the workshops in Baetica province}
-\label{workshop}
-\end{figure}
-
-The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
-
-The majority of olive oil amphorae produced in this province and shipped through the provinces belong to Dressel 20 type \citep{dressel_ricerche_1878,
-millet_anforas_1998}.
-This amphora is commonly associated with the transportation of Baetican olive oil during the Roman Empire \citep{berni_millet_epigrafianforica_2008}.
-Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti} with different information still unknown.
-Indeed, most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
-
-\xavi{Estás segura de esto?
-Porque aquí hay 2 cosas: 1.
-datar bien y 2 .
-período de uso.
-Yo no tengo claro que la mayoría de sellos se usen durante mucho tiempo.
-En qué está basada la idea?
-Deberías detallar un poco más este temita cronológico y explicar como sabemos el período de uso.}
-Some stamps show a more specific chronology while the majority of them display a large activity of production that it can be difficult for specifying an accurate chronology.
-This could be due to two reasons; firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible; secondly, Dressel 20 was produced during almost three centuries with apparently few changes \citep{berni_piero_chapter_2017}.
 
@@
 
-The frequency of stamps per workshop can be seen in Fig.
-\ref{stamps}.
-Most stamps are distributed in only one workshop whereas few workshops concentrate a high frequency of different amphoric stamps.
+The frequency of stamps per workshop can be seen in Fig. \ref{stamps}. Most 
+workshop contained only one stamp whereas one workshop (La Catria) concentrated a high frequency of amphoric stamps with a total number of 228 stamps.
 The type of distribution is also frequent in amphora production where we observed a self-organized complex system pattern with a major concentration of the number of stamps in a few workshops \citep{bayesian_2018,coto-sarmiento_identifying_2018}.
@@
 
-\xavi{que me quede claro...la mayoría de workshops solo tienen 1 sello??? Y hay un workshop con como 700 sellos distintos??? }
+%\xavi{que me quede claro...la mayoría de workshops solo tienen 1 sello??? Y hay un workshop con como 700 sellos distintos??? }
 
@@
 \includegraphics[width=\linewidth]{figs/frequencystamp.pdf}
-\caption{Histogram with the number of stamps (X axis) and the frequency of workshop (Y axis). The distribution is widely diverse with most workshops have only one stamp}
+\caption{Histogram on a log scale with base 10.
+X axis is represented by the number of stamps and Y axis is the frequency of workshop.
+The distribution is widely diverse with most workshops have only one stamp}
 \label{stamps}
@@
 
-\xavi{si el eje es logarítmico deberías mencionarlo}
+%\xavi{si el eje es logarítmico deberías mencionarlo}
 
@@
 The majority of stamps found are concentrated in \textit{Hispalis} with 574 stamps while \textit{Corduba} and \textit{Astigi} with 267 and 146 stamps, respectively.
-Mostly, workshops show a homogeneity on the frequency of stamps except for both La Catria and Arva that exhibit a big amount of stamps with 29 different stamps.
+Mostly, workshops show a homogeneity on the frequency of stamps except for both La Catria and Arva that exhibit a big amount of stamps with 29 different code stamps.
 According to previous studies, those workshops became in most important centres of amphora production although they could have more intensely prospected than others.

### 2020-05-11 Paper/jaccardbaetica.tex
add new changes britannia germania

@@
 The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}. 
+This fact is shown by the archaeological evidence that displays a highly specialised production with a long activity with apparently few changes \citep{remesal_anforas_2004}.
 
-The majority of olive oil amphorae produced in this province and shipped through the provinces belong to Dressel 20 type \citep{dressel_ricerche_1878,
-millet_anforas_1998}.
-In particular, the importance of olive oil production could be explained by the fact that we detected amphorae distributed in different Roman provinces such as \textit{Germania}, \textit{Britannia} and \textit{Gallia}.
-This amphora was commonly associated with the  transportation of Baetican olive oil for supplying military camps and civil settlements \citep{berni_millet_epigrafianforica_2008}.
+The majority of olive oil amphorae produced in this province belong to Dressel 20 type\citep{dressel_ricerche_1878,
+millet_anforas_1998}. Dressel 20 amphorae were shipped and distributed from Baetica throughout western Europe.
 
-Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti} with different information still unknown.
+Dressel 20 contains many information still unknown.
+A high percentage of  Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti}.
 Indeed, most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
@@
 Neither written records have been found that it could explain the economic role of \textit{Baetica} province in the Roman organisation.
-Indeed, archaeological evidence shows a highly specialised production with a long activity in this specific area with apparently few changes \citep{remesal_anforas_2004}.
 
@@
 \section{Case study: consumption centers}
+
+The conquest of new provinces allowed to the Roman Empire the arrival of new resources through Mediterranean and Atlantic routes.
+This lead to a progressive change in the economic and social structure creating new trade networks for supplying Roman conquest.
+As a result, Augusto's administration created the figure of the \textit{praefectura annonae} for the supply of wheat.
+The role of the \textit{praefectura annonae} was mainly to provide though \textit{frumentationes} a quantity per month of wheat to Roman citizens \citep{remesal_annona_1986,remesal_concierto}.
+
+\textit{Annona} could have also included other goods for the supply of Roman legions such as olive oil according for some hypotheses \citep{remesal_annona_1986,remesal_annona_1990}
+
+The importance of olive oil production could be explained by the fact that we detected a massive amphora distribution in different Roman provinces such as \textit{Germania} and \textit{Britannia}.
+This amphora has been commonly associated with the  transportation of Baetican olive oil for supplying military camps and civil settlements \citep{berni_millet_epigrafianforica_2008}.
+
+The concentration of amphorae in eminently militarised areas suggests a possible existence of a broad supply mechanism, being accurate to imagine a possible organisation in charge of its distribution from military camps to civil settlements\citep{remesal_annona_1986, carreras_britannia_1998}.
+
+The military consumption of olive oil in Roman provinces could be related to two aspects: a) cultural consumption whose product is consumed by cultural reasons such as to identity or habit and b) economical reasons where olive oil distribution is based on the transport costs unlike other products \citep[69-70]{carreras_britannia_1998}.
+
+\maria{falta conexión entre una cosa y hacer una breve introducción de lo siguiente britannia y germania}
+
 %\xavi{este título de sección debería cambiarse por algo más descriptivo. Case studies o algo asínnn? Por otra parte, igual podrías dividirlo en 2 subsections}
@@
 
-The conquest of new provinces allowed to the Roman Empire the arrival of new resources.
-This lead to a progressive change in the economic and social structure.
-As a result, Augusto's administration created the figure of the \textit{praefectura annonae} for the supply of wheat.
-The role of the \textit{praefectura annonae} was mainly to provide though \textit{frumentationes} a quantity per month of wheat to Roman citizens \citep{remesal_annona_1986,remesal_concierto}.
-
 The consumption of olive oil in \textit{Britannia} was residual until starting the Roman conquest \citep{funari_corpus_1996,
 carreras_abastecimiento_2003}.
-
 It is well known that this product was not frequently consumed by indigenous population \citep[161]{monfort_britanniaen_1998}.
-The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}.
+The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[1]{carreras_britannia_1998}.
 Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions.

### 2020-05-12 Paper/jaccardbaetica.tex
second draft finish

@@
 \newcommand{\memo}[2]{\textcolor{#1}{#2}}
-\newcommand{\maria}[1]{\memo{red}{#1\\}}
+\newcommand{\maria}[1]{\memo{red}{MC: #1\\}}
 \newcommand{\xavi}[1]{\memo{magenta}{XRC: #1\\}}
@@
 \begin{keyword}
-Roman Empire; amphora workshops; Dressel 20; social learning; cultural evolution
+Roman Empire; amphora production; Dressel 20; dissimilarity index; Roman provinces
 \end{keyword}
@@
 In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps.
-\textit{Baetica} became important production and distribution centre during the Roman Empire.
+\textit{Baetica} became an important production and distribution centre during the Roman Empire.
 However, it remains under debate about how this province was organised and whether it could be possible to identify patterns in the olive oil market.
@@
 They will be used to detect a connection between the distribution of amphoric stamps and the economic structure in both centres.
-Here we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance.
+Here, we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance.
 
@@
 Olive oil was considered the liquid gold since it was used in different aspect of daily life for cooking, hygiene or lighting  \citep{mattingly_d.j._oil_1988}.
-As a consequence, this product was distributed in a large number of amphorae along the provinces, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.
+As a consequence, this product was distributed in a large number of amphorae across the provinces, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.
 
@@
 
-The location of the workshops with a riverine connectivity allowed to set up a connection with the Mediterranean and Atlantic routes \citep{garcia_vargas_enrique_formal_2010}.
+The location of the workshops with a riverine connectivity allowed to arrange a connection with the Mediterranean and Atlantic routes \citep{garcia_vargas_enrique_formal_2010}.
 
@@
 The majority of olive oil amphorae produced in this province belong to Dressel 20 type\citep{dressel_ricerche_1878,
-millet_anforas_1998}. Dressel 20 amphorae were shipped and distributed from Baetica throughout western Europe.
+millet_anforas_1998}. Dressel 20 amphorae were shipped and distributed from \textit{Baetica} throughout western Europe.
 
-Dressel 20 contains many information still unknown.
+Dressel 20 contains much information still unknown.
 A high percentage of  Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti}.
@@
 Most stamps display a large activity of production that it can be difficult for specifying an accurate chronology in some cases.
-They have been frequently dated with the consultar year by studying \textit{tituli picti} found in \textit{Monte Testaccio} \citep{Testaccio1, berni_millet_epigrafianforica_2008}.
+They have been frequently dated with the consular dating by studying \textit{tituli picti} found in \textit{Monte Testaccio} \citep{Testaccio1, berni_millet_epigrafianforica_2008}.
 However, the chronology of stamps by using \textit{tituli picti} could be also biased in other sites taking into account that amphorae were deposited after being produced and consumed.  
-\maria{no sé cómo explicarlo mejor}
+\maria{no sé cómo explicarlo mejor ni como citarlo porque no hay citas}
 
 The fascination aroused by stamps has caused it to be one of the inscriptions with the most publications about its origin and meaning.
-Its endure in time become an essential element that were used to mark amphorae during almost three centuries \citep{remesal_sellar_2016} (Fig.\ref{amphora}).
+Its endure in time become an essential element that was used to mark amphorae during almost three centuries \citep{remesal_sellar_2016} (Fig.\ref{amphora}).
 
@@
 They could be identified based on three premises: content (olive oil), context (amphora workshop) and subject (individuals involved).
-On the one hand, it seems that stamps could have been identified as the land owner of the olive groves \citep{rodriguez_economioleicola_1977}.
+On the one hand, it seems that stamps could have been identified as the landowner of the olive groves \citep{rodriguez_economioleicola_1977}.
 On the other, they could belong to the owner of the making amphorae workshops or even a group of amphora workers \citep{berni_millet_epigrafianforica_2008}.
@@
 
-\section{Case study: consumption centers}
+\section{Case study: consumption centres}
 
-The conquest of new provinces allowed to the Roman Empire the arrival of new resources through Mediterranean and Atlantic routes.
+The conquest of new provinces allowed the Roman Empire the arrival of new resources through Mediterranean and Atlantic routes.
 This lead to a progressive change in the economic and social structure creating new trade networks for supplying Roman conquest.
@@
 
-The importance of olive oil production could be explained by the fact that we detected a massive amphora distribution in different Roman provinces such as \textit{Germania} and \textit{Britannia}.
+The importance of olive oil production could be explained by the fact that we detected a massive amphora distribution in different Roman provinces.
 This amphora has been commonly associated with the  transportation of Baetican olive oil for supplying military camps and civil settlements \citep{berni_millet_epigrafianforica_2008}.
 
-The concentration of amphorae in eminently militarised areas suggests a possible existence of a broad supply mechanism, being accurate to imagine a possible organisation in charge of its distribution from military camps to civil settlements\citep{remesal_annona_1986, carreras_britannia_1998}.
+The military consumption of olive oil in Roman provinces could be related to two aspects: a) cultural consumption whose product is consumed by cultural reasons such as to identity or habit and b) economical reasons where olive oil distribution is based on the transport costs, unlike other products \citep[69-70]{carreras_britannia_1998}.
 
-The military consumption of olive oil in Roman provinces could be related to two aspects: a) cultural consumption whose product is consumed by cultural reasons such as to identity or habit and b) economical reasons where olive oil distribution is based on the transport costs unlike other products \citep[69-70]{carreras_britannia_1998}.
-
-\maria{falta conexión entre una cosa y hacer una breve introducción de lo siguiente britannia y germania}
+Olive oil supply had a special impact in militarised provinces such as \textit{Britannia} and \textit{Germania}.
+The concentration of amphorae in eminently militarised areas suggests the existence of a broad supply mechanism, being accurate to imagine a possible organisation in charge of its distribution from military camps to civil settlements\citep{remesal_annona_1986, carreras_britannia_1998}.
 
@@
 
-\xavi{en general, echo a faltar un poco más de detalle en esta discusión y en los datos que tienes; en qué se parecen y diferencian las 2 provincias?
+%\xavi{en general, echo a faltar un poco más de detalle en esta discusión y en los datos que tienes; en qué se parecen y diferencian las 2 provincias?
 Si hablas del uso por las campañas o las guarniciones, donde están?
@@
 
-\xavi{Esta estructura no me convence...tenemos que pensar algo mejor, porque ahora es: 1.
+%\xavi{Esta estructura no me convence...tenemos que pensar algo mejor, porque ahora es: 1.
 hablo de britania/germania, 2.
@@
 
-The analysis of the amphora workshops in Baetica province resulted in a significant correlation between spatial distance and the similarity of the amphoric stamps.
+The analysis of the amphora workshops in \textit{Baetica} province resulted in a significant correlation between spatial distance and the similarity of the amphoric stamps.
 

### 2020-05-14 Paper/jaccardbaetica.tex
last review draft

@@
 
-Dressel 20 was the amphora most stamped during the Roman Empire \citep[18]{millet_anforas_1998}. The  period of use is associated with a large amphora production for almost three centuries.
+Dressel 20 was the amphora most stamped during the Roman Empire \citep[18]{millet_anforas_1998}. The usage period is associated with a large amphora production for almost three centuries.
 
@@
 They have been frequently dated with the consular dating by studying \textit{tituli picti} found in \textit{Monte Testaccio} \citep{Testaccio1, berni_millet_epigrafianforica_2008}.
-However, the chronology of stamps by using \textit{tituli picti} could be also biased in other sites taking into account that amphorae were deposited after being produced and consumed.  
+However, the chronology of stamps could be also biased in other sites taking into account that amphorae were deposited after being produced and consumed.  
 \maria{no sé cómo explicarlo mejor ni como citarlo porque no hay citas}
@@
 
-The goal of this study is analysing the effect of the production patterns between different centres.
-We are specially interested in identifying links between producers and consumption centres using amphoric stamps.
+The goal of this study is exploring the effect of the production patterns between different centres.
+We are specially interested in identifying links between producers and consumption centres by using amphoric stamps.
 To do that, we use the CEIPAC database to collect the stamps from different places.

### 2020-05-27 Paper/jaccardbaetica.tex
comentarios
new changes xavi correction

@@
 Amphoric stamps are used to identify the presence of different groups that might share similar stamps.
-To achieve this goal, we analyse a set of stamps from two centres: 1) producer centres by analysing different workshops in \textit{Baetica} province and 2) two Roman provinces such as \textit{Germania} and \textit{Britannia} as a receptor centres.
+To achieve this goal, we analyse a set of stamps from two centres: 1) production centres by analysing different workshops in \textit{Baetica} province and 2) two Roman provinces such as \textit{Germania} and \textit{Britannia} as consumption centres.
 They will be used to detect a connection between the distribution of amphoric stamps and the economic structure in both centres.
@@
 The role of the Mediterranean sea for the trade had an enormous impact during the Roman Empire.
-The spread of thousands of different goods along maritime shipping allowed a progressive exploitation of communication networks and a major interaction between communities \citep{rodriguez_baetican_1998,
+The spread of thousands of different goods along maritime shipping allowed progressive exploitation of communication networks and a major interaction between communities \citep{rodriguez_baetican_1998,
 temin_market_2001,
@@
 Olive oil was considered the liquid gold since it was used in different aspect of daily life for cooking, hygiene or lighting  \citep{mattingly_d.j._oil_1988}.
-As a consequence, this product was distributed in a large number of amphorae across the provinces, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.
+As a consequence of high demand, this product was distributed in a large number of amphorae across the provinces, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.
 
 The understanding of the olive oil production processes has been widely discussed over many decades\citep{rodriguez_economioleicola_1977, Chic_hispania_1997,millet_anforas_1998}.
-Despite this topic remains under debate: what patterns were followed for the distribution of olive oil to the different provinces?
+Despite this topic is still subject to debate: what patterns were followed for the distribution of olive oil to the different provinces?
 Did each province follow a different pattern for its distribution? Nor the lack of written records has made possible to detect any indication of patterns in the olive oil market.
 
-Advances in the research of the Roman world have currently led an environment with more diverse commercial dynamics \citep{duncan1982economy,
+Advances in the research of the Roman studies have currently led an environment with more diverse commercial dynamics \citep{duncan1982economy,
 temin_economy_2006,
@@
 
-%\xavi{le das demasiada importancia al similarity index. Es una simple herramienta de tu trabajo, y no un objetivo}.
 This paper aims to study the olive oil market connection between provinces by calculating the similarity of stamps.
@@
 In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces.
-This economic connection could be identified by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province.
+This economic connection could be identified by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific province.
 
-In particular, we study the distribution of amphoric stamp in order to identify a correlation between geographical distance and amphoric stamps.
-Based on this assumption, we proposed three hypotheses: a) we can identify a correlation between spatial distance and the distribution of stamps, b) stamps located in close workshops share similar traits and c) Low mobility of amphoric stamps to other regions: stamps always stay in the same region.
+In particular, we study the distribution of amphoric stamps to identify a correlation between geographical distance and similarity.
+Based on this assumption, we proposed three hypotheses: a) we can identify a correlation between spatial distance and the distribution of stamps, b) stamps located in close workshops share similar traits and c) low mobility of amphoric stamps to other regions: stamps always stay in the same region.
 
@@
 The production and distribution of olive oil were growing exponentially for almost three centuries \citep{remesal_concierto}.
-As a result, hundreds of amphora workshops were made to support this high demand of the Roman Empire.
+As a result, hundreds of amphora workshops were raised to support this high demand of the Roman Empire.
 The amphora workshops were situated in different locations in \textit{Baetica} province, along the rivers Guadalquivir and Genil (see Fig.\ref{workshop}).
 
-The location of the workshops with a riverine connectivity allowed to arrange a connection with the Mediterranean and Atlantic routes \citep{garcia_vargas_enrique_formal_2010}.
+The strategic location of the workshops close to the river allowed the transport of olive oil through river connectivity to the routes of the Mediterranean and the Atlantic \citep{garcia_vargas_enrique_formal_2010}.
 
@@
 \section{A potential indicator of Roman economy: Dressel 20 stamps}
-%\xavi{no sería mejor que esta sección esté antes de hablar de los casos de estudios? Así, podrías ya hablar de sellos encontrados en germania y britannia}
 
@@
 However, the chronology of stamps could be also biased in other sites taking into account that amphorae were deposited after being produced and consumed.
-\maria{no sé cómo explicarlo mejor ni como citarlo porque no hay citas}
 
@@
 Stamps are mostly formed by a code of three letters and they can appear in abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
-Despite researchers identify stamps as identity marks, there is not a general consensus about the meaning of the stamps \citep{rodriguez_baetican_1998}.
-Alike most Dressel 20 containers were not stamped being difficult to determinate.
+
+Despite researchers identify stamps as identity marks, there is not a consensus about the meaning of the stamps \citep{rodriguez_baetican_1998}.
+Alike a high percentage of Dressel 20 containers were not stamped being difficult to determinate.
 
@@
 Did stamps share the same workshop?
-Neither written records have been found that it could explain the economic role of \textit{Baetica} province in the Roman organisation.
+
+Neither the use of written records have allowed providing enough information that can explain the economic role of \textit{Baetica} province in the Roman organisation.
 
@@
 
-The conquest of new provinces allowed the Roman Empire the arrival of new resources through Mediterranean and Atlantic routes.
+The conquest of new provinces allowed the Roman Empire the arrival of resources through Mediterranean and Atlantic routes.
 This lead to a progressive change in the economic and social structure creating new trade networks for supplying Roman conquest.
@@
 
-The military consumption of olive oil in Roman provinces could be related to two aspects: a) cultural consumption whose product is consumed by cultural reasons such as to identity or habit and b) economical reasons where olive oil distribution is based on the transport costs, unlike other products \citep[69-70]{carreras_britannia_1998}.
+The military consumption of olive oil in Roman provinces may be related to two aspects: a) cultural consumption whose product is consumed by cultural reasons such as to identity or habit and b) economical reasons where olive oil distribution is based on the transport costs, unlike other products \citep[69-70]{carreras_britannia_1998}.
 
@@
 At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}.
-This fact will be spatial incidence in sites close at Hadrian Wall.
+This fact will have a spatial incidence in sites close at Hadrian Wall.
 Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic points \citep{carreras_atlantic_2012}.
 The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges.
-Thus, the network was mainly focused on the support of soldiers during the military campaigns.
+Thus, the network was mainly focused on the support of soldiers during military campaigns.
 
@@
 
-Finally, The increase in olive oil exports would experience a progressive slowdown from the third century A.D., coinciding with the change in market strategy in the Empire.
+Finally, the increase in olive oil exports would experience a progressive slowdown from the third century A.D., coinciding with the change in market strategy in the Empire.
 As a result, a progressive decrease of Dressel 20 is documented being gradually replaced for Dressel 23 amphorae \citep{rodriguez1991aceite,millet_anforas_1998}.
@@
 
-As the case of \textit{Britannia}, the consumption of olive oil in \textit{Germania} was not detected before the conquest.
-This is evidenced by the presence of Dressel 20 amphora in different sites close to limes.
-
-However, the Roman studies about the presence of Dressel 20 amphorae in \textit{Germania} has not had the same repercussion as the rest of the province.
+The Roman studies about the presence of Dressel 20 amphorae in \textit{Germania} has not had the same repercussion as the rest of the province.
 This could be explained due to the lack of archaeological sources \citep{horacio2010llegada}.
@@
 
-The presence of the Roman army encouraged the exchange in the province showed by the arrival of this product both civil settlement and military sites with a mayor concentration at German limes.
+As the case of \textit{Britannia}, the consumption of olive oil in \textit{Germania} was not detected before the conquest.
+This is evidenced by the presence of Dressel 20 amphora in different sites close to \textit{limes}.
+
+The presence of the Roman army encouraged the exchange in the province showed by the arrival of this product both civil settlement and military sites with a mayor concentration at German \textit{limes}.
 It seems that some Baetican centres would be assigned to the support of olive oil.
@@
 
-\section{Material}
-
-%\xavi{estas 2 cosas nunca van juntas. En Method explicas las técnicas, y Analysis en general no se usa porque es el todo del trabajo}
+\section{Material and Methods}
 
 The goal of this study is exploring the effect of the production patterns between different centres.
-We are specially interested in identifying links between producers and consumption centres by using amphoric stamps.
-To do that, we use the CEIPAC database to collect the stamps from different places.
+We are especially interested in identifying links between production and consumption centres by using amphoric stamps.
+To do that, we use the CEIPAC database to collect stamps from different places.
 The CEIPAC dataset contains over 50.000 of epigraphy records found in amphorae, mostly from \textit{Monte Testaccio}.
+
 Therefore, this study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps.
 A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps.
-Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
+Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potential spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
 
@@
 We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in \textit{Baetica} province.
-The stamp database was compiled by CEIPAC group \citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}).
-However, approximately the 70 \% of stamps cannot be tested due to fragmentation or incomplete information.
-Consequently, we discard integrate the fragmented stamps in our dataset.
+The database was compiled by CEIPAC group\citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}).
+However, approximately 70 \% of stamps cannot be tested due to the fragmentation of the dataset.
+Consequently, any stamps with incomplete information were discarded and not integrated into our dataset.
 We finally filtered a total sample of 987 stamps composed of 130 different stamps from 81 workshops.
@@
 We also created a new column with the area of the stamps.
-These areas, known as \textit{conventus}, were administrative centres for territorial organization in the Roman Empire.
+These areas, known as \textit{conventus}, were administrative centres for territorial organisation in the Roman Empire \citep[58]{ozcariz_gil_administracion_2013}.
 Dressel 20 stamps were found in three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter Hispalis), \textit{Cordubensis} (currently C\'ordoba, hereafter Corduba) and \textit{Astigi} (currently Écija, Sevilla, hereafter Astigi) \citep{rodriguez_economioleicola_1977,
-chicdatos2001,berni_millet_epigrafianforica_2008} .
-
-We calculated the frequency of distribution in amphora stamps with the purpose to carry out an exploratory analysis (EDA) to know the distribution and the number of stamps per each centre.
+chicdatos2001,berni_millet_epigrafianforica_2008} . It is important to mention that some workshops exhibited the same coordinates but they were differently catalogued as a different workshop. 
+\maria{seguir con esta mierda}
+We calculated the frequency of distribution in amphora stamps with the purpose of implementing an exploratory analysis (EDA) to detect the distribution and the number of stamps per each centre.
 
 The frequency of stamps per workshop can be seen in Fig. \ref{stamps}. Most 
-workshop contained only one stamp whereas one workshop (La Catria) concentrated a high frequency of amphoric stamps with a total number of 228 stamps.
-The type of distribution is also frequent in amphora production where we observed a self-organized complex system pattern with a major concentration of the number of stamps in a few workshops \citep{bayesian_2018,coto-sarmiento_identifying_2018}.
+workshops contained only one stamp whereas one workshop (La Catria) concentrated a high frequency of amphoric stamps with a total number of 228 stamps.
+The type of distribution is also frequent in amphora production where we observed a self-organised complex system pattern with a major concentration of the number of stamps in a few workshops \citep{bayesian_2018,coto-sarmiento_identifying_2018}.
 
@@
 \ref{frequency}.
-The majority of stamps found are concentrated in \textit{Hispalis} with 574 stamps while \textit{Corduba} and \textit{Astigi} with 267 and 146 stamps, respectively.
+The majority of stamps are concentrated in \textit{Hispalis} with 574 stamps while \textit{Corduba} and \textit{Astigi} with 267 and 146 stamps, respectively.
 Mostly, workshops show a homogeneity on the frequency of stamps except for both La Catria and Arva that exhibit a big amount of stamps with 29 different code stamps.
@@
 Each dot corresponds to a workshop sorted by different areas.
-Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
+Colours are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
 \label{frequency}
@@
 All the data was also compiled by CEIPAC database.
-We finally selected a total of 1621 stamps consisted of 850 different stamps from 46 sites (see Fig.
+We finally collected a total of 1621 stamps consisted of 850 different stamps from 46 sites (see Fig.
 \ref{germania}).
@@
 \includegraphics[width=\linewidth]{figs/germania5_2.pdf}
-\caption{Distribution of the sites in \textit{Germania}. Most sites were located in limes during the Roman Empire}
+\caption{Distribution of the sites in \textit{Germania}. Most sites were located in German \textit{limes}}
 \label{germania}
@@
 
-\section{Method: Measuring the dissimilarity}
-
-%\xavi{en stamps o en yacimientos? o en provincias?}
+\subsection{Measuring the dissimilarity}
 
@@
 
-%\xavi{pero este es el único método que usas?}
-
 \section{Results}
@@
 
-The analysis shows that amphoric stamps could be correlated with spatial distance.
-The correlation coefficients range from a minimum to a maximum.
+%The correlation coefficients range from a minimum to a maximum.
+
+The analysis shows that the similarity of amphoric stamps could be correlated with spatial distance.
 The dendrogram in Fig.
@@
 The dendrogram suggests that amphora workshops used different stamps for their production system.
-Nearby workshops show similarity on the stamps while most of them seem to display different stamps roles.
-Groups of workshops sharing similar stamps were not found in the cluster: the majority of stamp grouping was composed of no more than three workshops.
+Nearby workshops show similarity on the stamps while most workshops seem to display different stamps.
+Groups of workshops sharing similar stamps were not found in the dendrogram: the majority of stamp grouping was composed of no more than three workshops.
 Additionally, workshops that shared more similar amphoric stamps belonged to the same \textit{conventus} area, such as Picachos, Cerro de los Pesebres and El Castillejo.
@@
 	\centering
-\includegraphics[width=\linewidth]{figs/dendro}
+\includegraphics[angle=180, width=\linewidth]{figs/dendro}
 \caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in \textit{Baetica} area.
@@
 \subsection{Consumption centres: Britannia and Germania}
-%\xavi{mejor consumption?}
 
@@
 	\centering
-\includegraphics[width=\linewidth]{figs/dendrobrit5.pdf}
+\includegraphics[angle=180,width=\linewidth]{figs/dendrobrit5.pdf}
 \caption{Dendrogram obtained by Morisita-Horn algorithm of different sites in \textit{Britannia}.
@@
 
-A minor correlation can be explained by the fact that \textit{Britannia} spatial area is much wide than \textit{Baetica} where the centres were more concentrated.
+A minor correlation can be explained by the fact that \textit{Britannia} spatial area is much wider than \textit{Baetica} where the centres were more concentrated.
 
-Most military sites that showed greater similarity in the results were geographically close.
-We identified two factors in the results: 1) nearby sites tend to share similar stamps and 2) most sites have different stamps.
+Most military sites that showed a high similarity in the results were geographically close.
+We identified two factors: 1) nearby sites tend to share similar stamps and 2) most sites have different stamps.
 By contrast, most sites did not show a strong similarity in stamp correlation.
 Neither we found a grouping of similar stamps in a specific place.
-In general, we did not observe any production pattern that indicates a clear organisation between production centres and consumption centres.
-In other words, our results did not indicate the presence of a specific production center from \textit{Baetica} in \textit{Britannia} province.
-Rather, it seems that olive oil production was distributed by non-specific production centres.
+In general, the dendrogram did not observe any production pattern that indicates a clear organisation between production centres and consumption centres.
+In other words, our results did not indicate the presence of a specific production centre from \textit{Baetica} in \textit{Britannia} province.
+Rather, it would imply that olive oil production was distributed by non-specific production centres.
 
-The similarity of stamps shows results less significant than \textit{Britannia} province as it can be seen in the dendrogram (Fig. \ref{germap}).
+In \textit{Germania}, the similarity of stamps shows results less significant than \textit{Britannia} province as it can be seen in the dendrogram (Fig. \ref{germap}).
 
@@
 	\centering
-\includegraphics[width=\linewidth]{figs/dendroger5.pdf}
+\includegraphics[angle=180, width=\linewidth]{figs/dendroger5.pdf}
 \caption{Dendrogram obtained by Morisita-Horn algorithm of different sites in \textit{Germania}.
@@
 
-The sites with more similarity in stamps were closely located in limes.
-In this case, most sites have different stamps which means that we did not detect any pattern.
-On the other hand, there seems to be a greater concentration of the seals in areas eminently militarised and close to the German limes.
-However, It is still no possibility to determine the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.
+A higher concentration of sites sharing similar stamps were found in areas eminently militarised and close to German \textit{limes}, even if most sites mostly showed different stamps.
+The interpretation of the results suggests that there is still no possibility to determine the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.
 
@@
 
-The analysis suggests that there is  a connection between the same stamp code and nearby amphora workshops, excluding certain exceptions.
+The analysis suggests a connection between the same stamp code and nearby amphora workshops, excluding certain exceptions.
 Consequently, the majority of stamps are located in different amphora workshops and only similar stamps between closer amphora workshops were found.
 Indeed, our results show that most similar stamps were detected in the same \textit{conventus} area.
-This results can indicate a grouping of stamps with similar codes in the same \textit{conventus}.
-However, these stamps tend to share the same area of production but there is not a general relation between groups of amphora workshops and connection areas.
-In other words, we do not identify groups of more than three workshops or less sharing the same amphoric stamps as the dendrogram showed.
+This results may indicate a grouping of stamps with similar codes in the same \textit{conventus}.
+Despite these stamps tend to share the same area of production, we do not identify groups of any more than three workshops sharing the same amphoric stamps as the dendrogram showed.
 In general, the majority of stamps were located in different amphora workshops.
@@
 It is more probably than closer workshops tend to share more traits than distant workshops.
-While the role of the river was significant for the distribution of amphorae in consumption places, river connection amongst workshops does not seem to show relevancy for the distribution of stamps.
+While the role of the river was significant for the distribution of amphorae, river connection amongst workshops does not seem to show relevancy for the distribution of stamps.
 Finally, the distribution of stamps could have shown some research bias.
@@
 
-It is worth to mention that centres with a similarity correlation were located close to the Atlantic coast around the North Sea and the Celtic Sea.
-This could indicate that the Atlantic route could have played an essential role in transporting olive oil to the area of \textit{Britannia}, since in the places where there is greater similarity they are found in different strategic points near the sea.
+It should be noted that centres with a similarity correlation were located close to the Atlantic coast around the North Sea and the Celtic Sea.
+This may indicate that the Atlantic route could have played an essential role in transporting olive oil to the area of \textit{Britannia}, since in the places where there is greater similarity they are found in different strategic points near the sea.
 
-This fact seems to concur with the latest research that suggest the idea of a more important Atlantic route that would connect the provinces militarised\citep{remesal_annona_1986,
+This fact seems to concur with the latest research that suggests the idea of a more important Atlantic route that would connect the militarised provinces \citep{remesal_annona_1986,
 remesal_provincial_2008,
@@
 
-It is worth to mention that it was not detected a robust model of organisation in both cases with regard to the distribution of stamps in consumption centres. This means
-that it is unknown whether certain production centres went to one province or another or, at the
-less, that it can be clearly reflected in the data with a greater similarity in the amphoric stamps.
+It is worth to mention that it was not detected a robust model of organisation in both cases with regard to the distribution of stamps in consumption centres.
+This evidence means that it is unknown whether some production centres went to one province or another or, at least, that they can be clearly reflected in the data with a greater similarity in the amphoric stamps.
 Thus, production centres could have distributed randomly olive oil both \textit{Britannia} and \textit{Germania}.
-We do not detect production centres dedicated to the distribution of olive oil in a certain province.
+Neither we do not detect production centres dedicated to the distribution of olive oil in a certain province.
 
-In general, we observed a distribution of the stamps to military stamps.
-By contrast, judging by the results obtained, there does not seem to be a specific pattern in terms of geographical distribution.
+%In general, we observed a concentration of the stamps in military centres.
+
+Judging by the results obtained, there does not seem to be a specific pattern in terms of geographical distribution.
 Nor it was detected consumption areas where stamps are specified from an amphora workshop.
 
-\section{Conclusion}
+\section{Concluding remarks}
 
-The results of the case study could be interpreted by several reasons.
-On the one hand, the use of these amphoric stamps could have been exclusively running by the owner or owners of the workshop to distinguish the amphora workshop.
-This hypothesis would explain the fact that we do not find similar stamps in different workshops; however, we found different amphora stamps in the same workshops that they would be barely difficult to assign different owners.
+The results of this work can be interpreted for several reasons.
+On the one hand, the use of amphoric stamps could have been exclusively running by the owner or owners of the workshop to distinguish the amphora workshop.
+This hypothesis would explain the fact that we do not find similar stamps in different workshops; however, different code stamps have been detected in the same workshops that they would be barely difficult to assign to different owners.
 
-On the other hand, the existence of different stamps in the same workshop but which are only repeated in very close places it would imply some kind of organisation from within workshop affecting the closest centres.
+On the other hand, the presence of different stamps in the same workshop could imply some kind of organisation from within the workshop affecting the closest centres \citep{juanmorostesis}.
+The marked amphorae could have allowed to potters organise the production with a batch stamping for the posterior commercialisation.
+
 This could be explained such as batch systematic organisation between potters \citep{juanmorostesis}. 
-This method allowed to potters organise the production with the batch stamping for the posterior commercialisation.
+This method allowed to potters organises the production with the batch stamping for the posterior commercialisation.
 
-Considering that Dressel 20 was not marked in most cases, some researchers point out that potters marked amphorae to prepare and distribute the commodity to be shipped \citep{berni_millet_epigrafianforica_2008}.
+Considering that Dressel 20 was not marked in most cases, some researchers point out that potters marked the amphorae to prepare and distribute the product in order to be shipped \citep{berni_millet_epigrafianforica_2008}.
 This method would be used as an identifier to count the number of amphorae of a branch \citep{juanmorostesis}.
@@
 This fact would explain wherefore why we detect different stamps in the same workshop.
-In any case, we do not have enough archaeological evidence that can validate the interpretations presented here and our results are certainly valid only with the context of our case study.
+In any case, we do not have enough archaeological evidence that can validate the interpretations presented here and our results are only certainly valid with the context of our case study.
 
-As a summary, this method presented here provides a potential tool to understand mechanisms of production based on the similarity of artefacts.
+As a summary, this method presented here provides a potential tool to understand the mechanisms of production based on the similarity of artefacts.
 This work has identified differences in the case of the amphoric production within the Roman Empire.
-Accordingly, the results have highlighted for the interpretation of the complex economic processes connected with the archaeological evidence.
+The combination of archaeological data and quantitative methods have finally highlighted for the interpretation of the complex economic processes connected with the archaeological evidence.
 

### 2020-05-29 Paper/jaccardbaetica.tex
finish third draft

@@
 
-We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in \textit{Baetica} province.
-The database was compiled by CEIPAC group\citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}).
-However, approximately 70 \% of stamps cannot be tested due to the fragmentation of the dataset.
-Consequently, any stamps with incomplete information were discarded and not integrated into our dataset.
-We finally filtered a total sample of 987 stamps composed of 130 different stamps from 81 workshops.
+We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in \textit{Baetica} province. For the analysis, we used a database compiled by CEIPAC group 
+\citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}).
+CEIPAC database contains over 50.000 epigraphy records found on different types of amphorae.
+From the database, we collected the site where stamps were found, the stamp code and the geographical coordinates.
+
+The stamps with incomplete information were discarded and not integrated into our dataset. We finally filtered a total sample of 987 stamps composed of 130 different stamps from 81 workshops.
+
+%Approximately 70 \% of stamps cannot be tested due to the fragmentation of the dataset. Consequently, any stamps with incomplete information were discarded and not integrated into our dataset.
+
+We also created a new column with the area of the stamps.
+These areas, known as \textit{conventus}, were administrative centres for territorial organisation in the Roman Empire \citep[58]{ozcariz_gil_administracion_2013}.
+Dressel 20 stamps were found in three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter Hispalis), \textit{Cordubensis} (currently C\'ordoba, hereafter Corduba) and \textit{Astigi} (currently Écija, Sevilla, hereafter Astigi) \citep{rodriguez_economioleicola_1977,
+chicdatos2001,berni_millet_epigrafianforica_2008} .
 
@@
 
-From the database, we collected the site where stamps were found and the stamp code.
-We also created a new column with the area of the stamps.
-These areas, known as \textit{conventus}, were administrative centres for territorial organisation in the Roman Empire \citep[58]{ozcariz_gil_administracion_2013}.
-Dressel 20 stamps were found in three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter Hispalis), \textit{Cordubensis} (currently C\'ordoba, hereafter Corduba) and \textit{Astigi} (currently Écija, Sevilla, hereafter Astigi) \citep{rodriguez_economioleicola_1977,
-chicdatos2001,berni_millet_epigrafianforica_2008} . It is important to mention that some workshops exhibited the same coordinates but they were differently catalogued as a different workshop. 
-\maria{seguir con esta mierda}
+It is important to mention that some workshops exhibited the same coordinates but they were catalogued as different workshops in Roman Studies (i.e.
+Tesorillo de Doña Mencia and Doña Mencia).
+The analysis took into account this particularity but some biases can occur due to the archaeological dataset.
+
 We calculated the frequency of distribution in amphora stamps with the purpose of implementing an exploratory analysis (EDA) to detect the distribution and the number of stamps per each centre.
@@
 
-We analysed a dataset of 2219 stamps from different centres in \textit{Britannia}. 
+We analysed a dataset of 4271 stamps  from different centres in \textit{Britannia} (2219 stamps) and \textit{Germania} (2052 stamps).
+Both \textit{Germania} and \textit{Britannia} were analysed as regions and not as Roman provinces.
+This means that we selected centres beyond the borders of the provinces.
+We analyse the two provinces (Ulterior and Citerior) close to \textit{limes} in \textit{Germania} while the analysis for Britannia was extended to \textit{Caledonia}.
+
+We used the same CEIPAC database as previous before to select amphora stamps, centres and coordinates.
+The database contained anomalies that were eliminated using the same criteria cited above.
+We selected the centres with more or equal than five stamps.
+
+%We analysed a dataset of 2219 stamps from different centres in \textit{Britannia}. 
 %CITAR (Callender, 1965; Carreras Monfort y Funari,1998; Ayllón-Martı́ et al., 2018). 
-The stamp database was also compiled by CEIPAC database.
-The database contained anomalies that were eliminated using the same criteria cited above.
-We finally selected the centres with more or equal than five stamps.
-As a result, we studied a total of 1765 stamps composed of 968 different stamps from 46 centres.
-The centres located in \textit{Britannia} can be seen in Fig.\ref{britannia}.
+In \textit{Britannia}, we studied a total sample of 1765 stamps composed of 968 different stamps from 46 centres. The centres located in \textit{Britannia} can be seen in Fig.\ref{britannia}.
 
@@
 
-We analysed a dataset of 2052 stamps placed in \textit{Germania}.
+%We analysed a dataset of 2052 stamps placed in \textit{Germania}.
 All the data was also compiled by CEIPAC database.
-We finally collected a total of 1621 stamps consisted of 850 different stamps from 46 sites (see Fig.
-\ref{germania}).
 As previously before, the centres were selected with more or equal than five stamps.
+
+In \textit{Germania}, we finally collected a total of 1621 stamps consisted of 850 different stamps from 46 sites (see Fig. \ref{germania}).
 

### 2020-06-08 Paper/jaccardbaetica.tex
add correction finish draft

@@
 Mostly, workshops show a homogeneity on the frequency of stamps except for both La Catria and Arva that exhibit a big amount of stamps with 29 different code stamps.
-According to previous studies, those workshops became in most important centres of amphora production although they could have more intensely prospected than others.
-\citep{arva_1997}.
+According to previous studies, those workshops became in most important centres of amphora production although they could have more intensely prospected than others \citep{arva_1997}.
 
@@
 
-We used the same CEIPAC database as previous before to select amphora stamps, centres and coordinates.
+We used the same CEIPAC database as previous before to select amphora stamps, centres and geographical coordinates.
 The database contained anomalies that were eliminated using the same criteria cited above.

### 2020-06-23 Paper/jaccardbaetica.tex
updated maps

@@
 
-Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}).
+Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}) (see Figure~\ref{general}.
+
+\begin{figure}[htp]
+	\centering
+\includegraphics[width=\linewidth]{figs/general_map}
+\caption{Overview of the sites analyzed in this paper.
+The color defines the three different regionsu under study (Baetica as producer; Germania and Britannia as consumers) while the size of each dot highlights the number of Dressel-20 amphoric stamps found on each of these sitese}
+
+\label{general}
+\end{figure}
+
 In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces.
@@
 
-The strategic location of the workshops close to the river allowed the transport of olive oil through river connectivity to the routes of the Mediterranean and the Atlantic \citep{garcia_vargas_enrique_formal_2010}.
+The strategic location of the workshops at the riverbank allowed the transport of olive oil through riverine transport towards the maritime routes connecting both the Mediterranean and the Atlantic \citep{garcia_vargas_enrique_formal_2010}.
 
@@
 	\centering
-\includegraphics[width=\linewidth]{figs/workshop1}
-\caption{Distribution of the workshops (dots) in Baetica province}
+\includegraphics[width=\linewidth]{figs/baetica}
+\caption{Distribution of stamps found in Dressel-20 workshops along the Guadalquivir riverbank. The size of each point depends on the number of stamps found on each site}
+
 \label{workshop}
@@
 
-The frequency of stamps per workshop can be seen in Fig. \ref{stamps}. Most 
+The frequency of stamps per workshop can be seen in Fig.~\ref{stamps}. Most 
 workshops contained only one stamp whereas one workshop (La Catria) concentrated a high frequency of amphoric stamps with a total number of 228 stamps.
@@
 	\centering
-\includegraphics[width=\linewidth]{figs/britmap2.pdf}
-\caption{Distribution of centres in \textit{Britannia}}
+\includegraphics[width=\linewidth]{figs/britannia}
+\caption{Sites in Britannia where Dressel-20 amphoric stamps have been found.
+Each dot is a site while its size shows the sample size of stamps found in the site.
+A majority of stamps have been found in garrisons related to Hadrian's and Antonine walls}
 \label{britannia}
@@
 	\centering
-\includegraphics[width=\linewidth]{figs/germania5_2.pdf}
-\caption{Distribution of the sites in \textit{Germania}. Most sites were located in German \textit{limes}}
+\includegraphics[width=\linewidth]{figs/germania}
+\caption{Sites in Germania where Dressel-20 amphoric stamps have been found.
+Each dot is a site while its size shows the sample size of stamps found in the site.
+Most sites are located along the German \textit{limes}}
 \label{germania}

### 2020-06-30 Paper/jaccardbaetica.tex
comentarios y edits

@@
 
+\xavi{reescribir abstract, demasiado general y entonces demasiado específico}
 The aim of this study is to explore economics dynamics in the production and distribution of olive oil trade. 
@@
 
-The role of the Mediterranean sea for the trade had an enormous impact during the Roman Empire.
-The spread of thousands of different goods along maritime shipping allowed progressive exploitation of communication networks and a major interaction between communities \citep{rodriguez_baetican_1998,
-temin_market_2001,
-bevan_mediterranean_2014}. The exchange of different products between regions can be seen in both archaeological and written records as a frequent indicator of Roman trade.
+The intensification of long-range trade was one of the most important traits of the economy developed duriong the Roman Empire.
+The development of an extensive road network increased the connectivity between inland communities, but most shipping continued being based on maritime routes particularly in the Mediterranean basin~\citep{temin_market_2001,bevan_mediterranean_2014}.
+The intensity of this trade between distant regions can be observed both in archaeological and written evidence~\citep{rodriguez_baetican_1998}.
 
@@
 
-Given this fact, an important mechanism of production control under the Roman government was created spreading by different areas attending to the richness of different places.
-Roman provinces such as \textit{Baetica} (currently Andalusia, south Spain) became important production centres of olive oil during the Roman Empire.
-Olive oil was considered the liquid gold since it was used in different aspect of daily life for cooking, hygiene or lighting  \citep{mattingly_d.j._oil_1988}.
-As a consequence of high demand, this product was distributed in a large number of amphorae across the provinces, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.
+The Empire developed a series of structures to support and organize this long-range trade and specialized entire regions to massively produce specific goods.
+An example of this process was the province of \textit{Baetica} (currently Andalusia, southern Spain) which became an important olive oil production centre during\xavi{período?}.
+Olive oil was an essential good for Romans because it was used in almost every aspect of their daily life such as cooking, hygiene or lighting~\citep{mattingly_d.j._oil_1988}.
+This high demand required a huge increase on the production which was distributed through amphorae shipped via maritime routes to all the provinces and particularly to Italy and the hundreds of garrisons that the Roman army deployed along its borders~\citep{blazquez_exportacion_1980}.
 
-The understanding of the olive oil production processes has been widely discussed over many decades\citep{rodriguez_economioleicola_1977, Chic_hispania_1997,millet_anforas_1998}.
-Despite this topic is still subject to debate: what patterns were followed for the distribution of olive oil to the different provinces?
-Did each province follow a different pattern for its distribution? Nor the lack of written records has made possible to detect any indication of patterns in the olive oil market.
+The structure and processes associated to this massive olive oil production has been extensively discussed over the last decates\citep{rodriguez_economioleicola_1977, Chic_hispania_1997,millet_anforas_1998}.
+Our knowledge of this economic activity has benefitted from new findings and data sources, but despite all these advances several questions remain open: \xavi{no entiendo esta pregunta} what patterns were followed for the distribution of olive oil to the different provinces?
+\xavi{no es esta pregunta idéntica a la anterior?} Did each province follow a different pattern for its distribution?
+\xavi{esta frase no tiene sentido aquí; o la pones antes o en otro párrafo, pero es raro acabar un set de 2 preguntas con una frase que no sea otra pregunta} Nor the lack of written records has made possible to detect any indication of patterns in the olive oil market.
 
-Advances in the research of the Roman studies have currently led an environment with more diverse commercial dynamics \citep{duncan1982economy,
+\xavi{las preguntas no van sobre patrones; los patrones son las trazas en la evidencia que te ayudan a responder a las preguntas}
+
+Advances in the research of the Roman studies have currently led an environment with more diverse interpretations on commercial dynamics \citep{duncan1982economy,
 temin_economy_2006,
-quantifyingwilson2009}.
-The application of different quantitative approaches combined with the archaeological evidence has allowed us partly to deal with the complexity of the Roman production \citep{brughmans_roman_2016,
+quantifyingwilson2009}\xavi{more diverse que qué?}.
+The application of different quantitative approaches has allowed us to improve our interpretation of the growing amount of archaeological evidence and reveal the complexity of the Roman production \xavi{en qué sentido?
+qué quieres decir aquí con complexity?
+Si no especificas queda muy genérico (rollo "el mundo es muy complejo")}.\citep{brughmans_roman_2016,
 orengo_seeds_2016,bayesian_2018,
@@
 
+\xavi{me parece que aquí falta más chicha, porque saltas de una cosa super general (complejidad, economía, grandes dinámicas) a unos casos muy específicos.
+Deberías aquí explicar algo más sobre tu estudio y lo que quieres entender, por ejemplo qué preguntas específicas quieres responder?
+Como se enlazan con lo que has explicado hasta ahora?
+Cómo es que ahora saltas a casos de estudio regionales?
+Las preguntas específicas de los 3 casos de estudio deberían entrelazarse en una pregunta general del paper, que esté vinculada a lo que has explicado hasta ahora}
+
 This paper aims to study the olive oil market connection between provinces by calculating the similarity of stamps.
-Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}.
+Specifically, our work pretends to detect microeconomic processes\xavi{no me gusta lo de microeconomic, porque es un campo específico de la economía que nada tiene que ver con analizar un limes entero.
+Quizás regional?} focused on a commercial product from a specific province \citep{isaksen_network_2006}.
 We want to understand the pattern of olive oil production linked to amphora workshops and amphoric stamps used to mark them.
@@
 
-\section{The Amphoric production in Baetica province}
+\section{The Amphoric production of Baetica}
 
-The high demand of Roman provinces stimulated by the good condition of the \textit{Baetica} lands allowed to develop a massive infrastructure of olive oil production.
-The production and distribution of olive oil were growing exponentially for almost three centuries \citep{remesal_concierto}.
-As a result, hundreds of amphora workshops were raised to support this high demand of the Roman Empire.
-The amphora workshops were situated in different locations in \textit{Baetica} province, along the rivers Guadalquivir and Genil (see Fig.\ref{workshop}).
+The landscape of the \textit{Baetican} province was one of the best regions to face the increase in demand of olive oil across the Roman provinces.
+For this reason the area saw an increase in productivity as a massive infrastructure of olive oil production was gradually deplyoyed.
+The production and distribution of olive oil grew exponentially\xavi{seguro que exponencial?} for almost three centuries\xavi{qué centuries?} \citep{remesal_concierto}.
+Hundreds of presses and amphora-making workshops were build near large extensions covered by olive trees.
+The workshops where amphorae were made and filled with olive oil were places along the rivers Guadalquivir and Genil riverbanks (see Fig.\ref{workshop}).
 
-The strategic location of the workshops at the riverbank allowed the transport of olive oil through riverine transport towards the maritime routes connecting both the Mediterranean and the Atlantic \citep{garcia_vargas_enrique_formal_2010}.
+This strategic location allowed the transport of olive oil through riverine shipping towards the maritime routes that connected Baetica with Mediterranean and Atlantic trade routes towards the rest of the Empire\citep{garcia_vargas_enrique_formal_2010}.
 
@@
 
-The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}. 
-This fact is shown by the archaeological evidence that displays a highly specialised production with a long activity with apparently few changes \citep{remesal_anforas_2004}.
+The chronology of the workshops is widely diverse from the first to the third centuries AD\xavi{qué quiere decir que la cronología es diversa?} \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
+This fact is shown by the archaeological evidence that displays a highly specialised production with a long activity with apparently few changes\xavi{deberías desarrollar esto un poco más; qué tipo de evidencia demuestra esto?
+pocos cambios en qué?
+los talleres?
+las tipologías anfóricas?} \citep{remesal_anforas_2004}.
 
-The majority of olive oil amphorae produced in this province belong to Dressel 20 type\citep{dressel_ricerche_1878,
-millet_anforas_1998}. Dressel 20 amphorae were shipped and distributed from \textit{Baetica} throughout western Europe.
+A large majority of this Baetican amphorae production is composed by Dressel 20 types\citep{dressel_ricerche_1878,millet_anforas_1998}.
+Dressel 20 amphorae were shipped and distributed from \textit{Baetica} throughout western Europe\xavi{esto ya lo has mencionado; quizás podrías explicar en 2 líneas la capacidad y forma de la dressel 20}.
 
-Dressel 20 contains much information still unknown.
-A high percentage of  Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti}.
-Indeed, most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
+Dressel 20 amphorae display a lot of evidence, but some of it is still hard to interpret.
+A large percentage of Dressel 20 were marked with stamps, while they could also be inked with \textit{tituli picti} or incised with \textit{graffiti}.
+The interpretation for most of these inscriptions is still open due to the fragmentation of the material and the small sample size of well-preserved elements~\citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
 
@@
 
-Dressel 20 was the amphora most stamped during the Roman Empire \citep[18]{millet_anforas_1998}. The usage period is associated with a large amphora production for almost three centuries.
+Dressel 20 was the amphora most stamped during the Roman Empire \citep[18]{millet_anforas_1998}.
+The usage period is associated with a large amphora production for almost three centuries\xavi{esto está repetido no?}.
 
-Most stamps display a large activity of production that it can be difficult for specifying an accurate chronology in some cases.
-They have been frequently dated with the consular dating by studying \textit{tituli picti} found in \textit{Monte Testaccio} \citep{Testaccio1, berni_millet_epigrafianforica_2008}.
-However, the chronology of stamps could be also biased in other sites taking into account that amphorae were deposited after being produced and consumed.
-
-The fascination aroused by stamps has caused it to be one of the inscriptions with the most publications about its origin and meaning.
-Its endure in time become an essential element that was used to mark amphorae during almost three centuries \citep{remesal_sellar_2016} (Fig.\ref{amphora}).
+The large sample size of recovered Dressel 20 stamps is distributed across all the western provinces of the Roman Empire.
+For this reason there are several publications of this type of evidence discussing its origin, its meaning, its long period of use and its spatial distribution~\citep{remesal_sellar_2016} (Fig.\ref{amphora}).\xavi{si dices que es muy popular debería haber aquí más citas aparte de Remesal}.
 
@@
 
-Frequently, they were mainly marked in handles, but seldom in rims and body \citep{millet_anforas_1998}. 
-The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion.
-Stamps are mostly formed by a code of three letters and they can appear in abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
+Most stamp codes display a large activity of production that can be difficult to date using an accurate chronology.
+They have been frequently dated with the consular dating by studying \textit{tituli picti} found in \textit{Monte Testaccio} \citep{Testaccio1, berni_millet_epigrafianforica_2008}.
+However, the chronology of stamps could be also biased in other sites taking into account that amphorae were deposited after being produced and consumed\xavi{esto necesita un poco de desarrollo.
+quieres decir que el código se encuentra en contextos de cronologías diferentes no?
+Y lo del Testaccio, pos eso, que hay pocos y todos de un yacimioento, lo que añade un bias}.
 
-Despite researchers identify stamps as identity marks, there is not a consensus about the meaning of the stamps \citep{rodriguez_baetican_1998}.
-Alike a high percentage of Dressel 20 containers were not stamped being difficult to determinate.
+These stamps are typically found on handles but they were also imprinted on rims and the amphora body~\citep{millet_anforas_1998}.
+The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion\xavi{qué information?
+El código de letras?}.
+Stamps often displayed a code of three letters and they can appear in abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.\xavi{que quieres decir con abbreviated o complete?
+Como abrevias un código de 3 letras?}
 
-They could be identified based on three premises: content (olive oil), context (amphora workshop) and subject (individuals involved).
+Scholars agree that stamps are some type of identification mark, but there is no consensus on the meaning of these marks\citep{rodriguez_baetican_1998}.
+Additionally, a large percentage of Dressel 20 containers were not stamped thus complicating the interpretation~\xavi{esta última frase aquí es un poco random no?}.
+The stamp codes are interpreted based on three main ideas: content (olive oil), context (amphora workshop) and subject (individuals involved on the production).
 On the one hand, it seems that stamps could have been identified as the landowner of the olive groves \citep{rodriguez_economioleicola_1977}.
-On the other, they could belong to the owner of the making amphorae workshops or even a group of amphora workers \citep{berni_millet_epigrafianforica_2008}.
-In any case, the use of these stamps became in a good proxy to define somehow the system of working in the workshops.
+On the other hand, they could also belong to the owner of the workshops were the amphorae was made or even to a group of amphora workers~\citep{berni_millet_epigrafianforica_2008}.
+In any case, the use of these stamps is a good proxy to explore the system of olive oil production and trade developed in Baetica.
 
-Nevertheless, some challenges remain under discussion such as how this production was organised and whether it is possible to distinguish production patterns in the olive oil trade.
+Nevertheless, some challenges\xavi{acabas de decir que tampoco se sabe qué es el sello, así que esta frase queda rara} remain under discussion such as how this production was organised and whether it is possible to distinguish production patterns in the olive oil trade.
 Our questions will be focused on the distribution of amphoric stamps.
@@
 
-Neither the use of written records have allowed providing enough information that can explain the economic role of \textit{Baetica} province in the Roman organisation.
+Neither the use of written records have allowed providing enough information that can explain the economic role of \textit{Baetica} province in the Roman organisation.\xavi{esta frase esta fuera de contexto.
+Si la pones entonces deberías expandir la idea de frase a párrafo, y poner referencias}.
 
@@
 
-The conquest of new provinces allowed the Roman Empire the arrival of resources through Mediterranean and Atlantic routes.
-This lead to a progressive change in the economic and social structure creating new trade networks for supplying Roman conquest.
-As a result, Augusto's administration created the figure of the \textit{praefectura annonae} for the supply of wheat.
-The role of the \textit{praefectura annonae} was mainly to provide though \textit{frumentationes} a quantity per month of wheat to Roman citizens \citep{remesal_annona_1986,remesal_concierto}.
+\xavi{este título es muy raro, porque 1. que son casos de estudio ya lo sabes y 2. case study singular, consumption centres plural}
 
-\textit{Annona} could have also included other goods for the supply of Roman legions such as olive oil according for some hypotheses \citep{remesal_annona_1986,remesal_annona_1990}
+The creation of new provinces allowed the Roman Empire the arrival of resources through Mediterranean and Atlantic routes\xavi{no entiendo esto...allowed o required?}.
+This led to a gradual change in the economic and social structureas the trade networks that supplied the Roman army became more complex.
+Augustus's administration created the figure of the \textit{praefectura annonae} to organize wheat supply.
+The role of the \textit{praefectura annonae} was mainly focused on providing through \textit{frumentationes} a fixed monthly amount of wheat to each Roman citizen~\citep{remesal_annona_1986,remesal_concierto}\xavi{pero estabas hablando del ejército no?
+Has hecho un salto aquí a los ciudadanos}.
 
-The importance of olive oil production could be explained by the fact that we detected a massive amphora distribution in different Roman provinces.
-This amphora has been commonly associated with the  transportation of Baetican olive oil for supplying military camps and civil settlements \citep{berni_millet_epigrafianforica_2008}.
+Some authors argue elsewhere that this same \textit{Annona} could also have organized the supply of additional goods such as olive oil to the Roman legions \citep{remesal_annona_1986,remesal_annona_1990}\xavi{pero no hay evidencia?
+Las fuentes escritas no lo mencionan?}.
 
-The military consumption of olive oil in Roman provinces may be related to two aspects: a) cultural consumption whose product is consumed by cultural reasons such as to identity or habit and b) economical reasons where olive oil distribution is based on the transport costs, unlike other products \citep[69-70]{carreras_britannia_1998}.
+The importance of this olive oil trade is revealed by the massive amount of Dressel 20 amphorae found across the Roman provinces.
+This amphora has been commonly associated with the transportation of Baetican olive oil for supplying military camps and civil settlements \citep{berni_millet_epigrafianforica_2008}.\xavi{me parece repetitivo, porque unos párrafos antes ya hablabas de lo mismo.
+Quizás unificar texto aquí o allá?}
 
-Olive oil supply had a special impact in militarised provinces such as \textit{Britannia} and \textit{Germania}.
-The concentration of amphorae in eminently militarised areas suggests the existence of a broad supply mechanism, being accurate to imagine a possible organisation in charge of its distribution from military camps to civil settlements\citep{remesal_annona_1986, carreras_britannia_1998}.
+The military consumption of olive oil in Roman provinces may be related to two aspects: a) cultural consumption whose product is consumed by cultural reasons such as to identity or habit and b) economical reasons where olive oil distribution is based on the transport costs, unlike other products \citep[69-70]{carreras_britannia_1998}\xavi{no pillo esta frase}..
+
+Olive oil supply seems to be particularly intense in militarised provinces that functioned as the borders of the Empire such as \textit{Britannia} and \textit{Germania}.
+The high percentage of amphorae\xavi{amphorae o Dressel 20?} found in sites nearby military garrisons suggests the existence of diverse supply dynamics as there were trade processes flowing good from military sites to civil settlements\citep{remesal_annona_1986, carreras_britannia_1998}\xavi{lo que quieres decir es que tienes muchas ánforas en sitios civiles cerca de militares?
+no queda claro}.
+
+\xavi{faltaría un enlace a la siguiente sección, queda muy abrupto}
 
@@
 
-The consumption of olive oil in \textit{Britannia} was residual until starting the Roman conquest \citep{funari_corpus_1996,
-carreras_abastecimiento_2003}.
-It is well known that this product was not frequently consumed by indigenous population \citep[161]{monfort_britanniaen_1998}.
-The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[1]{carreras_britannia_1998}.
-Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions.
-This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman army.
+The evidence for olive oil consumption in \textit{Britannia} is scarce before the Roman conquest \citep{funari_corpus_1996,carreras_abastecimiento_2003}.
+Indigenous\xavi{no sé si indigenous es la mejor palabra...local?} population did not consume this product and it was not produced on the region as the landscape and climate of the British isles was not suitable for olive oil trees\citep[161]{monfort_britanniaen_1998}.
+Based on the archaeological record this absence of olive oil changed with the arrival of a large amount of Dressel 20 amphorae to \textit{Britannia} \citep[1]{carreras_britannia_1998}\xavi{en qué siglo?}.
+The demand of olive oil after the Roman conquest required of a new trade route due to the lack of local olive oil production; \textit{Baetica} was chosen as its main supply.
 
-At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}.
-This fact will have a spatial incidence in sites close at Hadrian Wall.
-Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic points \citep{carreras_atlantic_2012}.
+At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}\xavi{repetitivo?
+fusionar párrafos}.
+This fact will have a particular intensity within sites close to the Hadrian Wall's garrisons.
+Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic locations~\citep{carreras_atlantic_2012}.
 The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges.
-Thus, the network was mainly focused on the support of soldiers during military campaigns.
+Thus, the network was mainly focused on the support of soldiers during military campaigns\xavi{estas 2 frases repiten cosas ya explicadas}.
 
-The presence of Dressel 20 stamps in military camps in \textit{Britannia} have been widely studied in Roman archaeology \citep{carreras_britannia_1998}.
-This fact also indicates a possible governmental organisation allocate to reinforce and supply of olive oil in military camps, such as \textit{Germania} (Remesal Rodrıguez, 1986).
-However, it is unknown the type of economical system when managing or control the redistribution of olive oil in different sites in \textit{Britannia} but it seems that it could have been in the main cities \citep[45]{funari_economic_2005}.
+The presence of Dressel 20 stamps in military camps in \textit{Britannia} has been widely studied in Roman archaeology \citep{carreras_britannia_1998}.\xavi{si dices widely studied no puedes referencias solo 1 autor}.
+This fact\xavi{que fact?
+Que los sellos de dressel 20 han sido muy estudiados?
+No tengas miedo a repetir trozos rollo "This intense consumption of olive oil also indicatas blablabla...} also indicates a possible provincial structure designed to organize the supply of olive oil to military camps,m such as \textit{Germania} (Remesal Rodrıguez, 1986)\xavi{por qué such as Germania?
+Aún no has explicado qué hacen en Germania}.
+There are no written records explaining how this redistribution of essential goods was organised in \textit{Britannia}, but the archaeological evidence suggests that cities may have been the central nodes of this local trade network~\citep[45]{funari_economic_2005}.
 
-Finally, the increase in olive oil exports would experience a progressive slowdown from the third century A.D., coinciding with the change in market strategy in the Empire.
-As a result, a progressive decrease of Dressel 20 is documented being gradually replaced for Dressel 23 amphorae \citep{rodriguez1991aceite,millet_anforas_1998}.
+The consumption of olive oil would experience a progressive slowdown from the third century A.D.
+onwards.
+This date matches a change in market strategy in the Empire.
+This a gradual decrease on \textit{Baetican} olive oil exports can be observed as a decrease in the amount of Dressel 20 found in excavated sites as this typology is replaced by Dressel 23 amphorae \citep{rodriguez1991aceite,millet_anforas_1998}.\xavi{pero entonces la dressel 23 de donde viene?
+Cuidao que estás diciendo que se acaban las dressel 20, pero esto implica que hay menos aceite de oliva o que viene de otras partes o que va en recipientes distintos a la dressel 20?}
+
+\xavi{HE HECHO HASTA AQUÍ}
 

### 2020-07-01 Paper/jaccardbaetica.tex
más comentarios

@@
 
-\xavi{HE HECHO HASTA AQUÍ}
-
 \subsection{Germania}
 
-The Roman conquest in \textit{Germania} dates from the end of the 1st century BC, during Caesar campaign and subsequently continued by Augusto.
-The advance of the Roman legions toward \textit{Gallia} opened a path through the Atlantic Ocean to \textit{Germania} \citep{remesal_annona_1986,
-remesal_baetica_2002}.
-It is thought that this route would have been initiated previously through terrestrial transport, although recent studies question this position giving an important role to the Atlantic route for both the conquests of \textit{Germania} and those of \textit{Britannia} \citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
+Julius Caesar's campaigns in \textit{Gallia} during 1st BC opened the door to the invasion of \textit{Germania} during Augustus rule~\citep{remesal_annona_1986,remesal_baetica_2002}\xavi{no hay un resumen de la conquista similar para Britannia, queda raro}.
+Previous studies suggested that the supply to the German limes was mainly based on riverine transport, but recent works suggest that the maritime route through the Atlantic Ocean could have been more important than expected (both for Germania and Britannia).~\citep{remesal_germn_2010,rubio-campillo_ecology_2018}.\xavi{lo del océano atlántico está un poco descontextualizado; no sería mejor ponerlo al final de esta sección?}
 
-The Roman studies about the presence of Dressel 20 amphorae in \textit{Germania} has not had the same repercussion as the rest of the province.
+Roman studies exploring the presence of Dressel 20 amphorae in \textit{Germania} has not had the same impact as the rest of the province\xavi{no entiendo esta frase.
+Que qué provincia?
+Germania son 2 provincias no?}.
 This could be explained due to the lack of archaeological sources \citep{horacio2010llegada}.
+As a consequence, it is still unknown if trade agents participated on the distribution of Baetican olive oil in \textit{Germania} or the supply was exclusively focused on the Roman army garrisons\citep[156]{remesal_germn_2010}\xavi{quieres decir que aquí sólo encontramos dressel 20 en yacimientos militares porque no se han excavado muchos civiles?}.
 
-As a consequence, we do not the role that commerce agents in the participation of the distribution and exportation of olive oil in \textit{Germania} or, on the contrary, the Roman army was the responsible to supply the product in the own province \citep[156]{remesal_germn_2010}.
+\textit{Germania} presents a similar introduction pattern of olive oil than the one discussed for \textit{Britannia}.
+There is no archaeological evidence for olive oil consumption before the arrival of the Roman army to the region while a majority of recovered Dressel 20 amphorae are located in the military sites that formed the German \textit{limes}\xavi{cita?}.
 
-As the case of \textit{Britannia}, the consumption of olive oil in \textit{Germania} was not detected before the conquest.
-This is evidenced by the presence of Dressel 20 amphora in different sites close to \textit{limes}.
-
-The presence of the Roman army encouraged the exchange in the province showed by the arrival of this product both civil settlement and military sites with a mayor concentration at German \textit{limes}.
-It seems that some Baetican centres would be assigned to the support of olive oil.
-However, this hypothesis can not be proven by the lack of archaeological sources \citep[125]{remesal_concierto}.
+The presence of the Roman army encouraged the exchange in the province as shown by the arrival of this product both civil settlement and military sites with a mayor concentration at the German \textit{limes}\xavi{pero justo antes decías que no había en civiles no?}.
+It seems that some Baetican centres would be assigned to the support of olive oil\xavi{the support of olive oil no se entiende.
+Además, falta alguna cita y desarrollar esta idea}.
+However, this hypothesis is hard to assess given the current lack of archaeological evidence \xavi{related to BLABLABLA.
+Habría que detallar un poco porque hay un montón de evidencia arqueológica del limes, así que qué es exactamente lo que falta?}\citep[125]{remesal_concierto}.
 
@@
 The CEIPAC dataset contains over 50.000 of epigraphy records found in amphorae, mostly from \textit{Monte Testaccio}.
-
-Therefore, this study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps.
+This study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps.
 A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps.
 Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potential spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
+\xavi{estos 2 párrafos explican objetivos, que deberían ir en la intro y no aquí, donde deberías empezar a hablar directamente de la BBDD}
 
@@
 
-We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in \textit{Baetica} province. For the analysis, we used a database compiled by CEIPAC group 
-\citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}).
-CEIPAC database contains over 50.000 epigraphy records found on different types of amphorae.
-From the database, we collected the site where stamps were found, the stamp code and the geographical coordinates.
+To study the stamps found in the Baetican workshops we used the data collected in the CEIPAC database of amphoric epigraphy~\citep{remesal_centro_2015} (see the database here \url{http://romanopendata.eu}).
+This database contains over 50.000 epigraphy records found on different types of amphorae\xavi{podrías mover esta parte a los 2 párrafos anteriores y explicar la bbdd brevemente antes de empezar a hablar de la Bética (porque es info que también aplica a los otros 2 casos de estudio no?}.
 
-The stamps with incomplete information were discarded and not integrated into our dataset. We finally filtered a total sample of 987 stamps composed of 130 different stamps from 81 workshops.
+There databased allowed us to retrieve 3798 stamps found in Dressel 20 amphorae recovered from workshops within \textit{Baetica} province.
+Each of these stamps ha detailed records on the site where it was found, the inscribed code as well as its spatial coordinates.
+The stamps with incomplete information were discarded, thus finishing with a sample of 987 stamps from 81 sites and displaying 130 different code.
 
@@
 
-We also created a new column with the area of the stamps.
-These areas, known as \textit{conventus}, were administrative centres for territorial organisation in the Roman Empire \citep[58]{ozcariz_gil_administracion_2013}.
-Dressel 20 stamps were found in three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter Hispalis), \textit{Cordubensis} (currently C\'ordoba, hereafter Corduba) and \textit{Astigi} (currently Écija, Sevilla, hereafter Astigi) \citep{rodriguez_economioleicola_1977,
-chicdatos2001,berni_millet_epigrafianforica_2008} .
+Our dataset also contained a new categorical variable defining the \textit{conventus} of each site.
+The \textit{conventus} were administrative centres for territorial organisation in the Roman Empire \citep[58]{ozcariz_gil_administracion_2013}\xavi{estaría bien ampliar un poco esto; qué organizaban exactamente?}.
+The production area for Dressel 20 amphorae extended across three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter Hispalis), \textit{Cordubensis} (currently C\'ordoba, hereafter Corduba) and \textit{Astigi} (currently Écija, Sevilla, hereafter Astigi) \citep{rodriguez_economioleicola_1977,chicdatos2001,berni_millet_epigrafianforica_2008} .
 
@@
 Tesorillo de Doña Mencia and Doña Mencia).
-The analysis took into account this particularity but some biases can occur due to the archaeological dataset.
+The analysis took into account this particularity but some biases can occur due to the archaeological dataset\xavi{No se entiende; deberías explicarlo un poco mejor y citar algo si dices eso.
+Piensa que la mayoría de lectores no sabrán quién es Doña Mencia}.
 
-We calculated the frequency of distribution in amphora stamps with the purpose of implementing an exploratory analysis (EDA) to detect the distribution and the number of stamps per each centre.
-
-The frequency of stamps per workshop can be seen in Fig.~\ref{stamps}. Most 
-workshops contained only one stamp whereas one workshop (La Catria) concentrated a high frequency of amphoric stamps with a total number of 228 stamps.
-The type of distribution is also frequent in amphora production where we observed a self-organised complex system pattern with a major concentration of the number of stamps in a few workshops \citep{bayesian_2018,coto-sarmiento_identifying_2018}.
+The first step was to compute the frequency distribution of stamp codes and analyse it using Exploratory Data Analysis (EDA).
+This would allow us to study the distribution of the stamps across centres as can be seen in Figure~\ref{stamps}.
+Most workshops contained only one stamp~\xavi{el gráfico es dificil de entender...yo leo que hay unos 20 workshops con 1 solo stamp, por lo que entiendo que no es cierto que la mayoría solo tengan 1 sello (hay como 50 workshops con entre 2 y 30 sellos no?)} One workshop (La Catria) concentrated a large percentage of the amphoric stamps with a total number of 228 stamps.
+The type of distribution is also frequent in amphora production where we observed a self-organised complex system pattern with a major concentration of the number of stamps in a few workshops \citep{bayesian_2018,coto-sarmiento_identifying_2018}\xavi{no lo veo; recuerda que para el rollo bayesiano tenías una línea recta al hacedr el log de los 2 ejes, no solo 1, y este plot no está construído así no?}.
 
@@
 X axis is represented by the number of stamps and Y axis is the frequency of workshop.
-The distribution is widely diverse with most workshops have only one stamp}
+The distribution is widely diverse with most workshops having only one stamp}
 \label{stamps}
@@
 
-The distribution of amphora stamps in different \textit{conventus} can be seen in Fig.
-\ref{frequency}.
-The majority of stamps are concentrated in \textit{Hispalis} with 574 stamps while \textit{Corduba} and \textit{Astigi} with 267 and 146 stamps, respectively.
-Mostly, workshops show a homogeneity on the frequency of stamps except for both La Catria and Arva that exhibit a big amount of stamps with 29 different code stamps.
-According to previous studies, those workshops became in most important centres of amphora production although they could have more intensely prospected than others \citep{arva_1997}.
+The distribution of amphora stamps for each \textit{conventus} can be seen in Figure~\ref{frequency}.
+The majority of stamps are concentrated in \textit{Hispalis} (574 stamps) while \textit{Corduba} and \textit{Astigi} have roughly half this sample size (267 and 146 stamps).
+The workshops of the three conventus show a similar pattern on the stamp frequency distribution with the exceptions of 2 large workshops in \textit{Hispalis} (La Catria and Arva).
+On these two workshops a comparatively large amount of stamps were found (29 different code stamps on each of them)\xavi{una cosa, aquí has pasado de sellos a códigos, y los valores son distintos que en el gráfico anterior.
+Deberías clarificarlo antes de empezar a hablar de la figura}.
+According to previous studies those workshops became the most important centres of amphora production in the region\xavi{cita de estos estudios}.
+It is worth mentioning that a majority of these stamps were collected during ield surveys with no excavation, so this difference in sample size could be biased due to differences in intensity across the different sites~\citep{arva_1997}.
 
@@
 Quizás juntar esto con lo de los casos de estudio, o quizás hablar de los 2 casos de estudio en la intro, y dividir la sección anterior en 2 (lo más general a la intro, lo más concreto aquí)}
+The analysis of stamps in consumption areas also used the CEIPAC database and filtered in the same way than the previous dataset.
+We selected the centres with more or equal than five stamps\xavi{pero esto no se hizo en el de Baetica no?
+Por qué?}.
 
-We analysed a dataset of 4271 stamps  from different centres in \textit{Britannia} (2219 stamps) and \textit{Germania} (2052 stamps).
-Both \textit{Germania} and \textit{Britannia} were analysed as regions and not as Roman provinces.
-This means that we selected centres beyond the borders of the provinces.
-We analyse the two provinces (Ulterior and Citerior) close to \textit{limes} in \textit{Germania} while the analysis for Britannia was extended to \textit{Caledonia}.
-
-We used the same CEIPAC database as previous before to select amphora stamps, centres and geographical coordinates.
-The database contained anomalies that were eliminated using the same criteria cited above.
-We selected the centres with more or equal than five stamps.
+The output was a dataset of 4271 stamps found in sites belonging to \textit{Britannia} (2219 stamps) and \textit{Germania} (2052 stamps).
+Both \textit{Germania} and \textit{Britannia} were analysed as borders and not as Roman provinces.
+This means that we included in the sample some centres that are actually located beyond the administrative boundaries but were considered part of the same border.
+Specifically, for the German \textit{liomes} we included sites from the two provinces (Ulterior and Citerior\xavi{no es superior e inferior?}) while the analysis of Britannia extended to \textit{Caledonia}.
 
@@
 %CITAR (Callender, 1965; Carreras Monfort y Funari,1998; Ayllón-Martı́ et al., 2018). 
-In \textit{Britannia}, we studied a total sample of 1765 stamps composed of 968 different stamps from 46 centres. The centres located in \textit{Britannia} can be seen in Fig.\ref{britannia}.
+In \textit{Britannia}, we studied a total sample of 1765 stamps from 46 sites displaying 968 unique codes\xavi{pero esto no concuerda con las 2219 de antes...es por descartar los inciertos?
+Hay que dejarlo claro}.
+The centres located in \textit{Britannia} can be seen in Figure~\ref{britannia}.
 
@@
 
-In \textit{Germania}, we finally collected a total of 1621 stamps consisted of 850 different stamps from 46 sites (see Fig. \ref{germania}).
+In \textit{Germania}, we collected a total of 1621 stamps from 46 sites displaying 850 different stamp codes (see Figure~\ref{germania}).
 
@@
 
-The approach proposed here is based on the idea of measuring the similarity between amphora workshops by quantifying similar stamps.
-A measure of dissimilarity has been chosen to analyse the dataset.
-We use the statistical technique Morisita-Horn index \citep{morisita_measuring_1959, horn_measurement_1966}.
-This method was performed to measure the dissimilarity between different samples of sets.
-Generally, it describes the dissimilarity between the system of two communities based on the idea of inverse correlation between diversity and species \citep{magurran_why_1988}.
+The approach proposed here aimed at exploring links between production and consumption centers through the identification of common amphorae codes found in different sites.
+For this reason we measured the similarity between amphora workshops and/or consumption sites by quantifying a pairwise distance or dissimilarity index between places (i.e.
+to what extent the stamp codes found on these sites were different).
+The chosen dissimilarity measure was the Morisita-Horn index \citep{morisita_measuring_1959, horn_measurement_1966}.
+This method was applied to measure the dissimilarity between different samples of sets.
+Generally, it describes the dissimilarity between the system of two communities based on the idea of inverse correlation between diversity and species \citep{magurran_why_1988}\xavi{citar reforzar más esta idea}.
 
@@
 Considering our dataset as a non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity \citep{wolda_similarity_1981}.
-Morisita-Horn index can be expressed considering 0 as the total presence of similarity of stamps and 1 a total dissimilarity between stamps.
-In our case, it will be calculated the number of times that one stamp appears in an amphora workshop.
-This method allows to bear in mind a similar number of times for each repeated stamp per workshop.
-If two workshops have similar stamp codes, then the probability would be 0 whereas stamps codes are totally different when the results would be 1.
+Morisita-Horn index gives a value between 0 (sites have an exact stamp codes and frequency distribution) and 1 (complete difference between stamp codes).
+To apply Morisita-Horn we firstly calculated the number of times that each stamp code appeared in an amphora workshop.
+This method allowed us to bear in mind a similar number of times for each repeated stamp per workshop\xavi{no entiendo esta frase}.
+If two workshops hhad similar stamp code distribution then the distance index would be close to 0 whereas sites with completely unrelated stamp codes would give a distance close to 1.
 
@@
 
-Morisita-Horn index resulted in a numeric matrix of dissimilarities containing the distance value between the sites and the code stamps.
-The method was used to group sites with similar stamps codes.
-The matrix was then computed using hierarchical clustering.
+The Morisita-Horn index was used to generate a dissimilarity matrix containing the distance value between each pairwise site based on their code stamps.
+Hierarchical clustering was applied to this matrix in order to group sites with similar stamps codes distributions.
 The algorithm was selected to cluster similar groups in order to analyse the relationship between groups of sites and the distribution of similar stamp codes.
 The results were visualised using a dendrogram to detect groups of sites sharing similar stamp codes.
+
+\xavi{hasta aquí}
 

### 2020-07-01 Paper/jaccardbaetica.tex
comentarios finales

@@
 
-\xavi{hasta aquí}
-
 \section{Results}
@@
 
-The analysis shows that the similarity of amphoric stamps could be correlated with spatial distance.
-The dendrogram in Fig.
-\ref{dendro} was obtained with Morisita-Horn index.
-The dendrogram suggests that amphora workshops used different stamps for their production system.
-Nearby workshops show similarity on the stamps while most workshops seem to display different stamps.
-Groups of workshops sharing similar stamps were not found in the dendrogram: the majority of stamp grouping was composed of no more than three workshops.
-Additionally, workshops that shared more similar amphoric stamps belonged to the same \textit{conventus} area, such as Picachos, Cerro de los Pesebres and El Castillejo.
+The hierarchical clustering of the Morisita-Horn pairwise distance matrix can be seen as a dendrogram in Figure~\ref{dendro}.
+This visualization suggests that each amphora workshop often used unique stamps in their production system that have not been found on any other site.
+In fact, the similarity values are always low and a majority of stamp codes found on multiple sites are only present in two or three workshops.
+
+Nearby workshops display a higher similarity than distant workshops and the varying degrees of similarity between sites which could be correlated to their spatial distance; additionally, the workshops with higher similarity values belong to the same \textit{conventus} area, such as Picachos, Cerro de los Pesebres and El Castillejo.
 
@@
 \includegraphics[angle=180, width=\linewidth]{figs/dendro}
-\caption{Dendrogram obtained by Morisita-Horn algorithm of different amphora workshops in \textit{Baetica} area.
-Colours are represented by areas divided into \textit{Hispalis} (red), \textit{Astigi} (green) and \textit{Corduba} (blue)}
+\caption{Dendrogram displaying the results of hierarchical clustering based on Morisita-Horn metrics for Dressel 20 workshops located in the \textit{Baetica} area.
+Site names are color-coded by conventus: \textit{Hispalis} (red), \textit{Astigi} (green) and \textit{Corduba} (blue)}
 \label{dendro}
@@
 
-The results show a similarity with the results obtained in \textit{Baetica} province. However, the similarity is less pronounced as we observed in the dendrogram in Fig. \ref{britmap}
+Results for the consumption regions display similar patterns than the ones found for the production area of \textit{Baetica}.
+The most important difference is that the similarity values are lower as can be seen in Figure~\ref{britmap}
 
@@
 \includegraphics[angle=180,width=\linewidth]{figs/dendrobrit5.pdf}
-\caption{Dendrogram obtained by Morisita-Horn algorithm of different sites in \textit{Britannia}.
-Colours are represented by type of sites divided into military sites (red), civil sites (green) and no specific (grey)}
+\caption{Dendrogram displaying the results of hierarchical clustering based on Morisita-Horn metrics for sites in \textit{Britannia}.
+Site names are color-coded by typology: military sites (red), civilian sites (green) and unspecified (grey)}
 \label{britmap}
@@
 
-A minor correlation can be explained by the fact that \textit{Britannia} spatial area is much wider than \textit{Baetica} where the centres were more concentrated.
+The lower similarity of stamp codes in \textit{Britannia} compared to \textit{Baetica} could be explained by the spatial correlation as the area under study is much bigger and the sites are more distant between them.
+This can be observed more clearly on military sites as the ones with higher similarity are geographically closer.
+Two patterns can be observed on the dendrogram: 1) nearby sites tend to share more similar stamps and 2) a majority of stamp code are only found on one site.
 
-Most military sites that showed a high similarity in the results were geographically close.
-We identified two factors: 1) nearby sites tend to share similar stamps and 2) most sites have different stamps.
-By contrast, most sites did not show a strong similarity in stamp correlation.
-Neither we found a grouping of similar stamps in a specific place.
-In general, the dendrogram did not observe any production pattern that indicates a clear organisation between production centres and consumption centres.
-In other words, our results did not indicate the presence of a specific production centre from \textit{Baetica} in \textit{Britannia} province.
-Rather, it would imply that olive oil production was distributed by non-specific production centres.
+By contrast, most sites did not show a strong similarity in stamp correlation\xavi{qué quieres decir?
+porque lo de que hay una similaridad baja ya lo has repetido arriba}.
+We also did not find any groups of similar stamps in a specific place\xavi{tampoc entiendo esto.
+que no hay sellos similares en un solo sitio...?}.
 
-In \textit{Germania}, the similarity of stamps shows results less significant than \textit{Britannia} province as it can be seen in the dendrogram (Fig. \ref{germap}).
+These results suggest that there is no clear organizational pattern linking production centres and/or consumption centres beyond spatial distance.
+As a consequence, our analysis does not indicate the presence of a specific production centre from \textit{Baetica} to \textit{Britannia} province.
+Rather, it would imply that olive oil production was distributed by non-specific production centres\xavi{esto no iría en la discusión?
+Por otra parte, no entiendo qué quieres decir exactamente, igual hay que desarrollarlo}.
+
+The similarity of sites based on Dressel 20 stamp codes is also low in \textit{Germania} as can be seen in Figure~\ref{germap}.
 
@@
 \includegraphics[angle=180, width=\linewidth]{figs/dendroger5.pdf}
-\caption{Dendrogram obtained by Morisita-Horn algorithm of different sites in \textit{Germania}.
-Colours are represented by types of sites divided into military sites (red), civil sites (green) and no specific (grey)}
+\caption{Dendrogram displaying the results of hierarchical clustering based on Morisita-Horn metrics for sites in \textit{Germania}.
+Site names are color-coded by typology: military sites (red), civilian sites (green) and unspecified (grey)}
 \label{germap}
@@
 
-A higher concentration of sites sharing similar stamps were found in areas eminently militarised and close to German \textit{limes}, even if most sites mostly showed different stamps.
-The interpretation of the results suggests that there is still no possibility to determine the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.
+A higher concentration of sites sharing similar stamps were found in areas eminently militarised and close to German \textit{limes}, even if most sites mostly showed different stamps~\xavi{quieres decir que los sitios más cercanos al limes son más similares?
+no queda claro y hay que detallarlo un poco más}.
+
+\xavi{esta frase me la petaría; no aporta demasiado: The interpretation of the results suggests that there is still no possibility to determine the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.}
 
@@
 
-In this work, we aimed to analyse the distribution of the amphoric stamps in the organisation of the olive oil market both in the production and consumption areas.
+This work hasd explored to what extent the distribution of amphoric stamps can provide insights on the organisation of olive oil Roman markets both at the production and consumption areas.
 For this reason, an index of dissimilarity was used to detect differences between the distribution of the amphoric stamps and the spatial distance of producers and consumption centres.
-The general purpose was to explore if such differences found in the stamp codes could play an important role in the Roman market.
+The general purpose was to explore if such differences found in the stamp codes could play an important role in the Roman market\xavi{este párrafo es un poco reiterativo no?}.
 
@@
 
-The analysis of the amphora workshops in \textit{Baetica} province resulted in a significant correlation between spatial distance and the similarity of the amphoric stamps.
+The analysis of the amphora workshops in \textit{Baetica} province suggests that a majority of stamp codes were only used on a single workshop.
+Beyond this clear pattern it seems that the similarity of stamp codes is correlated with spatial distance and the amphoric production of spatially close sites shared some stamp codes.
+Particularly relevant here is the fact that our results show how similar stamp codes were found on sites belonging to the same \textit{conventus} area.
+This result could indicate that the workshops of the different \textit{conventus} used different stamp codes that were not shared between them.
+Despite these stamps tend to share the same area of production, we do not identify groups of any more than three workshops sharing the same amphoric stamps as the dendrogram showed.
+In general, the majority of stamps were located in different amphora workshops\xavi{estas 2 frases repiten cosas dichas antes}.
 
-The analysis suggests a connection between the same stamp code and nearby amphora workshops, excluding certain exceptions.
-Consequently, the majority of stamps are located in different amphora workshops and only similar stamps between closer amphora workshops were found.
-Indeed, our results show that most similar stamps were detected in the same \textit{conventus} area.
-This results may indicate a grouping of stamps with similar codes in the same \textit{conventus}.
-Despite these stamps tend to share the same area of production, we do not identify groups of any more than three workshops sharing the same amphoric stamps as the dendrogram showed.
-In general, the majority of stamps were located in different amphora workshops.
+Our results do not support previous working hypotheses suggesting that groups of workshopes used specific groups of stamp codes; while there is higher stamp similarity for nearby workshops this result does not suggest any kind of organisation beyond the simple fact that nearby workshops had a higher degree of interaction.
+In fact; the opposite seems true: each workshop used different stamp codes and only shared some of them with the closest ones.
+These stamp codes detected in closer workshops do not move from other distant workshops.
+In other words, the stamps tend to remain in the same area and different stamps were located in a same amphora workshop\xavi{de nuevo estas 2 frases repiten cosas de antes}.
 
-Our results indicate that the hypothesis about groups of amphora workshops sharing the same stamps seems do not match with the results of the analysis, even though there are similar stamps in closer workshops.
-Rather, it seems that each workshop was organised independently with different stamps.
-Those stamps detected in closer workshops do not move from other distant workshops.
-In other words, the stamps tend to remain in the same area and different stamps were located in a same amphora workshop.
+This unique link between workshop and stamp code can be interpreted in different ways.
+First, it is possible that each workshop operated independently and did not normally collaborate with other workshops.
+Second, stamp similarity in closer workshops could be linked to a spatial pattern.
+It is more probable than closer workshops tend to share more traits than distant workshops.
+While the role of the river was significant for the distribution of amphorae, river connectivity amongst workshops does not seem to show relevancy for the distribution of stamps\xavi{qué quieres decir con esto?
+que los diferentes ríos no muestran ningún patrón?}.
 
-This could be defined by several factors.
-First, each workshop had a different organisation involved in the use of stamps and they were not used in other workshops.
-Second, stamp similarity in closer workshops could be linked to a spatial pattern.
-It is more probably than closer workshops tend to share more traits than distant workshops.
-While the role of the river was significant for the distribution of amphorae, river connection amongst workshops does not seem to show relevancy for the distribution of stamps.
-Finally, the distribution of stamps could have shown some research bias.
-In some cases, workshops have been catalogued with different names in spite of belonging to the same workshops or being closer between each other.
-Additionally, most of the workshops were not widely excavated.
+It is worth mentioning that the performed analysis could be limited by a diversity of biases.
+Some workshops have been identified with different names despite the fact that they probably are sections of the same workshop; on the other hand, the amount of stamps collected on different sites is affected by different intensities for field surveys a lack of excavation.
+As a consequence, further fieldwork could provide a clearer picture of the links between workshops if the sample size of stamp codes is increased.
 
@@
 
-Both consumption centres showed a correlation between spatial distance and similarity in the amphoric stamps. In the case of\textit{Britannia}, the correlation was higher than \textit{Germania}.
+Both consumption areas showed a correlation between spatial distance of sites and their similarity of amphoric stamps.
+In the case of\textit{Britannia}, the correlation was higher than \textit{Germania}.
+In \textit{Britannia} the majority of similarity stamps were mainly found in military centres.
+This could be also interpreted by an intensity bias where military centres have been mostly excavated than civil centres\xavi{pero no decías que el aceite llegaba por las legione?
+Entonces por qué es un bias?
+En todo caso ponlo al final de la sección y no al inicio}.
 
-In \textit{Britannia}, the majority of similarity stamps were mainly found in military centres.
-This could be also interpreted by a intense bias where military centres have been mostly excavated than civil centres.
+A combined exploration of the map and the dendrogram also suggests an interesting pattern: centres with higher similarity values are closer to the coast (North Sea and Celtic Sea).
+This may indicate that the Atlantic route could have played an essential role in transporting olive oil to the area of \textit{Britannia}, since in the places where there is greater similarity they are found in different strategic points near the sea\xavi{por qué más conectividad implica más similaridad?
+no será porque encuentras más ánforas?}.
 
-It should be noted that centres with a similarity correlation were located close to the Atlantic coast around the North Sea and the Celtic Sea.
-This may indicate that the Atlantic route could have played an essential role in transporting olive oil to the area of \textit{Britannia}, since in the places where there is greater similarity they are found in different strategic points near the sea.
+This trend is aligned with recent works that have suggested that the Atlantic trade routes were mostly used for military supply~\citep{remesal_annona_1986,remesal_provincial_2008,carreras_atlantic_2012,morillo_hispania_2016,rubio-campillo_provincias_2018}.
+Therefore, a majority of centres with higher similarity are related to military activities and near the coasts, thus suggesting thast Dressel 20 containing olive oil were shipped to military areas and then redistributed to the civilian population using land-based transportation systems~\citep{carreras_britannia_1998,ayllon_olive_2018}.
 
-This fact seems to concur with the latest research that suggests the idea of a more important Atlantic route that would connect the militarised provinces \citep{remesal_annona_1986,
-remesal_provincial_2008,
-carreras_atlantic_2012,
-morillo_hispania_2016,rubio-campillo_provincias_2018}.
-
-Therefore, most part of the centres where similar stamps are shared can correspond to eminently military areas, which may mean that the transport could reach the military areas from the beginning and then be distributed by land routes along with other civil areas \citep{carreras_britannia_1998,
-ayllon_olive_2018}.
-
-In \textit{Germania}, the results followed the same pattern as \textit{Britannia} but with a minor correlation as we can observe in the dendrogram.
+The results for \textit{Germania} follow a similar pattern than \textit{Britannia} but with a minor correlation as we can observe in the dendrogram\xavi{esta frase está repetida de los resultados.
+quizá quitarla de ahí y dejarla aquí?}.
 The areas mostly militarised share similar stamps than civil areas.
@@
 
-It is worth to mention that it was not detected a robust model of organisation in both cases with regard to the distribution of stamps in consumption centres.
+It is worth mentioning that it was not detected a robust model of organisation in both cases with regard to the distribution of stamps in consumption centres\xavi{de nuevo esto es redundante no?}.
 This evidence means that it is unknown whether some production centres went to one province or another or, at least, that they can be clearly reflected in the data with a greater similarity in the amphoric stamps.
@@
 
-Judging by the results obtained, there does not seem to be a specific pattern in terms of geographical distribution.
+\xavi{me petaría esta frase, porque entre otras cosas arriba dices que hay correlación espacial, así que cómo es que aquí dices que no hay distribución geográfica aparente?}Judging by the results obtained, there does not seem to be a specific pattern in terms of geographical distribution.
 Nor it was detected consumption areas where stamps are specified from an amphora workshop.
@@
 
-The results of this work can be interpreted for several reasons.
-On the one hand, the use of amphoric stamps could have been exclusively running by the owner or owners of the workshop to distinguish the amphora workshop.
-This hypothesis would explain the fact that we do not find similar stamps in different workshops; however, different code stamps have been detected in the same workshops that they would be barely difficult to assign to different owners.
+This work has used quantitative analysis to explore long-range trade of olive oil using the stamps found in Dressel 20 vessels.
+The most interesting result is linked to the meaning of these inscriptions.
+The use of specific codes seems to be almost exclusively decided by each workshop, so it could be used to identify amphorae found in consumption centers (i.e.
+an amphorae with a given stamp was probably made on a specific workshop).
+This idea could explain why the similarity of stamps between workshops is generally very low.
+On the other hand, some code stamps have been identified in nearby workshops so this pattern suggests that the clustered workshops were somehow connected (e.g.
+they belonged to the same owner).
 
-On the other hand, the presence of different stamps in the same workshop could imply some kind of organisation from within the workshop affecting the closest centres \citep{juanmorostesis}.
-The marked amphorae could have allowed to potters organise the production with a batch stamping for the posterior commercialisation.
+Alternatively, the presence of different stamps in the same workshop could imply some kind of organisation from within the workshop only affecting nearby centres \citep{juanmorostesis}.
+The stamped amphorae could have allowed potters to organise the differnt production batches for posterior commercialisation\xavi{estos 2 párrafos igual deberían ir a la discusión no?
+La conclusión debería ser algo más general y aquí estás aún interpretando resultados}.
 
 This could be explained such as batch systematic organisation between potters \citep{juanmorostesis}.
-This method allowed to potters organises the production with the batch stamping for the posterior commercialisation.
+This method allowed to potters organises the production with the batch stamping for the posterior commercialisation\xavi{y esto para qué serviría?
+debería explicarse un poco más}.
 
-Considering that Dressel 20 was not marked in most cases, some researchers point out that potters marked the amphorae to prepare and distribute the product in order to be shipped \citep{berni_millet_epigrafianforica_2008}.
-This method would be used as an identifier to count the number of amphorae of a branch \citep{juanmorostesis}.
-This organisation could also have served to identify different groups of potters working in the same amphora workshop.
-Potters could have marked the amphorae to distinguish different groups working in parallel \citep{li_crossbows_2014}.
-This fact would explain wherefore why we detect different stamps in the same workshop.
-In any case, we do not have enough archaeological evidence that can validate the interpretations presented here and our results are only certainly valid with the context of our case study.
+A large number of Dressel 20 were not marked with any stamps, so some researchers suggest that potters marked the amphorae to prepare and distribute the product in order to be shipped \citep{berni_millet_epigrafianforica_2008}.
+This method would be used as an identifier to count the number of amphorae of a branch \citep{juanmorostesis}\xavi{a branch?
+que es una branch?}.
+This organisation\xavi{this organisation o unique stamp codes?} could have also served to identify different groups of potters working in the same amphora workshop.
+Potters could have marked the amphorae to distinguish different groups working at the same time in different orders~\citep{li_crossbows_2014}.
+This hypothesis could explain why we detect different stamps in the same workshop.
+In any case, at present we do not have enough archaeological evidence to validate which of these interpretations is more plausible in our case study.
 
-As a summary, this method presented here provides a potential tool to understand the mechanisms of production based on the similarity of artefacts.
-This work has identified differences in the case of the amphoric production within the Roman Empire.
-The combination of archaeological data and quantitative methods have finally highlighted for the interpretation of the complex economic processes connected with the archaeological evidence.
+To conclude, the method presented here provides a useful framework to improve our understanding of the organisation and processes that allowed the Roman Empire to manage a massive and highly specialized production of essential goods.
+This work has identified differences in the case of the amphoric production within the Roman Empire\xavi{qué diferencias?}.
+The growing amount of archaeological data requires of large-scale quantitative methods such as the one presented here to identify and explore complex patterns, thus allowing us to infer dynamics and improve our interpretation of complex economic processes of past societies.
 
@@
 
-The research was funded by European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros and Ignacio Morer for their useful suggestions.  
+This research was partly funded by the European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros and Ignacio Morer for their useful suggestions.  
 All data has been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}.

### 2020-07-01 Paper/jaccardbaetica.tex
add new changes xavi

--- 2020-06-30 Paper/jaccardbaetica.tex
@@
 
-The intensification of long-range trade was one of the most important traits of the economy developed duriong the Roman Empire.
+The intensification of long-range trade was one of the most important traits of the economy developed during the Roman Empire.
 The development of an extensive road network increased the connectivity between inland communities, but most shipping continued being based on maritime routes particularly in the Mediterranean basin~\citep{temin_market_2001,bevan_mediterranean_2014}.
@@
 In archaeology, they allow us to highlight a part of the mechanism of production and distribution of goods along the Mediterranean \citep{bevan_mediterranean_2014}.
-Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}\xavi{aquí faltan citas de obras más generales; esto es un ejemplo pero muy específico no?}.
+Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}
 %spread of goods and ideas had a enormous impact during the Roman Empire. The progressive exploitation of communication networks allowed a major interaction between communities
@@
 The Empire developed a series of structures to support and organize this long-range trade and specialized entire regions to massively produce specific goods.
-An example of this process was the province of \textit{Baetica} (currently Andalusia, southern Spain) which became an important olive oil production centre during\xavi{período?}.
+An example of this process was the province of \textit{Baetica} (currently Andalusia, southern Spain) which became an important olive oil production centre during the Roman Empire.
 Olive oil was an essential good for Romans because it was used in almost every aspect of their daily life such as cooking, hygiene or lighting~\citep{mattingly_d.j._oil_1988}.
@@
 
-The structure and processes associated to this massive olive oil production has been extensively discussed over the last decates\citep{rodriguez_economioleicola_1977, Chic_hispania_1997,millet_anforas_1998}.
-Our knowledge of this economic activity has benefitted from new findings and data sources, but despite all these advances several questions remain open: \xavi{no entiendo esta pregunta} what patterns were followed for the distribution of olive oil to the different provinces?
+The structure and processes associated to this massive olive oil production has been extensively discussed over the last decades\citep{rodriguez_economioleicola_1977, Chic_hispania_1997,millet_anforas_1998}.
+Our knowledge of this economic activity has benefited from new findings and data sources, but despite all these advances several questions remain open: \xavi{no entiendo esta pregunta} what patterns were followed for the distribution of olive oil to the different provinces?
 \xavi{no es esta pregunta idéntica a la anterior?} Did each province follow a different pattern for its distribution?
@@
 \caption{Overview of the sites analyzed in this paper.
-The color defines the three different regionsu under study (Baetica as producer; Germania and Britannia as consumers) while the size of each dot highlights the number of Dressel-20 amphoric stamps found on each of these sitese}
+The color defines the three different regions under study (Baetica as producer; Germania and Britannia as consumers) while the size of each dot highlights the number of Dressel-20 amphoric stamps found on each of these sites}
 
@@
 The landscape of the \textit{Baetican} province was one of the best regions to face the increase in demand of olive oil across the Roman provinces.
-For this reason the area saw an increase in productivity as a massive infrastructure of olive oil production was gradually deplyoyed.
+For this reason the area saw an increase in productivity as a massive infrastructure of olive oil production was gradually deployed.
 The production and distribution of olive oil grew exponentially\xavi{seguro que exponencial?} for almost three centuries\xavi{qué centuries?} \citep{remesal_concierto}.
@@
 The interpretation for most of these inscriptions is still open due to the fragmentation of the material and the small sample size of well-preserved elements~\citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
-
-%\xavi{Estás segura de esto?
-Porque aquí hay 2 cosas: 1.
-datar bien y 2 .
-período de uso.
-Yo no tengo claro que la mayoría de sellos se usen durante mucho tiempo.
-En qué está basada la idea?
-Deberías detallar un poco más este temita cronológico y explicar como sabemos el período de uso.}
 
@@
 The creation of new provinces allowed the Roman Empire the arrival of resources through Mediterranean and Atlantic routes\xavi{no entiendo esto...allowed o required?}.
-This led to a gradual change in the economic and social structureas the trade networks that supplied the Roman army became more complex.
+This led to a gradual change in the economic and social structures the trade networks that supplied the Roman army became more complex.
 Augustus's administration created the figure of the \textit{praefectura annonae} to organize wheat supply.
@@
 \xavi{faltaría un enlace a la siguiente sección, queda muy abrupto}
-
-%\xavi{este título de sección debería cambiarse por algo más descriptivo. Case studies o algo asínnn? Por otra parte, igual podrías dividirlo en 2 subsections}
-
-%\xavi{en general, echo a faltar un poco más de detalle en esta discusión y en los datos que tienes; en qué se parecen y diferencian las 2 provincias?
-Si hablas del uso por las campañas o las guarniciones, donde están?
-Qué strategic points de redistribución hay?
-Tampoco a saco, pero entender mejor cómo llega el aceite, cómo se redistribuye y quién y dónde se consume}
 

### 2020-07-01 Paper/jaccardbaetica.tex
add new changes correction xavi

@@
 For this reason the area saw an increase in productivity as a massive infrastructure of olive oil production was gradually deployed.
-The production and distribution of olive oil grew exponentially\xavi{seguro que exponencial?} for almost three centuries\xavi{qué centuries?} \citep{remesal_concierto}.
+The production and distribution of olive oil increased from the first to the third centuries AD ~\citep{remesal_concierto}.
 Hundreds of presses and amphora-making workshops were build near large extensions covered by olive trees.
@@
 The evidence for olive oil consumption in \textit{Britannia} is scarce before the Roman conquest \citep{funari_corpus_1996,carreras_abastecimiento_2003}.
-Indigenous\xavi{no sé si indigenous es la mejor palabra...local?} population did not consume this product and it was not produced on the region as the landscape and climate of the British isles was not suitable for olive oil trees\citep[161]{monfort_britanniaen_1998}.
+Local population did not consume this product and it was not produced on the region as the landscape and climate of the British isles was not suitable for olive oil trees\citep[161]{monfort_britanniaen_1998}.
 Based on the archaeological record this absence of olive oil changed with the arrival of a large amount of Dressel 20 amphorae to \textit{Britannia} \citep[1]{carreras_britannia_1998}\xavi{en qué siglo?}.
@@
 The presence of Dressel 20 stamps in military camps in \textit{Britannia} has been widely studied in Roman archaeology \citep{carreras_britannia_1998}.\xavi{si dices widely studied no puedes referencias solo 1 autor}.
-This fact\xavi{que fact?
-Que los sellos de dressel 20 han sido muy estudiados?
-No tengas miedo a repetir trozos rollo "This intense consumption of olive oil also indicatas blablabla...} also indicates a possible provincial structure designed to organize the supply of olive oil to military camps,m such as \textit{Germania} (Remesal Rodrıguez, 1986)\xavi{por qué such as Germania?
+
+This intense consumption of olive oil also indicates a possible provincial structure designed to organize the supply of olive oil to military camps, such as \textit{Germania} (Remesal Rodrıguez, 1986)\xavi{por qué such as Germania?
 Aún no has explicado qué hacen en Germania}.
@@
 
-%\xavi{si el eje es logarítmico deberías mencionarlo}
-
 The distribution of amphora stamps for each \textit{conventus} can be seen in Figure~\ref{frequency}.
@@
 
-%\xavi{vale pero qué es cada punto? un workshop?}
-
 \subsection{Consumption centres: Britannia and Germania}
 
-%\xavi{Esta estructura no me convence...tenemos que pensar algo mejor, porque ahora es: 1.
-hablo de britania/germania, 2.
-hablo de sellos en general, 3.
-hablo de sellos en britania/germania.
-Quizás juntar esto con lo de los casos de estudio, o quizás hablar de los 2 casos de estudio en la intro, y dividir la sección anterior en 2 (lo más general a la intro, lo más concreto aquí)}
 The analysis of stamps in consumption areas also used the CEIPAC database and filtered in the same way than the previous dataset.
@@
 This method allowed us to bear in mind a similar number of times for each repeated stamp per workshop\xavi{no entiendo esta frase}.
-If two workshops hhad similar stamp code distribution then the distance index would be close to 0 whereas sites with completely unrelated stamp codes would give a distance close to 1.
+If two workshops had similar stamp code distribution then the distance index would be close to 0 whereas sites with completely unrelated stamp codes would give a distance close to 1.
 
@@
 The most important difference is that the similarity values are lower as can be seen in Figure~\ref{britmap}
-
-%\xavi{no tengo claro que le puedas decir correlación, porque eso es un valor específico entre 2 distribuciones numéricas. Te refieres a similaridad?}
 
@@
 This trend is aligned with recent works that have suggested that the Atlantic trade routes were mostly used for military supply~\citep{remesal_annona_1986,remesal_provincial_2008,carreras_atlantic_2012,morillo_hispania_2016,rubio-campillo_provincias_2018}.
-Therefore, a majority of centres with higher similarity are related to military activities and near the coasts, thus suggesting thast Dressel 20 containing olive oil were shipped to military areas and then redistributed to the civilian population using land-based transportation systems~\citep{carreras_britannia_1998,ayllon_olive_2018}.
+Therefore, a majority of centres with higher similarity are related to military activities and near the coasts, thus suggesting that Dressel 20 containing olive oil were shipped to military areas and then redistributed to the civilian population using land-based transportation systems~\citep{carreras_britannia_1998,ayllon_olive_2018}.
 
@@
 Neither we do not detect production centres dedicated to the distribution of olive oil in a certain province.
-
-%In general, we observed a concentration of the stamps in military centres.
 
@@
 Alternatively, the presence of different stamps in the same workshop could imply some kind of organisation from within the workshop only affecting nearby centres \citep{juanmorostesis}.
-The stamped amphorae could have allowed potters to organise the differnt production batches for posterior commercialisation\xavi{estos 2 párrafos igual deberían ir a la discusión no?
+The stamped amphorae could have allowed potters to organise the different production batches for posterior commercialisation\xavi{estos 2 párrafos igual deberían ir a la discusión no?
 La conclusión debería ser algo más general y aquí estás aún interpretando resultados}.

### 2020-07-03 Paper/jaccardbaetica.tex
add new changes xavi

@@
 
-Advances in the research of the Roman studies have currently led an environment with more diverse interpretations on commercial dynamics \citep{duncan1982economy,
+Advances in the research of the Roman studies have currently led an environment with various interpretations on commercial dynamics \citep{duncan1982economy,
 temin_economy_2006,
-quantifyingwilson2009}\xavi{more diverse que qué?}.
-The application of different quantitative approaches has allowed us to improve our interpretation of the growing amount of archaeological evidence and reveal the complexity of the Roman production \xavi{en qué sentido?
+quantifyingwilson2009}.
+%\xavi{more diverse que qué?} 
+The application of different quantitative approaches has allowed us to improve our interpretation of the growing amount of archaeological evidence and reveal the dynamics of the Roman production %\xavi{en qué sentido?
 qué quieres decir aquí con complexity?
-Si no especificas queda muy genérico (rollo "el mundo es muy complejo")}.\citep{brughmans_roman_2016,
+Si no especificas queda muy genérico (rollo "el mundo es muy complejo")}.
+\citep{brughmans_roman_2016,
 orengo_seeds_2016,bayesian_2018,
@@
 This paper aims to study the olive oil market connection between provinces by calculating the similarity of stamps.
-Specifically, our work pretends to detect microeconomic processes\xavi{no me gusta lo de microeconomic, porque es un campo específico de la economía que nada tiene que ver con analizar un limes entero.
-Quizás regional?} focused on a commercial product from a specific province \citep{isaksen_network_2006}.
+Specifically, our work pretends to detect regional processes focused on a commercial product from a specific province \citep{isaksen_network_2006}.
 We want to understand the pattern of olive oil production linked to amphora workshops and amphoric stamps used to mark them.
@@
 
-The chronology of the workshops is widely diverse from the first to the third centuries AD\xavi{qué quiere decir que la cronología es diversa?} \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
+The chronology of the workshops is wide spanning from the first to the third centuries AD
+%\xavi{qué quiere decir que la cronología es diversa?} 
+\citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}. 
 This fact is shown by the archaeological evidence that displays a highly specialised production with a long activity with apparently few changes\xavi{deberías desarrollar esto un poco más; qué tipo de evidencia demuestra esto?
@@
 These stamps are typically found on handles but they were also imprinted on rims and the amphora body~\citep{millet_anforas_1998}.
-The information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion\xavi{qué information?
-El código de letras?}.
+The content information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion
+%\xavi{qué information? El código de letras?}. 
 Stamps often displayed a code of three letters and they can appear in abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.\xavi{que quieres decir con abbreviated o complete?
@@
 
-\section{Case study: consumption centres}
+\section{Consumption centres: Britannia and Germania}
 
-\xavi{este título es muy raro, porque 1. que son casos de estudio ya lo sabes y 2. case study singular, consumption centres plural}
+%\xavi{este título es muy raro, porque 1. que son casos de estudio ya lo sabes y 2. case study singular, consumption centres plural}
 
-The creation of new provinces allowed the Roman Empire the arrival of resources through Mediterranean and Atlantic routes\xavi{no entiendo esto...allowed o required?}.
+The creation of new provinces required the Roman Empire the arrival of resources through Mediterranean and Atlantic routes.
+%\xavi{no entiendo esto...allowed o required?}. 
 This led to a gradual change in the economic and social structures the trade networks that supplied the Roman army became more complex.
@@
 Olive oil supply seems to be particularly intense in militarised provinces that functioned as the borders of the Empire such as \textit{Britannia} and \textit{Germania}.
-The high percentage of amphorae\xavi{amphorae o Dressel 20?} found in sites nearby military garrisons suggests the existence of diverse supply dynamics as there were trade processes flowing good from military sites to civil settlements\citep{remesal_annona_1986, carreras_britannia_1998}\xavi{lo que quieres decir es que tienes muchas ánforas en sitios civiles cerca de militares?
+The high percentage of Dressel 20 amphorae found in sites nearby military garrisons suggests the existence of diverse supply dynamics as there were trade processes flowing good from military sites to civil settlements\citep{remesal_annona_1986, carreras_britannia_1998}\xavi{lo que quieres decir es que tienes muchas ánforas en sitios civiles cerca de militares?
 no queda claro}.
@@
 This means that we included in the sample some centres that are actually located beyond the administrative boundaries but were considered part of the same border.
-Specifically, for the German \textit{liomes} we included sites from the two provinces (Ulterior and Citerior\xavi{no es superior e inferior?}) while the analysis of Britannia extended to \textit{Caledonia}.
+Specifically, for the German \textit{limes} we included sites from the two provinces (Superior and Inferior) while the analysis of Britannia extended to \textit{Caledonia}.
 
@@
 
-\xavi{esta frase me la petaría; no aporta demasiado: The interpretation of the results suggests that there is still no possibility to determine the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.}
+%\xavi{esta frase me la petaría; no aporta demasiado: The interpretation of the results suggests that there is still no possibility to determine the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.}
 
@@
 
-This work hasd explored to what extent the distribution of amphoric stamps can provide insights on the organisation of olive oil Roman markets both at the production and consumption areas.
+This work has explored to what extent the distribution of amphoric stamps can provide insights on the organisation of olive oil Roman markets both at the production and consumption areas.
 For this reason, an index of dissimilarity was used to detect differences between the distribution of the amphoric stamps and the spatial distance of producers and consumption centres.
@@
 Both consumption areas showed a correlation between spatial distance of sites and their similarity of amphoric stamps.
-In the case of\textit{Britannia}, the correlation was higher than \textit{Germania}.
+In the case of  \textit{Britannia}, the correlation was higher than \textit{Germania}.
 In \textit{Britannia} the majority of similarity stamps were mainly found in military centres.
@@
 
-\xavi{me petaría esta frase, porque entre otras cosas arriba dices que hay correlación espacial, así que cómo es que aquí dices que no hay distribución geográfica aparente?}Judging by the results obtained, there does not seem to be a specific pattern in terms of geographical distribution.
+%\xavi{me petaría esta frase, porque entre otras cosas arriba dices que hay correlación espacial, así que cómo es que aquí dices que no hay distribución geográfica aparente?}Judging by the results obtained, there does not seem to be a specific pattern in terms of geographical distribution.
 Nor it was detected consumption areas where stamps are specified from an amphora workshop.
@@
 A large number of Dressel 20 were not marked with any stamps, so some researchers suggest that potters marked the amphorae to prepare and distribute the product in order to be shipped \citep{berni_millet_epigrafianforica_2008}.
-This method would be used as an identifier to count the number of amphorae of a branch \citep{juanmorostesis}\xavi{a branch?
-que es una branch?}.
-This organisation\xavi{this organisation o unique stamp codes?} could have also served to identify different groups of potters working in the same amphora workshop.
+This method would be used as an identifier to count the number of amphorae of a batch \citep{juanmorostesis}.
+
+Unique stamp codes could have also served to identify different groups of potters working in the same amphora workshop.
 Potters could have marked the amphorae to distinguish different groups working at the same time in different orders~\citep{li_crossbows_2014}.

### 2020-07-06 Paper/jaccardbaetica.tex
add new changes xavi

@@
 The structure and processes associated to this massive olive oil production has been extensively discussed over the last decades\citep{rodriguez_economioleicola_1977, Chic_hispania_1997,millet_anforas_1998}.
-Our knowledge of this economic activity has benefited from new findings and data sources, but despite all these advances several questions remain open: \xavi{no entiendo esta pregunta} what patterns were followed for the distribution of olive oil to the different provinces?
-\xavi{no es esta pregunta idéntica a la anterior?} Did each province follow a different pattern for its distribution?
-\xavi{esta frase no tiene sentido aquí; o la pones antes o en otro párrafo, pero es raro acabar un set de 2 preguntas con una frase que no sea otra pregunta} Nor the lack of written records has made possible to detect any indication of patterns in the olive oil market.
+However, the lack of written records has made possible to detect any indication of patterns in the olive oil market.
 
-\xavi{las preguntas no van sobre patrones; los patrones son las trazas en la evidencia que te ayudan a responder a las preguntas}
+Our knowledge of this economic activity has benefited from new findings and data sources, but despite all these advances several questions remain open: How was amphora production organized during the Roman Empire?
+Can we detected economic dynamics between provinces by analysing the distribution of amphorae?
+
+%Did each province follow a different pattern for its distribution? 
+%\xavi{no entiendo esta pregunta} what patterns were followed for the distribution of olive oil to the different provinces? 
+%\xavi{no es esta pregunta idéntica a la anterior?} 
+%\xavi{esta frase no tiene sentido aquí; o la pones antes o en otro párrafo, pero es raro acabar un set de 2 preguntas con una frase que no sea otra pregunta}
+
+%\xavi{las preguntas no van sobre patrones; los patrones son las trazas en la evidencia que te ayudan a responder a las preguntas}
 
@@
 These stamps are typically found on handles but they were also imprinted on rims and the amphora body~\citep{millet_anforas_1998}.
-The content information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion
+The content information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion.
 %\xavi{qué information? El código de letras?}. 
-Stamps often displayed a code of three letters and they can appear in abbreviated form or complete and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.\xavi{que quieres decir con abbreviated o complete?
-Como abrevias un código de 3 letras?}
+Stamps often displayed a code of three letters and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
+The three letters correspond to a initial forms of a name but they can also appear complete.
+%\xavi{que quieres decir con abbreviated o complete? Como abrevias un código de 3 letras?}
 
 Scholars agree that stamps are some type of identification mark, but there is no consensus on the meaning of these marks\citep{rodriguez_baetican_1998}.
-Additionally, a large percentage of Dressel 20 containers were not stamped thus complicating the interpretation~\xavi{esta última frase aquí es un poco random no?}.
+
+%Additionally, a large percentage of Dressel 20 containers were not stamped thus complicating the interpretation~\xavi{esta última frase aquí es un poco random no?}.
+
 The stamp codes are interpreted based on three main ideas: content (olive oil), context (amphora workshop) and subject (individuals involved on the production).
@@
 
-The output was a dataset of 4271 stamps found in sites belonging to \textit{Britannia} (2219 stamps) and \textit{Germania} (2052 stamps).
+The output was a dataset of 4271 stamps found in sites belonging to \textit{Britannia} (total: 2219 stamps, analysed: 1765 stamps ) and \textit{Germania} (total: 2052 stamps, analysed: 1621 stamps).
+Some stamps from \textit{Britannia} and \textit{Germania} were discarded for the analysis for incomplete information.
+
 Both \textit{Germania} and \textit{Britannia} were analysed as borders and not as Roman provinces.
@@
 %CITAR (Callender, 1965; Carreras Monfort y Funari,1998; Ayllón-Martı́ et al., 2018). 
-In \textit{Britannia}, we studied a total sample of 1765 stamps from 46 sites displaying 968 unique codes\xavi{pero esto no concuerda con las 2219 de antes...es por descartar los inciertos?
-Hay que dejarlo claro}.
+In \textit{Britannia}, we studied a total sample of 1765 stamps from 46 sites displaying 968 unique codes
+
+%\xavi{pero esto no concuerda con las 2219 de antes...es por descartar los inciertos? Hay que dejarlo claro}.
+
 The centres located in \textit{Britannia} can be seen in Figure~\ref{britannia}.
@@
 
+The results for \textit{Germania} follow a similar pattern than \textit{Britannia} but with a minor correlation as we can observe in the dendrogram.
+
 A higher concentration of sites sharing similar stamps were found in areas eminently militarised and close to German \textit{limes}, even if most sites mostly showed different stamps~\xavi{quieres decir que los sitios más cercanos al limes son más similares?
@@
 
-The results for \textit{Germania} follow a similar pattern than \textit{Britannia} but with a minor correlation as we can observe in the dendrogram\xavi{esta frase está repetida de los resultados.
+%The results for \textit{Germania} follow a similar pattern than \textit{Britannia} but with a minor correlation as we can observe in the dendrogram\xavi{esta frase está repetida de los resultados.
 quizá quitarla de ahí y dejarla aquí?}.
-The areas mostly militarised share similar stamps than civil areas.
-However, we do not detect a concrete pattern regarding the distribution of the stamps in the German limes \citep{xanten2018}.
+
+.The areas mostly militarised share similar stamps than civil areas. However, we do not detect a concrete pattern regarding the distribution of the stamps in the German limes \citep{xanten2018}.
 
@@
 To conclude, the method presented here provides a useful framework to improve our understanding of the organisation and processes that allowed the Roman Empire to manage a massive and highly specialized production of essential goods.
-This work has identified differences in the case of the amphoric production within the Roman Empire\xavi{qué diferencias?}.
+This work has identified similarities between amphoric stamps and spatial distribution in the case of the amphoric production within the Roman Empire
+%\xavi{qué diferencias?}. 
 The growing amount of archaeological data requires of large-scale quantitative methods such as the one presented here to identify and explore complex patterns, thus allowing us to infer dynamics and improve our interpretation of complex economic processes of past societies.

### 2020-07-13 Paper/jaccardbaetica.tex
add new changes correction

@@
 
+This study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps.
 To do this, a population approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}.
+A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps.
+Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potential spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
 Stamps will be used to identify economic patterns by analysing their similarity.
@@
 \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}. 
-This fact is shown by the archaeological evidence that displays a highly specialised production with a long activity with apparently few changes\xavi{deberías desarrollar esto un poco más; qué tipo de evidencia demuestra esto?
-pocos cambios en qué?
-los talleres?
-las tipologías anfóricas?} \citep{remesal_anforas_2004}.
+This fact is shown by the archaeological evidence that displays a highly specialised production with a long activity \citep{remesal_anforas_2004}.
+Thus, the specialisation of this production made amphorae with apparently few changes in their forms, reaching a high degree of standardisation.
+
+%\xavi{deberías desarrollar esto un poco más; qué tipo de evidencia demuestra esto? pocos cambios en qué? los talleres? las tipologías anfóricas?} .
 
 A large majority of this Baetican amphorae production is composed by Dressel 20 types\citep{dressel_ricerche_1878,millet_anforas_1998}.
-Dressel 20 amphorae were shipped and distributed from \textit{Baetica} throughout western Europe\xavi{esto ya lo has mencionado; quizás podrías explicar en 2 líneas la capacidad y forma de la dressel 20}.
+Dressel 20 amphorae have a recognisable globular form with short neck and oval-shaped handles.
+Its  capacity allowed to hold 70-75 litres of olive oil to be shipped and distributed throughout Empire \citep{berni_dressel_2016}.
+
+%were shipped and distributed from \textit{Baetica} throughout western Europe. 
+%\xavi{esto ya lo has mencionado; quizás podrías explicar en 2 líneas la capacidad y forma de la dressel 20}.
 
@@
 Dressel 20 was the amphora most stamped during the Roman Empire \citep[18]{millet_anforas_1998}.
-The usage period is associated with a large amphora production for almost three centuries\xavi{esto está repetido no?}.
+
+%The usage period is associated with a large amphora production for almost three centuries\xavi{esto está repetido no?}.
 
 The large sample size of recovered Dressel 20 stamps is distributed across all the western provinces of the Roman Empire.
-For this reason there are several publications of this type of evidence discussing its origin, its meaning, its long period of use and its spatial distribution~\citep{remesal_sellar_2016} (Fig.\ref{amphora}).\xavi{si dices que es muy popular debería haber aquí más citas aparte de Remesal}.
+For this reason there are several publications of this type of evidence discussing its origin, its meaning, its long period of use and its spatial distribution~\citep{dressel_ricerche_1878,
+rodriguez_economioleicola_1977,
+chicepi1985,millet_anforas_1998, remesal_sellar_2016} (Fig.\ref{amphora}).
+
+%\xavi{si dices que es muy popular debería haber aquí más citas aparte de Remesal}.
 
@@
 
-The presence of Dressel 20 stamps in military camps in \textit{Britannia} has been widely studied in Roman archaeology \citep{carreras_britannia_1998}.\xavi{si dices widely studied no puedes referencias solo 1 autor}.
+The presence of Dressel 20 stamps in military camps in \textit{Britannia} has been widely studied in Roman archaeology \citep{williams_importation_1983,
+funari_corpus_1996,
+carreras_britannia_1998,carreras_abastecimiento_2003}.%\xavi{si dices widely studied no puedes referencias solo 1 autor}.
 
-This intense consumption of olive oil also indicates a possible provincial structure designed to organize the supply of olive oil to military camps, such as \textit{Germania} (Remesal Rodrıguez, 1986)\xavi{por qué such as Germania?
-Aún no has explicado qué hacen en Germania}.
+This intense consumption of olive oil also indicates a possible provincial structure designed to organize the supply of olive oil to military camps,
+
+%such as \textit{Germania} (Remesal Rodrıguez, 1986)\xavi{por qué such as Germania? Aún no has explicado qué hacen en Germania}.
+
 There are no written records explaining how this redistribution of essential goods was organised in \textit{Britannia}, but the archaeological evidence suggests that cities may have been the central nodes of this local trade network~\citep[45]{funari_economic_2005}.
@@
 This date matches a change in market strategy in the Empire.
-This a gradual decrease on \textit{Baetican} olive oil exports can be observed as a decrease in the amount of Dressel 20 found in excavated sites as this typology is replaced by Dressel 23 amphorae \citep{rodriguez1991aceite,millet_anforas_1998}.\xavi{pero entonces la dressel 23 de donde viene?
+This a gradual decrease on \textit{Baetican} olive oil exports can be observed as a decrease in the amount of Dressel 20 found in excavated sites as this typology is replaced by other types of amphorae from other places \citep{rodriguez1991aceite,millet_anforas_1998}.
+
+%\xavi{pero entonces la dressel 23 de donde viene?
 Cuidao que estás diciendo que se acaban las dressel 20, pero esto implica que hay menos aceite de oliva o que viene de otras partes o que va en recipientes distintos a la dressel 20?}
@@
 \textit{Germania} presents a similar introduction pattern of olive oil than the one discussed for \textit{Britannia}.
-There is no archaeological evidence for olive oil consumption before the arrival of the Roman army to the region while a majority of recovered Dressel 20 amphorae are located in the military sites that formed the German \textit{limes}\xavi{cita?}.
+There is no archaeological evidence for olive oil consumption before the arrival of the Roman army to the region while a majority of recovered Dressel 20 amphorae are located in the military sites that formed the German \textit{limes} \citep{remesal_germaniaengl_2002}.
+
+%\xavi{cita?}
 
@@
 We are especially interested in identifying links between production and consumption centres by using amphoric stamps.
-To do that, we use the CEIPAC database to collect stamps from different places.
-The CEIPAC dataset contains over 50.000 of epigraphy records found in amphorae, mostly from \textit{Monte Testaccio}.
-This study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps.
+
+%To do that, we use the CEIPAC database to collect stamps from different places. The CEIPAC dataset contains over 50.000 of epigraphy records found in amphorae, mostly from \textit{Monte Testaccio}.
+
+To study the amphoric stamps, we used the data collected in the CEIPAC database of amphoric epigraphy from different places~\citep{remesal_centro_2015} (see the database here \url{http://romanopendata.eu}).
+The CEIPAC dataset contains over 50.000 epigraphy records found on different types of amphorae, mostly from \textit{Monte Testaccio}.
+
+%\xavi{podrías mover esta parte a los 2 párrafos anteriores y explicar la bbdd brevemente antes de empezar a hablar de la Bética (porque es info que también aplica a los otros 2 casos de estudio no?}.
+
+%This study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps.
 A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps.
 Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potential spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
-\xavi{estos 2 párrafos explican objetivos, que deberían ir en la intro y no aquí, donde deberías empezar a hablar directamente de la BBDD}
+%\xavi{estos 2 párrafos explican objetivos, que deberían ir en la intro y no aquí, donde deberías empezar a hablar directamente de la BBDD}
 
 \subsection{Production centres: Baetica province}
-
-To study the stamps found in the Baetican workshops we used the data collected in the CEIPAC database of amphoric epigraphy~\citep{remesal_centro_2015} (see the database here \url{http://romanopendata.eu}).
-This database contains over 50.000 epigraphy records found on different types of amphorae\xavi{podrías mover esta parte a los 2 párrafos anteriores y explicar la bbdd brevemente antes de empezar a hablar de la Bética (porque es info que también aplica a los otros 2 casos de estudio no?}.
 
@@
 Our dataset also contained a new categorical variable defining the \textit{conventus} of each site.
-The \textit{conventus} were administrative centres for territorial organisation in the Roman Empire \citep[58]{ozcariz_gil_administracion_2013}\xavi{estaría bien ampliar un poco esto; qué organizaban exactamente?}.
+The \textit{conventus} were administrative centres for territorial organisation under the Roman Empire \citep[58]{ozcariz_gil_administracion_2013}.
+The creation of the \textit{conventus} in some Roman provinces was a form of control with the purpose of organising the administration in the provinces \citep{albertini_les_1923}.
+
+%\xavi{estaría bien ampliar un poco esto; qué organizaban exactamente?}
+
 The production area for Dressel 20 amphorae extended across three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter Hispalis), \textit{Cordubensis} (currently C\'ordoba, hereafter Corduba) and \textit{Astigi} (currently Écija, Sevilla, hereafter Astigi) \citep{rodriguez_economioleicola_1977,chicdatos2001,berni_millet_epigrafianforica_2008} .
@@
 It is important to mention that some workshops exhibited the same coordinates but they were catalogued as different workshops in Roman Studies (i.e.
-Tesorillo de Doña Mencia and Doña Mencia).
-The analysis took into account this particularity but some biases can occur due to the archaeological dataset\xavi{No se entiende; deberías explicarlo un poco mejor y citar algo si dices eso.
-Piensa que la mayoría de lectores no sabrán quién es Doña Mencia}.
+nearby amphora workshops in \textit{Baetica} named as Tesorillo de Doña Mencia and Doña Mencia).
+The analysis took into account this particularity but some biases can occur due to the archaeological dataset.
+
+\xavi{No se entiende; deberías explicarlo un poco mejor y citar algo si dices eso. Piensa que la mayoría de lectores no sabrán quién es Doña Mencia}.
 
@@
 
-The distribution of amphora stamps for each \textit{conventus} can be seen in Figure~\ref{frequency}.
-The majority of stamps are concentrated in \textit{Hispalis} (574 stamps) while \textit{Corduba} and \textit{Astigi} have roughly half this sample size (267 and 146 stamps).
-The workshops of the three conventus show a similar pattern on the stamp frequency distribution with the exceptions of 2 large workshops in \textit{Hispalis} (La Catria and Arva).
-On these two workshops a comparatively large amount of stamps were found (29 different code stamps on each of them)\xavi{una cosa, aquí has pasado de sellos a códigos, y los valores son distintos que en el gráfico anterior.
-Deberías clarificarlo antes de empezar a hablar de la figura}.
+The distribution of code stamps for each \textit{conventus} can be seen in Figure~\ref{frequency}.
+The majority of stamps are concentrated in \textit{Hispalis} (574 code stamps) while \textit{Corduba} and \textit{Astigi} have roughly half this sample size (267 and 146 code stamps).
+The workshops of the three \textit{conventus} show a similar pattern on the stamp frequency distribution with the exceptions of 2 large workshops in \textit{Hispalis} (La Catria and Arva).
+On these two workshops a comparatively large amount of code stamps were found (29 different code stamps on each of them)
+
+%\xavi{una cosa, aquí has pasado de sellos a códigos, y los valores son distintos que en el gráfico anterior. Deberías clarificarlo antes de empezar a hablar de la figura}.
+
 According to previous studies those workshops became the most important centres of amphora production in the region\xavi{cita de estos estudios}.
@@
 To apply Morisita-Horn we firstly calculated the number of times that each stamp code appeared in an amphora workshop.
-This method allowed us to bear in mind a similar number of times for each repeated stamp per workshop\xavi{no entiendo esta frase}.
+This method allowed us to bear in mind the number of frequency of stamps for each workshop. 
+%This method allowed us to bear in mind a similar number of times for each repeated stamp per workshop\xavi{no entiendo esta frase}.
+
 If two workshops had similar stamp code distribution then the distance index would be close to 0 whereas sites with completely unrelated stamp codes would give a distance close to 1.
@@
 For this reason, an index of dissimilarity was used to detect differences between the distribution of the amphoric stamps and the spatial distance of producers and consumption centres.
-The general purpose was to explore if such differences found in the stamp codes could play an important role in the Roman market\xavi{este párrafo es un poco reiterativo no?}.
+
+%The general purpose was to explore if such differences found in the stamp codes could play an important role in the Roman market\xavi{este párrafo es un poco reiterativo no?}.
 
@@
 
+Alternatively, the presence of different stamps in the same workshop could imply some kind of organisation from within the workshop only affecting nearby centres \citep{juanmorostesis}.
+The stamped amphorae could have allowed potters to organise the different production batches for posterior commercialisation.
+%\xavi{estos 2 párrafos igual deberían ir a la discusión no? La conclusión debería ser algo más general y aquí estás aún interpretando resultados}.
+
+This could be explained such as batch systematic organisation between potters \citep{juanmorostesis}.
+This method allowed to potters organises the production with the batch stamping for the posterior commercialisation\xavi{y esto para qué serviría?
+debería explicarse un poco más}.
+
 It is worth mentioning that the performed analysis could be limited by a diversity of biases.
@@
 In \textit{Britannia} the majority of similarity stamps were mainly found in military centres.
-This could be also interpreted by an intensity bias where military centres have been mostly excavated than civil centres\xavi{pero no decías que el aceite llegaba por las legione?
-Entonces por qué es un bias?
-En todo caso ponlo al final de la sección y no al inicio}.
 
@@
 
-.The areas mostly militarised share similar stamps than civil areas. However, we do not detect a concrete pattern regarding the distribution of the stamps in the German limes \citep{xanten2018}.
+The areas mostly militarised share similar stamps than civil areas.
+However, we do not detect a concrete pattern regarding the distribution of the stamps in the German limes \citep{xanten2018}.
+This could be also interpreted by an intensity bias where military centres have been mostly excavated than civil centres
 
-It is worth mentioning that it was not detected a robust model of organisation in both cases with regard to the distribution of stamps in consumption centres\xavi{de nuevo esto es redundante no?}.
+%\xavi{pero no decías que el aceite llegaba por las legione? Entonces por qué es un bias? En todo caso ponlo al final de la sección y no al inicio}.
+
+%It is worth mentioning that it was not detected a robust model of organisation in both cases with regard to the distribution of stamps in consumption centres\xavi{de nuevo esto es redundante no?}.
+
 This evidence means that it is unknown whether some production centres went to one province or another or, at least, that they can be clearly reflected in the data with a greater similarity in the amphoric stamps.
@@
 
-Alternatively, the presence of different stamps in the same workshop could imply some kind of organisation from within the workshop only affecting nearby centres \citep{juanmorostesis}.
+%Alternatively, the presence of different stamps in the same workshop could imply some kind of organisation from within the workshop only affecting nearby centres \citep{juanmorostesis}.
 The stamped amphorae could have allowed potters to organise the different production batches for posterior commercialisation\xavi{estos 2 párrafos igual deberían ir a la discusión no?
 La conclusión debería ser algo más general y aquí estás aún interpretando resultados}.
-
-This could be explained such as batch systematic organisation between potters \citep{juanmorostesis}.
-This method allowed to potters organises the production with the batch stamping for the posterior commercialisation\xavi{y esto para qué serviría?
-debería explicarse un poco más}.
 

### 2020-07-15 Paper/jaccardbaetica.tex
add new changes xavi review

@@
 The production and distribution of olive oil increased from the first to the third centuries AD ~\citep{remesal_concierto}.
+
 Hundreds of presses and amphora-making workshops were build near large extensions covered by olive trees.
@@
 
-A large majority of this Baetican amphorae production is composed by Dressel 20 types\citep{dressel_ricerche_1878,millet_anforas_1998}.
+The importance of this olive oil trade is revealed by the massive amount of Dressel 20 type amphorae found across the Roman provinces \citep{dressel_ricerche_1878,
+millet_anforas_1998}.
+This amphora has been commonly associated with the transportation of Baetican olive oil for supplying military camps and civil settlements \citep{berni_millet_epigrafianforica_2008}.%\xavi{me parece repetitivo, porque unos párrafos antes ya hablabas de lo mismo.
+Quizás unificar texto aquí o allá?}.
+
 Dressel 20 amphorae have a recognisable globular form with short neck and oval-shaped handles.
 Its capacity allowed to hold 70-75 litres of olive oil to be shipped and distributed throughout Empire \citep{berni_dressel_2016}.
+
+%A large majority of this Baetican amphorae production is composed by Dressel 20 types\citep{dressel_ricerche_1878,
+%millet_anforas_1998}.
 
@@
 
-%\xavi{este título es muy raro, porque 1. que son casos de estudio ya lo sabes y 2. case study singular, consumption centres plural}
-
 The creation of new provinces required the Roman Empire the arrival of resources through Mediterranean and Atlantic routes.
@@
 
-Some authors argue elsewhere that this same \textit{Annona} could also have organized the supply of additional goods such as olive oil to the Roman legions \citep{remesal_annona_1986,remesal_annona_1990}\xavi{pero no hay evidencia?
-Las fuentes escritas no lo mencionan?}.
+Some hypotheses argue elsewhere that this same \textit{Annona} could also have organized the supply of additional goods such as olive oil to the Roman legions \citep{remesal_annona_1986,remesal_annona_1990}
 
-The importance of this olive oil trade is revealed by the massive amount of Dressel 20 amphorae found across the Roman provinces.
+%\xavi{pero no hay evidencia? Las fuentes escritas no lo mencionan? - He cambiado authors por hipotesis porque me quería referir a investigadores zorry}.
+
+%The importance of this olive oil trade is revealed by the massive amount of Dressel 20 amphorae found across the Roman provinces.
 This amphora has been commonly associated with the transportation of Baetican olive oil for supplying military camps and civil settlements \citep{berni_millet_epigrafianforica_2008}.\xavi{me parece repetitivo, porque unos párrafos antes ya hablabas de lo mismo.
-Quizás unificar texto aquí o allá?}
-
-The military consumption of olive oil in Roman provinces may be related to two aspects: a) cultural consumption whose product is consumed by cultural reasons such as to identity or habit and b) economical reasons where olive oil distribution is based on the transport costs, unlike other products \citep[69-70]{carreras_britannia_1998}\xavi{no pillo esta frase}..
+Quizás unificar texto aquí o allá?} LO HE PUESTO MÁS ARRIBA!
 
 Olive oil supply seems to be particularly intense in militarised provinces that functioned as the borders of the Empire such as \textit{Britannia} and \textit{Germania}.
-The high percentage of Dressel 20 amphorae found in sites nearby military garrisons suggests the existence of diverse supply dynamics as there were trade processes flowing good from military sites to civil settlements\citep{remesal_annona_1986, carreras_britannia_1998}\xavi{lo que quieres decir es que tienes muchas ánforas en sitios civiles cerca de militares?
-no queda claro}.
+The high percentage of Dressel 20 amphorae found in sites nearby military garrisons suggests the existence of diverse supply dynamics as there were trade processes flowing from military sites to civil settlements\citep{remesal_annona_1986, carreras_britannia_1998}
+\xavi{lo que quieres decir es que tienes muchas ánforas en sitios civiles cerca de militares? no queda claro}.
+\maria{sí, pero creo que yo no he escrito está frase xD}
+
+An explanation for the interest to export olive oil for the military garrison in the Roman provinces is still unclear.
+The origin of this exportation may be related to two aspects: a) cultural consumption whose olive oil is used by cultural reasons such as group identification or by habit and b) economical and political reasons to create a redistributive system mechanism to supply military camps.
+
+%may be related to two aspects according with some hypotheses: a) cultural consumption whose product is consumed by cultural reasons such as to identity or habit and b) economical reasons where olive oil distribution is based on the transport costs, unlike other products \citep[69-70]{carreras_britannia_1998}\xavi{no pillo esta frase}..
 
@@
 Local population did not consume this product and it was not produced on the region as the landscape and climate of the British isles was not suitable for olive oil trees\citep[161]{monfort_britanniaen_1998}.
-Based on the archaeological record this absence of olive oil changed with the arrival of a large amount of Dressel 20 amphorae to \textit{Britannia} \citep[1]{carreras_britannia_1998}\xavi{en qué siglo?}.
+Based on the archaeological records, this absence of olive oil changed with the arrival of a large amount of Dressel 20 amphorae to \textit{Britannia} from the first century A.D \citep{peacock_amphorae_1991,
+carreras_britannia_1998}.
+At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}
+%\xavi{en qué siglo?}
+
+The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges.
 The demand of olive oil after the Roman conquest required of a new trade route due to the lack of local olive oil production; \textit{Baetica} was chosen as its main supply.
 
-At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}\xavi{repetitivo?
+%At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}\xavi{repetitivo?
 fusionar párrafos}.
+
 This fact will have a particular intensity within sites close to the Hadrian Wall's garrisons.
 Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic locations~\citep{carreras_atlantic_2012}.
-The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges.
+
+%The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges.
 Thus, the network was mainly focused on the support of soldiers during military campaigns\xavi{estas 2 frases repiten cosas ya explicadas}.
@@
 
-According to previous studies those workshops became the most important centres of amphora production in the region\xavi{cita de estos estudios}.
-It is worth mentioning that a majority of these stamps were collected during ield surveys with no excavation, so this difference in sample size could be biased due to differences in intensity across the different sites~\citep{arva_1997}.
+According to previous studies those workshops became the most important centres of amphora production in the region \citep{rodriguez_economioleicola_1977,
+arva_1997}.
+%\xavi{cita de estos estudios}
+
+It is worth mentioning that a majority of these stamps were collected during field surveys with no excavation, so this difference in sample size could be biased due to differences in intensity across the different sites~\citep{arva_1997}.
 
@@
 The analysis of stamps in consumption areas also used the CEIPAC database and filtered in the same way than the previous dataset.
-We selected the centres with more or equal than five stamps\xavi{pero esto no se hizo en el de Baetica no?
-Por qué?}.
+However, in this case, we selected the centres with more or equal than five stamps in order to maximize the analysis with more stamps in the dataset.
+
+%\xavi{pero esto no se hizo en el de Baetica no? Por qué?}.
 

### 2020-07-22 Paper/jaccardbaetica.tex
add new changes xavi

@@
 Has hecho un salto aquí a los ciudadanos}.
+\maria{Pero porque la annona estaba enfocada a los ciudadanos romanos y luego hablo después que hay otras hipótesis que dicen que iba más allá a los militares}
 
-Some hypotheses argue elsewhere that this same \textit{Annona} could also have organized the supply of additional goods such as olive oil to the Roman legions \citep{remesal_annona_1986,remesal_annona_1990}
+Some hypotheses argue elsewhere that this same \textit{Annona} could also have organized the supply of additional goods such as olive oil to the Roman legions \citep{remesal_annona_1986,
+remesal_annona_1990}
 
-%\xavi{pero no hay evidencia? Las fuentes escritas no lo mencionan? - He cambiado authors por hipotesis porque me quería referir a investigadores zorry}.
+\xavi{pero no hay evidencia? Las fuentes escritas no lo mencionan?}
+\maria{He cambiado authors por hipotesis porque me quería referir a investigadores}.
 
@@
 
-Roman studies exploring the presence of Dressel 20 amphorae in \textit{Germania} has not had the same impact as the rest of the province\xavi{no entiendo esta frase.
-Que qué provincia?
-Germania son 2 provincias no?}.
+Roman studies exploring the presence of Dressel 20 amphorae in the provinces Inferior and Superior have not had the same impact as the rest of the provinces.
+%\xavi{no entiendo esta frase. Que qué provincia? Germania son 2 provincias no?}
+
 This could be explained due to the lack of archaeological sources \citep{horacio2010llegada}.
@@
 Además, falta alguna cita y desarrollar esta idea}.
-However, this hypothesis is hard to assess given the current lack of archaeological evidence \xavi{related to BLABLABLA.
-Habría que detallar un poco porque hay un montón de evidencia arqueológica del limes, así que qué es exactamente lo que falta?}\citep[125]{remesal_concierto}.
+However, this hypothesis is hard to assess given the current lack of archaeological evidence related to
+\xavi{related to BLABLABLA. Habría que detallar un poco porque hay un montón de evidencia arqueológica del limes, así que qué es exactamente lo que falta?}\citep[125]{remesal_concierto}.
 
@@
 This would allow us to study the distribution of the stamps across centres as can be seen in Figure~\ref{stamps}.
-Most workshops contained only one stamp~\xavi{el gráfico es dificil de entender...yo leo que hay unos 20 workshops con 1 solo stamp, por lo que entiendo que no es cierto que la mayoría solo tengan 1 sello (hay como 50 workshops con entre 2 y 30 sellos no?)} One workshop (La Catria) concentrated a large percentage of the amphoric stamps with a total number of 228 stamps.
+Most workshops concentrated few stamps while few workshops contained more than 100 stamps.
+For instance, one workshop (La Catria) concentrated a large percentage of the amphoric stamps with a total number of 228 stamps.
+
+%\xavi{el gráfico es dificil de entender...yo leo que hay unos 20 workshops con 1 solo stamp, por lo que entiendo que no es cierto que la mayoría solo tengan 1 sello (hay como 50 workshops con entre 2 y 30 sellos no?)}
+\maria{he cambiado la frase a ver si se entiende mejor}
+
 The type of distribution is also frequent in amphora production where we observed a self-organised complex system pattern with a major concentration of the number of stamps in a few workshops \citep{bayesian_2018,coto-sarmiento_identifying_2018}\xavi{no lo veo; recuerda que para el rollo bayesiano tenías una línea recta al hacedr el log de los 2 ejes, no solo 1, y este plot no está construído así no?}.
-
-%The type of distribution is also frequent in the Roman economy where we observed a self-organised complex system pattern typical of a free market
-
-%\xavi{que me quede claro...la mayoría de workshops solo tienen 1 sello??? Y hay un workshop con como 700 sellos distintos??? }
+\maria{son dos ejes, no? ¿aquí no se ve línea recta? Lo mismo si lo hago de puntitos...}
 
@@
 
-By contrast, most sites did not show a strong similarity in stamp correlation\xavi{qué quieres decir?
+%By contrast, most sites did not show a strong similarity in stamp correlation\xavi{qué quieres decir?
 porque lo de que hay una similaridad baja ya lo has repetido arriba}.
@@
 que no hay sellos similares en un solo sitio...?}.
-
-These results suggest that there is no clear organizational pattern linking production centres and/or consumption centres beyond spatial distance.
-As a consequence, our analysis does not indicate the presence of a specific production centre from \textit{Baetica} to \textit{Britannia} province.
-Rather, it would imply that olive oil production was distributed by non-specific production centres\xavi{esto no iría en la discusión?
-Por otra parte, no entiendo qué quieres decir exactamente, igual hay que desarrollarlo}.
 
@@
 
-A higher concentration of sites sharing similar stamps were found in areas eminently militarised and close to German \textit{limes}, even if most sites mostly showed different stamps~\xavi{quieres decir que los sitios más cercanos al limes son más similares?
-no queda claro y hay que detallarlo un poco más}.
+A higher concentration of sites sharing similar stamps were found in areas eminently militarised and close to German \textit{limes}, although most sites mostly showed different stamps. 
+\maria{he cambiado esta frase}
+%even if most sites mostly showed different stamps~\xavi{quieres decir que los sitios más cercanos al limes son más similares? no queda claro y hay que detallarlo un poco más}.
 
@@
 Despite these stamps tend to share the same area of production, we do not identify groups of any more than three workshops sharing the same amphoric stamps as the dendrogram showed.
-In general, the majority of stamps were located in different amphora workshops\xavi{estas 2 frases repiten cosas dichas antes}.
 
-Our results do not support previous working hypotheses suggesting that groups of workshopes used specific groups of stamp codes; while there is higher stamp similarity for nearby workshops this result does not suggest any kind of organisation beyond the simple fact that nearby workshops had a higher degree of interaction.
+%In general, the majority of stamps were located in different amphora workshops\xavi{estas 2 frases repiten cosas dichas antes}. 
+\maria{he quitado la última frase pero la anterior no se repite}
+
+Our results do not support previous working hypotheses suggesting that groups of workshops used specific groups of stamp codes; while there is higher stamp similarity for nearby workshops this result does not suggest any kind of organisation beyond the simple fact that nearby workshops had a higher degree of interaction.
 In fact; the opposite seems true: each workshop used different stamp codes and only shared some of them with the closest ones.
-These stamp codes detected in closer workshops do not move from other distant workshops.
+
+%These stamp codes detected in closer workshops do not move from other distant workshops.
 In other words, the stamps tend to remain in the same area and different stamps were located in a same amphora workshop\xavi{de nuevo estas 2 frases repiten cosas de antes}.
@@
 It is more probable than closer workshops tend to share more traits than distant workshops.
-While the role of the river was significant for the distribution of amphorae, river connectivity amongst workshops does not seem to show relevancy for the distribution of stamps\xavi{qué quieres decir con esto?
-que los diferentes ríos no muestran ningún patrón?}.
+While the role of the river was significant for the distribution of amphorae, river connectivity amongst workshops does not seem to show relevancy for the distribution or concentration of stamps in a specific river area.
+\xavi{qué quieres decir con esto? que los diferentes ríos no muestran ningún patrón?}
+\maria{he añadido cosas a la frase para ver si se entiende}
 
@@
 This could be explained such as batch systematic organisation between potters \citep{juanmorostesis}.
-This method allowed to potters organises the production with the batch stamping for the posterior commercialisation\xavi{y esto para qué serviría?
-debería explicarse un poco más}.
+This work method could have allowed to potters to stamp amphorae in order to identify the production for batches for posterior commercialization.
+High production of amphorae had to maintain a specialize system organization to control this production \citep[104]{juanmorostesis}.
+
+%This method allowed to potters organises the production with the batch stamping for the posterior commercialisation
+%\xavi{y esto para qué serviría? debería explicarse un poco más}.
 
@@
 A combined exploration of the map and the dendrogram also suggests an interesting pattern: centres with higher similarity values are closer to the coast (North Sea and Celtic Sea).
-This may indicate that the Atlantic route could have played an essential role in transporting olive oil to the area of \textit{Britannia}, since in the places where there is greater similarity they are found in different strategic points near the sea\xavi{por qué más conectividad implica más similaridad?
-no será porque encuentras más ánforas?}.
+This may indicate that the Atlantic route could have played an essential role in transporting olive oil to the area of \textit{Britannia}, since in the places where there is greater similarity they are found in different strategic points near the sea
+\xavi{por qué más conectividad implica más similaridad? no será porque encuentras más ánforas?}
+\maria{no entiendo este comentario :-(}.
 
@@
 %It is worth mentioning that it was not detected a robust model of organisation in both cases with regard to the distribution of stamps in consumption centres\xavi{de nuevo esto es redundante no?}.
+
+Finally, the results suggest that there is no clear organizational pattern linking production centres and/or consumption centres beyond the spatial distance.
+As a consequence, our analysis does not indicate the presence of a specific workshop from \textit{Baetica} supplying to some specific province.
+
+\xavi{esto no iría en la discusión? Por otra parte, no entiendo qué quieres decir exactamente, igual hay que desarrollarlo}.
+\maria{lo he incluido aquí y lo he cambiado a ver si se entiende}
+%Rather, it would imply that olive oil production was distributed from non-specific production centres.
 

### 2020-07-23 Paper/jaccardbaetica.tex
add new changes xavi

@@
 
-Some hypotheses argue elsewhere that this same \textit{Annona} could also have organized the supply of additional goods such as olive oil to the Roman legions \citep{remesal_annona_1986,
+However, some hypotheses argue elsewhere that this same \textit{praefectura annonae} could also have organized the supply of additional goods such as olive oil to the Roman legions  \citep{remesal_annona_1986,
 remesal_annona_1990}
@@
 The origin of this exportation may be related to two aspects: a) cultural consumption whose olive oil is used by cultural reasons such as group identification or by habit and b) economical and political reasons to create a redistributive system mechanism to supply military camps.
+\maria{esta frase está explicada mejor}
 
@@
 
-Julius Caesar's campaigns in \textit{Gallia} during 1st BC opened the door to the invasion of \textit{Germania} during Augustus rule~\citep{remesal_annona_1986,remesal_baetica_2002}\xavi{no hay un resumen de la conquista similar para Britannia, queda raro}.
-Previous studies suggested that the supply to the German limes was mainly based on riverine transport, but recent works suggest that the maritime route through the Atlantic Ocean could have been more important than expected (both for Germania and Britannia).~\citep{remesal_germn_2010,rubio-campillo_ecology_2018}.\xavi{lo del océano atlántico está un poco descontextualizado; no sería mejor ponerlo al final de esta sección?}
+%Julius Caesar's campaigns in \textit{Gallia} during 1st BC opened the door to the invasion of \textit{Germania} during Augustus rule~\citep{remesal_annona_1986,remesal_baetica_2002}\xavi{no hay un resumen de la conquista similar para Britannia, queda raro}.
+\maria{lo he eliminado}
 
-Roman studies exploring the presence of Dressel 20 amphorae in the provinces Inferior and Superior have not had the same impact as the rest of the provinces.
+Roman studies exploring the presence of Dressel 20 amphorae in the provinces Inferior and Superior during Roman Empire have not had the same impact as the rest of the provinces \citep[293]{remesal_baetica_2002}.
 %\xavi{no entiendo esta frase. Que qué provincia? Germania son 2 provincias no?}
 
-This could be explained due to the lack of archaeological sources \citep{horacio2010llegada}.
-As a consequence, it is still unknown if trade agents participated on the distribution of Baetican olive oil in \textit{Germania} or the supply was exclusively focused on the Roman army garrisons\citep[156]{remesal_germn_2010}\xavi{quieres decir que aquí sólo encontramos dressel 20 en yacimientos militares porque no se han excavado muchos civiles?}.
+The massive importation of Dressel 20 in both \textit{Germania} provinces could be explained due to the archaeological sources found mostly in German \textit{limes}. 
+By contrast, it is still unknown if trade agents participated on the distribution of Baetican olive oil in \textit{Germania} or the supply was exclusively focused on the Roman army garrisons\citep[156]{remesal_germn_2010}
+\xavi{quieres decir que aquí sólo encontramos dressel 20 en yacimientos militares porque no se han excavado muchos civiles?}.
+\maria{lo he mejorado a ver si se entiende}
+
+%This could be explained due to the lack of archaeological sources \citep{horacio2010llegada}
 
@@
 
-The presence of the Roman army encouraged the exchange in the province as shown by the arrival of this product both civil settlement and military sites with a mayor concentration at the German \textit{limes}\xavi{pero justo antes decías que no había en civiles no?}.
-It seems that some Baetican centres would be assigned to the support of olive oil\xavi{the support of olive oil no se entiende.
-Además, falta alguna cita y desarrollar esta idea}.
-However, this hypothesis is hard to assess given the current lack of archaeological evidence related to
-\xavi{related to BLABLABLA. Habría que detallar un poco porque hay un montón de evidencia arqueológica del limes, así que qué es exactamente lo que falta?}\citep[125]{remesal_concierto}.
+The presence of the Roman army encouraged the exchange in the province as shown by the arrival of this product both civil settlement and military sites with a mayor concentration at the German \textit{limes}.
+
+\xavi{pero justo antes decías que no había en civiles no?}. 
+\maria{yo no he dicho eso :-[}
+
+It seems that some Baetican centres could have been assigned to a specific province for the olive oil supply to military garrisons \citep[125]{remesal_concierto}. 
+\xavi{the support of olive oil no se entiende. Además, falta alguna cita y desarrollar esta idea}. 
+\maria{mejorado, creo}
+However, this hypothesis is hard to assess given the current lack of written evidence related to the meaning of the amphoric stamps. 
+\xavi{related to BLABLABLA. Habría que detallar un poco porque hay un montón de evidencia arqueológica del limes, así que qué es exactamente lo que falta?}. 
+\maria{mejorado,creo}
+
+Previous studies suggested that the supply to the German limes was mainly based on riverine transport, but recent works suggest that the maritime route through the Atlantic Ocean could have been more important than expected (both for Germania and Britannia).~\citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
+%\xavi{lo del océano atlántico está un poco descontextualizado; no sería mejor ponerlo al final de esta sección?}
 

### 2020-07-27 Paper/jaccardbaetica.tex
add new comments xavi almost

@@
 \xavi{reescribir abstract, demasiado general y entonces demasiado específico}
+\maria{lo he reeformulado a ver si ahora..}
+
 The aim of this study is to explore economics dynamics in the production and distribution of olive oil trade. 
-Our case of study has been focused on the production processes located in \textit{Baetica} province (currently Andalusia) from 1st to 3rd AD.
+Baetica became an important production and distribution centre during the Roman Empire.
+However, it remains under debate about how this province was organised and whether it could be possible to identify patterns in the olive oil market.
+
+Our case of study has been focused on the production processes located in Baetica province (currently Andalusia) from 1st to 3rd AD.
+In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps.
+Amphoric stamps are used to identify the presence of different groups that might share similar stamps.
+To achieve this goal, we analyse a set of stamps from two centres: 1) production centres by analysing different workshops in Baetica province and 2) two Roman provinces such as \textit{Germania} (Inferior and Superior) and \textit{Britannia} as consumption centres.
+They will be used to detect a connection between the distribution of amphoric stamps and the economic structure in both centres.
+Here, we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance.
+The analysis explores how the quantitative approach provides a useful tool for the interpretation of the economic processes.
+Finally, results pretend to highlight the organization of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.
+
+%Our case of study has been focused on the production processes located in \textit{Baetica} province (currently Andalusia) from 1st to 3rd AD.
 In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps.
@@
 Here, we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance.
-
 The analysis explores how the quantitative approach provides a useful tool for the interpretation of the economic processes.
@@
 They have been frequently dated with the consular dating by studying \textit{tituli picti} found in \textit{Monte Testaccio} \citep{Testaccio1, berni_millet_epigrafianforica_2008}.
-However, the chronology of stamps could be also biased in other sites taking into account that amphorae were deposited after being produced and consumed\xavi{esto necesita un poco de desarrollo.
+Despite \textit{tituli picti} contain important information about chronology, they have been still found in few quantities mostly concentrated in a only settlement such as \textit{Monte Testaccio}.
+In addition, the chronology of stamps could be also biased taking into account that amphorae were deposited after being produced and consumed\xavi{esto necesita un poco de desarrollo.
 quieres decir que el código se encuentra en contextos de cronologías diferentes no?
 Y lo del Testaccio, pos eso, que hay pocos y todos de un yacimioento, lo que añade un bias}.
+\maria{desarrollado}
 
+The use of amphoric stamps can be a good proxy to explore the system of olive oil production and trade developed in Baetica province.
 These stamps are typically found on handles but they were also imprinted on rims and the amphora body~\citep{millet_anforas_1998}.
@@
 On the other hand, they could also belong to the owner of the workshops were the amphorae was made or even to a group of amphora workers~\citep{berni_millet_epigrafianforica_2008}.
-In any case, the use of these stamps is a good proxy to explore the system of olive oil production and trade developed in Baetica.
+%In any case, the use of these stamps is a good proxy to explore the system of olive oil production and trade developed in Baetica.
 
-Nevertheless, some challenges\xavi{acabas de decir que tampoco se sabe qué es el sello, así que esta frase queda rara} remain under discussion such as how this production was organised and whether it is possible to distinguish production patterns in the olive oil trade.
+\xavi{acabas de decir que tampoco se sabe qué es el sello, así que esta frase queda rara} 
+\maria{he cambiado la frase de sitio para que se contextualice mejor}
+Nevertheless, some challenges remain under discussion such as how this production was organised and whether it is possible to distinguish production patterns in the olive oil trade.
 Our questions will be focused on the distribution of amphoric stamps.
@@
 
-Neither the use of written records have allowed providing enough information that can explain the economic role of \textit{Baetica} province in the Roman organisation.\xavi{esta frase esta fuera de contexto.
+%Neither the use of written records have allowed providing enough information that can explain the economic role of \textit{Baetica} province in the Roman organisation.\xavi{esta frase esta fuera de contexto.
 Si la pones entonces deberías expandir la idea de frase a párrafo, y poner referencias}.
+\maria{he eliminado la última frase}
 
@@
 
+The Roman conquest in \textit{Britannia} and \textit{Germania} was a new opening to foreign trade in order to supply the military campaign.
+A high demand for olive oil in both provinces stimulated an important trade network involving different agents and the Empire itself.
+Baetica became a strategic province for the production and distribution in both provinces since it can be detected by the high presence of amphoric stamps found in civil and military areas during this period.
+
 \xavi{faltaría un enlace a la siguiente sección, queda muy abrupto}
+\maria{he añadido varios párrafos a ver si se enlaza}
 
@@
 
-It is important to mention that some workshops exhibited the same coordinates but they were catalogued as different workshops in Roman Studies (i.e.
+It is important to mention that some workshops exhibited the same geographical coordinates while they were catalogued differently as workshops.
+For instance, two nearby amphora workshops such as Tesorillo de Doña Mencia and Doña Mencia were catalogued as a different name in spite of having the same geographical coordinates.
+Our work considered this particularity in the archaeological dataset but some biases can occur in the analysis.
+
+%It is important to mention that some workshops exhibited the same coordinates but they were catalogued as different workshops in Roman Studies (i.e.
 nearby amphora workshops in \textit{Baetica} named as Tesorillo de Doña Mencia and Doña Mencia).
@@
 \xavi{No se entiende; deberías explicarlo un poco mejor y citar algo si dices eso. Piensa que la mayoría de lectores no sabrán quién es Doña Mencia}. 
+\maria{he reformulado la frase a ver si ahora se entiende}
 
@@
 
-This research was partly funded by the European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros and Ignacio Morer for their useful suggestions.  
+This research was partly funded by the European Research Council Advanced Grant EPNet (340828).
+We are grateful to Simon Carrignon, Juan Moros, Víctor M.
+Lozano and Ignacio Morer for their useful suggestions.
 All data has been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}.

### 2020-07-29 Paper/jaccardbaetica.tex
add parrafo intro corr xavi

@@
 rubio-campillo_ecology_2018}.
+
+This work focuses on the debate about the dynamics of the distribution of olive oil market in Roman provinces.
+Amphoric production in Baetica province will help us to identify links between production centres and consumption centres in order to detect an economic pattern in the commerce by analysing stamps amphorae finding in the provinces.
+
+The finding of amphoric stamps in different provinces has allowed us to know the importance of the distribution of olive oil within the Empire.
+Despite having archaeological evidence, no consensus has been reached: How was the economical relation between production and consumption centers?
+Did they follow a economical pattern to distribute olive oil in those provinces?
 
@@
 This method was applied to measure the dissimilarity between different samples of sets.
-Generally, it describes the dissimilarity between the system of two communities based on the idea of inverse correlation between diversity and species \citep{magurran_why_1988}\xavi{citar reforzar más esta idea}.
+Generally, it describes the dissimilarity between the system of two communities based on the idea of inverse correlation between diversity and species \citep{magurran_why_1988}.
+In our case, stamps from different sites will be used as a proxy to measure their dissimilarity by analyzing their code.
+\xavi{citar reforzar más esta idea}.
+\maria{he añadido unas líneas a ver si están bien o no xD}
 

### 2020-08-03 Paper/jaccardbaetica.tex
add new changes xavi

@@
 
-The aim of this study is to explore economics dynamics in the production and distribution of olive oil trade. 
-Baetica became an important production and distribution centre during the Roman Empire.
+The aim of this study is to explore economics dynamics in the production and distribution of olive oil trade in \textit{Baetica} province (currently Andalusia).
+\textit{Baetica} became an important production and distribution centre during the Roman Empire.
 However, it remains under debate about how this province was organised and whether it could be possible to identify patterns in the olive oil market.
 
-Our case of study has been focused on the production processes located in Baetica province (currently Andalusia) from 1st to 3rd AD.
+Our case of study has been focused on the amphoric production from 1st to 3rd AD.
 In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps.
 Amphoric stamps are used to identify the presence of different groups that might share similar stamps.
-To achieve this goal, we analyse a set of stamps from two centres: 1) production centres by analysing different workshops in Baetica province and 2) two Roman provinces such as \textit{Germania} (Inferior and Superior) and \textit{Britannia} as consumption centres.
+To achieve this goal, we analyse a set of stamps from two centres: 1) production centres by analysing different workshops in \textit{Baetica} province and 2) two Roman provinces such as \textit{Germania} (Inferior and Superior) and \textit{Britannia} as consumption centres.
 They will be used to detect a connection between the distribution of amphoric stamps and the economic structure in both centres.
@@
 This work focuses on the debate about the dynamics of the distribution of olive oil market in Roman provinces.
-Amphoric production in Baetica province will help us to identify links between production centres and consumption centres in order to detect an economic pattern in the commerce by analysing stamps amphorae finding in the provinces.
+Amphoric production in \textit{Baetica} province will help us to identify links between production centres and consumption centres in order to detect an economic pattern in the commerce by analysing stamps amphorae finding in the provinces.
 
@@
 
-Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}) (see Figure~\ref{general}.
+Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}) (see Figure~\ref{general}).
 
@@
 \caption{Overview of the sites analyzed in this paper.
-The color defines the three different regions under study (Baetica as producer; Germania and Britannia as consumers) while the size of each dot highlights the number of Dressel-20 amphoric stamps found on each of these sites}
+The color defines the three different regions under study (\textit{Baetica} as producer; \textit{Germania} and \textit{Britannia} as consumers) while the size of each dot highlights the number of Dressel-20 amphoric stamps found on each of these sites}
 
@@
 
-This strategic location allowed the transport of olive oil through riverine shipping towards the maritime routes that connected Baetica with Mediterranean and Atlantic trade routes towards the rest of the Empire\citep{garcia_vargas_enrique_formal_2010}.
+This strategic location allowed the transport of olive oil through riverine shipping towards the maritime routes that connected \textit{Baetica} with Mediterranean and Atlantic trade routes towards the rest of the Empire\citep{garcia_vargas_enrique_formal_2010}.
 
@@
 
-The use of amphoric stamps can be a good proxy to explore the system of olive oil production and trade developed in Baetica province.
+The use of amphoric stamps can be a good proxy to explore the system of olive oil production and trade developed in \textit{Baetica} province.
 These stamps are typically found on handles but they were also imprinted on rims and the amphora body~\citep{millet_anforas_1998}.
@@
 Olive oil supply seems to be particularly intense in militarised provinces that functioned as the borders of the Empire such as \textit{Britannia} and \textit{Germania}.
-The high percentage of Dressel 20 amphorae found in sites nearby military garrisons suggests the existence of diverse supply dynamics as there were trade processes flowing from military sites to civil settlements\citep{remesal_annona_1986, carreras_britannia_1998}
+The high percentage of Dressel 20 amphorae found in sites nearby military garrisons suggests the existence of diverse supply dynamics as there were trade processes flowing from military sites to civil settlements~\citep{remesal_annona_1986, carreras_britannia_1998}
 \xavi{lo que quieres decir es que tienes muchas ánforas en sitios civiles cerca de militares? no queda claro}.
@@
 A high demand for olive oil in both provinces stimulated an important trade network involving different agents and the Empire itself.
-Baetica became a strategic province for the production and distribution in both provinces since it can be detected by the high presence of amphoric stamps found in civil and military areas during this period.
+\textit{Baetica} became a strategic province for the production and distribution in both provinces since it can be detected by the high presence of amphoric stamps found in civil and military areas during this period.
 
@@
 
-Previous studies suggested that the supply to the German limes was mainly based on riverine transport, but recent works suggest that the maritime route through the Atlantic Ocean could have been more important than expected (both for Germania and Britannia).~\citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
+Previous studies suggested that the supply to the German limes was mainly based on riverine transport, but recent works suggest that the maritime route through the Atlantic Ocean could have been more important than expected (both for \textit{Germania} and \textit{Britannia}).~\citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
 %\xavi{lo del océano atlántico está un poco descontextualizado; no sería mejor ponerlo al final de esta sección?}
@@
 This means that we included in the sample some centres that are actually located beyond the administrative boundaries but were considered part of the same border.
-Specifically, for the German \textit{limes} we included sites from the two provinces (Superior and Inferior) while the analysis of Britannia extended to \textit{Caledonia}.
+Specifically, for the German \textit{limes} we included sites from the two provinces (Superior and Inferior) while the analysis of \textit{Britannia} extended to \textit{Caledonia}.
 
@@
 \includegraphics[width=\linewidth]{figs/britannia}
-\caption{Sites in Britannia where Dressel-20 amphoric stamps have been found.
+\caption{Sites in \textit{Britannia} where Dressel-20 amphoric stamps have been found.
 Each dot is a site while its size shows the sample size of stamps found in the site.

### 2020-08-05 Paper/jaccardbaetica.tex
eng correct intro

@@
 
-Our case of study has been focused on the amphoric production from 1st to 3rd AD.
+Our case of study has been focused on amphoric production from 1st to 3rd AD.
 In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps.
 Amphoric stamps are used to identify the presence of different groups that might share similar stamps.
+
 To achieve this goal, we analyse a set of stamps from two centres: 1) production centres by analysing different workshops in \textit{Baetica} province and 2) two Roman provinces such as \textit{Germania} (Inferior and Superior) and \textit{Britannia} as consumption centres.
@@
 The analysis explores how the quantitative approach provides a useful tool for the interpretation of the economic processes.
-Finally, results pretend to highlight the organization of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.
+Finally, results pretend to highlight the organisation of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.
 
@@
 
-The Empire developed a series of structures to support and organize this long-range trade and specialized entire regions to massively produce specific goods.
+The Empire developed a series of structures to support and organise this long-range trade and specialised entire regions to massively produce specific goods.
 An example of this process was the province of \textit{Baetica} (currently Andalusia, southern Spain) which became an important olive oil production centre during the Roman Empire.
@@
 
-The structure and processes associated to this massive olive oil production has been extensively discussed over the last decades\citep{rodriguez_economioleicola_1977, Chic_hispania_1997,millet_anforas_1998}.
+The structure and processes associated with this massive olive oil production have been extensively discussed over the last decades\citep{rodriguez_economioleicola_1977, Chic_hispania_1997,millet_anforas_1998}.
 However, the lack of written records has made possible to detect any indication of patterns in the olive oil market.
 
-Our knowledge of this economic activity has benefited from new findings and data sources, but despite all these advances several questions remain open: How was amphora production organized during the Roman Empire?
-Can we detected economic dynamics between provinces by analysing the distribution of amphorae?
+Our knowledge of this economic activity has benefited from new findings and data sources, but despite all these advances several questions remain open: How was amphora production organised during the Roman Empire?
+Can we detect economic dynamics between provinces by analysing the distribution of amphorae?
 
@@
 %\xavi{more diverse que qué?} 
-The application of different quantitative approaches has allowed us to improve our interpretation of the growing amount of archaeological evidence and reveal the dynamics of the Roman production %\xavi{en qué sentido?
+The application of different quantitative approaches has allowed us to improve our interpretation of the growing amount of archaeological evidence and reveal the dynamics of Roman production %\xavi{en qué sentido?
 qué quieres decir aquí con complexity?
@@
 The finding of amphoric stamps in different provinces has allowed us to know the importance of the distribution of olive oil within the Empire.
-Despite having archaeological evidence, no consensus has been reached: How was the economical relation between production and consumption centers?
-Did they follow a economical pattern to distribute olive oil in those provinces?
+Despite having archaeological evidence, no consensus has been reached: How was the economical relation between production and consumption centres?
+Did they follow an economical pattern to distribute olive oil in those provinces?
 
@@
 \includegraphics[width=\linewidth]{figs/general_map}
-\caption{Overview of the sites analyzed in this paper.
+\caption{Overview of the sites analysed in this paper.
 The color defines the three different regions under study (\textit{Baetica} as producer; \textit{Germania} and \textit{Britannia} as consumers) while the size of each dot highlights the number of Dressel-20 amphoric stamps found on each of these sites}
@@
 A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps.
-Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potential spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
+Here, we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
 Stamps will be used to identify economic patterns by analysing their similarity.

### 2020-08-21 Paper/jaccardbaetica.tex
correction until measuring the dissimilarity

@@
 Las preguntas específicas de los 3 casos de estudio deberían entrelazarse en una pregunta general del paper, que esté vinculada a lo que has explicado hasta ahora}
+\maria{he añadido unos párrafos :-)}
 
@@
 
-The landscape of the \textit{Baetican} province was one of the best regions to face the increase in demand of olive oil across the Roman provinces.
-For this reason the area saw an increase in productivity as a massive infrastructure of olive oil production was gradually deployed.
+The landscape of the \textit{Baetican} province was one of the best regions to face the increased demand for olive oil across the Roman provinces.
+For this reason, the area saw an increase in productivity as a massive infrastructure of olive oil production was gradually deployed.
 The production and distribution of olive oil increased from the first to the third centuries AD ~\citep{remesal_concierto}.
 
-Hundreds of presses and amphora-making workshops were build near large extensions covered by olive trees.
+Hundreds of presses and amphora-making workshops were built near large extensions covered by olive trees.
 The workshops where amphorae were made and filled with olive oil were places along the rivers Guadalquivir and Genil riverbanks (see Fig.\ref{workshop}).
@@
 \includegraphics[width=\linewidth]{figs/baetica}
-\caption{Distribution of stamps found in Dressel-20 workshops along the Guadalquivir riverbank. The size of each point depends on the number of stamps found on each site}
+\caption{Distribution of stamps found in Dressel 20 workshops along the Guadalquivir riverbank. The size of each point depends on the number of stamps found on each site}
 
@@
 
-%\xavi{deberías desarrollar esto un poco más; qué tipo de evidencia demuestra esto? pocos cambios en qué? los talleres? las tipologías anfóricas?} .
+\xavi{deberías desarrollar esto un poco más; qué tipo de evidencia demuestra esto? pocos cambios en qué? los talleres? las tipologías anfóricas?}
+\maria{hecho}
 
@@
 
-Dressel 20 amphorae have a recognisable globular form with short neck and oval-shaped handles.
+They have a recognisable globular form with a short neck and oval-shaped handles.
 Its capacity allowed to hold 70-75 litres of olive oil to be shipped and distributed throughout Empire \citep{berni_dressel_2016}.
 
-%A large majority of this Baetican amphorae production is composed by Dressel 20 types\citep{dressel_ricerche_1878,
-%millet_anforas_1998}.
-
-%were shipped and distributed from \textit{Baetica} throughout western Europe. 
-%\xavi{esto ya lo has mencionado; quizás podrías explicar en 2 líneas la capacidad y forma de la dressel 20}.
+\xavi{esto ya lo has mencionado; quizás podrías explicar en 2 líneas la capacidad y forma de la dressel 20}.
+\maria{hecho y borrado párrafo}
 
@@
 
-Dressel 20 was the amphora most stamped during the Roman Empire \citep[18]{millet_anforas_1998}.
+Dressel 20 was the amphora most stamped during the Roman Empire ~\citep[18]{millet_anforas_1998}.
 
@@
 The large sample size of recovered Dressel 20 stamps is distributed across all the western provinces of the Roman Empire.
-For this reason there are several publications of this type of evidence discussing its origin, its meaning, its long period of use and its spatial distribution~\citep{dressel_ricerche_1878,
+For this reason, there are several publications of this type of evidence discussing its origin, its meaning, its long period of use and its spatial distribution~\citep{dressel_ricerche_1878,
 rodriguez_economioleicola_1977,
@@
 
-%\xavi{si dices que es muy popular debería haber aquí más citas aparte de Remesal}.
+\xavi{si dices que es muy popular debería haber aquí más citas aparte de Remesal}
+\maria{he incluido nuevas citas}
 
@@
 
-Most stamp codes display a large activity of production that can be difficult to date using an accurate chronology.
+Most stamps display a large activity of production that can be difficult to date using an accurate chronology.
 They have been frequently dated with the consular dating by studying \textit{tituli picti} found in \textit{Monte Testaccio} \citep{Testaccio1, berni_millet_epigrafianforica_2008}.
-Despite \textit{tituli picti} contain important information about chronology, they have been still found in few quantities mostly concentrated in a only settlement such as \textit{Monte Testaccio}.
+Despite \textit{tituli picti} contain important information about chronology, they have been still found in few quantities mostly concentrated in an only settlement such as \textit{Monte Testaccio}.
 In addition, the chronology of stamps could be also biased taking into account that amphorae were deposited after being produced and consumed\xavi{esto necesita un poco de desarrollo.
@@
 Stamps often displayed a code of three letters and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
-The three letters correspond to a initial forms of a name but they can also appear complete.
-%\xavi{que quieres decir con abbreviated o complete? Como abrevias un código de 3 letras?}
+The three letters correspond to initial forms of a name but they can also appear complete.
+\xavi{que quieres decir con abbreviated o complete? Como abrevias un código de 3 letras?}
+\maria{corregido}
 
@@
 
-The stamp codes are interpreted based on three main ideas: content (olive oil), context (amphora workshop) and subject (individuals involved on the production).
+The stamp codes are interpreted based on three main ideas: content (olive oil), context (amphora workshop) and subject (individuals involved in the production).
 On the one hand, it seems that stamps could have been identified as the landowner of the olive groves \citep{rodriguez_economioleicola_1977}.
-On the other hand, they could also belong to the owner of the workshops were the amphorae was made or even to a group of amphora workers~\citep{berni_millet_epigrafianforica_2008}.
+On the other hand, they could also belong to the owner of the workshops were the amphorae were made or even to a group of amphora workers~\citep{berni_millet_epigrafianforica_2008}.
 %In any case, the use of these stamps is a good proxy to explore the system of olive oil production and trade developed in Baetica.
@@
 \maria{he cambiado la frase de sitio para que se contextualice mejor}
+
 Nevertheless, some challenges remain under discussion such as how this production was organised and whether it is possible to distinguish production patterns in the olive oil trade.
@@
 Si la pones entonces deberías expandir la idea de frase a párrafo, y poner referencias}.
-\maria{he eliminado la última frase}
+%\maria{he eliminado la última frase}
 
@@
 The Roman conquest in \textit{Britannia} and \textit{Germania} was a new opening to foreign trade in order to supply the military campaign.
-A high demand for olive oil in both provinces stimulated an important trade network involving different agents and the Empire itself.
+High demand for olive oil in both provinces stimulated an important trade network involving different agents and the Empire itself.
 \textit{Baetica} became a strategic province for the production and distribution in both provinces since it can be detected by the high presence of amphoric stamps found in civil and military areas during this period.
@@
 The evidence for olive oil consumption in \textit{Britannia} is scarce before the Roman conquest \citep{funari_corpus_1996,carreras_abastecimiento_2003}.
-Local population did not consume this product and it was not produced on the region as the landscape and climate of the British isles was not suitable for olive oil trees\citep[161]{monfort_britanniaen_1998}.
+The local population did not consume this product and it was not produced on the region as the landscape and climate of the British Isles was not suitable for olive oil trees\citep[161]{monfort_britanniaen_1998}.
 Based on the archaeological records, this absence of olive oil changed with the arrival of a large amount of Dressel 20 amphorae to \textit{Britannia} from the first century A.D \citep{peacock_amphorae_1991,
@@
 The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges.
-The demand of olive oil after the Roman conquest required of a new trade route due to the lack of local olive oil production; \textit{Baetica} was chosen as its main supply.
+The demand for olive oil after the Roman conquest required of a new trade route due to the lack of local olive oil production; \textit{Baetica} was chosen as its main supply.
 
@@
 funari_corpus_1996,
-carreras_britannia_1998,carreras_abastecimiento_2003}.%\xavi{si dices widely studied no puedes referencias solo 1 autor}.
+carreras_britannia_1998,carreras_abastecimiento_2003}.\xavi{si dices widely studied no puedes referencias solo 1 autor}
+\maria{he añadido más autores}
 
-This intense consumption of olive oil also indicates a possible provincial structure designed to organize the supply of olive oil to military camps,
+This intense consumption of olive oil also indicates a possible provincial structure designed to organize the supply of olive oil to military camps.
 
@@
 
-%\xavi{pero entonces la dressel 23 de donde viene?
+\xavi{pero entonces la dressel 23 de donde viene?
 Cuidao que estás diciendo que se acaban las dressel 20, pero esto implica que hay menos aceite de oliva o que viene de otras partes o que va en recipientes distintos a la dressel 20?}
+\maria{he eliminado esta frase para que quede más claro}
 
@@
 
-%Julius Caesar's campaigns in \textit{Gallia} during 1st BC opened the door to the invasion of \textit{Germania} during Augustus rule~\citep{remesal_annona_1986,remesal_baetica_2002}\xavi{no hay un resumen de la conquista similar para Britannia, queda raro}.
+%Julius Caesar's campaigns in \textit{Gallia} during 1st BC opened the door to the invasion of \textit{Germania} during Augustus rule~\citep{remesal_annona_1986,remesal_baetica_2002ç
+\xavi{no hay un resumen de la conquista similar para Britannia, queda raro}. 
 \maria{lo he eliminado}
@@
 The massive importation of Dressel 20 in both \textit{Germania} provinces could be explained due to the archaeological sources found mostly in German \textit{limes}. 
-By contrast, it is still unknown if trade agents participated on the distribution of Baetican olive oil in \textit{Germania} or the supply was exclusively focused on the Roman army garrisons\citep[156]{remesal_germn_2010}
+By contrast, it is still unknown if trade agents participated in the distribution of Baetican olive oil in \textit{Germania} or the supply was exclusively focused on the Roman army garrisons\citep[156]{remesal_germn_2010}
 \xavi{quieres decir que aquí sólo encontramos dressel 20 en yacimientos militares porque no se han excavado muchos civiles?}.
-\maria{lo he mejorado a ver si se entiende}
+\maria{lo he mejorado, a ver si se entiende}
 
@@
 
-There databased allowed us to retrieve 3798 stamps found in Dressel 20 amphorae recovered from workshops within \textit{Baetica} province.
+The database allowed us to retrieve 3798 stamps found in Dressel 20 amphorae recovered from workshops within \textit{Baetica} province.
 Each of these stamps ha detailed records on the site where it was found, the inscribed code as well as its spatial coordinates.
@@
 
-According to previous studies those workshops became the most important centres of amphora production in the region \citep{rodriguez_economioleicola_1977,
+According to previous studies, those workshops became the most important centres of amphora production in the region \citep{rodriguez_economioleicola_1977,
 arva_1997}.
-%\xavi{cita de estos estudios}
+\xavi{cita de estos estudios}
+\maria{citados}
 
@@
 
-%\xavi{pero esto no se hizo en el de Baetica no? Por qué?}.
+\xavi{pero esto no se hizo en el de Baetica no? Por qué?}
+\maria{lo he especificado}
 
@@
 %CITAR (Callender, 1965; Carreras Monfort y Funari,1998; Ayllón-Martı́ et al., 2018). 
-In \textit{Britannia}, we studied a total sample of 1765 stamps from 46 sites displaying 968 unique codes
+In \textit{Britannia}, we studied a total sample of 1765 stamps from 46 sites displaying 968 unique codes.
 
-%\xavi{pero esto no concuerda con las 2219 de antes...es por descartar los inciertos? Hay que dejarlo claro}.
+\xavi{pero esto no concuerda con las 2219 de antes...es por descartar los inciertos? Hay que dejarlo claro}
+\maria{clarificado}
 
@@
 \includegraphics[width=\linewidth]{figs/britannia}
-\caption{Sites in \textit{Britannia} where Dressel-20 amphoric stamps have been found.
-Each dot is a site while its size shows the sample size of stamps found in the site.
+\caption{Sites in \textit{Britannia} where Dressel 20 amphoric stamps have been found.
+Each dot is a site while its size shows the sample size of stamps found on the site.
 A majority of stamps have been found in garrisons related to Hadrian's and Antonine walls}
@@
 \caption{Sites in Germania where Dressel-20 amphoric stamps have been found.
-Each dot is a site while its size shows the sample size of stamps found in the site.
+Each dot is a site while its size shows the sample size of stamps found on the site.
 Most sites are located along the German \textit{limes}}

### 2020-09-29 Paper/jaccardbaetica.tex
correction finished

@@
 Considering our dataset as a non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity \citep{wolda_similarity_1981}.
-Morisita-Horn index gives a value between 0 (sites have an exact stamp codes and frequency distribution) and 1 (complete difference between stamp codes).
-To apply Morisita-Horn we firstly calculated the number of times that each stamp code appeared in an amphora workshop.
+Morisita-Horn index gives a value between 0 (sites have exact stamp codes and frequency distribution) and 1 (complete difference between stamp codes).
+To apply Morisita-Horn we first calculated the number of times that each stamp code appeared in an amphora workshop.
 This method allowed us to bear in mind the number of frequency of stamps for each workshop. 
@@
 This can be observed more clearly on military sites as the ones with higher similarity are geographically closer.
-Two patterns can be observed on the dendrogram: 1) nearby sites tend to share more similar stamps and 2) a majority of stamp code are only found on one site.
+Two patterns can be observed on the dendrogram: 1) nearby sites tend to share more similar stamps and 2) a majority of stamp codes are only found in one site.
 
@@
 
-A higher concentration of sites sharing similar stamps were found in areas eminently militarised and close to German \textit{limes}, although most sites mostly showed different stamps. 
+A higher concentration of sites sharing similar stamps was found in areas eminently militarised and close to German \textit{limes}, although most sites mostly showed different stamps. 
 \maria{he cambiado esta frase}
@@
 
-This work has explored to what extent the distribution of amphoric stamps can provide insights on the organisation of olive oil Roman markets both at the production and consumption areas.
+This work has explored to what extent the distribution of amphoric stamps can provide insights into the organisation of olive oil Roman markets both at the production and consumption areas.
 For this reason, an index of dissimilarity was used to detect differences between the distribution of the amphoric stamps and the spatial distance of producers and consumption centres.
@@
 The analysis of the amphora workshops in \textit{Baetica} province suggests that a majority of stamp codes were only used on a single workshop.
-Beyond this clear pattern it seems that the similarity of stamp codes is correlated with spatial distance and the amphoric production of spatially close sites shared some stamp codes.
+Beyond this clear pattern, it seems that the similarity of stamp codes is correlated with spatial distance and the amphoric production of spatially close sites shared some stamp codes.
 Particularly relevant here is the fact that our results show how similar stamp codes were found on sites belonging to the same \textit{conventus} area.
@@
 
-Both consumption areas showed a correlation between spatial distance of sites and their similarity of amphoric stamps.
+Both consumption areas showed a correlation between the spatial distance of sites and their similarity of amphoric stamps.
 In the case of  \textit{Britannia}, the correlation was higher than \textit{Germania}.
@@
 The use of specific codes seems to be almost exclusively decided by each workshop, so it could be used to identify amphorae found in consumption centers (i.e.
-an amphorae with a given stamp was probably made on a specific workshop).
+an amphora with a given stamp was probably made on a specific workshop).
 This idea could explain why the similarity of stamps between workshops is generally very low.
@@
 
-This research was partly funded by the European Research Council Advanced Grant EPNet (340828).
-We are grateful to Simon Carrignon, Juan Moros, Víctor M.
-Lozano and Ignacio Morer for their useful suggestions.
+This research was partly funded by the European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros and Ignacio Morer for their useful suggestions.  
 All data has been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}.

### 2020-11-02 Paper/jaccardbaetica.tex
review bef submit

@@
 
-%\journal{Journal of Archaeological Science}
+\journal{Journal of Archaeological Method and Theory}
 
@@
 
-\title{The markings of the trade: exploring the patterns of olive oil production in Roman Baetica}
+%\title{The markings of the trade: exploring the patterns of olive oil production in Roman Baetica}
+\title{The tracing of trade: exploring the patterns of olive oil production and distribution from Roman Baetica}
 
-\author[ceipacadress]{Maria Coto-Sarmiento\corref{mycorrespondingauthor}}
+\author[ceipacadress]{María Coto-Sarmiento\corref{mycorrespondingauthor}}
 \cortext[mycorrespondingauthor]{Corresponding author}
@@
 
-\author[edadress]{Xavier Rubio-Campillo}
+\author[edadress,ubadress]{Xavier Rubio-Campillo}
 
-\address[ceipacadress]{CEIPAC, Department of Prehistory and Archaeology, Montalegre, 6-8, 08001, University of Barcelona, Barcelona, Spain}
+\address[ceipacadress]{Department of Prehistory and Archaeology, Montalegre, 6-8, 08001, University of Barcelona, Barcelona, Spain}
 \address[edadress]{School of History, Classic \& Archaeology, Room OOM.33, William Robertson Wing, Old Medical School, Teviot Place, University of Edinburgh, UK}
+\address[ubadress]{DIDPATRI, Passeig de la Vall d'Hebrón, 171, Barcelona, Spain}
 
@@
 \begin{abstract}
-
-\xavi{reescribir abstract, demasiado general y entonces demasiado específico}
-\maria{lo he reeformulado a ver si ahora..}
 
@@
 
-%Did each province follow a different pattern for its distribution? 
-%\xavi{no entiendo esta pregunta} what patterns were followed for the distribution of olive oil to the different provinces? 
-%\xavi{no es esta pregunta idéntica a la anterior?} 
-%\xavi{esta frase no tiene sentido aquí; o la pones antes o en otro párrafo, pero es raro acabar un set de 2 preguntas con una frase que no sea otra pregunta}
-
-%\xavi{las preguntas no van sobre patrones; los patrones son las trazas en la evidencia que te ayudan a responder a las preguntas}
-
 Advances in the research of the Roman studies have currently led an environment with various interpretations on commercial dynamics \citep{duncan1982economy,
@@
 %\xavi{more diverse que qué?} 
-The application of different quantitative approaches has allowed us to improve our interpretation of the growing amount of archaeological evidence and reveal the dynamics of Roman production %\xavi{en qué sentido?
-qué quieres decir aquí con complexity?
-Si no especificas queda muy genérico (rollo "el mundo es muy complejo")}.
+The application of different quantitative approaches has allowed us to improve our interpretation of the growing amount of archaeological evidence and reveal the dynamics of Roman production
 \citep{brughmans_roman_2016,
@@
 Did they follow an economical pattern to distribute olive oil in those provinces?
-
-\xavi{me parece que aquí falta más chicha, porque saltas de una cosa super general (complejidad, economía, grandes dinámicas) a unos casos muy específicos.
-Deberías aquí explicar algo más sobre tu estudio y lo que quieres entender, por ejemplo qué preguntas específicas quieres responder?
-Como se enlazan con lo que has explicado hasta ahora?
-Cómo es que ahora saltas a casos de estudio regionales?
-Las preguntas específicas de los 3 casos de estudio deberían entrelazarse en una pregunta general del paper, que esté vinculada a lo que has explicado hasta ahora}
-\maria{he añadido unos párrafos :-)}
 
@@
 For this reason, the area saw an increase in productivity as a massive infrastructure of olive oil production was gradually deployed.
-The production and distribution of olive oil increased from the first to the third centuries AD ~\citep{remesal_concierto}.
+The production and distribution of olive oil increased from the first to the third centuries AD~\citep{remesal_concierto}.
 
@@
 
-This strategic location allowed the transport of olive oil through riverine shipping towards the maritime routes that connected \textit{Baetica} with Mediterranean and Atlantic trade routes towards the rest of the Empire\citep{garcia_vargas_enrique_formal_2010}.
+This strategic location allowed the transport of olive oil through riverine shipping towards the maritime routes that connected \textit{Baetica} with Mediterranean and Atlantic trade routes towards the rest of the Empire~\citep{garcia_vargas_enrique_formal_2010}.
 
@@
 
-The chronology of the workshops is wide spanning from the first to the third centuries AD
-%\xavi{qué quiere decir que la cronología es diversa?} 
-\citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}. 
+The chronology of the workshops is wide spanning from the first to the third centuries AD\citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}. 
 This fact is shown by the archaeological evidence that displays a highly specialised production with a long activity \citep{remesal_anforas_2004}.
@@
 
-\xavi{deberías desarrollar esto un poco más; qué tipo de evidencia demuestra esto? pocos cambios en qué? los talleres? las tipologías anfóricas?}
-\maria{hecho}
-
 The importance of this olive oil trade is revealed by the massive amount of Dressel 20 type amphorae found across the Roman provinces \citep{dressel_ricerche_1878,
 millet_anforas_1998}.
-This amphora has been commonly associated with the transportation of Baetican olive oil for supplying military camps and civil settlements \citep{berni_millet_epigrafianforica_2008}.%\xavi{me parece repetitivo, porque unos párrafos antes ya hablabas de lo mismo.
-Quizás unificar texto aquí o allá?}.
+This amphora has been commonly associated with the transportation of Baetican olive oil for supplying military camps and civil settlements \citep{berni_millet_epigrafianforica_2008}.
 
@@
 Its capacity allowed to hold 70-75 litres of olive oil to be shipped and distributed throughout Empire \citep{berni_dressel_2016}.
-
-\xavi{esto ya lo has mencionado; quizás podrías explicar en 2 líneas la capacidad y forma de la dressel 20}.
-\maria{hecho y borrado párrafo}
 
@@
 Dressel 20 was the amphora most stamped during the Roman Empire ~\citep[18]{millet_anforas_1998}.
-
-%The usage period is associated with a large amphora production for almost three centuries\xavi{esto está repetido no?}.
-
 The large sample size of recovered Dressel 20 stamps is distributed across all the western provinces of the Roman Empire.
@@
 chicepi1985,millet_anforas_1998, remesal_sellar_2016} (Fig.\ref{amphora}).
-
-\xavi{si dices que es muy popular debería haber aquí más citas aparte de Remesal}
-\maria{he incluido nuevas citas}
 
@@
 Despite \textit{tituli picti} contain important information about chronology, they have been still found in few quantities mostly concentrated in an only settlement such as \textit{Monte Testaccio}.
-In addition, the chronology of stamps could be also biased taking into account that amphorae were deposited after being produced and consumed\xavi{esto necesita un poco de desarrollo.
-quieres decir que el código se encuentra en contextos de cronologías diferentes no?
-Y lo del Testaccio, pos eso, que hay pocos y todos de un yacimioento, lo que añade un bias}.
-\maria{desarrollado}
+In addition, the chronology of stamps could be also biased taking into account that amphorae were deposited after being produced and consumed.
 
@@
 The content information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion.
-%\xavi{qué information? El código de letras?}. 
+
 Stamps often displayed a code of three letters and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
 The three letters correspond to initial forms of a name but they can also appear complete.
-\xavi{que quieres decir con abbreviated o complete? Como abrevias un código de 3 letras?}
-\maria{corregido}
 
@@
 
-%Additionally, a large percentage of Dressel 20 containers were not stamped thus complicating the interpretation~\xavi{esta última frase aquí es un poco random no?}.
+%Additionally, a large percentage of Dressel 20 containers were not stamped thus complicating the interpretation
 
@@
 
-\xavi{acabas de decir que tampoco se sabe qué es el sello, así que esta frase queda rara} 
-\maria{he cambiado la frase de sitio para que se contextualice mejor}
-
 Nevertheless, some challenges remain under discussion such as how this production was organised and whether it is possible to distinguish production patterns in the olive oil trade.
@@
 Did stamps share the same workshop?
-
-%Neither the use of written records have allowed providing enough information that can explain the economic role of \textit{Baetica} province in the Roman organisation.\xavi{esta frase esta fuera de contexto.
-Si la pones entonces deberías expandir la idea de frase a párrafo, y poner referencias}.
-%\maria{he eliminado la última frase}
 
@@
 The creation of new provinces required the Roman Empire the arrival of resources through Mediterranean and Atlantic routes.
-%\xavi{no entiendo esto...allowed o required?}. 
+
 This led to a gradual change in the economic and social structures the trade networks that supplied the Roman army became more complex.
 Augustus's administration created the figure of the \textit{praefectura annonae} to organize wheat supply.
-The role of the \textit{praefectura annonae} was mainly focused on providing through \textit{frumentationes} a fixed monthly amount of wheat to each Roman citizen~\citep{remesal_annona_1986,remesal_concierto}\xavi{pero estabas hablando del ejército no?
-Has hecho un salto aquí a los ciudadanos}.
-\maria{Pero porque la annona estaba enfocada a los ciudadanos romanos y luego hablo después que hay otras hipótesis que dicen que iba más allá a los militares}
+The role of the \textit{praefectura annonae} was mainly focused on providing through \textit{frumentationes} a fixed monthly amount of wheat to each Roman citizen~\citep{remesal_annona_1986,remesal_concierto}
 
@@
 remesal_annona_1990}
-
-\xavi{pero no hay evidencia? Las fuentes escritas no lo mencionan?}
-\maria{He cambiado authors por hipotesis porque me quería referir a investigadores}.
 
@@
 Olive oil supply seems to be particularly intense in militarised provinces that functioned as the borders of the Empire such as \textit{Britannia} and \textit{Germania}.
-The high percentage of Dressel 20 amphorae found in sites nearby military garrisons suggests the existence of diverse supply dynamics as there were trade processes flowing from military sites to civil settlements~\citep{remesal_annona_1986, carreras_britannia_1998}
-\xavi{lo que quieres decir es que tienes muchas ánforas en sitios civiles cerca de militares? no queda claro}.
-\maria{sí, pero creo que yo no he escrito está frase xD}
+The high percentage of Dressel 20 amphorae found in sites nearby military garrisons suggests the existence of diverse supply dynamics as there were trade processes flowing from military sites to civil settlements~\citep{remesal_annona_1986, carreras_britannia_1998}.
 
 An explanation for the interest to export olive oil for the military garrison in the Roman provinces is still unclear.
-The origin of this exportation may be related to two aspects: a) cultural consumption whose olive oil is used by cultural reasons such as group identification or by habit and b) economical and political reasons to create a redistributive system mechanism to supply military camps.
-\maria{esta frase está explicada mejor}
-
-%may be related to two aspects according with some hypotheses: a) cultural consumption whose product is consumed by cultural reasons such as to identity or habit and b) economical reasons where olive oil distribution is based on the transport costs, unlike other products \citep[69-70]{carreras_britannia_1998}\xavi{no pillo esta frase}..
+The origin of this exportation may be related to two aspects: a) cultural consumption whose olive oil is used by cultural reasons such as group identification or by habit and b) economical and political reasons to create a redistributive system mechanism to supply military camps \citep[69-70]{carreras_britannia_1998}.
 
@@
 \textit{Baetica} became a strategic province for the production and distribution in both provinces since it can be detected by the high presence of amphoric stamps found in civil and military areas during this period.
-
-\xavi{faltaría un enlace a la siguiente sección, queda muy abrupto}
-\maria{he añadido varios párrafos a ver si se enlaza}
 
@@
 carreras_britannia_1998}.
-At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}
-%\xavi{en qué siglo?}
+At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}.
 
@@
 funari_corpus_1996,
-carreras_britannia_1998,carreras_abastecimiento_2003}.\xavi{si dices widely studied no puedes referencias solo 1 autor}
-\maria{he añadido más autores}
+carreras_britannia_1998,carreras_abastecimiento_2003}.
 
 This intense consumption of olive oil also indicates a possible provincial structure designed to organize the supply of olive oil to military camps.
-
-%such as \textit{Germania} (Remesal Rodrıguez, 1986)\xavi{por qué such as Germania? Aún no has explicado qué hacen en Germania}.
 
@@
 
-\xavi{pero entonces la dressel 23 de donde viene?
-Cuidao que estás diciendo que se acaban las dressel 20, pero esto implica que hay menos aceite de oliva o que viene de otras partes o que va en recipientes distintos a la dressel 20?}
-\maria{he eliminado esta frase para que quede más claro}
-
 \subsection{Germania}
 
-%Julius Caesar's campaigns in \textit{Gallia} during 1st BC opened the door to the invasion of \textit{Germania} during Augustus rule~\citep{remesal_annona_1986,remesal_baetica_2002ç
-\xavi{no hay un resumen de la conquista similar para Britannia, queda raro}. 
-\maria{lo he eliminado}
-
 Roman studies exploring the presence of Dressel 20 amphorae in the provinces Inferior and Superior during Roman Empire have not had the same impact as the rest of the provinces \citep[293]{remesal_baetica_2002}.
-%\xavi{no entiendo esta frase. Que qué provincia? Germania son 2 provincias no?}
 
 The massive importation of Dressel 20 in both \textit{Germania} provinces could be explained due to the archaeological sources found mostly in German \textit{limes}. 
-By contrast, it is still unknown if trade agents participated in the distribution of Baetican olive oil in \textit{Germania} or the supply was exclusively focused on the Roman army garrisons\citep[156]{remesal_germn_2010}
-\xavi{quieres decir que aquí sólo encontramos dressel 20 en yacimientos militares porque no se han excavado muchos civiles?}.
-\maria{lo he mejorado, a ver si se entiende}
+By contrast, it is still unknown if trade agents participated in the distribution of Baetican olive oil in \textit{Germania} or the supply was exclusively focused on the Roman army garrisons\citep[156]{remesal_germn_2010}.
 
@@
 
-%\xavi{cita?}
-
 The presence of the Roman army encouraged the exchange in the province as shown by the arrival of this product both civil settlement and military sites with a mayor concentration at the German \textit{limes}.
 
-\xavi{pero justo antes decías que no había en civiles no?}. 
-\maria{yo no he dicho eso :-[}
+It seems that some Baetican centres could have been assigned to a specific province for the olive oil supply to military garrisons \citep[125]{remesal_concierto}.
 
-It seems that some Baetican centres could have been assigned to a specific province for the olive oil supply to military garrisons \citep[125]{remesal_concierto}. 
-\xavi{the support of olive oil no se entiende. Además, falta alguna cita y desarrollar esta idea}. 
-\maria{mejorado, creo}
 However, this hypothesis is hard to assess given the current lack of written evidence related to the meaning of the amphoric stamps.
-\xavi{related to BLABLABLA. Habría que detallar un poco porque hay un montón de evidencia arqueológica del limes, así que qué es exactamente lo que falta?}. 
-\maria{mejorado,creo}
 
 Previous studies suggested that the supply to the German limes was mainly based on riverine transport, but recent works suggest that the maritime route through the Atlantic Ocean could have been more important than expected (both for \textit{Germania} and \textit{Britannia}).~\citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
-%\xavi{lo del océano atlántico está un poco descontextualizado; no sería mejor ponerlo al final de esta sección?}
 
@@
 We are especially interested in identifying links between production and consumption centres by using amphoric stamps.
-
-%To do that, we use the CEIPAC database to collect stamps from different places. The CEIPAC dataset contains over 50.000 of epigraphy records found in amphorae, mostly from \textit{Monte Testaccio}.
 
@@
 
-%\xavi{estaría bien ampliar un poco esto; qué organizaban exactamente?}
-
 The production area for Dressel 20 amphorae extended across three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter Hispalis), \textit{Cordubensis} (currently C\'ordoba, hereafter Corduba) and \textit{Astigi} (currently Écija, Sevilla, hereafter Astigi) \citep{rodriguez_economioleicola_1977,chicdatos2001,berni_millet_epigrafianforica_2008} .
@@
 
-%It is important to mention that some workshops exhibited the same coordinates but they were catalogued as different workshops in Roman Studies (i.e.
-nearby amphora workshops in \textit{Baetica} named as Tesorillo de Doña Mencia and Doña Mencia).
-The analysis took into account this particularity but some biases can occur due to the archaeological dataset.
-
-\xavi{No se entiende; deberías explicarlo un poco mejor y citar algo si dices eso. Piensa que la mayoría de lectores no sabrán quién es Doña Mencia}. 
-\maria{he reformulado la frase a ver si ahora se entiende}
-
 The first step was to compute the frequency distribution of stamp codes and analyse it using Exploratory Data Analysis (EDA).
@@
 
-%\xavi{el gráfico es dificil de entender...yo leo que hay unos 20 workshops con 1 solo stamp, por lo que entiendo que no es cierto que la mayoría solo tengan 1 sello (hay como 50 workshops con entre 2 y 30 sellos no?)}
-\maria{he cambiado la frase a ver si se entiende mejor}
+The type of distribution is also frequent in amphora production where we observed a major concentration of the number of stamps in few workshops\citep{bayesian_2018,coto-sarmiento_identifying_2018}.
 
-The type of distribution is also frequent in amphora production where we observed a self-organised complex system pattern with a major concentration of the number of stamps in a few workshops \citep{bayesian_2018,coto-sarmiento_identifying_2018}\xavi{no lo veo; recuerda que para el rollo bayesiano tenías una línea recta al hacedr el log de los 2 ejes, no solo 1, y este plot no está construído así no?}.
-\maria{son dos ejes, no? ¿aquí no se ve línea recta? Lo mismo si lo hago de puntitos...}
+%\xavi{no lo veo; recuerda que para el rollo bayesiano tenías una línea recta al hacedr el log de los 2 ejes, no solo 1, y este plot no está construído así no?}.
+%\maria{son dos ejes, no? ¿aquí no se ve línea recta? Lo mismo si lo hago de puntitos...}
 
@@
 
-%\xavi{una cosa, aquí has pasado de sellos a códigos, y los valores son distintos que en el gráfico anterior. Deberías clarificarlo antes de empezar a hablar de la figura}.
-
 According to previous studies, those workshops became the most important centres of amphora production in the region \citep{rodriguez_economioleicola_1977,
 arva_1997}.
-\xavi{cita de estos estudios}
-\maria{citados}
 
@@
 Each dot corresponds to a workshop sorted by different areas.
-Colours are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
+Colours are represented by areas divided into \textit{Hispalis} (red), \textit{Astigi} (green) and \textit{Corduba} (blue)}
 \label{frequency}
@@
 However, in this case, we selected the centres with more or equal than five stamps in order to maximize the analysis with more stamps in the dataset.
-
-\xavi{pero esto no se hizo en el de Baetica no? Por qué?}
-\maria{lo he especificado}
 
@@
 %CITAR (Callender, 1965; Carreras Monfort y Funari,1998; Ayllón-Martı́ et al., 2018). 
-In \textit{Britannia}, we studied a total sample of 1765 stamps from 46 sites displaying 968 unique codes.
-
-\xavi{pero esto no concuerda con las 2219 de antes...es por descartar los inciertos? Hay que dejarlo claro}
-\maria{clarificado}
-
-The centres located in \textit{Britannia} can be seen in Figure~\ref{britannia}.
+In \textit{Britannia}, we studied a total sample of 1765 stamps from 46 sites displaying 968 unique codes. The centres located in \textit{Britannia} can be seen in Figure~\ref{britannia}.
 
@@
 
-The approach proposed here aimed at exploring links between production and consumption centers through the identification of common amphorae codes found in different sites.
+The approach proposed here aimed at exploring links between production and consumption centres through the identification of common amphorae codes found in different sites.
 For this reason we measured the similarity between amphora workshops and/or consumption sites by quantifying a pairwise distance or dissimilarity index between places (i.e.
@@
 In our case, stamps from different sites will be used as a proxy to measure their dissimilarity by analyzing their code.
-\xavi{citar reforzar más esta idea}.
-\maria{he añadido unas líneas a ver si están bien o no xD}
 
@@
 This method allowed us to bear in mind the number of frequency of stamps for each workshop.
-%This method allowed us to bear in mind a similar number of times for each repeated stamp per workshop\xavi{no entiendo esta frase}.
 
@@
 
-%By contrast, most sites did not show a strong similarity in stamp correlation\xavi{qué quieres decir?
-porque lo de que hay una similaridad baja ya lo has repetido arriba}.
-We also did not find any groups of similar stamps in a specific place\xavi{tampoc entiendo esto.
-que no hay sellos similares en un solo sitio...?}.
-
 The similarity of sites based on Dressel 20 stamp codes is also low in \textit{Germania} as can be seen in Figure~\ref{germap}.
@@
 A higher concentration of sites sharing similar stamps was found in areas eminently militarised and close to German \textit{limes}, although most sites mostly showed different stamps.
-\maria{he cambiado esta frase}
-%even if most sites mostly showed different stamps~\xavi{quieres decir que los sitios más cercanos al limes son más similares? no queda claro y hay que detallarlo un poco más}.
-
-%\xavi{esta frase me la petaría; no aporta demasiado: The interpretation of the results suggests that there is still no possibility to determine the existence of a defined pattern that reflects in more detail the route of this production in the area of the German limes.}
 
@@
 For this reason, an index of dissimilarity was used to detect differences between the distribution of the amphoric stamps and the spatial distance of producers and consumption centres.
-
-%The general purpose was to explore if such differences found in the stamp codes could play an important role in the Roman market\xavi{este párrafo es un poco reiterativo no?}.
 
@@
 
-%In general, the majority of stamps were located in different amphora workshops\xavi{estas 2 frases repiten cosas dichas antes}. 
-\maria{he quitado la última frase pero la anterior no se repite}
-
 Our results do not support previous working hypotheses suggesting that groups of workshops used specific groups of stamp codes; while there is higher stamp similarity for nearby workshops this result does not suggest any kind of organisation beyond the simple fact that nearby workshops had a higher degree of interaction.
 In fact; the opposite seems true: each workshop used different stamp codes and only shared some of them with the closest ones.
-
-%These stamp codes detected in closer workshops do not move from other distant workshops.
-In other words, the stamps tend to remain in the same area and different stamps were located in a same amphora workshop\xavi{de nuevo estas 2 frases repiten cosas de antes}.
 
@@
 While the role of the river was significant for the distribution of amphorae, river connectivity amongst workshops does not seem to show relevancy for the distribution or concentration of stamps in a specific river area.
-\xavi{qué quieres decir con esto? que los diferentes ríos no muestran ningún patrón?}
-\maria{he añadido cosas a la frase para ver si se entiende}
 
@@
 The stamped amphorae could have allowed potters to organise the different production batches for posterior commercialisation.
-%\xavi{estos 2 párrafos igual deberían ir a la discusión no? La conclusión debería ser algo más general y aquí estás aún interpretando resultados}.
 
@@
 A combined exploration of the map and the dendrogram also suggests an interesting pattern: centres with higher similarity values are closer to the coast (North Sea and Celtic Sea).
-This may indicate that the Atlantic route could have played an essential role in transporting olive oil to the area of \textit{Britannia}, since in the places where there is greater similarity they are found in different strategic points near the sea
-\xavi{por qué más conectividad implica más similaridad? no será porque encuentras más ánforas?}
-\maria{no entiendo este comentario :-(}.
+This may indicate that the Atlantic route could have played an essential role in transporting olive oil to the area of \textit{Britannia}, since in the places where there is greater similarity they are found in different strategic points near the sea.
 
@@
 However, we do not detect a concrete pattern regarding the distribution of the stamps in the German limes \citep{xanten2018}.
-This could be also interpreted by an intensity bias where military centres have been mostly excavated than civil centres
+This could be also interpreted by an intensity bias where military centres have been mostly excavated than civil centres.
 
@@
 As a consequence, our analysis does not indicate the presence of a specific workshop from \textit{Baetica} supplying to some specific province.
-
-\xavi{esto no iría en la discusión? Por otra parte, no entiendo qué quieres decir exactamente, igual hay que desarrollarlo}.
-\maria{lo he incluido aquí y lo he cambiado a ver si se entiende}
-%Rather, it would imply that olive oil production was distributed from non-specific production centres.
 
@@
 To conclude, the method presented here provides a useful framework to improve our understanding of the organisation and processes that allowed the Roman Empire to manage a massive and highly specialized production of essential goods.
-This work has identified similarities between amphoric stamps and spatial distribution in the case of the amphoric production within the Roman Empire
-%\xavi{qué diferencias?}. 
+This work has identified similarities between amphoric stamps and spatial distribution in the case of the amphoric production within the Roman Empire.
+
 The growing amount of archaeological data requires of large-scale quantitative methods such as the one presented here to identify and explore complex patterns, thus allowing us to infer dynamics and improve our interpretation of complex economic processes of past societies.

### 2020-11-04 Paper/jaccardbaetica.tex
add review VML and git repository

@@
 %\title{The markings of the trade: exploring the patterns of olive oil production in Roman Baetica}
-\title{The tracing of trade: exploring the patterns of olive oil production and distribution from Roman Baetica}
+\title{The tracing of trade: exploring the patterns of olive oil production and distribution from Roman \textit{Baetica}}
 
@@
 The structure and processes associated with this massive olive oil production have been extensively discussed over the last decades\citep{rodriguez_economioleicola_1977, Chic_hispania_1997,millet_anforas_1998}.
-However, the lack of written records has made possible to detect any indication of patterns in the olive oil market.
+However, the lack of written records has not made possible to detect any indication of patterns in the olive oil market.
 
@@
 \caption{Overview of the sites analysed in this paper.
-The color defines the three different regions under study (\textit{Baetica} as producer; \textit{Germania} and \textit{Britannia} as consumers) while the size of each dot highlights the number of Dressel-20 amphoric stamps found on each of these sites}
+The colour defines the three different regions under study (\textit{Baetica} as producer; \textit{Germania} and \textit{Britannia} as consumers) while the size of each dot highlights the number of Dressel 20 amphoric stamps found on each of these sites}
 
@@
 In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces.
-This economic connection could be identified by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific province.
+This economic connection could be identified by different aspects: a) correlation between spatial distance and the sites based on the idea that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific province.
 
@@
 On the one hand, it seems that stamps could have been identified as the landowner of the olive groves \citep{rodriguez_economioleicola_1977}.
-On the other hand, they could also belong to the owner of the workshops were the amphorae were made or even to a group of amphora workers~\citep{berni_millet_epigrafianforica_2008}.
+On the other hand, they could also belong to the owner of the workshops where the amphorae were made or even to a group of amphora workers~\citep{berni_millet_epigrafianforica_2008}.
 %In any case, the use of these stamps is a good proxy to explore the system of olive oil production and trade developed in Baetica.
@@
 This date matches a change in market strategy in the Empire.
-This a gradual decrease on \textit{Baetican} olive oil exports can be observed as a decrease in the amount of Dressel 20 found in excavated sites as this typology is replaced by other types of amphorae from other places \citep{rodriguez1991aceite,millet_anforas_1998}.
+This gradual decrease on \textit{Baetican} olive oil exports can be observed as a decrease in the amount of Dressel 20 found in excavated sites as this typology is replaced by other types of amphorae from other places \citep{rodriguez1991aceite,millet_anforas_1998}.
 
@@
 
-The presence of the Roman army encouraged the exchange in the province as shown by the arrival of this product both civil settlement and military sites with a mayor concentration at the German \textit{limes}.
+The presence of the Roman army encouraged the exchange in the province as shown by the arrival of this product both civil settlement and military sites with a major concentration at the German \textit{limes}.
 
@@
 
-The type of distribution is also frequent in amphora production where we observed a major concentration of the number of stamps in few workshops\citep{bayesian_2018,coto-sarmiento_identifying_2018}.
+The type of distribution is also frequent in amphora production where we observed a major concentration of the number of stamps in few workshops~\citep{bayesian_2018,coto-sarmiento_identifying_2018}.
 
@@
 The workshops of the three \textit{conventus} show a similar pattern on the stamp frequency distribution with the exceptions of 2 large workshops in \textit{Hispalis} (La Catria and Arva).
-On these two workshops a comparatively large amount of code stamps were found (29 different code stamps on each of them)
+On these two workshops a comparatively large amount of code stamps were found (29 different code stamps on each of them).
 
@@
 \includegraphics[width=\linewidth]{figs/germania}
-\caption{Sites in Germania where Dressel-20 amphoric stamps have been found.
+\caption{Sites in Germania where Dressel 20 amphoric stamps have been found.
 Each dot is a site while its size shows the sample size of stamps found on the site.
@@
 Morisita-Horn index gives a value between 0 (sites have exact stamp codes and frequency distribution) and 1 (complete difference between stamp codes).
-To apply Morisita-Horn we first calculated the number of times that each stamp code appeared in an amphora workshop.
+To apply Morisita-Horn, we first calculated the number of times that each stamp code appeared in an amphora workshop.
 This method allowed us to bear in mind the number of frequency of stamps for each workshop.
@@
 Results for the consumption regions display similar patterns than the ones found for the production area of \textit{Baetica}.
-The most important difference is that the similarity values are lower as can be seen in Figure~\ref{britmap}
+The most important difference is that the similarity values are lower as can be seen in Figure~\ref{britmap}.
 
@@
 This could be explained such as batch systematic organisation between potters \citep{juanmorostesis}.
-This work method could have allowed to potters to stamp amphorae in order to identify the production for batches for posterior commercialization.
+This work method could have allowed potters to identify batches by stamping amphorae for posterior commercialization.
+
+%This work method could have allowed to potters to stamp amphorae in order to identify the production for batches for posterior commercialization.
+
 High production of amphorae had to maintain a specialize system organization to control this production \citep[104]{juanmorostesis}.
@@
 %This method allowed to potters organises the production with the batch stamping for the posterior commercialisation
-%\xavi{y esto para qué serviría? debería explicarse un poco más}.
 
@@
 This evidence means that it is unknown whether some production centres went to one province or another or, at least, that they can be clearly reflected in the data with a greater similarity in the amphoric stamps.
-Thus, production centres could have distributed randomly olive oil both \textit{Britannia} and \textit{Germania}.
+Thus, production centres could have distributed randomly olive oil to both \textit{Britannia} and \textit{Germania}.
 Neither we do not detect production centres dedicated to the distribution of olive oil in a certain province.
@@
 
-This research was partly funded by the European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros and Ignacio Morer for their useful suggestions.  
-All data has been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}.
-Source and code are available at \maria{incluir github}.
+This research was partly funded by the European Research Council Advanced Grant EPNet (340828).
+We are grateful to Simon Carrignon, Juan Moros, Ignacio Morer and Víctor M.
+Lozano for their useful suggestions.
+Data have been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}.
+Datasets and source code are freely available at https://github.com/Mcotsar/Baetica\_stamps.
 

### 2020-11-05 Paper/jaccardbaetica.tex
submission ready

@@
 %\title{The markings of the trade: exploring the patterns of olive oil production in Roman Baetica}
-\title{The tracing of trade: exploring the patterns of olive oil production and distribution from Roman \textit{Baetica}}
+\title{The tracing of trade: exploring the patterns of olive oil production and distribution from Roman Baetica}
 
@@
 
-Our case of study has been focused on amphoric production from 1st to 3rd AD.
+Our case of study has been focused on amphoric production from 1st to 3rd centuries AD.
 In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps.
@@
 This work focuses on the debate about the dynamics of the distribution of olive oil market in Roman provinces.
-Amphoric production in \textit{Baetica} province will help us to identify links between production centres and consumption centres in order to detect an economic pattern in the commerce by analysing stamps amphorae finding in the provinces.
+Amphoric production in \textit{Baetica} province will help us identify links between production centres and consumption centres in order to detect an economic pattern in the commerce by analysing stamps amphorae finding in the provinces.
 
@@
 
-The chronology of the workshops is wide spanning from the first to the third centuries AD\citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}. 
+The chronology of the workshops is wide spanning from the first to the third centuries AD~\citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}. 
 This fact is shown by the archaeological evidence that displays a highly specialised production with a long activity \citep{remesal_anforas_2004}.
@@
 
-Scholars agree that stamps are some type of identification mark, but there is no consensus on the meaning of these marks\citep{rodriguez_baetican_1998}.
+Scholars agree that stamps are some type of identification mark, but there is no consensus on the meaning of these marks~\citep{rodriguez_baetican_1998}.
 
@@
 
-However, some hypotheses argue elsewhere that this same \textit{praefectura annonae} could also have organized the supply of additional goods such as olive oil to the Roman legions  \citep{remesal_annona_1986,
+However, some hypotheses argue elsewhere that this same \textit{praefectura annonae} could also have organised the supply of additional goods such as olive oil to the Roman legions  \citep{remesal_annona_1986,
 remesal_annona_1990}
@@
 
-This intense consumption of olive oil also indicates a possible provincial structure designed to organize the supply of olive oil to military camps.
+This intense consumption of olive oil also indicates a possible provincial structure designed to organise the supply of olive oil to military camps.
 
@@
 It seems that some Baetican centres could have been assigned to a specific province for the olive oil supply to military garrisons \citep[125]{remesal_concierto}.
-
 However, this hypothesis is hard to assess given the current lack of written evidence related to the meaning of the amphoric stamps.
 
-Previous studies suggested that the supply to the German limes was mainly based on riverine transport, but recent works suggest that the maritime route through the Atlantic Ocean could have been more important than expected (both for \textit{Germania} and \textit{Britannia}).~\citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
+Previous studies suggested that the supply to the German limes was mainly based on riverine transport, but recent works suggest that the maritime route through the Atlantic Ocean could have been more important than expected (both for \textit{Germania} and \textit{Britannia})~\citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
 
@@
 
-The production area for Dressel 20 amphorae extended across three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter Hispalis), \textit{Cordubensis} (currently C\'ordoba, hereafter Corduba) and \textit{Astigi} (currently Écija, Sevilla, hereafter Astigi) \citep{rodriguez_economioleicola_1977,chicdatos2001,berni_millet_epigrafianforica_2008} .
+The production area for Dressel 20 amphorae extended across three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter \textit{Hispalis}), \textit{Cordubensis} (currently C\'ordoba, hereafter \textit{Corduba}) and \textit{Astigi} (currently Écija, Sevilla, hereafter \textit{Astigi}) \citep{rodriguez_economioleicola_1977,chicdatos2001,berni_millet_epigrafianforica_2008} .
 
@@
 The analysis of stamps in consumption areas also used the CEIPAC database and filtered in the same way than the previous dataset.
-However, in this case, we selected the centres with more or equal than five stamps in order to maximize the analysis with more stamps in the dataset.
+However, in this case, we selected the centres with more or equal than five stamps in order to maximise the analysis with more stamps in the dataset.
 
@@
 Generally, it describes the dissimilarity between the system of two communities based on the idea of inverse correlation between diversity and species \citep{magurran_why_1988}.
-In our case, stamps from different sites will be used as a proxy to measure their dissimilarity by analyzing their code.
+In our case, stamps from different sites will be used as a proxy to measure their dissimilarity by analysing their code.
 
@@
 Morisita-Horn index gives a value between 0 (sites have exact stamp codes and frequency distribution) and 1 (complete difference between stamp codes).
-To apply Morisita-Horn, we first calculated the number of times that each stamp code appeared in an amphora workshop.
+To apply Morisita-Horn index, we first calculated the number of times that each stamp code appeared in an amphora workshop.
 This method allowed us to bear in mind the number of frequency of stamps for each workshop.
@@
 The hierarchical clustering of the Morisita-Horn pairwise distance matrix can be seen as a dendrogram in Figure~\ref{dendro}.
-This visualization suggests that each amphora workshop often used unique stamps in their production system that have not been found on any other site.
+This visualisation suggests that each amphora workshop often used unique stamps in their production system that have not been found on any other site.
 In fact, the similarity values are always low and a majority of stamp codes found on multiple sites are only present in two or three workshops.
@@
 \caption{Dendrogram displaying the results of hierarchical clustering based on Morisita-Horn metrics for Dressel 20 workshops located in the \textit{Baetica} area.
-Site names are color-coded by conventus: \textit{Hispalis} (red), \textit{Astigi} (green) and \textit{Corduba} (blue)}
+Site names are colour-coded by \textit{conventus}: \textit{Hispalis} (red), \textit{Astigi} (green) and \textit{Corduba} (blue)}
 \label{dendro}
@@
 \caption{Dendrogram displaying the results of hierarchical clustering based on Morisita-Horn metrics for sites in \textit{Britannia}.
-Site names are color-coded by typology: military sites (red), civilian sites (green) and unspecified (grey)}
+Site names are colour-coded by typology: military sites (red), civilian sites (green) and unspecified (grey)}
 \label{britmap}
@@
 \caption{Dendrogram displaying the results of hierarchical clustering based on Morisita-Horn metrics for sites in \textit{Germania}.
-Site names are color-coded by typology: military sites (red), civilian sites (green) and unspecified (grey)}
+Site names are colour-coded by typology: military sites (red), civilian sites (green) and unspecified (grey)}
 \label{germap}
@@
 This could be explained such as batch systematic organisation between potters \citep{juanmorostesis}.
-This work method could have allowed potters to identify batches by stamping amphorae for posterior commercialization.
+This work method could have allowed potters to identify batches by stamping amphorae for posterior commercialisation.
+High production of amphorae had to maintain a specialise system organisation to control this production \citep[104]{juanmorostesis}.
 
 %This work method could have allowed to potters to stamp amphorae in order to identify the production for batches for posterior commercialization.
-
-High production of amphorae had to maintain a specialize system organization to control this production \citep[104]{juanmorostesis}.
 
@@
 
-Finally, the results suggest that there is no clear organizational pattern linking production centres and/or consumption centres beyond the spatial distance.
+Finally, the results suggest that there is no clear organisational pattern linking production centres and/or consumption centres beyond the spatial distance.
 As a consequence, our analysis does not indicate the presence of a specific workshop from \textit{Baetica} supplying to some specific province.
@@
 The most interesting result is linked to the meaning of these inscriptions.
-The use of specific codes seems to be almost exclusively decided by each workshop, so it could be used to identify amphorae found in consumption centers (i.e.
+The use of specific codes seems to be almost exclusively decided by each workshop, so it could be used to identify amphorae found in consumption centres (i.e.
 an amphora with a given stamp was probably made on a specific workshop).
@@
 
-To conclude, the method presented here provides a useful framework to improve our understanding of the organisation and processes that allowed the Roman Empire to manage a massive and highly specialized production of essential goods.
+To conclude, the method presented here provides a useful framework to improve our understanding of the organisation and processes that allowed the Roman Empire to manage a massive and highly specialised production of essential goods.
 This work has identified similarities between amphoric stamps and spatial distribution in the case of the amphoric production within the Roman Empire.
@@
 Data have been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}.
-Datasets and source code are freely available at https://github.com/Mcotsar/Baetica\_stamps.
+Maps were performed using QGIS software v.3.14 `Pi'.
+Datasets and source code are freely available under Open licenses from https://github.com/Mcotsar/Baetica\_stamps.
 

### 2020-11-05 Paper/jaccardbaetica.tex
finished and submit

@@
 
-\address[ceipacadress]{Department of Prehistory and Archaeology, Montalegre, 6-8, 08001, University of Barcelona, Barcelona, Spain}
+\address[ceipacadress]{Department of Prehistory and Archaeology, Montalegre, 6-8, 08001, Universitat de Barcelona, Barcelona, Spain}
 \address[edadress]{School of History, Classic \& Archaeology, Room OOM.33, William Robertson Wing, Old Medical School, Teviot Place, University of Edinburgh, UK}
-\address[ubadress]{DIDPATRI, Passeig de la Vall d'Hebrón, 171, Barcelona, Spain}
+\address[ubadress]{DIDPATRI, Universitat de Barcelona, Passeig de la Vall d'Hebrón, 171, Barcelona, Spain}
 
@@
 
-To achieve this goal, we analyse a set of stamps from two centres: 1) production centres by analysing different workshops in \textit{Baetica} province and 2) two Roman provinces such as \textit{Germania} (Inferior and Superior) and \textit{Britannia} as consumption centres.
+To achieve this goal, we analyse a set of stamps from two types of centres (production and consumption): 1) production centres by analysing different workshops in \textit{Baetica} province and 2) two Roman provinces such as \textit{Germania} (Inferior and Superior) and \textit{Britannia} as consumption centres.
 They will be used to detect a connection between the distribution of amphoric stamps and the economic structure in both centres.
@@
 This research was partly funded by the European Research Council Advanced Grant EPNet (340828).
-We are grateful to Simon Carrignon, Juan Moros, Ignacio Morer and Víctor M.
-Lozano for their useful suggestions.
+XRC is funded by the Ramon y Cajal programme RYC2018-024050-I (Fondo Social Europeo – Agencia Estatal de Investigación).
+We are grateful to Simon Carrignon, Juan Moros, Ignacio Morer and Víctor M. Lozano for their useful suggestions.  
 Data have been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}.

### 2021-01-29 Manuscript/Manuscript.tex
manuscript corrected

--- 2020-11-05 Paper/jaccardbaetica.tex
@@
-\documentclass[review]{elsarticle}
-\usepackage{hyperref,lineno}
+%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% This is a general template file for the LaTeX package SVJour3
+% for Springer journals.          Springer Heidelberg 2010/09/16
+%
+% Copy it to a new file with a new name and use it as the basis
+% for your article. Delete % signs as needed.
+%
+% This template includes a few options for different layouts and
+% content for various journals. Please consult a previous issue of
+% your journal as needed.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% First comes an example EPS file -- just ignore it and
+% proceed on the \documentclass line
+% your LaTeX will extract the file if required
+\begin{filecontents*}{example.eps}
+%!PS-Adobe-3.0 EPSF-3.0
+%%BoundingBox: 19 19 221 221
+%%CreationDate: Mon Sep 29 1997
+%%Creator: programmed by hand (JK)
+%%EndComments
+gsave
+newpath
+  20 20 moveto
+  20 220 lineto
+  220 220 lineto
+  220 20 lineto
+closepath
+2 setlinewidth
+gsave
+  .4 setgray fill
+grestore
+stroke
+grestore
+\end{filecontents*}
+%
+\RequirePackage{fix-cm}
+%
+%\documentclass{svjour3}                     % onecolumn (standard format)
+%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
+\documentclass[smallextended]{svjour3}       % onecolumn (second format)
+%\documentclass[twocolumn]{svjour3}          % twocolumn
+%
+\smartqed  % flush right qed marks, e.g. at end of proof
+%
+
+\usepackage{hyperref}
+\usepackage{graphicx}
 \usepackage{wrapfig}
@@
 \usepackage{rotating}
-\usepackage{xcolor}
-\usepackage[utf8]{inputenc}
-\modulolinenumbers[5]
+%
+% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
 
-\newcommand{\memo}[2]{\textcolor{#1}{#2}}
-\newcommand{\maria}[1]{\memo{red}{MC: #1\\}}
-\newcommand{\xavi}[1]{\memo{magenta}{XRC: #1\\}}
+\usepackage{natbib}
+%\bibliographystyle{plainnat}
+\bibliographystyle{agsm}
 
-\journal{Journal of Archaeological Method and Theory}
-
-\bibliographystyle{model2-names.bst}\biboptions{authoryear}
+%\bibliographystyle{model2-names.bst}%\biboptions{authoryear}
+%\bibliographystyle{spbasic}
+%\bibliographystyle{spphys}
+%\bibliographystyle{spmpsci}
+% insert here the call for the packages your document requires
+%\usepackage{latexsym}
+% etc.
+%
+% please place your own definitions here and don't use \def but
+% \newcommand{}{}
+%
+% Insert the name of "your journal" with
+\journalname{Archaeological and Anthropological Sciences}
 
@@
 
-\begin{frontmatter}
+\title{The tracing of trade%\thanks{Grants or other notes
+%about the article that should go on the front page should be
+%placed here. General acknowledgments should be placed at the end of the article.}
+}
+\subtitle{Exploring the patterns of olive oil production and distribution from Roman Baetica}
 
-%\title{The markings of the trade: exploring the patterns of olive oil production in Roman Baetica}
-\title{The tracing of trade: exploring the patterns of olive oil production and distribution from Roman Baetica}
+%\titlerunning{Short form of title}        % if too long for running head
 
-\author[ceipacadress]{María Coto-Sarmiento\corref{mycorrespondingauthor}}
-\cortext[mycorrespondingauthor]{Corresponding author}
-\ead{mcotsar@gmail.com}
+\author{Mar\'ia Coto-Sarmiento       \and Xavier Rubio-Campillo}
 
-\author[edadress,ubadress]{Xavier Rubio-Campillo}
+%\authorrunning{Short form of author list} % if too long for running head
 
-\address[ceipacadress]{Department of Prehistory and Archaeology, Montalegre, 6-8, 08001, Universitat de Barcelona, Barcelona, Spain}
-\address[edadress]{School of History, Classic \& Archaeology, Room OOM.33, William Robertson Wing, Old Medical School, Teviot Place, University of Edinburgh, UK}
-\address[ubadress]{DIDPATRI, Universitat de Barcelona, Passeig de la Vall d'Hebrón, 171, Barcelona, Spain}
+\institute{Mar\'ia Coto-Sarmiento \at Department of Prehistory and Archaeology, Montalegre, 6-8, 08001, Universitat de Barcelona, Barcelona, Spain
+\email{mcotsar@gmail.com}            
+\emph{mcotsar@gmail.com} 
+%of F. Author  %  if needed
+           \and
+\\
+Xavier Rubio-Campillo \at
+School of History, Classic \& Archaeology, Room OOM.33, William Robertson Wing, Old Medical School, Teviot Place, University of Edinburgh, UK \at
+DIDPATRI, Universitat de Barcelona, Passeig de la Vall d'Hebr\'on, 171, Barcelona, Spain             
+}
 
-\begin{keyword}
-Roman Empire; amphora production; Dressel 20; dissimilarity index; Roman provinces
-\end{keyword}
+\date{Received: date / Accepted: date}
+% The correct dates will be entered by the editor
+
+\maketitle
 
@@
 
-%Our case of study has been focused on the production processes located in \textit{Baetica} province (currently Andalusia) from 1st to 3rd AD.
-In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps.
-\textit{Baetica} became an important production and distribution centre during the Roman Empire.
-However, it remains under debate about how this province was organised and whether it could be possible to identify patterns in the olive oil market.
-Amphoric stamps are used to identify the presence of different groups that might share similar stamps.
-To achieve this goal, we analyse a set of stamps from two centres: 1) production centres by analysing different workshops in \textit{Baetica} province and 2) two Roman provinces such as \textit{Germania} and \textit{Britannia} as consumption centres.
-They will be used to detect a connection between the distribution of amphoric stamps and the economic structure in both centres.
-Here, we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance.
-The analysis explores how the quantitative approach provides a useful tool for the interpretation of the economic processes.
-Finally, results pretend to highlight the organisation of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.
-
+%Insert your abstract here. Include keywords, PACS and mathematical subject classification numbers as needed.
+\keywords{Roman Empire \and amphora production \and Dressel 20 \and dissimilarity index \and Roman provinces}
+% \PACS{PACS code1 \and PACS code2 \and more}
+% \subclass{MSC code1 \and MSC code2 \and more}
 \end{abstract}
 
-\end{frontmatter}
-
 \section{Introduction}
+\label{intro}
 
@@
 The intensity of this trade between distant regions can be observed both in archaeological and written evidence~\citep{rodriguez_baetican_1998}.
-
-%Material culture is one of the most frequent indicators of trade in the archaeological record.
-In archaeology, they allow us to highlight a part of the mechanism of production and distribution of goods along the Mediterranean \citep{bevan_mediterranean_2014}.
-Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}
-%spread of goods and ideas had a enormous impact during the Roman Empire. The progressive exploitation of communication networks allowed a major interaction between communities
-%This can be seen by the fact that material culture is found in different regions along the Mediterranean as a frequent indicator of trade exchange in the archaeological record.
 
@@
 
-Advances in the research of the Roman studies have currently led an environment with various interpretations on commercial dynamics \citep{duncan1982economy,
+Advances in Roman studies have currently led to an environment with various interpretations on commercial dynamics \citep{duncan1982economy,
 temin_economy_2006,
 quantifyingwilson2009}.
-%\xavi{more diverse que qué?} 
+
 The application of different quantitative approaches has allowed us to improve our interpretation of the growing amount of archaeological evidence and reveal the dynamics of Roman production
@@
 This work focuses on the debate about the dynamics of the distribution of olive oil market in Roman provinces.
-Amphoric production in \textit{Baetica} province will help us identify links between production centres and consumption centres in order to detect an economic pattern in the commerce by analysing stamps amphorae finding in the provinces.
+Amphoric production in \textit{Baetica} province will help us identify links between production centres and consumption centres in order to detect an economic pattern in the commerce by analysing amphoric stamps finds in the provinces.
 
 The finding of amphoric stamps in different provinces has allowed us to know the importance of the distribution of olive oil within the Empire.
-Despite having archaeological evidence, no consensus has been reached: How was the economical relation between production and consumption centres?
-Did they follow an economical pattern to distribute olive oil in those provinces?
+Despite having archaeological evidence, no consensus has been reached: How was the economic relation between production and consumption centres?
+Did they follow an economic pattern to distribute olive oil in those provinces?
 
-This paper aims to study the olive oil market connection between provinces by calculating the similarity of stamps.
-Specifically, our work pretends to detect regional processes focused on a commercial product from a specific province \citep{isaksen_network_2006}.
-We want to understand the pattern of olive oil production linked to amphora workshops and amphoric stamps used to mark them.
-We focus here on exploring the economic relation between stamps and amphora production and distribution centres.
+This paper aims to study the olive oil market connection between provinces by calculating the similarity of amphoric stamps.
+Specifically, our work aims to detect regional processes focused on a commercial product (i.e.
+olive oil) from a specific province \citep{isaksen_network_2006}.
+
+%We want to understand the pattern of olive oil production linked to amphora workshops and stamps used to mark them. 
+%We focus here on exploring the economic relation between stamps and amphora production and distribution centres.
 
@@
 	\centering
-\includegraphics[width=\linewidth]{figs/general_map}
+\includegraphics[width=\linewidth]{general_map}
 \caption{Overview of the sites analysed in this paper.
@@
 
-In particular, we study the distribution of amphoric stamps to identify a correlation between geographical distance and similarity.
-Based on this assumption, we proposed three hypotheses: a) we can identify a correlation between spatial distance and the distribution of stamps, b) stamps located in close workshops share similar traits and c) low mobility of amphoric stamps to other regions: stamps always stay in the same region.
+%In particular, we study the distribution of amphoric stamps to identify a correlation between geographical distance and similarity.
+
+Based on this assumption, we proposed three hypotheses: a) we can identify a correlation between spatial distance and the distribution of amphoric stamps, b) stamps located in close workshops share similar traits and c) low mobility of amphoric stamps to other regions: stamps always stay in the same region.
 
@@
 To do this, a population approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}.
-A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps.
+Quantitative approach provides one way to measure the similarity between amphoric stamps.
 Here, we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
-Stamps will be used to identify economic patterns by analysing their similarity.
-If workshops and provinces share stamps with similar traits, then we can identify connections.
+
+%Stamps will be used to identify economic patterns by analysing their similarity.
+
+For example, if workshops and provinces share stamps with similar traits, then we can identify connections.
 By contrast, if we do not detect similar stamps between workshops, then workshops worked independently.
@@
 \section{The Amphoric production of Baetica}
+\label{sec:1}
 
@@
 For this reason, the area saw an increase in productivity as a massive infrastructure of olive oil production was gradually deployed.
-The production and distribution of olive oil increased from the first to the third centuries AD~\citep{remesal_concierto}.
+The production and distribution of olive oil increased from the first to the third century AD~\citep{remesal_concierto}.
 
 Hundreds of presses and amphora-making workshops were built near large extensions covered by olive trees.
-The workshops where amphorae were made and filled with olive oil were places along the rivers Guadalquivir and Genil riverbanks (see Fig.\ref{workshop}).
+The workshops where amphorae were made and filled with olive oil were places along the rivers Guadalquivir and Genil riverbanks (see Fig.\ref{workshop}).The proximity distance between workshops was relatively close which allowed for more efficient distribution.
 
@@
 	\centering
-\includegraphics[width=\linewidth]{figs/baetica}
+\includegraphics[width=\linewidth]{baetica}
 \caption{Distribution of stamps found in Dressel 20 workshops along the Guadalquivir riverbank. The size of each point depends on the number of stamps found on each site}
@@
 
-%Some stamps show a more specific chronology while the majority of them display a large activity of production that it can be difficult for specifying an accurate chronology in some cases.
-This could be due to two reasons; firstly, most of the workshops were partially excavated and focused on archaeological surveys in order to collect the maximum stamps as possible; secondly, Dressel 20 was produced during almost three centuries with apparently few changes \citep{berni_piero_chapter_2017}.
-
 \section{A potential indicator of Roman economy: Dressel 20 stamps}
+\label{sec:2}
 
@@
 	\centering
-\includegraphics[scale=0.5]{figs/dressel20}
+\includegraphics[scale=0.5]{dressel20}
 \caption{Dressel 20 were mostly marked with stamps of three letters called \textit{tria nomina}}
@@
 These stamps are typically found on handles but they were also imprinted on rims and the amphora body~\citep{millet_anforas_1998}.
-The content information of the stamps is shown in different forms and letter content and it seems that there was not a unique criterion.
-
+The content information of the stamps had not a unique criterion.
+They showed a code with different letters and forms.
 Stamps often displayed a code of three letters and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
@@
 
-%Additionally, a large percentage of Dressel 20 containers were not stamped thus complicating the interpretation
-
 The stamp codes are interpreted based on three main ideas: content (olive oil), context (amphora workshop) and subject (individuals involved in the production).
@@
 On the other hand, they could also belong to the owner of the workshops where the amphorae were made or even to a group of amphora workers~\citep{berni_millet_epigrafianforica_2008}.
-%In any case, the use of these stamps is a good proxy to explore the system of olive oil production and trade developed in Baetica.
 
@@
 
-%\subsubsection{Jaccard distance}
-%The dataset was analysed using a statistic method as Jaccard distance.
-This method allows to measure the dissimilarity by calculating the presence of sets (CITAR).
-In our case, Jaccard distance was used to compute the mutual presence of traits in the amphora stamps but it does not consider the number of absences.
-A comparison was done with the distance of the workshops to identify whether there was an association between stamps and spatial distance amongst workshops.
-
 \section{Consumption centres: Britannia and Germania}
+\label{sec:3}
 
 The creation of new provinces required the Roman Empire the arrival of resources through Mediterranean and Atlantic routes.
+This effectively led to a gradual change in the economic and social structures when the trade networks that supplied the Roman army became more complex.
 
-This led to a gradual change in the economic and social structures the trade networks that supplied the Roman army became more complex.
 Augustus's administration created the figure of the \textit{praefectura annonae} to organize wheat supply.
@@
 
-%The importance of this olive oil trade is revealed by the massive amount of Dressel 20 amphorae found across the Roman provinces.
-This amphora has been commonly associated with the transportation of Baetican olive oil for supplying military camps and civil settlements \citep{berni_millet_epigrafianforica_2008}.\xavi{me parece repetitivo, porque unos párrafos antes ya hablabas de lo mismo.
-Quizás unificar texto aquí o allá?} LO HE PUESTO MÁS ARRIBA!
-
 Olive oil supply seems to be particularly intense in militarised provinces that functioned as the borders of the Empire such as \textit{Britannia} and \textit{Germania}.
@@
 An explanation for the interest to export olive oil for the military garrison in the Roman provinces is still unclear.
-The origin of this exportation may be related to two aspects: a) cultural consumption whose olive oil is used by cultural reasons such as group identification or by habit and b) economical and political reasons to create a redistributive system mechanism to supply military camps \citep[69-70]{carreras_britannia_1998}.
+The origin of this exportation may be related to two aspects: a) consumption based on cultural reasons (i.e.
+social identification or habit  by a group of consumers) and b) economical and political reasons to create a redistributive system mechanism to supply military camps \citep[69-70]{carreras_britannia_1998}.
 
@@
 \subsection{Britannia}
+\label{sec:4}
 
 The evidence for olive oil consumption in \textit{Britannia} is scarce before the Roman conquest \citep{funari_corpus_1996,carreras_abastecimiento_2003}.
-The local population did not consume this product and it was not produced on the region as the landscape and climate of the British Isles was not suitable for olive oil trees\citep[161]{monfort_britanniaen_1998}.
+The local population did not consume this product and it was not produced on the region as the landscape and climate of the British Isles was not suitable for the growth of olive trees~\citep[161]{monfort_britanniaen_1998}.
 Based on the archaeological records, this absence of olive oil changed with the arrival of a large amount of Dressel 20 amphorae to \textit{Britannia} from the first century A.D \citep{peacock_amphorae_1991,
@@
 
-%At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}\xavi{repetitivo?
-fusionar párrafos}.
-
 This fact will have a particular intensity within sites close to the Hadrian Wall's garrisons.
 Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic locations~\citep{carreras_atlantic_2012}.
-
-%The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges.
-Thus, the network was mainly focused on the support of soldiers during military campaigns\xavi{estas 2 frases repiten cosas ya explicadas}.
 
@@
 
-There are no written records explaining how this redistribution of essential goods was organised in \textit{Britannia}, but the archaeological evidence suggests that cities may have been the central nodes of this local trade network~\citep[45]{funari_economic_2005}.
+There are no written records explaining how this redistribution of essential goods was organised in \textit{Britannia}, but the archaeological evidence suggests that cities may have been the central nodes of this local trade network~\citep{funari_economic_2005,orengo_seeds_2016}.
 
@@
 \subsection{Germania}
+\label{sec:4}
 
@@
 By contrast, it is still unknown if trade agents participated in the distribution of Baetican olive oil in \textit{Germania} or the supply was exclusively focused on the Roman army garrisons\citep[156]{remesal_germn_2010}.
-
-%This could be explained due to the lack of archaeological sources \citep{horacio2010llegada}
 
@@
 
-Previous studies suggested that the supply to the German limes was mainly based on riverine transport, but recent works suggest that the maritime route through the Atlantic Ocean could have been more important than expected (both for \textit{Germania} and \textit{Britannia})~\citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
+Previous studies suggested that the supply to the German \textit{limes} was mainly based on riverine transport, but recent works suggest that the maritime route through the Atlantic Ocean could have been more important than expected (both for \textit{Germania} and \textit{Britannia})~\citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
 
 \section{Material and Methods}
+\label{sec:5}
 
@@
 The CEIPAC dataset contains over 50.000 epigraphy records found on different types of amphorae, mostly from \textit{Monte Testaccio}.
+Most stamps used in the analysis did not have a specific chronology or were not available in the dataset.
+This fact did that we had to discard the possibility to analyse a distribution of stamps per chronologies.
 
-%\xavi{podrías mover esta parte a los 2 párrafos anteriores y explicar la bbdd brevemente antes de empezar a hablar de la Bética (porque es info que también aplica a los otros 2 casos de estudio no?}.
-
-%This study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps.
-A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps.
-Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potential spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
-%\xavi{estos 2 párrafos explican objetivos, que deberían ir en la intro y no aquí, donde deberías empezar a hablar directamente de la BBDD}
+In order to enable reproducibility of the results, the selected datasets and source codes are available on the following GitHub repository: \url{https://github.com/Mcotsar/Baetica\_stamps}.
 
 \subsection{Production centres: Baetica province}
+\label{sec:5}
 
 The database allowed us to retrieve 3798 stamps found in Dressel 20 amphorae recovered from workshops within \textit{Baetica} province.
-Each of these stamps ha detailed records on the site where it was found, the inscribed code as well as its spatial coordinates.
-The stamps with incomplete information were discarded, thus finishing with a sample of 987 stamps from 81 sites and displaying 130 different code.
-
-%Approximately 70 \% of stamps cannot be tested due to the fragmentation of the dataset. Consequently, any stamps with incomplete information were discarded and not integrated into our dataset.
+Each of these stamps has detailed records at the site where it was found, the inscribed code as well as its spatial coordinates.
+The stamps with incomplete information were discarded, thus finishing with a sample of 987 stamps from 81 sites and displaying 130 different codes.
+Query dataset from Baetica province can be found on the following GitHub repository: \url{https://github.com/Mcotsar/Baetica_stamps/tree/main/Databases/Baetica}.
 
@@
 
-The production area for Dressel 20 amphorae extended across three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter \textit{Hispalis}), \textit{Cordubensis} (currently C\'ordoba, hereafter \textit{Corduba}) and \textit{Astigi} (currently Écija, Sevilla, hereafter \textit{Astigi}) \citep{rodriguez_economioleicola_1977,chicdatos2001,berni_millet_epigrafianforica_2008} .
+The production area for Dressel 20 amphorae extended across three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter \textit{Hispalis}), \textit{Cordubensis} (currently C\'ordoba, hereafter \textit{Corduba}) and \textit{Astigi} (currently \'Ecija, Sevilla, hereafter \textit{Astigi}) \citep{rodriguez_economioleicola_1977,chicdatos2001,berni_millet_epigrafianforica_2008} .
 
@@
 It is important to mention that some workshops exhibited the same geographical coordinates while they were catalogued differently as workshops.
-For instance, two nearby amphora workshops such as Tesorillo de Doña Mencia and Doña Mencia were catalogued as a different name in spite of having the same geographical coordinates.
+For instance, two nearby amphora workshops such as Tesorillo de Do\~na Mencia and Do\~na Mencia were catalogued as a different name in spite of having the same geographical coordinates.
 Our work considered this particularity in the archaeological dataset but some biases can occur in the analysis.
@@
 
-%\xavi{no lo veo; recuerda que para el rollo bayesiano tenías una línea recta al hacedr el log de los 2 ejes, no solo 1, y este plot no está construído así no?}.
-%\maria{son dos ejes, no? ¿aquí no se ve línea recta? Lo mismo si lo hago de puntitos...}
-
 \begin{figure}[htp]
 	\centering
-\includegraphics[width=\linewidth]{figs/frequencystamp.pdf}
+\includegraphics[width=\linewidth]{frequencystamp.pdf}
 \caption{Histogram on a log scale with base 10.
@@
 
-The distribution of code stamps for each \textit{conventus} can be seen in Figure~\ref{frequency}.
-The majority of stamps are concentrated in \textit{Hispalis} (574 code stamps) while \textit{Corduba} and \textit{Astigi} have roughly half this sample size (267 and 146 code stamps).
+The distribution of codes in stamps for each \textit{conventus} can be seen in Figure~\ref{frequency}.
+The majority of stamps are concentrated in \textit{Hispalis} (574 stamps) while \textit{Corduba} and \textit{Astigi} have roughly half this sample size (267 and 146 stamps respectively).
 The workshops of the three \textit{conventus} show a similar pattern on the stamp frequency distribution with the exceptions of 2 large workshops in \textit{Hispalis} (La Catria and Arva).
-On these two workshops a comparatively large amount of code stamps were found (29 different code stamps on each of them).
+On these two workshops a comparatively large amount of stamp codes were found (29 different codes on each of them).
 
@@
 	\centering
-\includegraphics[width=\linewidth]{figs/frequency}
-\caption{Distribution of the number of different code stamps (X axis) for each \textit{conventus} (Y axis).
+\includegraphics[width=\linewidth]{frequency}
+\caption{Distribution of the number of different codes in stamps (X axis) for each \textit{conventus} (Y axis).
 Each dot corresponds to a workshop sorted by different areas.
@@
 \subsection{Consumption centres: Britannia and Germania}
+\label{sec:5}
 
@@
 The output was a dataset of 4271 stamps found in sites belonging to \textit{Britannia} (total: 2219 stamps, analysed: 1765 stamps ) and \textit{Germania} (total: 2052 stamps, analysed: 1621 stamps).
-Some stamps from \textit{Britannia} and \textit{Germania} were discarded for the analysis for incomplete information.
+The stamps with incomplete information were also not taken into account for the analysis.
+
+Query datasets from both provinces can be found on the following GitHub repository: Britannia~ \url{https://github.com/Mcotsar/Baetica_stamps/tree/main/Databases/Britannia} and Germania~ \url{https://github.com/Mcotsar/Baetica_stamps/tree/main/Databases/Germania}.
 
@@
 	\centering
-\includegraphics[width=\linewidth]{figs/britannia}
+\includegraphics[width=\linewidth]{britannia}
 \caption{Sites in \textit{Britannia} where Dressel 20 amphoric stamps have been found.
@@
 
-%We analysed a dataset of 2052 stamps placed in \textit{Germania}.
-All the data was also compiled by CEIPAC database.
-As previously before, the centres were selected with more or equal than five stamps.
-
 In \textit{Germania}, we collected a total of 1621 stamps from 46 sites displaying 850 different stamp codes (see Figure~\ref{germania}).
@@
 	\centering
-\includegraphics[width=\linewidth]{figs/germania}
+\includegraphics[width=\linewidth]{germania}
 \caption{Sites in Germania where Dressel 20 amphoric stamps have been found.
@@
 \subsection{Measuring the dissimilarity}
-
-%quizás habría que hablar también del filtrado de códigos que se hizo en python porque antes había 3783 stamps pero se filtró por el número de letras...
-%\xavi{lo del filtrado lo puedes poner de supplementary information con referencia al antiquity}
+\label{sec:5}
 
@@
 The chosen dissimilarity measure was the Morisita-Horn index \citep{morisita_measuring_1959, horn_measurement_1966}.
-This method was applied to measure the dissimilarity between different samples of sets.
+This method was applied to measure the dissimilarity between different samples of stamps.
 Generally, it describes the dissimilarity between the system of two communities based on the idea of inverse correlation between diversity and species \citep{magurran_why_1988}.
@@
 This method allowed us to bear in mind the number of frequency of stamps for each workshop.
-
-If two workshops had similar stamp code distribution then the distance index would be close to 0 whereas sites with completely unrelated stamp codes would give a distance close to 1.
+For example, if two workshops had similar stamp code distribution then the distance index would be close to 0 whereas sites with completely unrelated stamp codes would give a distance close to 1.
 
 \subsection{Hierarchical clustering}
+\label{sec:5}
 
-The Morisita-Horn index was used to generate a dissimilarity matrix containing the distance value between each pairwise site based on their code stamps.
-Hierarchical clustering was applied to this matrix in order to group sites with similar stamps codes distributions.
+The Morisita-Horn index was used to generate a dissimilarity matrix containing the distance value between each pairwise site based on their codes stamped in amphorae.
+Hierarchical clustering was applied to this matrix in order to group sites with similar stamp codes distributions.
 The algorithm was selected to cluster similar groups in order to analyse the relationship between groups of sites and the distribution of similar stamp codes.
@@
 \section{Results}
+\label{sec:6}
 
 \subsection{Production centres: Baetica Province}
-
-%The correlation coefficients range from a minimum to a maximum.
+\label{sec:6}
 
@@
 
-\begin{sidewaysfigure}[htp]
+%\begin{sidewaysfigure}[htp]
+\begin{figure}[htp]
 	\centering
-\includegraphics[angle=180, width=\linewidth]{figs/dendro}
+\includegraphics[angle=270, width=0.7\linewidth]{dendro}
 \caption{Dendrogram displaying the results of hierarchical clustering based on Morisita-Horn metrics for Dressel 20 workshops located in the \textit{Baetica} area.
@@
 \label{dendro}
-\end{sidewaysfigure}
+%\end{sidewaysfigure}
+
+\end{figure}
 
 \subsection{Consumption centres: Britannia and Germania}
+\label{sec:6}
 
@@
 
-\begin{sidewaysfigure}[htp]
+%\begin{sidewaysfigure}[htp]
+\begin{figure}[htp]
 	\centering
-\includegraphics[angle=180,width=\linewidth]{figs/dendrobrit5.pdf}
+\includegraphics[angle=270,
+width=0.7\linewidth]{dendrobrit5.pdf}
 \caption{Dendrogram displaying the results of hierarchical clustering based on Morisita-Horn metrics for sites in \textit{Britannia}.
@@
 \label{britmap}
-\end{sidewaysfigure}
+%\end{sidewaysfigure}
+\end{figure}
 
@@
 
-\begin{sidewaysfigure}[htp]
+%\begin{sidewaysfigure}[htp]
+\begin{figure}
 	\centering
-\includegraphics[angle=180, width=\linewidth]{figs/dendroger5.pdf}
+\includegraphics[angle=270, width=0.7\linewidth]{dendroger5.pdf}
 \caption{Dendrogram displaying the results of hierarchical clustering based on Morisita-Horn metrics for sites in \textit{Germania}.
@@
 \label{germap}
-\end{sidewaysfigure}
+\end{figure}
+%\end{sidewaysfigure}
 
@@
 \section{Discussion}
+\label{sec:7}
 
@@
 \subsection{Production centres}
+\label{sec:7}
 
@@
 Beyond this clear pattern, it seems that the similarity of stamp codes is correlated with spatial distance and the amphoric production of spatially close sites shared some stamp codes.
-Particularly relevant here is the fact that our results show how similar stamp codes were found on sites belonging to the same \textit{conventus} area.
+Particularly relevant here is the fact that our results show that each \textit{conventus} area had similar stamps codes and they were not shared in other areas.
 This result could indicate that the workshops of the different \textit{conventus} used different stamp codes that were not shared between them.
@@
 
-%This work method could have allowed to potters to stamp amphorae in order to identify the production for batches for posterior commercialization.
-
-%This method allowed to potters organises the production with the batch stamping for the posterior commercialisation
-
 It is worth mentioning that the performed analysis could be limited by a diversity of biases.
@@
 \subsection{Consumption centres}
+\label{sec:7}
 
@@
 
-This trend is aligned with recent works that have suggested that the Atlantic trade routes were mostly used for military supply~\citep{remesal_annona_1986,remesal_provincial_2008,carreras_atlantic_2012,morillo_hispania_2016,rubio-campillo_provincias_2018}.
-Therefore, a majority of centres with higher similarity are related to military activities and near the coasts, thus suggesting that Dressel 20 containing olive oil were shipped to military areas and then redistributed to the civilian population using land-based transportation systems~\citep{carreras_britannia_1998,ayllon_olive_2018}.
-
-%The results for \textit{Germania} follow a similar pattern than \textit{Britannia} but with a minor correlation as we can observe in the dendrogram\xavi{esta frase está repetida de los resultados.
-quizá quitarla de ahí y dejarla aquí?}.
+This trend is aligned with recent works that have suggested that the Atlantic trade routes were mostly used for military supply~\citep{remesal_annona_1986,
+remesal_provincial_2008,
+carreras_atlantic_2012,
+morillo_hispania_2016}.
+Therefore, a majority of centres with higher similarity are related to military activities and near the coasts, thus suggesting that Dressel 20 containing olive oil were shipped to military areas and then redistributed to the civilian population using land-based transportation systems~\citep{carreras_britannia_1998,
+orengo_seeds_2016,
+ayllon_olive_2018}.
 
 The areas mostly militarised share similar stamps than civil areas.
-However, we do not detect a concrete pattern regarding the distribution of the stamps in the German limes \citep{xanten2018}.
+However, we do not detect a concrete pattern regarding the distribution of the stamps in the German \textit{limes} \citep{xanten2018}.
 This could be also interpreted by an intensity bias where military centres have been mostly excavated than civil centres.
-
-%\xavi{pero no decías que el aceite llegaba por las legione? Entonces por qué es un bias? En todo caso ponlo al final de la sección y no al inicio}.
-
-%It is worth mentioning that it was not detected a robust model of organisation in both cases with regard to the distribution of stamps in consumption centres\xavi{de nuevo esto es redundante no?}.
 
@@
 
-%\xavi{me petaría esta frase, porque entre otras cosas arriba dices que hay correlación espacial, así que cómo es que aquí dices que no hay distribución geográfica aparente?}Judging by the results obtained, there does not seem to be a specific pattern in terms of geographical distribution.
-Nor it was detected consumption areas where stamps are specified from an amphora workshop.
-
 \section{Concluding remarks}
+\label{sec:8}
 
@@
 This idea could explain why the similarity of stamps between workshops is generally very low.
-On the other hand, some code stamps have been identified in nearby workshops so this pattern suggests that the clustered workshops were somehow connected (e.g.
+On the other hand, some codes in stamps have been identified in nearby workshops so this pattern suggests that the clustered workshops were somehow connected (e.g.
 they belonged to the same owner).
-
-%Alternatively, the presence of different stamps in the same workshop could imply some kind of organisation from within the workshop only affecting nearby centres \citep{juanmorostesis}.
-The stamped amphorae could have allowed potters to organise the different production batches for posterior commercialisation\xavi{estos 2 párrafos igual deberían ir a la discusión no?
-La conclusión debería ser algo más general y aquí estás aún interpretando resultados}.
 
@@
 
-\section{Acknowledgements}
+%\section{Acknowledgements}
+%\label{sec:9}
+
+\begin{acknowledgements}
 
 This research was partly funded by the European Research Council Advanced Grant EPNet (340828).
-XRC is funded by the Ramon y Cajal programme RYC2018-024050-I (Fondo Social Europeo – Agencia Estatal de Investigación).
-We are grateful to Simon Carrignon, Juan Moros, Ignacio Morer and Víctor M. Lozano for their useful suggestions.  
+XRC is funded by the Ram\'on y Cajal programme RYC2018-024050-I (Fondo Social Europeo – Agencia Estatal de Investigaci\'on).
+We would like to thank the editor and reviewers (Stephen Shennan
+and two anonymous reviewers) for their constructive comments that helped improve the original manuscript.
+We are grateful to Simon Carrignon, Juan Moros, Jordi P\'erez-Gonz\'alez and V\'ictor M.
+Lozano for their useful suggestions.
 Data have been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}.
 Maps were performed using QGIS software v.3.14 `Pi'.
-Datasets and source code are freely available under Open licenses from https://github.com/Mcotsar/Baetica\_stamps.
+Datasets and source code are freely available under Open licenses from \url{https://github.com/Mcotsar/Baetica\_stamps}.
 
-\section{References}
+\end{acknowledgements}
 
-%\bibliography{bibliotex}
+%\subsection{Subsection title}
+%\label{sec:2}
+%as required. Don't forget to give each section and subsection a unique label (see Sect.~\ref{sec:1}).
+%\paragraph{Paragraph headings} Use paragraph headings as needed.
+
+% For one-column wide figures use
+%\begin{figure}
+% Use the relevant command to insert your figure file.
+% For example, with the graphicx package use
+%\includegraphics{example.eps}
+% figure caption is below the figure
+%\caption{Please write your figure caption here}
+%\label{fig:1}       % Give a unique label
+%\end{figure}
+%
+% For two-column wide figures use
+%\begin{figure*}
+% Use the relevant command to insert your figure file.
+% For example, with the graphicx package use
+%\includegraphics[width=0.75\textwidth]{example.eps}
+% figure caption is below the figure
+%\caption{Please write your figure caption here}
+%\label{fig:2}       % Give a unique label
+%\end{figure*}
+
+%\begin{acknowledgements}
+%If you'd like to thank anyone, place your comments here
+%and remove the percent signs.
+%\end{acknowledgements}
+
+% Authors must disclose all relationships or interests that 
+% could have direct or potential influence or impart bias on 
+% the work: 
+%
+\section*{Conflict of interest}
+
+The authors declare that they have no conflict of interest.
+
+% BibTeX users please use one of
+%\bibliographystyle{spbasic}      % basic style, author-year citations
+%\bibliographystyle{spmpsci}      % mathematics and physical sciences
+%\bibliographystyle{spphys}       % APS-like style for physics
+%\bibliography{}   % name your BibTeX data base
+
+% Non-BibTeX users please use
+%\begin{thebibliography}{}
+%
+% and use \bibitem to create references. Consult the Instructions
+% for authors for reference list style.
 \bibliography{bibtesis}
 
+%\bibliographystyle{model2-names.bst}
+%\biboptions{authoryear}
+%\bibitem{RefJ}
+% Format for Journal Reference
+%Author, Article title, Journal, Volume, page numbers (year)
+% Format for books
+%\bibitem{RefB}
+%Author, Book title, page numbers. Publisher, place (year)
+% etc
+%\end{thebibliography}
+
 \end{document}
+% end of file template.tex

### 2021-02-10 Manuscript/Manuscript.tex
revision

@@
 Xavier Rubio-Campillo \at
+DIDPATRI, Universitat de Barcelona, Passeig de la Vall d'Hebr\'on, 171, Barcelona, Spain
 School of History, Classic \& Archaeology, Room OOM.33, William Robertson Wing, Old Medical School, Teviot Place, University of Edinburgh, UK \at
-DIDPATRI, Universitat de Barcelona, Passeig de la Vall d'Hebr\'on, 171, Barcelona, Spain             
 }
@@
 
-The aim of this study is to explore economics dynamics in the production and distribution of olive oil trade in \textit{Baetica} province (currently Andalusia).
-\textit{Baetica} became an important production and distribution centre during the Roman Empire.
-However, it remains under debate about how this province was organised and whether it could be possible to identify patterns in the olive oil market.
+The aim of this study is to explore the trade dynamics involved on the large-scale distribution of olive oil produced in \textit{Baetica} province (currently Andalusia).
+\textit{Baetica} became an important production and distribution centre during the Roman Empire and its goods have been found in thousands of sites, but it is still unclear how exactly this large scale production was organised.
 
-Our case of study has been focused on amphoric production from 1st to 3rd centuries AD.
-In particular, we want to detect patterns of olive oil production that link amphora workshops and amphoric stamps.
-Amphoric stamps are used to identify the presence of different groups that might share similar stamps.
+We explore here connections between production and consumption sites during the period of most intense activity (1st to 3rd centuries AD).
+Specifically, we use amphoric stamps as proxies to identify the workshops where amphorae found on consumer sites were produced.
 
-To achieve this goal, we analyse a set of stamps from two types of centres (production and consumption): 1) production centres by analysing different workshops in \textit{Baetica} province and 2) two Roman provinces such as \textit{Germania} (Inferior and Superior) and \textit{Britannia} as consumption centres.
-They will be used to detect a connection between the distribution of amphoric stamps and the economic structure in both centres.
-Here, we use methods borrowed from Ecology that allow us to identify if amphora workshops share similar amphoric stamps depending on the spatial distance.
-The analysis explores how the quantitative approach provides a useful tool for the interpretation of the economic processes.
-Finally, results pretend to highlight the organisation of Baetican olive oil production in the Roman Empire linked to the differences observed in the archaeological evidence.
+To achieve this goal, we analyse a dataset of stamps found in two types of sites: 1) workshops from \textit{Baetica} province and 2) consumption centres located in two provinces \textit{Germania} (Inferior and Superior) and \textit{Britannia}.
+This dataset is used to detect links between specific sites that may share a large percentage of common stamps.
+The analysis is performed using quantitative methods borrowed from Ecology that allows us to group sites based on the amount of stamps shared between them.
+Results highlight the structure of economic processes of the Roman Empire using the archaeological evidence of olive oil large-scale distribution.
 
 %Insert your abstract here. Include keywords, PACS and mathematical subject classification numbers as needed.
-\keywords{Roman Empire \and amphora production \and Dressel 20 \and dissimilarity index \and Roman provinces}
+\keywords{Roman Empire \and amphora production \and Dressel 20 \and similarity index \and Roman provinces}
 % \PACS{PACS code1 \and PACS code2 \and more}
@@
 The intensification of long-range trade was one of the most important traits of the economy developed during the Roman Empire.
-The development of an extensive road network increased the connectivity between inland communities, but most shipping continued being based on maritime routes particularly in the Mediterranean basin~\citep{temin_market_2001,bevan_mediterranean_2014}.
-The intensity of this trade between distant regions can be observed both in archaeological and written evidence~\citep{rodriguez_baetican_1998}.
+The development of an extensive road network increased the connectivity between inland communities while maritime and riverine shipping continued to be the most common trade methods, particularly in the Mediterranean basin~\citep{temin_market_2001,bevan_mediterranean_2014}.
+The intensity of this long-range trade between distant regions is observed both in archaeological and textual evidence~\citep{rodriguez_baetican_1998}.
 
-The Empire developed a series of structures to support and organise this long-range trade and specialised entire regions to massively produce specific goods.
-An example of this process was the province of \textit{Baetica} (currently Andalusia, southern Spain) which became an important olive oil production centre during the Roman Empire.
-Olive oil was an essential good for Romans because it was used in almost every aspect of their daily life such as cooking, hygiene or lighting~\citep{mattingly_d.j._oil_1988}.
-This high demand required a huge increase on the production which was distributed through amphorae shipped via maritime routes to all the provinces and particularly to Italy and the hundreds of garrisons that the Roman army deployed along its borders~\citep{blazquez_exportacion_1980}.
+The Empire developed a series of structures to support and organise this long-range trade and specialised entire regions to produce specific goods at a massive scale.
+The province of \textit{Baetica} (currently Andalusia, southern Spain) is an example of this specialization process.
+\textit{Baetica} became one of the most important olive oil production centres during the Roman Empire; olive oil was an essential good for Romans because it was used in almost every aspect of their daily life such as cooking, hygiene or lighting~\citep{mattingly_d.j._oil_1988}.
+This high demand required a huge increase on the production which needed to be distributed to cities across thousands of kilometers.
+This large-scale trade was organised via the shipment of millions of amphorae using maritime and riverine routes to all the provinces and particularly to Italy and the military garrisons guarding the borders of the Empire~\citep{blazquez_exportacion_1980}.
 
-The structure and processes associated with this massive olive oil production have been extensively discussed over the last decades\citep{rodriguez_economioleicola_1977, Chic_hispania_1997,millet_anforas_1998}.
-However, the lack of written records has not made possible to detect any indication of patterns in the olive oil market.
+The structure and processes associated with this massive olive oil production have been extensively discussed over the last decades\citep{rodriguez_economioleicola_1977, duncan1982economy, Chic_hispania_1997,millet_anforas_1998}.
+However, the complexity and sheer scale of the process makes its understanding challenging despite several archaeologicale xcavations and the existence of some written records discussing its structure.
+Our knowledge of this economic activity has benefited from new findings and data sources, but despite all these advances several questions remain open: How was this massive amphora production organised?
+Can we detect the economic ties between provinces?
+Was this distribution centralized or each region organised it differently?
 
-Our knowledge of this economic activity has benefited from new findings and data sources, but despite all these advances several questions remain open: How was amphora production organised during the Roman Empire?
-Can we detect economic dynamics between provinces by analysing the distribution of amphorae?
+Recent advances in Roman studies have currently led to an environment with various interpretations on these commercial dynamics~\citep{temin_economy_2006,quantifyingwilson2009}.
+The application of diverse novel quantitative approaches to new accessible datasets has allowed us to improve our interpretation of the growing amount of archaeological evidence and contribute to a better understanding of the dynamics of Roman production\citep{brughmans_roman_2016,orengo_seeds_2016,bayesian_2018,coto-sarmiento_identifying_2018,rubio-campillo_ecology_2018}.
+This work aligns with these efforts and focuses on the distribution of olive oil towards Roman provinces in the borders.
+Amphoric production in \textit{Baetica} province will help us identify links between production centres and consumption centres in order to detect an economic pattern in the commerce by analysing amphoric stamps finds in these provinces.
 
-Advances in Roman studies have currently led to an environment with various interpretations on commercial dynamics \citep{duncan1982economy,
-temin_economy_2006,
-quantifyingwilson2009}.
+The finding of identical amphoric stamps in different provinces has allowed us to explore patterns of olive oil distribution within the Empire.
+Despite this archaeological evidence there is still no consensus on the structure that generated this pattern; how was the economic relation between production and consumption centres?
+Did they follow a centralised process to distribute olive oil in those provinces or each workshop organised its own trade networks?
 
-The application of different quantitative approaches has allowed us to improve our interpretation of the growing amount of archaeological evidence and reveal the dynamics of Roman production
-\citep{brughmans_roman_2016,
-orengo_seeds_2016,bayesian_2018,
-coto-sarmiento_identifying_2018,
-rubio-campillo_ecology_2018}.
-
-This work focuses on the debate about the dynamics of the distribution of olive oil market in Roman provinces.
-Amphoric production in \textit{Baetica} province will help us identify links between production centres and consumption centres in order to detect an economic pattern in the commerce by analysing amphoric stamps finds in the provinces.
-
-The finding of amphoric stamps in different provinces has allowed us to know the importance of the distribution of olive oil within the Empire.
-Despite having archaeological evidence, no consensus has been reached: How was the economic relation between production and consumption centres?
-Did they follow an economic pattern to distribute olive oil in those provinces?
-
-This paper aims to study the olive oil market connection between provinces by calculating the similarity of amphoric stamps.
-Specifically, our work aims to detect regional processes focused on a commercial product (i.e.
-olive oil) from a specific province \citep{isaksen_network_2006}.
+This paper aims to explore these questions regarding the olive oil market by identifying links between productionn centres and consumption provinces using similarity indexes for amphoric stamps.
+In this way, our work aims to detect regional patterns for a specific commercial product (i.e.
+olive oil) which was shipped between distant provinces~\citep{isaksen_network_2006}.
 
@@
 %We focus here on exploring the economic relation between stamps and amphora production and distribution centres. 
-
-Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}) (see Figure~\ref{general}).
+This relation between the production centre (i.e.
+\textit{Baetica}) and consumption sites have been studied here for two case studies: \textit{Britannia} and \textit{Germania} (see Figure~\ref{general}).
 
@@
 \includegraphics[width=\linewidth]{general_map}
-\caption{Overview of the sites analysed in this paper.
-The colour defines the three different regions under study (\textit{Baetica} as producer; \textit{Germania} and \textit{Britannia} as consumers) while the size of each dot highlights the number of Dressel 20 amphoric stamps found on each of these sites}
+\caption{Sites analysed in this paper.
+The colour defines the three different regions under study (\textit{Baetica} (yellow) as producer; \textit{Germania} (blue) and \textit{Britannia} (green) as consumers) while the size of each dot is proportional to the number of Dressel 20 amphoric stamps found on each site}
 
@@
 
-In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces.
-This economic connection could be identified by different aspects: a) correlation between spatial distance and the sites based on the idea that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific province.
+In \textit{Baetica} province we want to identify links between workshops based on the similarity of their stamps.
+In the consumption Roman provinces our aim is to detect groups of nearby sites sharing stamps and to what extent stamps are clustered based on provinces.
+This economic connection could be identified with two ideas : a) nearby sites have more similar amphoric stamps and b) sets of stamps can only be found on specific provinces because they were shipped together.
 
@@
 
-Based on this assumption, we proposed three hypotheses: a) we can identify a correlation between spatial distance and the distribution of amphoric stamps, b) stamps located in close workshops share similar traits and c) low mobility of amphoric stamps to other regions: stamps always stay in the same region.
+We proposed three working hypotheses based on these ideas: a) we can identify a correlation between spatial distance and the distribution of amphoric stamps, b) stamp sets found in nearby workshops should be more similar and c) there is a low mobility of amphoric stamps from consumption centres to other regions (i.e.
+stamps always stay in the same region).
 
-This study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps.
-To do this, a population approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}.
-Quantitative approach provides one way to measure the similarity between amphoric stamps.
-Here, we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
+This study proposes a robust baseline to explore the distribution of Baetican olive oil production by identifying which centres have more similar stamps and then confront this measure against their proximity.
+To do this, a population-based approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}.
+This quantitative method provides a way to measure the similarity between centre based on the amphoric stamps found on each of them.
+This ecological approach is based on three steps: a) to quantify stamp similarity between sites, b) to group sites based on this similarity and c) to find correlation patterns between stamp similarity and proximity.
+The rationale behind the method is as follows: if trade structure was centrally planned then we should be able to stablish links between workshops and consumption sites based on the similarity of its stamps; if this is not the case then results would suggest that workshops worked independently and agreements were made at an individual level.
 
@@
 
-For example, if workshops and provinces share stamps with similar traits, then we can identify connections.
-By contrast, if we do not detect similar stamps between workshops, then workshops worked independently.
-
-The paper addresses these questions as follows: the next section introduces the historical context.
-Section two displays the dataset and the methods used for the analysis.
-Section three presents the results and the last section shows the discussion and the main conclusions of this work.
+The paper discusses these questions, the analysis and results as follows: next section introduces the historical context.
+Section two defines the dataset and the methods used for the analysis.
+Section three presents the results and section 4 discusses them; the paper finishes with some concluding remarks on the research performed.
 
@@
 
-The landscape of the \textit{Baetican} province was one of the best regions to face the increased demand for olive oil across the Roman provinces.
-For this reason, the area saw an increase in productivity as a massive infrastructure of olive oil production was gradually deployed.
-The production and distribution of olive oil increased from the first to the third century AD~\citep{remesal_concierto}.
+The \textit{Baetican} province was one of the most suitable regions to tackle the challenge of the huge demand for olive oil across the Roman provinces.
+For this reason, the area saw an increase in productivity as a massive specialized infrastructure of olive oil production was gradually deployed.
+This production of olive oil had the largest peark from the first to the third century AD~\citep{remesal_concierto}.
 
-Hundreds of presses and amphora-making workshops were built near large extensions covered by olive trees.
-The workshops where amphorae were made and filled with olive oil were places along the rivers Guadalquivir and Genil riverbanks (see Fig.\ref{workshop}).The proximity distance between workshops was relatively close which allowed for more efficient distribution.
+Hundreds of presses and amphora-making workshops were built near a landscape covered by olive trees.
+These workshops made the amphorae which were filled with olive oil from the presses; they were built at the banks of the rivers Guadalquivir and Genil (see Fig.\ref{workshop}).
 
-This strategic location allowed the transport of olive oil through riverine shipping towards the maritime routes that connected \textit{Baetica} with Mediterranean and Atlantic trade routes towards the rest of the Empire~\citep{garcia_vargas_enrique_formal_2010}.
+This proximity between production area, olive processing infrastructure, amphorae production and navigable rivers made the entire process extremely efficient.
+The strategic location allowed the transport of olive oil through riverine shipping towards the maritime routes that connected \textit{Baetica} with Mediterranean and Atlantic trade routes towards the rest of the Empire~\citep{garcia_vargas_enrique_formal_2010}.
 
@@
 
-The chronology of the workshops is wide spanning from the first to the third centuries AD~\citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}. 
-This fact is shown by the archaeological evidence that displays a highly specialised production with a long activity \citep{remesal_anforas_2004}.
-Thus, the specialisation of this production made amphorae with apparently few changes in their forms, reaching a high degree of standardisation.
+The chronology of the Baetican workshops is typically defined between the first to the third centuries AD~\citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
+Archaeological evidence suggests that this highly specialised production was developed across a long activity period\citep{remesal_anforas_2004}, but at the same time the process seemed to have been planned at some level as is observed by the relatively minor changes in amphoric typologies and the high degree of standardisation on the amphorae found across at least 50 workshops.
 
-The importance of this olive oil trade is revealed by the massive amount of Dressel 20 type amphorae found across the Roman provinces \citep{dressel_ricerche_1878,
-millet_anforas_1998}.
-This amphora has been commonly associated with the transportation of Baetican olive oil for supplying military camps and civil settlements \citep{berni_millet_epigrafianforica_2008}.
-
+The importance of this olive oil trade is revealed by the massive amount of Dressel 20 type amphorae found in all the Empire and particularly in the Western provinces~\citep{dressel_ricerche_1878,millet_anforas_1998}.
+This amphora has been commonly associated with the transportation of Baetican olive oil for supplying military camps and civil settlements~\citep{berni_millet_epigrafianforica_2008}.
 They have a recognisable globular form with a short neck and oval-shaped handles.
@@
 
-Dressel 20 amphorae display a lot of evidence, but some of it is still hard to interpret.
-A large percentage of Dressel 20 were marked with stamps, while they could also be inked with \textit{tituli picti} or incised with \textit{graffiti}.
-The interpretation for most of these inscriptions is still open due to the fragmentation of the material and the small sample size of well-preserved elements~\citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
+Dressel 20 amphorae display a lot of interesting information beyond their shape, even though some of it is hard to interpret.
+A large percentage of Dressel 20 were marked with stamps, while some of them were also inked with \textit{tituli picti} or incised with \textit{graffiti}.
+The interpretation for most of these inscriptions is still open due to the fragmentation of the material, the lack of written records clarifying its use and for \textit{tituli} and \textit{graffiti} a small sample size of well-preserved elements~\citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.
 
-\section{A potential indicator of Roman economy: Dressel 20 stamps}
+\section{Amphoric stamps as a proxy for economic structures}
 \label{sec:2}
@@
 Dressel 20 was the amphora most stamped during the Roman Empire ~\citep[18]{millet_anforas_1998}.
-The large sample size of recovered Dressel 20 stamps is distributed across all the western provinces of the Roman Empire.
-For this reason, there are several publications of this type of evidence discussing its origin, its meaning, its long period of use and its spatial distribution~\citep{dressel_ricerche_1878,
-rodriguez_economioleicola_1977,
-chicepi1985,millet_anforas_1998, remesal_sellar_2016} (Fig.\ref{amphora}).
+The large sample size of recovered Dressel 20 stamps is distributed across all Western provinces of the Roman Empire.
+This wide distribution is observed in the large volume of publications discussing the origin, use, chronology and spatial distribution of the stamped codes~\citep{dressel_ricerche_1878,rodriguez_economioleicola_1977,chicepi1985,millet_anforas_1998, remesal_sellar_2016} (Fig.\ref{amphora}).
 
@@
 \includegraphics[scale=0.5]{dressel20}
-\caption{Dressel 20 were mostly marked with stamps of three letters called \textit{tria nomina}}
+\caption{Dressel 20 were typically marked with stamps of three letters called \textit{tria nomina}}
 \label{amphora}
@@
 
-Most stamps display a large activity of production that can be difficult to date using an accurate chronology.
-They have been frequently dated with the consular dating by studying \textit{tituli picti} found in \textit{Monte Testaccio} \citep{Testaccio1, berni_millet_epigrafianforica_2008}.
-Despite \textit{tituli picti} contain important information about chronology, they have been still found in few quantities mostly concentrated in an only settlement such as \textit{Monte Testaccio}.
-In addition, the chronology of stamps could be also biased taking into account that amphorae were deposited after being produced and consumed.
+Most stamps seemed to have been used across a long activity period which can be difficult to date with the required level of accuracy.
+They have been frequently dated using as reference consular dates found on nearby \textit{tituli picti} recovered from \textit{Monte Testaccio}~\citep{Testaccio1, berni_millet_epigrafianforica_2008}.
+The chronological information of \textit{tituli picti} is very useful, but the small amount of dates recovered and the fact that a vast majority of them come from the same site poses some challenges to this precise dating.
+In addition, the fact that \textit{Monte Testaccio} was a dump suggests that the diposition of amphorae vessels may have been altered at a later phase due to changes in the long-term dynamics of creation of this artificial mound.
 
 The use of amphoric stamps can be a good proxy to explore the system of olive oil production and trade developed in \textit{Baetica} province.
-These stamps are typically found on handles but they were also imprinted on rims and the amphora body~\citep{millet_anforas_1998}.
-The content information of the stamps had not a unique criterion.
-They showed a code with different letters and forms.
-Stamps often displayed a code of three letters and they are known as \textit{Tria Nomina} \citep{berni_millet_amphora_1996}.
-The three letters correspond to initial forms of a name but they can also appear complete.
-
-Scholars agree that stamps are some type of identification mark, but there is no consensus on the meaning of these marks~\citep{rodriguez_baetican_1998}.
+These stamps are typically found on handles but they could also be imprinted on rims and the amphora body~\citep{millet_anforas_1998}.
+The content information of the stamps was not defined by a single criteria, but they often displayed a code composed by letters and additional symbols.
+One of the most common forms is a code of three letters showing a \textit{Tria Nomina}~\citep{berni_millet_amphora_1996}; in this case the three letters are the initials of a Roman name.
+In some cases the full name may appear, while the meaning of other stamps is difficult to interpret~\citep{rodriguez_baetican_1998}.
 
@@
 On the other hand, they could also belong to the owner of the workshops where the amphorae were made or even to a group of amphora workers~\citep{berni_millet_epigrafianforica_2008}.
-
-Nevertheless, some challenges remain under discussion such as how this production was organised and whether it is possible to distinguish production patterns in the olive oil trade.
-Our questions will be focused on the distribution of amphoric stamps.
-Did they follow a distribution pattern?
-Did stamps share the same workshop?
+Despite the unsolved question of ownership some interesting ideas can be explored based on the finding of similar stamps across the Empire.
+Observable patterns can be used to explore how this production was organised and whether it is possible to identify specific production processes fdor olive oil trade.
+Our questions here will be focused on the distribution of amphoric stamps; were different stamps shipped through different distribution routes?
+Were specific stamp codes used in different workshops?
 
@@
 
-The creation of new provinces required the Roman Empire the arrival of resources through Mediterranean and Atlantic routes.
-This effectively led to a gradual change in the economic and social structures when the trade networks that supplied the Roman army became more complex.
+The expansion of the Roman Empire to increasingly distant regions required an effort to intensify trade activity through Mediterranean and Atlantic routes.
+These needs effectively led to a gradual change in the economic and social structures as the trade networks that supplied the provinces and particularly the Roman army became more complex.
+An interesting example of this system is the figure of the \textit{praefectura annonae} which was created during Augustus's administration to organize wheat supply.
+The role of the \textit{praefectura annonae} was mainly focused on providing through \textit{frumentationes} a fixed monthly amount of wheat to each Roman citizen~\citep{remesal_annona_1986,remesal_concierto}.
 
-Augustus's administration created the figure of the \textit{praefectura annonae} to organize wheat supply.
-The role of the \textit{praefectura annonae} was mainly focused on providing through \textit{frumentationes} a fixed monthly amount of wheat to each Roman citizen~\citep{remesal_annona_1986,remesal_concierto}
+Some authors argue elsewhere that this same \textit{praefectura annonae} could also have organised the supply of additional goods such as olive oil to the Roman legions~\citep{remesal_annona_1986,remesal_annona_1990}.
+From this perspective, olive oil supply seems to be particularly intense in militarised provinces that functioned as the borders of the Empire such as \textit{Britannia} and \textit{Germania}.
+The high percentage of Dressel 20 amphorae found in civilian sites close to the military garrisons also suggests the existence of diverse supply dynamics involving systematic trade between military sites to civil settlements~\citep{remesal_annona_1986, carreras_britannia_1998}.
+An explanation for the interest for exporting olive oil to the military garrison in the Roman provinces is still unclear.
+The reasons behind this effort to supply olive oil to all soldiers may be related to two aspects: a) consumption based on cultural reasons (i.e.
+social identification or habit  by a group of consumers) and b) economical and political reasons to create a redistributive system mechanism to supply military camps~\citep[69-70]{carreras_britannia_1998}.
 
-However, some hypotheses argue elsewhere that this same \textit{praefectura annonae} could also have organised the supply of additional goods such as olive oil to the Roman legions  \citep{remesal_annona_1986,
-remesal_annona_1990}
-
-Olive oil supply seems to be particularly intense in militarised provinces that functioned as the borders of the Empire such as \textit{Britannia} and \textit{Germania}.
-The high percentage of Dressel 20 amphorae found in sites nearby military garrisons suggests the existence of diverse supply dynamics as there were trade processes flowing from military sites to civil settlements~\citep{remesal_annona_1986, carreras_britannia_1998}.
-
-An explanation for the interest to export olive oil for the military garrison in the Roman provinces is still unclear.
-The origin of this exportation may be related to two aspects: a) consumption based on cultural reasons (i.e.
-social identification or habit  by a group of consumers) and b) economical and political reasons to create a redistributive system mechanism to supply military camps \citep[69-70]{carreras_britannia_1998}.
-
-The Roman conquest in \textit{Britannia} and \textit{Germania} was a new opening to foreign trade in order to supply the military campaign.
+The Roman conquest of \textit{Britannia} and \textit{Germania} was a new opportunity to expand economical activity to support military efforts.
 High demand for olive oil in both provinces stimulated an important trade network involving different agents and the Empire itself.
-\textit{Baetica} became a strategic province for the production and distribution in both provinces since it can be detected by the high presence of amphoric stamps found in civil and military areas during this period.
+\textit{Baetica} became a strategic province for the production and distribution in both provinces as can be observed by the large volume of Dressel 20 amphora vessels found both in civil and military sites of these areas during and after the conquest.
 
@@
 The evidence for olive oil consumption in \textit{Britannia} is scarce before the Roman conquest \citep{funari_corpus_1996,carreras_abastecimiento_2003}.
-The local population did not consume this product and it was not produced on the region as the landscape and climate of the British Isles was not suitable for the growth of olive trees~\citep[161]{monfort_britanniaen_1998}.
-Based on the archaeological records, this absence of olive oil changed with the arrival of a large amount of Dressel 20 amphorae to \textit{Britannia} from the first century A.D \citep{peacock_amphorae_1991,
-carreras_britannia_1998}.
-At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}.
+The local population did not consume this product and it was not produced in the region as the landscape and climate of the British Isles was not suitable for the growth of olive trees~\citep[161]{monfort_britanniaen_1998}.
+Archaeological records suggest that this absence of olive oil changed with the arrival of a large amount of Dressel 20 amphorae to \textit{Britannia} from the first century A.D.
+onwards \citep{peacock_amphorae_1991,carreras_britannia_1998}.
+At this moment, we detect an increase of the olive oil exports linked to the arrival of legions caused by different military campaigns~\citep[161]{monfort_britanniaen_1998}.
 
-The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges.
+The increase of Dressel 20 exports created an important commercial network for exchanges.
 The demand for olive oil after the Roman conquest required of a new trade route due to the lack of local olive oil production; \textit{Baetica} was chosen as its main supply.
+This economic relation will be particularly intense for sites close to Hadrian Wall's garrisons.
+Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic locations~\citep{carreras_atlantic_2012}.
+This presence of Dressel 20 stamps in military camps in \textit{Britannia} has been widely studied in Roman archaeology \citep{williams_importation_1983,funari_corpus_1996,carreras_britannia_1998,carreras_abastecimiento_2003}.
+This intense consumption of olive oil also suggests the existance of provincial structures created to organise the supply of olive oil across military camps.
+There are no written records detailing how this redistribution of essential goods was organised in \textit{Britannia}, but the archaeological evidence suggests that cities may have been the central nodes of this local trade network~\citep{funari_economic_2005,orengo_seeds_2016}.
 
-This fact will have a particular intensity within sites close to the Hadrian Wall's garrisons.
-Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic locations~\citep{carreras_atlantic_2012}.
-
-The presence of Dressel 20 stamps in military camps in \textit{Britannia} has been widely studied in Roman archaeology \citep{williams_importation_1983,
-funari_corpus_1996,
-carreras_britannia_1998,carreras_abastecimiento_2003}.
-
-This intense consumption of olive oil also indicates a possible provincial structure designed to organise the supply of olive oil to military camps.
-
-There are no written records explaining how this redistribution of essential goods was organised in \textit{Britannia}, but the archaeological evidence suggests that cities may have been the central nodes of this local trade network~\citep{funari_economic_2005,orengo_seeds_2016}.
-
-The consumption of olive oil would experience a progressive slowdown from the third century A.D.
+The consumption of Baetican olive oil would experience a progressive slowdown from the third century A.D.
 onwards.
 This date matches a change in market strategy in the Empire.
-This gradual decrease on \textit{Baetican} olive oil exports can be observed as a decrease in the amount of Dressel 20 found in excavated sites as this typology is replaced by other types of amphorae from other places \citep{rodriguez1991aceite,millet_anforas_1998}.
+This gradual decrease on \textit{Baetican} olive oil exports can be observed as a decrease in the amount of Dressel 20 found in excavated sites as this typology is replaced by other types of amphorae from other places and particularly the African provinces~\citep{rodriguez1991aceite,millet_anforas_1998}.
 
@@
 
-Roman studies exploring the presence of Dressel 20 amphorae in the provinces Inferior and Superior during Roman Empire have not had the same impact as the rest of the provinces \citep[293]{remesal_baetica_2002}.
+The distribution of Dressel 20 amphorae in the provinces of Germania Inferior and Germania Superior is not as well interpreted as the rest of provinces discussed in this work~\citep[293]{remesal_baetica_2002}.
+It is clear that there was a massive import of Dressel 20 in both provinces as is observed in archaeological excavations across the German \textit{limes}.
+However, it is still unknown to what extent trade agents participated in the distribution of Baetican olive oil in \textit{Germania} or the supply was exclusively focused on the Roman army garrisons\citep[156]{remesal_germn_2010}.
 
-The massive importation of Dressel 20 in both \textit{Germania} provinces could be explained due to the archaeological sources found mostly in German \textit{limes}. 
-By contrast, it is still unknown if trade agents participated in the distribution of Baetican olive oil in \textit{Germania} or the supply was exclusively focused on the Roman army garrisons\citep[156]{remesal_germn_2010}.
+In any case, the introduction of olive oil in \textit{Germania} shows similar patterns than the one discussed for \textit{Britannia}.
+There is no archaeological evidence for olive oil consumption before the arrival of the Roman army to the region, while a majority of recovered Dressel 20 amphorae are located in the military sites built over the German \textit{limes} \citep{remesal_germaniaengl_2002}.
+The presence of the Roman army encouraged the exchange in the province as shown by the arrival of this product both civil settlement and military sites with a major concentration at the German \textit{limes}.
+It seems that some Baetican centres could have been assigned to a specific province for the olive oil supply to military garrisons~\citep[125]{remesal_concierto}.
+However, this hypothesis is hard to assess given the current lack of written evidence clarifying the meaning of the amphoric stamps.
 
-\textit{Germania} presents a similar introduction pattern of olive oil than the one discussed for \textit{Britannia}.
-There is no archaeological evidence for olive oil consumption before the arrival of the Roman army to the region while a majority of recovered Dressel 20 amphorae are located in the military sites that formed the German \textit{limes} \citep{remesal_germaniaengl_2002}.
-
-The presence of the Roman army encouraged the exchange in the province as shown by the arrival of this product both civil settlement and military sites with a major concentration at the German \textit{limes}.
-
-It seems that some Baetican centres could have been assigned to a specific province for the olive oil supply to military garrisons \citep[125]{remesal_concierto}.
-However, this hypothesis is hard to assess given the current lack of written evidence related to the meaning of the amphoric stamps.
-
-Previous studies suggested that the supply to the German \textit{limes} was mainly based on riverine transport, but recent works suggest that the maritime route through the Atlantic Ocean could have been more important than expected (both for \textit{Germania} and \textit{Britannia})~\citep{remesal_germn_2010,rubio-campillo_ecology_2018}.
+Some studies assumed that Roman naval technology was not suited for oceanic transport and for this reason defined riverine shipping across Rhône and Rhine rivers as the basic supply route for the German \textit{limes}.
+However, more recent large-scale computational analysis as well as new excavations on the Iberian Atlantic coast suggest that the Atlantic route, which was far more efficient, may have played a major role on the supply of \textit{Germania} and \textit{Britannia}~\citep{rubio-campillo_ecology_2018}.
 
@@
 
-The goal of this study is exploring the effect of the production patterns between different centres.
-We are especially interested in identifying links between production and consumption centres by using amphoric stamps.
+The goal of this study is to explore these economic and long-range trade dynamics using amphoric stamps as proxies for common supply networks.
+We are especially interested on identifying links between Baetican production centres and consumption centres in the different \textit{limes}.
 
-To study the amphoric stamps, we used the data collected in the CEIPAC database of amphoric epigraphy from different places~\citep{remesal_centro_2015} (see the database here \url{http://romanopendata.eu}).
-The CEIPAC dataset contains over 50.000 epigraphy records found on different types of amphorae, mostly from \textit{Monte Testaccio}.
-Most stamps used in the analysis did not have a specific chronology or were not available in the dataset.
-This fact did that we had to discard the possibility to analyse a distribution of stamps per chronologies.
+For this analysis we used data collected in the CEIPAC database of amphoric epigraphy~\citep{remesal_centro_2015}.
+The CEIPAC dataset contains over 50.000 epigraphic records found on different types of amphorae (the database is openly accessible from \url{http://romanopendata.eu}).
+A large percentage of these records come from from \textit{Monte Testaccio}, but the rest is distributed mainly across the Western provinces of the Empire.
+It is worth mentioning that the database does not provide temporal information due to the reasons previously discussed, and for this reason this work can only analyze the spatial distributions within a coarse chronology of 2 centuries (3rd-1st AD).
 
-In order to enable reproducibility of the results, the selected datasets and source codes are available on the following GitHub repository: \url{https://github.com/Mcotsar/Baetica\_stamps}.
+In order to enable reproducibility of the results, this dataset as well as the source code are available on the following GitHub repository: \url{https://github.com/Mcotsar/Baetica\_stamps}.
 
@@
 
-The database allowed us to retrieve 3798 stamps found in Dressel 20 amphorae recovered from workshops within \textit{Baetica} province.
-Each of these stamps has detailed records at the site where it was found, the inscribed code as well as its spatial coordinates.
-The stamps with incomplete information were discarded, thus finishing with a sample of 987 stamps from 81 sites and displaying 130 different codes.
-Query dataset from Baetica province can be found on the following GitHub repository: \url{https://github.com/Mcotsar/Baetica_stamps/tree/main/Databases/Baetica}.
+We retrieved information on 3798 stamps found in Dressel 20 amphorae recovered from workshops within \textit{Baetica} province.
+Each record provides information about the spatial coordinates of the site and the inscribed code.
+Fragmented stamps or code not able to be read were discarded, thus finishing with a sample of 987 stamps from 81 sites and displaying 130 different codes.
 
-Our dataset also contained a new categorical variable defining the \textit{conventus} of each site.
+Our dataset also contained a categorical variable defining the \textit{conventus} for each production site.
 The \textit{conventus} were administrative centres for territorial organisation under the Roman Empire \citep[58]{ozcariz_gil_administracion_2013}.
-The creation of the \textit{conventus} in some Roman provinces was a form of control with the purpose of organising the administration in the provinces \citep{albertini_les_1923}.
-
-The production area for Dressel 20 amphorae extended across three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter \textit{Hispalis}), \textit{Cordubensis} (currently C\'ordoba, hereafter \textit{Corduba}) and \textit{Astigi} (currently \'Ecija, Sevilla, hereafter \textit{Astigi}) \citep{rodriguez_economioleicola_1977,chicdatos2001,berni_millet_epigrafianforica_2008} .
+The creation of the \textit{conventus} in some Roman provinces was a form of control with the purpose of organising provincial administration~\citep{albertini_les_1923}.
+The production area for Dressel 20 amphorae extended across three different \textit{conventus}: \textit{Hispalensis} (currently Seville, hereafter \textit{Hispalis}), \textit{Cordubensis} (currently C\'ordoba, hereafter \textit{Corduba}) and \textit{Astigi} (currently \'Ecija, Sevilla, hereafter \textit{Astigi})~\citep{rodriguez_economioleicola_1977,chicdatos2001,berni_millet_epigrafianforica_2008} .
 
@@
 
-It is important to mention that some workshops exhibited the same geographical coordinates while they were catalogued differently as workshops.
-For instance, two nearby amphora workshops such as Tesorillo de Do\~na Mencia and Do\~na Mencia were catalogued as a different name in spite of having the same geographical coordinates.
-Our work considered this particularity in the archaeological dataset but some biases can occur in the analysis.
+It is worth mentioning that some workshops were coded under the same geographical coordinates.
+For instance, the stamps of two nearby amphora workshops such as Tesorillo de Do\~na Mencia and Do\~na Mencia were catalogued under different names while sharing the same geographical coordinates.
+The interpretation of our analysis considered this bias in the archaeological dataset which may have a minor effect in the final results.
 
-The first step was to compute the frequency distribution of stamp codes and analyse it using Exploratory Data Analysis (EDA).
+The first step of this analysis was to compute the frequency distribution of stamp codes and analyse it using Exploratory Data Analysis (EDA).
 This would allow us to study the distribution of the stamps across centres as can be seen in Figure~\ref{stamps}.
-Most workshops concentrated few stamps while few workshops contained more than 100 stamps.
-For instance, one workshop (La Catria) concentrated a large percentage of the amphoric stamps with a total number of 228 stamps.
+As can be seen the distribution is widely diverse with most workshops having only one stamp, while a few workshops contained over 100 stamps.
+Particularly, one workshop (La Catria) concentrated a large percentage of the amphoric stamps with a total number of 228 stamps.
 
-The type of distribution is also frequent in amphora production where we observed a major concentration of the number of stamps in few workshops~\citep{bayesian_2018,coto-sarmiento_identifying_2018}.
+This uneven distribution of stamps is typical for amphora production as it is observed that a major concentration on the number of stamps is focused on a few workshops~\citep{coto-sarmiento_identifying_2018}.
 
@@
 \includegraphics[width=\linewidth]{frequencystamp.pdf}
-\caption{Histogram on a log scale with base 10.
-X axis is represented by the number of stamps and Y axis is the frequency of workshop.
-The distribution is widely diverse with most workshops having only one stamp}
+\caption{Histogram of stamps found in different workshops.
+Y axis shows the number of workshops where a given number of stamps have been found (X Axis).
+X axis is shown at a logarithmic scale given the long right tail}
 \label{stamps}
@@
 The distribution of codes in stamps for each \textit{conventus} can be seen in Figure~\ref{frequency}.
-The majority of stamps are concentrated in \textit{Hispalis} (574 stamps) while \textit{Corduba} and \textit{Astigi} have roughly half this sample size (267 and 146 stamps respectively).
-The workshops of the three \textit{conventus} show a similar pattern on the stamp frequency distribution with the exceptions of 2 large workshops in \textit{Hispalis} (La Catria and Arva).
+The majority of stamps are concentrated in \textit{Hispalis} (574 stamps) while \textit{Corduba} and \textit{Astigi} have roughly half this number (267 and 146 stamps respectively).
+The workshops of the three \textit{conventus} show a similar pattern on the stamp frequency distribution with the exception of 2 large workshops in \textit{Hispalis} (La Catria and Arva).
 On these two workshops a comparatively large amount of stamp codes were found (29 different codes on each of them).
+According to previous studies, those workshops became the most important centres of amphora production in the region \citep{rodriguez_economioleicola_1977,arva_1997}.
 
-According to previous studies, those workshops became the most important centres of amphora production in the region \citep{rodriguez_economioleicola_1977,
-arva_1997}.
-
-It is worth mentioning that a majority of these stamps were collected during field surveys with no excavation, so this difference in sample size could be biased due to differences in intensity across the different sites~\citep{arva_1997}.
+It is worth mentioning that a majority of these stamps were collected during field surveys while only a hanfdul of workshops have been excavated; this may explain some of the differences in sample size as there could be intensity bias across the different sites~\citep{arva_1997}.
 
@@
 \caption{Distribution of the number of different codes in stamps (X axis) for each \textit{conventus} (Y axis).
-Each dot corresponds to a workshop sorted by different areas.
+Each dot corresponds to a workshop.
 Colours are represented by areas divided into \textit{Hispalis} (red), \textit{Astigi} (green) and \textit{Corduba} (blue)}
@@
 
-The analysis of stamps in consumption areas also used the CEIPAC database and filtered in the same way than the previous dataset.
-However, in this case, we selected the centres with more or equal than five stamps in order to maximise the analysis with more stamps in the dataset.
+The analysis of stamps in consumption areas also used the CEIPAC database and filters performed for the previous dataset.
+However, in this case there was a large amount of sites with only one or two stamps which may introduce noise in the similarity analysis.
+This is the reason why in this case an additional filter was performed and we only selected sites having at least five stamps.
 
-The output was a dataset of 4271 stamps found in sites belonging to \textit{Britannia} (total: 2219 stamps, analysed: 1765 stamps ) and \textit{Germania} (total: 2052 stamps, analysed: 1621 stamps).
-The stamps with incomplete information were also not taken into account for the analysis.
-
-Query datasets from both provinces can be found on the following GitHub repository: Britannia~ \url{https://github.com/Mcotsar/Baetica_stamps/tree/main/Databases/Britannia} and Germania~ \url{https://github.com/Mcotsar/Baetica_stamps/tree/main/Databases/Germania}.
-
-Both \textit{Germania} and \textit{Britannia} were analysed as borders and not as Roman provinces.
-This means that we included in the sample some centres that are actually located beyond the administrative boundaries but were considered part of the same border.
+The output was a dataset of 3386 stamps with similar sample size: 1765 from \textit{Britannia} and 1621 for the \textit{Germania} provinces.
+Both \textit{Germania} and \textit{Britannia} were analysed as frontiers and not as administrative Roman provinces; we included in the sample some centres that are actually located beyond the administrative boundaries but were considered part of the same border.
 Specifically, for the German \textit{limes} we included sites from the two provinces (Superior and Inferior) while the analysis of \textit{Britannia} extended to \textit{Caledonia}.
@@
 %CITAR (Callender, 1965; Carreras Monfort y Funari,1998; Ayllón-Martı́ et al., 2018). 
-In \textit{Britannia}, we studied a total sample of 1765 stamps from 46 sites displaying 968 unique codes. The centres located in \textit{Britannia} can be seen in Figure~\ref{britannia}.
+The 1765 stamps from \textit{Britannia} showed 986 unique codes from 46 sites. These consumption centres can be seen in Figure~\ref{britannia}.
 
@@
 
-In \textit{Germania}, we collected a total of 1621 stamps from 46 sites displaying 850 different stamp codes (see Figure~\ref{germania}).
+The 1621 stamps from \textit{Germania} displayed 850 different code from the same number of sites (46) which can be seen in Figure~\ref{germania}.
 
@@
 
-\subsection{Measuring the dissimilarity}
+\subsection{Quantifying similarity}
 \label{sec:5}
 
-The approach proposed here aimed at exploring links between production and consumption centres through the identification of common amphorae codes found in different sites.
-For this reason we measured the similarity between amphora workshops and/or consumption sites by quantifying a pairwise distance or dissimilarity index between places (i.e.
+Similarity between amphora workshops and/or consumption sites was measured by quantifying a distance (also called dissimilarity) between the set of stamp codes found for each pair of sites (i.e.
 to what extent the stamp codes found on these sites were different).
-The chosen dissimilarity measure was the Morisita-Horn index \citep{morisita_measuring_1959, horn_measurement_1966}.
-This method was applied to measure the dissimilarity between different samples of stamps.
+The chosen measure for this study was the Morisita-Horn index \citep{morisita_measuring_1959, horn_measurement_1966}.
+This method was applied to measure the dissimilarity between different samples of stamps while taking into account the frequency that each code is present on each site.
 Generally, it describes the dissimilarity between the system of two communities based on the idea of inverse correlation between diversity and species \citep{magurran_why_1988}.
@@
 
-where $N_{a}$ is the total number of stamps in workshop A; $N_{b}$ is the total number of stamps in workshop B; $a_{i}$ is the number of different stamps for workshop A and $b_{i}$ is the number of different stamps for workshop B.
+where $N_{a}$ is the total number of stamps found in workshop A; $N_{b}$ is the total number of stamps found in workshop B; $a_{i}$ is the number of different stamps for workshop A and $b_{i}$ is the number of different stamps for workshop B.
 
 Considering our dataset as a non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity \citep{wolda_similarity_1981}.
-Morisita-Horn index gives a value between 0 (sites have exact stamp codes and frequency distribution) and 1 (complete difference between stamp codes).
+Morisita-Horn index gives a distance value between 0 (sites have exact stamp codes and frequency distribution) and 1 (complete difference between stamp codes).
 To apply Morisita-Horn index, we first calculated the number of times that each stamp code appeared in an amphora workshop.
 This method allowed us to bear in mind the number of frequency of stamps for each workshop.
-For example, if two workshops had similar stamp code distribution then the distance index would be close to 0 whereas sites with completely unrelated stamp codes would give a distance close to 1.
+For example, if two workshops had similar stamp code distribution then the distance index would be close to 0 whereas sites with no overlap between stamp codes would give a distance close to 1.
 
@@
 
-The Morisita-Horn index was used to generate a dissimilarity matrix containing the distance value between each pairwise site based on their codes stamped in amphorae.
-Hierarchical clustering was applied to this matrix in order to group sites with similar stamp codes distributions.
-The algorithm was selected to cluster similar groups in order to analyse the relationship between groups of sites and the distribution of similar stamp codes.
-The results were visualised using a dendrogram to detect groups of sites sharing similar stamp codes.
+Morisita-Horn was used to generate a pairwise matrix of distances between sites based on the overlap of the amphorae stamp codes recovered from them.
+Hierarchical clustering was then applied to this matrix in order to group sites based on the distance of their stamp codes.
+The construction of this similarity tree would allow us to analyse the relationship between groups of sites and the distribution of similar stamp codes.
+The results were visualised using a dendrogram to identify patterns within the different clusters.
 
@@
 
-The hierarchical clustering of the Morisita-Horn pairwise distance matrix can be seen as a dendrogram in Figure~\ref{dendro}.
-This visualisation suggests that each amphora workshop often used unique stamps in their production system that have not been found on any other site.
-In fact, the similarity values are always low and a majority of stamp codes found on multiple sites are only present in two or three workshops.
+The result of the hierarchical clustering inferred from Morisita-Horn measures can be seen in Figure~\ref{dendro}.
+This visualisation suggests that most amphora workshops used a single stamp code which has not been found in any other site.
+Simultaneously the majority of stamp codes found on multiple sites are only present in two or three workshops.
+The combination of these two patterns generates a map of low similarity values for the entire sample.
 
-Nearby workshops display a higher similarity than distant workshops and the varying degrees of similarity between sites which could be correlated to their spatial distance; additionally, the workshops with higher similarity values belong to the same \textit{conventus} area, such as Picachos, Cerro de los Pesebres and El Castillejo.
+Nearby workshops seem to display a higher similarity than distant workshops and the higher degrees of similarity between sites could be related to their spatial closeness; additionally, the workshops with higher similarity values belong to the same \textit{conventus} area, such as Picachos, Cerro de los Pesebres and El Castillejo.
 
@@
 \includegraphics[angle=270, width=0.7\linewidth]{dendro}
-\caption{Dendrogram displaying the results of hierarchical clustering based on Morisita-Horn metrics for Dressel 20 workshops located in the \textit{Baetica} area.
+\caption{Dendrogram displaying hierarchical clustering of \textit{Baetican} Dressel 20 workshops based on Morisita-Horn metrics.
 Site names are colour-coded by \textit{conventus}: \textit{Hispalis} (red), \textit{Astigi} (green) and \textit{Corduba} (blue)}
@@
 
-Results for the consumption regions display similar patterns than the ones found for the production area of \textit{Baetica}.
-The most important difference is that the similarity values are lower as can be seen in Figure~\ref{britmap}.
+Results for the consumption regions display similar patterns than the ones found in the production area of \textit{Baetica}.
+The most significant difference is the lower similarity values as can be seen in Figure~\ref{britmap}.
 
@@
 width=0.7\linewidth]{dendrobrit5.pdf}
-\caption{Dendrogram displaying the results of hierarchical clustering based on Morisita-Horn metrics for sites in \textit{Britannia}.
+\caption{Dendrogram displaying hierarchical clustering for \textit{Britannia} sites based on Morisita-Horn metrics.
 Site names are colour-coded by typology: military sites (red), civilian sites (green) and unspecified (grey)}
@@
 
-The lower similarity of stamp codes in \textit{Britannia} compared to \textit{Baetica} could be explained by the spatial correlation as the area under study is much bigger and the sites are more distant between them.
-This can be observed more clearly on military sites as the ones with higher similarity are geographically closer.
-Two patterns can be observed on the dendrogram: 1) nearby sites tend to share more similar stamps and 2) a majority of stamp codes are only found in one site.
+The lower similarity of stamp codes in \textit{Britannia} compared to \textit{Baetica} could be explained by a change of scale as the area under study is much bigger and the sites are more distant between them.
+The pattern can be more clearly observed for military sites which are often more well researched; in this case the ones with higher similarity tend to be geographically closer.
+To summarize, the same patterns than before can be observed on the dendrogram: 1) nearby sites tend to share more similar stamps and 2) a majority of stamp codes are only found in one site.
 
@@
 \includegraphics[angle=270, width=0.7\linewidth]{dendroger5.pdf}
-\caption{Dendrogram displaying the results of hierarchical clustering based on Morisita-Horn metrics for sites in \textit{Germania}.
+\caption{Dendrogram displaying hierarchical clustering for \textit{Germania} sites based on Morisita-Horn metrics.
 Site names are colour-coded by typology: military sites (red), civilian sites (green) and unspecified (grey)}
@@
 
-The results for \textit{Germania} follow a similar pattern than \textit{Britannia} but with a minor correlation as we can observe in the dendrogram.
-
-A higher concentration of sites sharing similar stamps was found in areas eminently militarised and close to German \textit{limes}, although most sites mostly showed different stamps.
+Again, results in \textit{Germania} exhibit similar patterns with lower similarities.
+The only difference is that a higher concentration of sites sharing similar stamps was found in militarised areas near the \textit{limes}.
 
@@
 
-The analysis of the amphora workshops in \textit{Baetica} province suggests that a majority of stamp codes were only used on a single workshop.
-Beyond this clear pattern, it seems that the similarity of stamp codes is correlated with spatial distance and the amphoric production of spatially close sites shared some stamp codes.
+The analysis of stamp similarity between workshops in \textit{Baetica} province suggests that a majority of stamp codes were only used in a single workshop.
+Beyond this clear pattern, it seems that the similarity of stamp codes is linked to proximity and the amphoric production of spatially close sites shared some stamp codes.
 Particularly relevant here is the fact that our results show that each \textit{conventus} area had similar stamps codes and they were not shared in other areas.
 This result could indicate that the workshops of the different \textit{conventus} used different stamp codes that were not shared between them.
-Despite these stamps tend to share the same area of production, we do not identify groups of any more than three workshops sharing the same amphoric stamps as the dendrogram showed.
+Within each conventus we can identify small groups of three/four workshops sharing the same amphoric stamps as seen in the dendrogam, but there are no stamp codes that may be found across several workshops.
+This pattern of similarity based on spatial closeness is closely aligned with previous studies that focused on morphometric distances between amphorae made on different workshops which also showed spatial correlation and large differences between \textit{conventus}~\citep{coto-sarmiento_identifying_2018}.
+Finally, while the role of the river was significant for the distribution of amphorae, river connectivity amongst workshops does not seem to show relevancy for the distribution or concentration of stamps in a specific river area.
 
-Our results do not support previous working hypotheses suggesting that groups of workshops used specific groups of stamp codes; while there is higher stamp similarity for nearby workshops this result does not suggest any kind of organisation beyond the simple fact that nearby workshops had a higher degree of interaction.
-In fact; the opposite seems true: each workshop used different stamp codes and only shared some of them with the closest ones.
+On the other hand, our results do not seem to support the hypothesis that groups of workshops collectively used specific sets of stamp codes.
+There is higher similarities between nearby workshops but this could be simply explained by spatial proximity which would enable more frequent interactions.
+The opposite seems valid: each workshop used different stamp codes and only shared some of them with the closest ones.
 
@@
 First, it is possible that each workshop operated independently and did not normally collaborate with other workshops.
-Second, stamp similarity in closer workshops could be linked to a spatial pattern.
-It is more probable than closer workshops tend to share more traits than distant workshops.
-While the role of the river was significant for the distribution of amphorae, river connectivity amongst workshops does not seem to show relevancy for the distribution or concentration of stamps in a specific river area.
+Second, as discussedd stamp similarity in closer workshops could be linked to more intense interactions.
+Finally, the presence of different stamps in the same workshop could suggest some degree of internal organisation that would only affect nearby centres~\citep[104]{juanmorostesis}.
+For example, the stamping process could have allowed potters to sort out the different production batches and plan the shipments.
+Systematic procedures ike this one would probably be required to organise such a massive specialized production.
 
-Alternatively, the presence of different stamps in the same workshop could imply some kind of organisation from within the workshop only affecting nearby centres \citep{juanmorostesis}.
-The stamped amphorae could have allowed potters to organise the different production batches for posterior commercialisation.
-
-This could be explained such as batch systematic organisation between potters \citep{juanmorostesis}.
-This work method could have allowed potters to identify batches by stamping amphorae for posterior commercialisation.
-High production of amphorae had to maintain a specialise system organisation to control this production \citep[104]{juanmorostesis}.
-
-It is worth mentioning that the performed analysis could be limited by a diversity of biases.
-Some workshops have been identified with different names despite the fact that they probably are sections of the same workshop; on the other hand, the amount of stamps collected on different sites is affected by different intensities for field surveys a lack of excavation.
-As a consequence, further fieldwork could provide a clearer picture of the links between workshops if the sample size of stamp codes is increased.
+As we previously discussed some bias on the intensity of archaeological works and workshop identification may have an impact on this analysis.
+Further fieldwork could provide a clearer picture of the links between workshops if they allow to increase the sample size of stamp codes is increased.
 
@@
 
-Both consumption areas showed a correlation between the spatial distance of sites and their similarity of amphoric stamps.
-In the case of  \textit{Britannia}, the correlation was higher than \textit{Germania}.
-In \textit{Britannia} the majority of similarity stamps were mainly found in military centres.
+Both consumption areas showed a similar link between spatial proximity and similarity of amphoric stamps.
+In the case of  \textit{Britannia} this pattern was stronger than \textit{Germania}; this could be related to the fact that a majority of shared stamp codes in \textit{Britannia} were found in military centres.
 
 A combined exploration of the map and the dendrogram also suggests an interesting pattern: centres with higher similarity values are closer to the coast (North Sea and Celtic Sea).
-This may indicate that the Atlantic route could have played an essential role in transporting olive oil to the area of \textit{Britannia}, since in the places where there is greater similarity they are found in different strategic points near the sea.
+This may indicate that the Atlantic route could have played an essential role in transporting olive oil to the area of \textit{Britannia}, since sites with higher degrees of similarity were strategic points near the sea.
 
-This trend is aligned with recent works that have suggested that the Atlantic trade routes were mostly used for military supply~\citep{remesal_annona_1986,
-remesal_provincial_2008,
-carreras_atlantic_2012,
-morillo_hispania_2016}.
-Therefore, a majority of centres with higher similarity are related to military activities and near the coasts, thus suggesting that Dressel 20 containing olive oil were shipped to military areas and then redistributed to the civilian population using land-based transportation systems~\citep{carreras_britannia_1998,
-orengo_seeds_2016,
-ayllon_olive_2018}.
+This trend is aligned with recent works that have suggested that the Atlantic trade routes were mostly used for military supply~\citep{remesal_annona_1986,remesal_provincial_2008,carreras_atlantic_2012,morillo_hispania_2016}.
+Therefore, a majority of centres with higher similarity are related to military activities and near the coasts, thus suggesting that Dressel 20 containing olive oil were shipped to military areas and then redistributed to the civilian population using land-based transportation systems~\citep{carreras_britannia_1998,orengo_seeds_2016,ayllon_olive_2018}.
 
-The areas mostly militarised share similar stamps than civil areas.
-However, we do not detect a concrete pattern regarding the distribution of the stamps in the German \textit{limes} \citep{xanten2018}.
-This could be also interpreted by an intensity bias where military centres have been mostly excavated than civil centres.
+The areas in \textit{Britannia} with most intense military activity have some degree of overlap with civil areas.
+However, we do not detect this overlap for the the German \textit{limes} \citep{xanten2018}.
+This could be interpreted by parallel supply systems but it also could be caused by intensity bias as military centres have been more excavated than civil centres.
 
-Finally, the results suggest that there is no clear organisational pattern linking production centres and/or consumption centres beyond the spatial distance.
-As a consequence, our analysis does not indicate the presence of a specific workshop from \textit{Baetica} supplying to some specific province.
-
-This evidence means that it is unknown whether some production centres went to one province or another or, at least, that they can be clearly reflected in the data with a greater similarity in the amphoric stamps.
-Thus, production centres could have distributed randomly olive oil to both \textit{Britannia} and \textit{Germania}.
-Neither we do not detect production centres dedicated to the distribution of olive oil in a certain province.
+Finally, the results suggest that there is no clear organisational pattern linking production centres and/or consumption centres beyond local spatial proximity within regions.
+As a consequence, our analysis do not indicate the presence of specific Baetican amphoric workshop consistently supplying specific regions or even provinces.
+This outcome suggests that stamp codes cannot be used to identify whether production centres had fixed clients and other proxies would be needed to identify this link in case it existed.
+Based on our analysis production centres could have distributed olive oil to both \textit{Britannia} and \textit{Germania} at a short-term basis of demand and without any long-term planning or centralised decision-making system.
 
@@
 
-This work has used quantitative analysis to explore long-range trade of olive oil using the stamps found in Dressel 20 vessels.
+This work has used quantitative analysis to explore the structure of olive oil long-range trade using the stamps found in Dressel 20 amphora vessels.
 The most interesting result is linked to the meaning of these inscriptions.
@@
 an amphora with a given stamp was probably made on a specific workshop).
-This idea could explain why the similarity of stamps between workshops is generally very low.
-On the other hand, some codes in stamps have been identified in nearby workshops so this pattern suggests that the clustered workshops were somehow connected (e.g.
-they belonged to the same owner).
+This hypothesis is sustained by the fact that the similarity of stamps between workshops is very low and a majority of stamp codes are only found at one workshop.
+On the other hand, the presence of stamp codes in small groups of nearby workshops could suggest that the production of these workshops was somehow connected (e.g.
+they belonged to the same owner or were collectively producing a batch for the same producer).
 
-A large number of Dressel 20 were not marked with any stamps, so some researchers suggest that potters marked the amphorae to prepare and distribute the product in order to be shipped \citep{berni_millet_epigrafianforica_2008}.
-This method would be used as an identifier to count the number of amphorae of a batch \citep{juanmorostesis}.
+One interesting idea from this reasoning is that there is a reason behind the fact that a large number of Dressel 20 were not marked with any stamps; some researchers suggest that potters marked some amphorae of each batch to prepare and distribute the product in order to be shipped \citep{berni_millet_epigrafianforica_2008}.
+As a consequence, this method could potentially be used to count the number of amphorae of a batch or even the number of batches~\citep{juanmorostesis}.
 
@@
 This hypothesis could explain why we detect different stamps in the same workshop.
-In any case, at present we do not have enough archaeological evidence to validate which of these interpretations is more plausible in our case study.
+We will need additional archaeological evidence and computational methods such as the one presented here to test which of these interpretations is more plausible.
 
-To conclude, the method presented here provides a useful framework to improve our understanding of the organisation and processes that allowed the Roman Empire to manage a massive and highly specialised production of essential goods.
-This work has identified similarities between amphoric stamps and spatial distribution in the case of the amphoric production within the Roman Empire.
+To conclude, this work provides a useful framework to improve our understanding of the organisation and processes that allowed the Roman Empire to manage a massive and highly specialised production of essential goods.
+The analysis has identified similarities between amphoric stamps and spatial distribution in the case of the amphoric production within the Roman Empire.
+Beyond olive oil, similarity analysis could be used to identify trade links between other goods and contexts provided that an effective archaeological proxy is identified and collected with enough sample size.
 
-The growing amount of archaeological data requires of large-scale quantitative methods such as the one presented here to identify and explore complex patterns, thus allowing us to infer dynamics and improve our interpretation of complex economic processes of past societies.
+The growing amount of archaeological data we see today requires of large-scale quantitative methods to identify and explore complex patterns.
+This emrgence of new computational tools should allow us to infer multiscale dynamics and improve our interpretation of the economic processes of past societies.
 
@@
 XRC is funded by the Ram\'on y Cajal programme RYC2018-024050-I (Fondo Social Europeo – Agencia Estatal de Investigaci\'on).
-We would like to thank the editor and reviewers (Stephen Shennan
-and two anonymous reviewers) for their constructive comments that helped improve the original manuscript.
+We would like to thank the editor and two anonymous reviewers for their constructive comments that helped improve the original manuscript.
 We are grateful to Simon Carrignon, Juan Moros, Jordi P\'erez-Gonz\'alez and V\'ictor M.

### 2021-02-10 Manuscript/Manuscript.tex
review hasta 3

@@
 
-The aim of this study is to explore the trade dynamics involved on the large-scale distribution of olive oil produced in \textit{Baetica} province (currently Andalusia).
+The aim of this study is to explore the trade dynamics involved in the large-scale distribution of olive oil produced in \textit{Baetica} province (currently Andalusia).
 \textit{Baetica} became an important production and distribution centre during the Roman Empire and its goods have been found in thousands of sites, but it is still unclear how exactly this large scale production was organised.
@@
 We explore here connections between production and consumption sites during the period of most intense activity (1st to 3rd centuries AD).
-Specifically, we use amphoric stamps as proxies to identify the workshops where amphorae found on consumer sites were produced.
+Specifically, we use amphoric stamps as proxies to identify the workshops where amphorae located in consumer sites were produced.
 
@@
 This dataset is used to detect links between specific sites that may share a large percentage of common stamps.
-The analysis is performed using quantitative methods borrowed from Ecology that allows us to group sites based on the amount of stamps shared between them.
-Results highlight the structure of economic processes of the Roman Empire using the archaeological evidence of olive oil large-scale distribution.
+The analysis is performed using quantitative methods borrowed from Ecology that allows us to group sites based on the number of stamps shared between them.
+
+%The analysis explores how the quantitative approach provides a useful tool for the interpretation of the economic processes.
+
+Finally, results highlight the structure of economic processes of the Roman Empire using the archaeological evidence of olive oil large-scale distribution.
 
@@
 The Empire developed a series of structures to support and organise this long-range trade and specialised entire regions to produce specific goods at a massive scale.
-The province of \textit{Baetica} (currently Andalusia, southern Spain) is an example of this specialization process.
+The province of \textit{Baetica} (currently Andalusia, southern Spain) is an example of this specialisation process.
 \textit{Baetica} became one of the most important olive oil production centres during the Roman Empire; olive oil was an essential good for Romans because it was used in almost every aspect of their daily life such as cooking, hygiene or lighting~\citep{mattingly_d.j._oil_1988}.
-This high demand required a huge increase on the production which needed to be distributed to cities across thousands of kilometers.
+This high demand required a huge increase in the production which needed to be distributed to cities across thousands of kilometres.
 This large-scale trade was organised via the shipment of millions of amphorae using maritime and riverine routes to all the provinces and particularly to Italy and the military garrisons guarding the borders of the Empire~\citep{blazquez_exportacion_1980}.
@@
 The structure and processes associated with this massive olive oil production have been extensively discussed over the last decades\citep{rodriguez_economioleicola_1977, duncan1982economy, Chic_hispania_1997,millet_anforas_1998}.
-However, the complexity and sheer scale of the process makes its understanding challenging despite several archaeologicale xcavations and the existence of some written records discussing its structure.
+However, the complexity and sheer scale of the process makes its understanding challenging,
+%se repite despite aquí dos veces
+despite several archaeological excavations and the existence of some written records discussing its structure.
 Our knowledge of this economic activity has benefited from new findings and data sources, but despite all these advances several questions remain open: How was this massive amphora production organised?
 Can we detect the economic ties between provinces?
-Was this distribution centralized or each region organised it differently?
+Was this distribution centralised or each region organised it differently?
 
 Recent advances in Roman studies have currently led to an environment with various interpretations on these commercial dynamics~\citep{temin_economy_2006,quantifyingwilson2009}.
-The application of diverse novel quantitative approaches to new accessible datasets has allowed us to improve our interpretation of the growing amount of archaeological evidence and contribute to a better understanding of the dynamics of Roman production\citep{brughmans_roman_2016,orengo_seeds_2016,bayesian_2018,coto-sarmiento_identifying_2018,rubio-campillo_ecology_2018}.
+%The application of diverse novel quantitative approaches to new accessible datasets has allowed us to improve our interpretation of the growing amount of archaeological evidence and contribute to a better understanding of the dynamics of Roman production
+The application of novel quantitative approaches has allowed us to improve our interpretation of the growing amount of archaeological evidence and contribute to a better understanding of the Roman production
+\citep{brughmans_roman_2016,orengo_seeds_2016,bayesian_2018,coto-sarmiento_identifying_2018,rubio-campillo_ecology_2018}.
 This work aligns with these efforts and focuses on the distribution of olive oil towards Roman provinces in the borders.
@@
 The finding of identical amphoric stamps in different provinces has allowed us to explore patterns of olive oil distribution within the Empire.
-Despite this archaeological evidence there is still no consensus on the structure that generated this pattern; how was the economic relation between production and consumption centres?
+Despite this archaeological evidence, there is still no consensus on the structure that generated this pattern; how was the economic relation between production and consumption centres?
 Did they follow a centralised process to distribute olive oil in those provinces or each workshop organised its own trade networks?
-
-This paper aims to explore these questions regarding the olive oil market by identifying links between productionn centres and consumption provinces using similarity indexes for amphoric stamps.
+%repite aims dos veces  
+This paper aims to explore these questions regarding the olive oil market by identifying links between production centres and consumption provinces using similarity indexes for amphoric stamps.
 In this way, our work aims to detect regional patterns for a specific commercial product (i.e.
@@
 In \textit{Baetica} province we want to identify links between workshops based on the similarity of their stamps.
-In the consumption Roman provinces our aim is to detect groups of nearby sites sharing stamps and to what extent stamps are clustered based on provinces.
-This economic connection could be identified with two ideas : a) nearby sites have more similar amphoric stamps and b) sets of stamps can only be found on specific provinces because they were shipped together.
+In the consumption Roman provinces, our aim is to detect groups of nearby sites sharing stamps and to what extent stamps are clustered based on provinces.
+This economic connection could be identified with two ideas: a) nearby sites have more similar amphoric stamps and b) sets of stamps can only be found on specific provinces because they were shipped together.
+%find in or find on
 
-%In particular, we study the distribution of amphoric stamps to identify a correlation between geographical distance and similarity.
-
-We proposed three working hypotheses based on these ideas: a) we can identify a correlation between spatial distance and the distribution of amphoric stamps, b) stamp sets found in nearby workshops should be more similar and c) there is a low mobility of amphoric stamps from consumption centres to other regions (i.e.
+We proposed three working hypotheses based on these ideas: a) we can identify a correlation between spatial distance and the distribution of amphoric stamps, b) stamp sets found in nearby workshops should be more similar and c) there is low mobility of amphoric stamps from consumption centres to other regions (i.e.
 stamps always stay in the same region).
@@
 This ecological approach is based on three steps: a) to quantify stamp similarity between sites, b) to group sites based on this similarity and c) to find correlation patterns between stamp similarity and proximity.
-The rationale behind the method is as follows: if trade structure was centrally planned then we should be able to stablish links between workshops and consumption sites based on the similarity of its stamps; if this is not the case then results would suggest that workshops worked independently and agreements were made at an individual level.
+The rationale behind the method is as follows: if trade structure was centrally planned then we should be able to establish links between workshops and consumption sites based on the similarity of its stamps; if this is not the case then results would suggest that workshops worked independently and agreements were made at an individual level.
 
-%Stamps will be used to identify economic patterns by analysing their similarity.
-
-The paper discusses these questions, the analysis and results as follows: next section introduces the historical context.
+The paper addresses these questions, the analysis and results as follows: next section introduces the historical context.
 Section two defines the dataset and the methods used for the analysis.
-Section three presents the results and section 4 discusses them; the paper finishes with some concluding remarks on the research performed.
+Section three presents the results and section four discusses them; the paper finishes with some concluding remarks on the research performed.
 
@@
 The \textit{Baetican} province was one of the most suitable regions to tackle the challenge of the huge demand for olive oil across the Roman provinces.
-For this reason, the area saw an increase in productivity as a massive specialized infrastructure of olive oil production was gradually deployed.
-This production of olive oil had the largest peark from the first to the third century AD~\citep{remesal_concierto}.
+For this reason, the area saw an increase in productivity as a massive specialised infrastructure of olive oil production was gradually deployed.
+This production of olive oil had the largest peak from the first to the third century AD~\citep{remesal_concierto}.
 
@@
 The chronology of the Baetican workshops is typically defined between the first to the third centuries AD~\citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
-Archaeological evidence suggests that this highly specialised production was developed across a long activity period\citep{remesal_anforas_2004}, but at the same time the process seemed to have been planned at some level as is observed by the relatively minor changes in amphoric typologies and the high degree of standardisation on the amphorae found across at least 50 workshops.
+Archaeological evidence suggests that this highly specialised production was developed across a long activity period\citep{remesal_anforas_2004}.
+The process seemed to have been planned at some level as is observed by the relatively minor changes in amphoric typologies and the high degree of standardisation in the production.
 
@@
 
-Dressel 20 was the amphora most stamped during the Roman Empire ~\citep[18]{millet_anforas_1998}.
+Dressel 20 was the amphora most stamped during the Roman Empire~\citep[18]{millet_anforas_1998}.
 The large sample size of recovered Dressel 20 stamps is distributed across all Western provinces of the Roman Empire.
@@
 \end{figure} 
-
+%found in find in
 Most stamps seemed to have been used across a long activity period which can be difficult to date with the required level of accuracy.
@@
 The chronological information of \textit{tituli picti} is very useful, but the small amount of dates recovered and the fact that a vast majority of them come from the same site poses some challenges to this precise dating.
-In addition, the fact that \textit{Monte Testaccio} was a dump suggests that the diposition of amphorae vessels may have been altered at a later phase due to changes in the long-term dynamics of creation of this artificial mound.
+In addition, the fact that \textit{Monte Testaccio} was a dump suggests that the disposition of amphorae vessels may have been altered at a later phase due to changes in the long-term dynamics of creation of this artificial mound.
 
@@
 On the other hand, they could also belong to the owner of the workshops where the amphorae were made or even to a group of amphora workers~\citep{berni_millet_epigrafianforica_2008}.
-Despite the unsolved question of ownership some interesting ideas can be explored based on the finding of similar stamps across the Empire.
-Observable patterns can be used to explore how this production was organised and whether it is possible to identify specific production processes fdor olive oil trade.
+Despite the unsolved question of ownership, some interesting ideas can be explored based on the finding of similar stamps across the Empire.
+Observable patterns can be used to explore how this production was organised and whether it is possible to identify specific production processes for olive oil trade.
 Our questions here will be focused on the distribution of amphoric stamps; were different stamps shipped through different distribution routes?

### 2021-02-11 Manuscript/Manuscript.tex
second review ready

@@
 
-To achieve this goal, we analyse a dataset of stamps found in two types of sites: 1) workshops from \textit{Baetica} province and 2) consumption centres located in two provinces \textit{Germania} (Inferior and Superior) and \textit{Britannia}.
+To achieve this goal, we analyse a dataset of stamps found in two types of sites: 1) workshops from \textit{Baetica} province and 2) consumption centres located in two provinces: \textit{Germania} (Inferior and Superior) and \textit{Britannia}.
 This dataset is used to detect links between specific sites that may share a large percentage of common stamps.
@@
 The finding of identical amphoric stamps in different provinces has allowed us to explore patterns of olive oil distribution within the Empire.
-Despite this archaeological evidence, there is still no consensus on the structure that generated this pattern; how was the economic relation between production and consumption centres?
+Despite this archaeological evidence, there is still no consensus on the structure that generated this pattern; how were the economic relations between production and consumption centres?
 Did they follow a centralised process to distribute olive oil in those provinces or each workshop organised its own trade networks?
@@
 
-%We want to understand the pattern of olive oil production linked to amphora workshops and stamps used to mark them. 
-%We focus here on exploring the economic relation between stamps and amphora production and distribution centres. 
-This relation between the production centre (i.e.
+This relationship between the production centre (i.e.
 \textit{Baetica}) and consumption sites have been studied here for two case studies: \textit{Britannia} and \textit{Germania} (see Figure~\ref{general}).
@@
 \caption{Sites analysed in this paper.
-The colour defines the three different regions under study (\textit{Baetica} (yellow) as producer; \textit{Germania} (blue) and \textit{Britannia} (green) as consumers) while the size of each dot is proportional to the number of Dressel 20 amphoric stamps found on each site}
+The colour defines the three different regions under study (\textit{Baetica} (yellow) as producer; \textit{Germania} (blue) and \textit{Britannia} (green) as consumers) while the size of each dot is proportional to the number of Dressel 20 amphoric stamps found in each site}
 
@@
 In the consumption Roman provinces, our aim is to detect groups of nearby sites sharing stamps and to what extent stamps are clustered based on provinces.
-This economic connection could be identified with two ideas: a) nearby sites have more similar amphoric stamps and b) sets of stamps can only be found on specific provinces because they were shipped together.
+This economic connection could be identified with two ideas: a) nearby sites have more similar amphoric stamps and b) sets of stamps can only be found in specific provinces because they were shipped together.
 %find in or find on
@@
 To do this, a population-based approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}.
-This quantitative method provides a way to measure the similarity between centre based on the amphoric stamps found on each of them.
+This quantitative method provides a way to measure the similarity between centre based on the amphoric stamps found in each of them.
 This ecological approach is based on three steps: a) to quantify stamp similarity between sites, b) to group sites based on this similarity and c) to find correlation patterns between stamp similarity and proximity.
@@
 
-The paper addresses these questions, the analysis and results as follows: next section introduces the historical context.
-Section two defines the dataset and the methods used for the analysis.
-Section three presents the results and section four discusses them; the paper finishes with some concluding remarks on the research performed.
+The paper addresses these questions, the analysis and results as follows: next sections introduce the historical context.
+Section five defines the dataset and the methods used for the analysis.
+Section six presents the results and section seven discusses them; the paper finishes with some concluding remarks on the research performed.
 
@@
 
-The \textit{Baetican} province was one of the most suitable regions to tackle the challenge of the huge demand for olive oil across the Roman provinces.
+The province of \textit{Baetica} was one of the most suitable regions to tackle the challenge of the huge demand for olive oil across the Roman provinces.
 For this reason, the area saw an increase in productivity as a massive specialised infrastructure of olive oil production was gradually deployed.
@@
 \includegraphics[width=\linewidth]{baetica}
-\caption{Distribution of stamps found in Dressel 20 workshops along the Guadalquivir riverbank. The size of each point depends on the number of stamps found on each site}
+\caption{Distribution of stamps found in Dressel 20 workshops along the Guadalquivir riverbank. The size of each point depends on the number of stamps found in each site}
 
@@
 Most stamps seemed to have been used across a long activity period which can be difficult to date with the required level of accuracy.
-They have been frequently dated using as reference consular dates found on nearby \textit{tituli picti} recovered from \textit{Monte Testaccio}~\citep{Testaccio1, berni_millet_epigrafianforica_2008}.
+They have been frequently dated using as reference consular dates found in nearby \textit{tituli picti} recovered from \textit{Monte Testaccio}~\citep{Testaccio1, berni_millet_epigrafianforica_2008}.
 The chronological information of \textit{tituli picti} is very useful, but the small amount of dates recovered and the fact that a vast majority of them come from the same site poses some challenges to this precise dating.
@@
 The use of amphoric stamps can be a good proxy to explore the system of olive oil production and trade developed in \textit{Baetica} province.
-These stamps are typically found on handles but they could also be imprinted on rims and the amphora body~\citep{millet_anforas_1998}.
+These stamps are typically found in handles but they could also be imprinted on rims and the amphora body~\citep{millet_anforas_1998}.
 The content information of the stamps was not defined by a single criteria, but they often displayed a code composed by letters and additional symbols.
@@
 These needs effectively led to a gradual change in the economic and social structures as the trade networks that supplied the provinces and particularly the Roman army became more complex.
-An interesting example of this system is the figure of the \textit{praefectura annonae} which was created during Augustus's administration to organize wheat supply.
+An interesting example of this system is the figure of the \textit{praefectura annonae} which was created during Augustus's administration to organise wheat supply.
 The role of the \textit{praefectura annonae} was mainly focused on providing through \textit{frumentationes} a fixed monthly amount of wheat to each Roman citizen~\citep{remesal_annona_1986,remesal_concierto}.
@@
 The local population did not consume this product and it was not produced in the region as the landscape and climate of the British Isles was not suitable for the growth of olive trees~\citep[161]{monfort_britanniaen_1998}.
-Archaeological records suggest that this absence of olive oil changed with the arrival of a large amount of Dressel 20 amphorae to \textit{Britannia} from the first century A.D.
-onwards \citep{peacock_amphorae_1991,carreras_britannia_1998}.
-At this moment, we detect an increase of the olive oil exports linked to the arrival of legions caused by different military campaigns~\citep[161]{monfort_britanniaen_1998}.
+Archaeological records suggest that this absence of olive oil changed with the arrival of a large amount of Dressel 20 amphorae to \textit{Britannia} from the first century AD onwards \citep{peacock_amphorae_1991,carreras_britannia_1998}.
+At this moment, we detect an increase in the olive oil exports linked to the arrival of legions caused by different military campaigns~\citep[161]{monfort_britanniaen_1998}.
 
-The increase of Dressel 20 exports created an important commercial network for exchanges.
-The demand for olive oil after the Roman conquest required of a new trade route due to the lack of local olive oil production; \textit{Baetica} was chosen as its main supply.
-This economic relation will be particularly intense for sites close to Hadrian Wall's garrisons.
+The increase in Dressel 20 exports created an important commercial network for exchanges.
+The demand for olive oil after the Roman conquest required a new trade route due to the lack of local olive oil production; \textit{Baetica} was chosen as its main supply.
+This economic relationship will be particularly intense for sites close to Hadrian Wall's garrisons.
 Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic locations~\citep{carreras_atlantic_2012}.
 This presence of Dressel 20 stamps in military camps in \textit{Britannia} has been widely studied in Roman archaeology \citep{williams_importation_1983,funari_corpus_1996,carreras_britannia_1998,carreras_abastecimiento_2003}.
-This intense consumption of olive oil also suggests the existance of provincial structures created to organise the supply of olive oil across military camps.
-There are no written records detailing how this redistribution of essential goods was organised in \textit{Britannia}, but the archaeological evidence suggests that cities may have been the central nodes of this local trade network~\citep{funari_economic_2005,orengo_seeds_2016}.
+This intense consumption of olive oil also indicates the existence of provincial structures created to organise the supply of olive oil across military camps.
+No written records detailing how this redistribution of essential goods was organised in \textit{Britannia}, but the archaeological evidence suggests that cities may have been the central nodes of this local trade network~\citep{funari_economic_2005,orengo_seeds_2016}.
 
-The consumption of Baetican olive oil would experience a progressive slowdown from the third century A.D.
-onwards.
+The consumption of Baetican olive oil would experience a progressive slowdown from the third century AD onwards.
 This date matches a change in market strategy in the Empire.
-This gradual decrease on \textit{Baetican} olive oil exports can be observed as a decrease in the amount of Dressel 20 found in excavated sites as this typology is replaced by other types of amphorae from other places and particularly the African provinces~\citep{rodriguez1991aceite,millet_anforas_1998}.
+This gradual decrease in \textit{Baetican} olive oil exports can be observed as a decrease in the amount of Dressel 20 found in excavated sites as this typology is replaced by other types of amphorae from other places and particularly the African provinces~\citep{rodriguez1991aceite,millet_anforas_1998}.
 
@@
 
-The distribution of Dressel 20 amphorae in the provinces of Germania Inferior and Germania Superior is not as well interpreted as the rest of provinces discussed in this work~\citep[293]{remesal_baetica_2002}.
+The distribution of Dressel 20 amphorae in the provinces of \textit{Germania Inferior} and \textit{Germania Superior} is not as well interpreted as the rest of the provinces discussed in this work~\citep[293]{remesal_baetica_2002}.
+%it is clear no se si ponerlo
 It is clear that there was a massive import of Dressel 20 in both provinces as is observed in archaeological excavations across the German \textit{limes}.
@@
 
-In any case, the introduction of olive oil in \textit{Germania} shows similar patterns than the one discussed for \textit{Britannia}.
+In any case, the introduction of olive oil in \textit{Germania} shows similar patterns to the one discussed for\textit{Britannia}.
 There is no archaeological evidence for olive oil consumption before the arrival of the Roman army to the region, while a majority of recovered Dressel 20 amphorae are located in the military sites built over the German \textit{limes} \citep{remesal_germaniaengl_2002}.
@@
 
-Some studies assumed that Roman naval technology was not suited for oceanic transport and for this reason defined riverine shipping across Rhône and Rhine rivers as the basic supply route for the German \textit{limes}.
-However, more recent large-scale computational analysis as well as new excavations on the Iberian Atlantic coast suggest that the Atlantic route, which was far more efficient, may have played a major role on the supply of \textit{Germania} and \textit{Britannia}~\citep{rubio-campillo_ecology_2018}.
+Some studies assumed that Roman naval technology was not suited for oceanic transport and for this reason defined riverine shipping across Rh\^one and Rhine rivers as the basic supply route for the German \textit{limes} \citep{marliere2001tonneau}.
+However, more recent large-scale computational analysis, as well as new excavations on the Iberian Atlantic coast suggest that the Atlantic route, which was far more efficient, may have played a major role in the supply of \textit{Germania} and \textit{Britannia}~\citep{rubio-campillo_ecology_2018}.
 
@@
 The goal of this study is to explore these economic and long-range trade dynamics using amphoric stamps as proxies for common supply networks.
-We are especially interested on identifying links between Baetican production centres and consumption centres in the different \textit{limes}.
+We are especially interested in identifying links between Baetican production centres and consumption centres in the different \textit{limes}.
 
-For this analysis we used data collected in the CEIPAC database of amphoric epigraphy~\citep{remesal_centro_2015}.
+For this analysis, we used data collected in the CEIPAC database of amphoric epigraphy~\citep{remesal_centro_2015}.
 The CEIPAC dataset contains over 50.000 epigraphic records found on different types of amphorae (the database is openly accessible from \url{http://romanopendata.eu}).
-A large percentage of these records come from from \textit{Monte Testaccio}, but the rest is distributed mainly across the Western provinces of the Empire.
-It is worth mentioning that the database does not provide temporal information due to the reasons previously discussed, and for this reason this work can only analyze the spatial distributions within a coarse chronology of 2 centuries (3rd-1st AD).
+A large percentage of these records come from \textit{Monte Testaccio}, but the rest is distributed mainly across the Western provinces of the Empire.
+It must be mentioned that the database does not provide temporal information due to the reasons previously discussed, and for this reason, this work can only analyse the spatial distributions within a coarse chronology of three centuries (1st-3rd AD).
 
@@
 
-It is worth mentioning that some workshops were coded under the same geographical coordinates.
+It is worth noting that some workshops were coded under the same geographical coordinates.
 For instance, the stamps of two nearby amphora workshops such as Tesorillo de Do\~na Mencia and Do\~na Mencia were catalogued under different names while sharing the same geographical coordinates.
-The interpretation of our analysis considered this bias in the archaeological dataset which may have a minor effect in the final results.
+The interpretation of our analysis considered this bias in the archaeological dataset which may have a minor effect on the final results.
 
@@
 This would allow us to study the distribution of the stamps across centres as can be seen in Figure~\ref{stamps}.
-As can be seen the distribution is widely diverse with most workshops having only one stamp, while a few workshops contained over 100 stamps.
+The distribution is widely diverse with most workshops having only one stamp, while a few workshops contained over 100 stamps.
 Particularly, one workshop (La Catria) concentrated a large percentage of the amphoric stamps with a total number of 228 stamps.
 
-This uneven distribution of stamps is typical for amphora production as it is observed that a major concentration on the number of stamps is focused on a few workshops~\citep{coto-sarmiento_identifying_2018}.
+This uneven distribution of stamps is typical for amphora production as it is observed that a major concentration on the number of stamps is focused on a few workshops~\citep{bayesian_2018,coto-sarmiento_identifying_2018}.
 
@@
 The majority of stamps are concentrated in \textit{Hispalis} (574 stamps) while \textit{Corduba} and \textit{Astigi} have roughly half this number (267 and 146 stamps respectively).
-The workshops of the three \textit{conventus} show a similar pattern on the stamp frequency distribution with the exception of 2 large workshops in \textit{Hispalis} (La Catria and Arva).
-On these two workshops a comparatively large amount of stamp codes were found (29 different codes on each of them).
+The workshops of the three \textit{conventus} show a similar pattern on the stamp frequency distribution with the exception of two large workshops in \textit{Hispalis} (La Catria and Arva).
+In these two workshops a comparatively large amount of stamp codes were found (29 different codes on each of them).
 According to previous studies, those workshops became the most important centres of amphora production in the region \citep{rodriguez_economioleicola_1977,arva_1997}.
 
-It is worth mentioning that a majority of these stamps were collected during field surveys while only a hanfdul of workshops have been excavated; this may explain some of the differences in sample size as there could be intensity bias across the different sites~\citep{arva_1997}.
+It is worth mentioning that a majority of these stamps were collected during field surveys while only a handful of workshops have been excavated; this may explain some of the differences in sample size as there could be intensity bias across the different sites~\citep{arva_1997}.
 
@@
 
-The analysis of stamps in consumption areas also used the CEIPAC database and filters performed for the previous dataset.
-However, in this case there was a large amount of sites with only one or two stamps which may introduce noise in the similarity analysis.
+The analysis of stamps in consumption areas used also the CEIPAC database and filters performed for the previous dataset.
+However, in this case there was a large number of sites with only one or two stamps which may introduce noise in the similarity analysis.
 This is the reason why in this case an additional filter was performed and we only selected sites having at least five stamps.
@@
 The output was a dataset of 3386 stamps with similar sample size: 1765 from \textit{Britannia} and 1621 for the \textit{Germania} provinces.
-Both \textit{Germania} and \textit{Britannia} were analysed as frontiers and not as administrative Roman provinces; we included in the sample some centres that are actually located beyond the administrative boundaries but were considered part of the same border.
+Both \textit{Germania} and \textit{Britannia} were analysed as frontiers and not as administrative Roman provinces; we included in the sample some centres that are located beyond the administrative boundaries but were considered part of the same border.
 Specifically, for the German \textit{limes} we included sites from the two provinces (Superior and Inferior) while the analysis of \textit{Britannia} extended to \textit{Caledonia}.
 
-%We analysed a dataset of 2219 stamps from different centres in \textit{Britannia}. 
-%CITAR (Callender, 1965; Carreras Monfort y Funari,1998; Ayllón-Martı́ et al., 2018). 
 The 1765 stamps from \textit{Britannia} showed 986 unique codes from 46 sites. These consumption centres can be seen in Figure~\ref{britannia}.
@@
 \caption{Sites in \textit{Britannia} where Dressel 20 amphoric stamps have been found.
-Each dot is a site while its size shows the sample size of stamps found on the site.
+Each dot is a site while its size shows the sample size of stamps found in the site.
 A majority of stamps have been found in garrisons related to Hadrian's and Antonine walls}
@@
 \includegraphics[width=\linewidth]{germania}
-\caption{Sites in Germania where Dressel 20 amphoric stamps have been found.
-Each dot is a site while its size shows the sample size of stamps found on the site.
+\caption{Sites in \textit{Germania} where Dressel 20 amphoric stamps have been found.
+Each dot is a site while its size shows the sample size of stamps found in the site.
 Most sites are located along the German \textit{limes}}
@@
 
-Similarity between amphora workshops and/or consumption sites was measured by quantifying a distance (also called dissimilarity) between the set of stamp codes found for each pair of sites (i.e.
-to what extent the stamp codes found on these sites were different).
+Similarity between amphora workshops and/or consumption sites was measured by quantifying a distance (also called dissimilarity) between the set of codes of stamps found for each pair of sites (i.e.
+to what extent the stamp codes detected in these sites were different).
 The chosen measure for this study was the Morisita-Horn index \citep{morisita_measuring_1959, horn_measurement_1966}.
@@
 This visualisation suggests that most amphora workshops used a single stamp code which has not been found in any other site.
-Simultaneously the majority of stamp codes found on multiple sites are only present in two or three workshops.
+Simultaneously the majority of stamp codes found in multiple sites are only present in two or three workshops.
 The combination of these two patterns generates a map of low similarity values for the entire sample.
@@
 The pattern can be more clearly observed for military sites which are often more well researched; in this case the ones with higher similarity tend to be geographically closer.
-To summarize, the same patterns than before can be observed on the dendrogram: 1) nearby sites tend to share more similar stamps and 2) a majority of stamp codes are only found in one site.
+To summarise, the same patterns than before can be observed in the dendrogram: 1) nearby sites tend to share more similar stamps and 2) a majority of stamp codes are only found in one site.
 
@@
 
-Again, results in \textit{Germania} exhibit similar patterns with lower similarities.
+Additionally, results in \textit{Germania} exhibit similar patterns with lower similarities.
 The only difference is that a higher concentration of sites sharing similar stamps was found in militarised areas near the \textit{limes}.
@@
 This result could indicate that the workshops of the different \textit{conventus} used different stamp codes that were not shared between them.
-Within each conventus we can identify small groups of three/four workshops sharing the same amphoric stamps as seen in the dendrogam, but there are no stamp codes that may be found across several workshops.
+Within each \textit{conventus} we can identify small groups of three/four workshops sharing the same amphoric stamps as seen in the dendrogram, but there are no stamp codes that may be found across several workshops.
+
 This pattern of similarity based on spatial closeness is closely aligned with previous studies that focused on morphometric distances between amphorae made on different workshops which also showed spatial correlation and large differences between \textit{conventus}~\citep{coto-sarmiento_identifying_2018}.
@@
 On the other hand, our results do not seem to support the hypothesis that groups of workshops collectively used specific sets of stamp codes.
-There is higher similarities between nearby workshops but this could be simply explained by spatial proximity which would enable more frequent interactions.
+There are higher similarities between nearby workshops but this could be simply explained by spatial proximity which would enable more frequent interactions.
 The opposite seems valid: each workshop used different stamp codes and only shared some of them with the closest ones.
@@
 First, it is possible that each workshop operated independently and did not normally collaborate with other workshops.
-Second, as discussedd stamp similarity in closer workshops could be linked to more intense interactions.
+Second, as discussed stamp similarity in closer workshops could be linked to more intense interactions.
 Finally, the presence of different stamps in the same workshop could suggest some degree of internal organisation that would only affect nearby centres~\citep[104]{juanmorostesis}.
 For example, the stamping process could have allowed potters to sort out the different production batches and plan the shipments.
-Systematic procedures ike this one would probably be required to organise such a massive specialized production.
+Systematic procedures like this one would probably be required to organise and control such a massive specialised production.
 
 As we previously discussed some bias on the intensity of archaeological works and workshop identification may have an impact on this analysis.
-Further fieldwork could provide a clearer picture of the links between workshops if they allow to increase the sample size of stamp codes is increased.
+Further fieldwork could provide a clearer picture of the links between workshops if they allow increasing the sample size of stamp codes is increased.
+%mirar el increased que creo que está mal puesto
 
@@
 A combined exploration of the map and the dendrogram also suggests an interesting pattern: centres with higher similarity values are closer to the coast (North Sea and Celtic Sea).
-This may indicate that the Atlantic route could have played an essential role in transporting olive oil to the area of \textit{Britannia}, since sites with higher degrees of similarity were strategic points near the sea.
+This may indicate that the Atlantic route could have played an essential role in transporting olive oil to the area of \textit{Britannia} since sites with higher degrees of similarity were strategic points near the sea.
 
@@
 The areas in \textit{Britannia} with most intense military activity have some degree of overlap with civil areas.
-However, we do not detect this overlap for the the German \textit{limes} \citep{xanten2018}.
+However, we do not detect this overlap for the German \textit{limes} \citep{xanten2018}.
 This could be interpreted by parallel supply systems but it also could be caused by intensity bias as military centres have been more excavated than civil centres.
-
+%civil centres no sé si está bien
 Finally, the results suggest that there is no clear organisational pattern linking production centres and/or consumption centres beyond local spatial proximity within regions.
-As a consequence, our analysis do not indicate the presence of specific Baetican amphoric workshop consistently supplying specific regions or even provinces.
+As a consequence, our analysis does not indicate the presence of specific Baetican amphoric workshop consistently supplying specific regions or even provinces.
 This outcome suggests that stamp codes cannot be used to identify whether production centres had fixed clients and other proxies would be needed to identify this link in case it existed.
@@
 The analysis has identified similarities between amphoric stamps and spatial distribution in the case of the amphoric production within the Roman Empire.
-Beyond olive oil, similarity analysis could be used to identify trade links between other goods and contexts provided that an effective archaeological proxy is identified and collected with enough sample size.
+Beyond olive oil, similarity analysis could be used to identify trade links between other goods and contexts provided that an effective archaeological proxy is detected and collected with enough sample size.
 
-The growing amount of archaeological data we see today requires of large-scale quantitative methods to identify and explore complex patterns.
-This emrgence of new computational tools should allow us to infer multiscale dynamics and improve our interpretation of the economic processes of past societies.
+The growing amount of archaeological data we see today requires large-scale quantitative methods to identify and explore complex patterns.
+This emergence of new computational tools should allow us to infer multiscale dynamics and improve our interpretation of the economic processes of past societies.
 

### 2021-02-12 Manuscript/Manuscript.tex
final review

@@
 \usepackage{rotating}
+\usepackage{color}
 %
@@
 
+%\textcolor{red}{We have integrated all these changes and we have improved the manuscript with an extensive revision of the text by a native reader.
+So it means that we cannot track every single change because the manuscript has been entirely changed}
+
 \begin{abstract}
@@
 The analysis is performed using quantitative methods borrowed from Ecology that allows us to group sites based on the number of stamps shared between them.
-
-%The analysis explores how the quantitative approach provides a useful tool for the interpretation of the economic processes.
 
@@
 
-The structure and processes associated with this massive olive oil production have been extensively discussed over the last decades\citep{rodriguez_economioleicola_1977, duncan1982economy, Chic_hispania_1997,millet_anforas_1998}.
-However, the complexity and sheer scale of the process makes its understanding challenging,
-%se repite despite aquí dos veces
-despite several archaeological excavations and the existence of some written records discussing its structure.
-Our knowledge of this economic activity has benefited from new findings and data sources, but despite all these advances several questions remain open: How was this massive amphora production organised?
+The structure and processes associated with this massive olive oil production have been extensively discussed over the last decades~\citep{rodriguez_economioleicola_1977, duncan1982economy, Chic_hispania_1997,millet_anforas_1998}.
+However, the complexity and sheer scale of the process makes its understanding challenging, despite several archaeological excavations and the existence of some written records discussing its structure.
+Our knowledge of this economic activity has benefited from new findings and data sources, although all these advances several questions remain open: How was this massive amphora production organised?
 Can we detect the economic ties between provinces?
@@
 Recent advances in Roman studies have currently led to an environment with various interpretations on these commercial dynamics~\citep{temin_economy_2006,quantifyingwilson2009}. 
-%The application of diverse novel quantitative approaches to new accessible datasets has allowed us to improve our interpretation of the growing amount of archaeological evidence and contribute to a better understanding of the dynamics of Roman production
-The application of novel quantitative approaches has allowed us to improve our interpretation of the growing amount of archaeological evidence and contribute to a better understanding of the Roman production
-\citep{brughmans_roman_2016,orengo_seeds_2016,bayesian_2018,coto-sarmiento_identifying_2018,rubio-campillo_ecology_2018}.
+The application of novel quantitative approaches has allowed us to improve our interpretation of the growing amount of archaeological evidence and contribute to a better understanding of the dynamics of Roman production~\citep{brughmans_roman_2016,orengo_seeds_2016,bayesian_2018,coto-sarmiento_identifying_2018,rubio-campillo_ecology_2018}.
 This work aligns with these efforts and focuses on the distribution of olive oil towards Roman provinces in the borders.
@@
 Did they follow a centralised process to distribute olive oil in those provinces or each workshop organised its own trade networks?
-%repite aims dos veces  
-This paper aims to explore these questions regarding the olive oil market by identifying links between production centres and consumption provinces using similarity indexes for amphoric stamps.
+
+This paper proposes to explore these questions regarding the olive oil market by identifying links between production centres and consumption provinces using similarity indexes for amphoric stamps.
 In this way, our work aims to detect regional patterns for a specific commercial product (i.e.
@@
 In the consumption Roman provinces, our aim is to detect groups of nearby sites sharing stamps and to what extent stamps are clustered based on provinces.
-This economic connection could be identified with two ideas: a) nearby sites have more similar amphoric stamps and b) sets of stamps can only be found in specific provinces because they were shipped together.
-%find in or find on
+This economic connection could be identified with two ideas: a) nearby sites have more similar amphoric stamps and b) sets of stamps can only be detected in specific provinces because they were shipped together.
 
@@
 To do this, a population-based approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}.
-This quantitative method provides a way to measure the similarity between centre based on the amphoric stamps found in each of them.
+This quantitative method provides a way to measure the similarity between centres based on the amphoric stamps found in each of them.
 This ecological approach is based on three steps: a) to quantify stamp similarity between sites, b) to group sites based on this similarity and c) to find correlation patterns between stamp similarity and proximity.
@@
 The chronology of the Baetican workshops is typically defined between the first to the third centuries AD~\citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.
-Archaeological evidence suggests that this highly specialised production was developed across a long activity period\citep{remesal_anforas_2004}.
+Archaeological evidence suggests that this highly specialised production was developed across a long activity period~\citep{remesal_anforas_2004}.
 The process seemed to have been planned at some level as is observed by the relatively minor changes in amphoric typologies and the high degree of standardisation in the production.
 
-The importance of this olive oil trade is revealed by the massive amount of Dressel 20 type amphorae found in all the Empire and particularly in the Western provinces~\citep{dressel_ricerche_1878,millet_anforas_1998}.
-This amphora has been commonly associated with the transportation of Baetican olive oil for supplying military camps and civil settlements~\citep{berni_millet_epigrafianforica_2008}.
+The importance of this olive oil trade is revealed by the massive number of Dressel 20 type amphorae found in all the Empire and particularly in the Western provinces~\citep{dressel_ricerche_1878,millet_anforas_1998}.
+This amphora has been commonly associated with the transportation of Baetican olive oil for supplying military camps and civilian settlements~\citep{berni_millet_epigrafianforica_2008}.
 They have a recognisable globular form with a short neck and oval-shaped handles.
-Its capacity allowed to hold 70-75 litres of olive oil to be shipped and distributed throughout Empire \citep{berni_dressel_2016}.
+Its capacity allowed to hold 70-75 litres of olive oil to be shipped and distributed throughout the Empire~\citep{berni_dressel_2016}.
 
@@
 \end{figure}
-%found in find in
+
 Most stamps seemed to have been used across a long activity period which can be difficult to date with the required level of accuracy.
@@
 The stamp codes are interpreted based on three main ideas: content (olive oil), context (amphora workshop) and subject (individuals involved in the production).
-On the one hand, it seems that stamps could have been identified as the landowner of the olive groves \citep{rodriguez_economioleicola_1977}.
+On the one hand, it seems that stamps could have been identified as the landowner of the olive groves~ \citep{rodriguez_economioleicola_1977}.
 On the other hand, they could also belong to the owner of the workshops where the amphorae were made or even to a group of amphora workers~\citep{berni_millet_epigrafianforica_2008}.
@@
 From this perspective, olive oil supply seems to be particularly intense in militarised provinces that functioned as the borders of the Empire such as \textit{Britannia} and \textit{Germania}.
-The high percentage of Dressel 20 amphorae found in civilian sites close to the military garrisons also suggests the existence of diverse supply dynamics involving systematic trade between military sites to civil settlements~\citep{remesal_annona_1986, carreras_britannia_1998}.
+The high percentage of Dressel 20 amphorae found in civilian sites close to the military garrisons also suggests the existence of diverse supply dynamics involving systematic trade between military sites to civilian settlements~\citep{remesal_annona_1986, carreras_britannia_1998}.
 An explanation for the interest for exporting olive oil to the military garrison in the Roman provinces is still unclear.
@@
 High demand for olive oil in both provinces stimulated an important trade network involving different agents and the Empire itself.
-\textit{Baetica} became a strategic province for the production and distribution in both provinces as can be observed by the large volume of Dressel 20 amphora vessels found both in civil and military sites of these areas during and after the conquest.
+\textit{Baetica} became a strategic province for the production and distribution in both provinces as can be observed by the large volume of Dressel 20 amphora vessels found both in civilian and military sites of these areas during and after the conquest.
 
@@
 The local population did not consume this product and it was not produced in the region as the landscape and climate of the British Isles was not suitable for the growth of olive trees~\citep[161]{monfort_britanniaen_1998}.
-Archaeological records suggest that this absence of olive oil changed with the arrival of a large amount of Dressel 20 amphorae to \textit{Britannia} from the first century AD onwards \citep{peacock_amphorae_1991,carreras_britannia_1998}.
+Archaeological records suggest that this absence of olive oil changed with the arrival of a large number of Dressel 20 amphorae to \textit{Britannia} from the first century AD onwards \citep{peacock_amphorae_1991,carreras_britannia_1998}.
 At this moment, we detect an increase in the olive oil exports linked to the arrival of legions caused by different military campaigns~\citep[161]{monfort_britanniaen_1998}.
@@
 This date matches a change in market strategy in the Empire.
-This gradual decrease in \textit{Baetican} olive oil exports can be observed as a decrease in the amount of Dressel 20 found in excavated sites as this typology is replaced by other types of amphorae from other places and particularly the African provinces~\citep{rodriguez1991aceite,millet_anforas_1998}.
+This gradual decrease in \textit{Baetican} olive oil exports can be observed as a decrease in the number of Dressel 20 detected in excavated sites as this typology is replaced by other types of amphorae from other places and particularly the African provinces~\citep{rodriguez1991aceite,millet_anforas_1998}.
 
@@
 The distribution of Dressel 20 amphorae in the provinces of \textit{Germania Inferior} and \textit{Germania Superior} is not as well interpreted as the rest of the provinces discussed in this work~\citep[293]{remesal_baetica_2002}.
-%it is clear no se si ponerlo
+
 It is clear that there was a massive import of Dressel 20 in both provinces as is observed in archaeological excavations across the German \textit{limes}.
-However, it is still unknown to what extent trade agents participated in the distribution of Baetican olive oil in \textit{Germania} or the supply was exclusively focused on the Roman army garrisons\citep[156]{remesal_germn_2010}.
+However, it is still unknown to what extent trade agents participated in the distribution of Baetican olive oil in \textit{Germania} or the supply was exclusively focused on the Roman army garrisons~\citep[156]{remesal_germn_2010}.
 
-In any case, the introduction of olive oil in \textit{Germania} shows similar patterns to the one discussed for\textit{Britannia}.
-There is no archaeological evidence for olive oil consumption before the arrival of the Roman army to the region, while a majority of recovered Dressel 20 amphorae are located in the military sites built over the German \textit{limes} \citep{remesal_germaniaengl_2002}.
-The presence of the Roman army encouraged the exchange in the province as shown by the arrival of this product both civil settlement and military sites with a major concentration at the German \textit{limes}.
+In any case, the introduction of olive oil in \textit{Germania} shows similar patterns to the one discussed for \textit{Britannia}.
+There is no archaeological evidence for olive oil consumption before the arrival of the Roman army to the region, while a majority of recovered Dressel 20 amphorae are located in the military sites built over the German \textit{limes}~\citep{remesal_germaniaengl_2002}.
+The presence of the Roman army encouraged the exchange in the province as shown by the arrival of this product both civilian settlement and military sites with a major concentration at the German \textit{limes}.
 It seems that some Baetican centres could have been assigned to a specific province for the olive oil supply to military garrisons~\citep[125]{remesal_concierto}.
@@
 
-Some studies assumed that Roman naval technology was not suited for oceanic transport and for this reason defined riverine shipping across Rh\^one and Rhine rivers as the basic supply route for the German \textit{limes} \citep{marliere2001tonneau}.
+Some studies assumed that Roman naval technology was not suited for oceanic transport and for this reason defined riverine shipping across Rh\^one and Rhine rivers as the basic supply route for the German \textit{limes}~\citep{marliere2001tonneau}.
 However, more recent large-scale computational analysis, as well as new excavations on the Iberian Atlantic coast suggest that the Atlantic route, which was far more efficient, may have played a major role in the supply of \textit{Germania} and \textit{Britannia}~\citep{rubio-campillo_ecology_2018}.
@@
 Our dataset also contained a categorical variable defining the \textit{conventus} for each production site.
-The \textit{conventus} were administrative centres for territorial organisation under the Roman Empire \citep[58]{ozcariz_gil_administracion_2013}.
+The \textit{conventus} were administrative centres for territorial organisation under the Roman Empire~\citep[58]{ozcariz_gil_administracion_2013}.
 The creation of the \textit{conventus} in some Roman provinces was a form of control with the purpose of organising provincial administration~\citep{albertini_les_1923}.
@@
 The workshops of the three \textit{conventus} show a similar pattern on the stamp frequency distribution with the exception of two large workshops in \textit{Hispalis} (La Catria and Arva).
-In these two workshops a comparatively large amount of stamp codes were found (29 different codes on each of them).
-According to previous studies, those workshops became the most important centres of amphora production in the region \citep{rodriguez_economioleicola_1977,arva_1997}.
+In these two workshops a comparatively large number of stamp codes were found (29 different codes on each of them).
+According to previous studies, those workshops became the most important centres of amphora production in the region~\citep{rodriguez_economioleicola_1977,arva_1997}.
 
@@
 
-The analysis of stamps in consumption areas used also the CEIPAC database and filters performed for the previous dataset.
+The analysis of stamps in consumption areas also used the CEIPAC database and filters performed for the previous dataset.
 However, in this case there was a large number of sites with only one or two stamps which may introduce noise in the similarity analysis.
@@
 to what extent the stamp codes detected in these sites were different).
-The chosen measure for this study was the Morisita-Horn index \citep{morisita_measuring_1959, horn_measurement_1966}.
+The chosen measure for this study was the Morisita-Horn index~\citep{morisita_measuring_1959, horn_measurement_1966}.
 This method was applied to measure the dissimilarity between different samples of stamps while taking into account the frequency that each code is present on each site.
@@
 As we previously discussed some bias on the intensity of archaeological works and workshop identification may have an impact on this analysis.
-Further fieldwork could provide a clearer picture of the links between workshops if they allow increasing the sample size of stamp codes is increased.
-%mirar el increased que creo que está mal puesto
+Further fieldwork could provide a clearer picture of the links between workshops if these contributions increase the sample size of the stamp code dataset.
 
@@
 
-The areas in \textit{Britannia} with most intense military activity have some degree of overlap with civil areas.
+The areas in \textit{Britannia} with most intense military activity have some degree of overlap with civilian areas.
 However, we do not detect this overlap for the German \textit{limes} \citep{xanten2018}.
-This could be interpreted by parallel supply systems but it also could be caused by intensity bias as military centres have been more excavated than civil centres.
-%civil centres no sé si está bien
+This could be interpreted by parallel supply systems but it also could be caused by intensity bias as military centres have been more excavated than civilian centres.
+
 Finally, the results suggest that there is no clear organisational pattern linking production centres and/or consumption centres beyond local spatial proximity within regions.
@@
 an amphora with a given stamp was probably made on a specific workshop).
-This hypothesis is sustained by the fact that the similarity of stamps between workshops is very low and a majority of stamp codes are only found at one workshop.
+This hypothesis is sustained by the fact that the similarity of stamps between workshops is very low and a majority of stamp codes are only detected in one workshop.
 On the other hand, the presence of stamp codes in small groups of nearby workshops could suggest that the production of these workshops was somehow connected (e.g.
